### 2013-05-01 paper.tex
First commit. Need to fix toy2

\documentclass{llncs}

\usepackage{amsmath,stmaryrd,natbib}
\usepackage[usenames,dvipsnames]{color}

\newcommand{\Scribtexttt}[1]{{\texttt{#1}}}
\newcommand{\SColorize}[2]{\color{#1}{#2}}
\newcommand{\inColor}[2]{{\Scribtexttt{\SColorize{#1}{#2}}}}
\definecolor{PaleBlue}{rgb}{0.90,0.90,1.0}
\newcommand{\rackett}[1]{\inColor{black}{#1}}

\title{Concrete Pushdown Summarization}
\author{J. Ian Johnson}
\institute{Northeastern University \\
           \email{ianj@ccs.neu.edu}}

\begin{document}
\maketitle

% Outline:
% Introduction.
% High level
% PDCFA
% - Concrete
% - Abstract
% CFA2
% - Concrete
% - Abstract
% WCM machine
% - Concrete
% - Abstract
% Related Work
% Future Work
% Conclusion

\section{What to expect}
This paper lays out a common framework to talk about pushdown analysis
for higher-order languages. In particular, we extract the ``essence''
of the existing techniques into a quality of the machine semantics
that specify a language's meaning, much in the same way as
~\citet{dvanhorn:VanHorn2012Systematic}. Once the machine semantics
is in this form, simple pointwise abstraction leads to the
summarization algorithms that we see in the literature. In effect, we
give a \emph{concrete} semantics to pushdown analysis.

Summarization algorithms need not be restricted to languages with
well-bracketed calls and returns. We can adopt the technique for
higher precision in the common case but still handle difficult cases
such as first-class control. This was shown for the
\rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc})
operator in ~\citet{ianjohnson:Vardoulakis2011Pushdown}. This
impressive work illuminated the fact that we can harness the enhanced
technology of pushdown analyses in non-pushdown models of
computation. Doing this sacrifices call/return matching in the general
case, but in practice the precision is much better than the
alternative regular model. A downside of the work is that it was an
algorithmic change to the already complicated CFA2 - there was no
recipe for how to do this for one's favorite control operator. Instead
of deriving a ``pushdown'' analysis for a language with
\rackett{call/cc}, we will show a new analysis for a language with the
all the control operators in \citet{ianjohnson:Flatt:2007:ADC:1291151.1291178} (call it the
PLT machine) in order to demonstrate the applicability of this
viewpoint even in the context of complex control operators.

The remaining sections of the paper are
\begin{itemize}
\item{section \ref{sec:pdcfa}: we derive a cousin of PDCFA}
\item{section \ref{sec:cfa2}: we make additions to the previous semantics to get a direct-style CFA2 without first-class control}
\item{section \ref{sec:plt}: we give a novel analysis of the PLT machine, a core calculus of Racket's control operations.}
\end{itemize}

\section{Deriving PDCFA}
\label{sec:pdcfa}

\section{Deriving CFA2}
\label{sec:cfa2}

\section{Analysis of the PLT machine}
\label{sec:plt}

\bibliographystyle{chicago}
\bibliography{bibliography}

\end{document}

### 2013-05-06 paper.tex
More prose. Need to get down to business on the PLT machine.

@@
 \newcommand{\rackett}[1]{\inColor{black}{#1}}
+\input{local-macros}
 
-\title{Concrete Pushdown Summarization}
-\author{J. Ian Johnson}
+\title{Concrete Semantics for Pushdown Analysis:\\The Essence of Summarization}
+\author{J. Ian Johnson and David Van Horn}
 \institute{Northeastern University \\
-           \email{ianj@ccs.neu.edu}}
+           \email{\{ianj,dvanhorn\}@ccs.neu.edu}}
 
+\usepackage{hyperref}
 \begin{document}
@@
 that specify a language's meaning, much in the same way as
-~\citet{dvanhorn:VanHorn2012Systematic}. Once the machine semantics
-is in this form, simple pointwise abstraction leads to the
-summarization algorithms that we see in the literature. In effect, we
-give a \emph{concrete} semantics to pushdown analysis.
+~\citet{dvanhorn:VanHorn2012Systematic} (a method we will refer to as
+AAM: abstracting abstract machines). Once the machine semantics is in
+this form, simple pointwise abstraction leads to the summarization
+algorithms that we see in the literature. In effect, we give a
+\emph{concrete} semantics to pushdown analysis.
 
@@
 case, but in practice the precision is much better than the
-alternative regular model. A downside of the work is that it was an
+alternative regular\footnote{Regular as in regular language.} model
+that, say, AAM would provide. A downside of the work is that it was an
 algorithmic change to the already complicated CFA2 - there was no
@@
 \rackett{call/cc}, we will show a new analysis for a language with the
-all the control operators in \citet{ianjohnson:Flatt:2007:ADC:1291151.1291178} (call it the
-PLT machine) in order to demonstrate the applicability of this
-viewpoint even in the context of complex control operators.
+all the control operators in
+\citet{ianjohnson:Flatt:2007:ADC:1291151.1291178} (call it the PLT
+machine) in order to demonstrate the applicability of this viewpoint
+even in the context of complex control operators.
 
@@
 \label{sec:pdcfa}
+PDCFA is a simpler analysis than CFA2, and thus enjoys a polynomial
+complexity in the monovariant case. Its final ${\mathcal O}(n^6)$ form
+unfortunately is not a complete abstraction of an unbounded stack
+model. In this section, we show how to derive a ${\mathcal O}(n^9)$
+complete abstraction\footnote{Interestingly, the available
+  \emph{implementations} of PDCFA incorporate the fix and are
+  themselves ${\mathcal O}(n^9)$.}. These bounds are absolute worst
+case that would never ever happen in practice. The complete
+abstraction restricts value flow to fewer spurious paths, often
+leading to a faster analysis than a typical 0CFA.
+
+The magic of the method is in keeping a table of (local) continuations
+for each function $\times$ store pair. This closely mirrors the
+technique of store-allocating continuations used in AAM. Instead of
+deferring to the table (store) for the remainder of a continuation for
+each frame, we only have indirections at function call
+boundaries. Note that since AAM is about finitizing the state space,
+this step itself fits within the AAM method, since continuations
+within functions truncated at call boundaries are bounded by the
+nesting depth of those functions, thereby making the continuation
+space finite. In the monovariant case, the number of continuations is
+still linear in the size of the program.
+
+The tail of a continuation in the context of a function will contain
+the stackless context in which the function was called, in order to
+link up with the proper callsite(s). The context includes the function
+(or unique label of the function), the environment, and the store. If
+doing a flow analysis that takes advantage of the pushdown model, the
+context would also have a lattice element.
+
+Once we finitize the address space of the store, this computes a
+characteristic finite state machine that has only reachable states of
+the pushdown system the unbounded stack model embodies.
+
+A ``summary edge'' that is in the literature, in terms of pushdown
+systems, is an edge from the source of a push edge to the target of
+the matching pop edge. ``Matching'' here means there is a path through
+machine edges that don't change the stack, or through summary
+edges. There is an analogy to something more operational: summary
+edges embody memoization. Instead of following an entire path through
+a call to return a value, we simply jump from the call to the return
+(the result of the call). We are very conservative with how much of
+the context is necessary to memoize, so we use the entire heap. This
+conservativeness isn't useful for skipping any work in computing the
+flow analysis (so much so that we don't even bother consulting the
+memo table in the semantics), but it is useful for interpreting the
+results to talk about the behavior of the body of a function without
+detours to other functions. We will return to this analogy and what it
+means in the presence of first-class control in section \ref{sec:plt}.
+
+\begin{figure}
+  \begin{align*}
+    \mexpr \in \Expr &= \svar{\mvar} \mid \sapp{\mexpr}{\mexpr} \mid \slam{\mvar}{\mexpr} \\
+    \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \\
+    \mval \in \Value &::= (\slam{\mvar}{\mexpr},\ \menv) \\
+    \mkont \in \Kont &::= \kmt \mid \kar{\mexpr,\menv,\mkont} \mid \kfn{\mval,\mkont} \\
+    \menv \in \Env &= \Var \to \Addr \\
+    \mstore \in \Store &= \Addr \to \wp(\Value) \\
+    \Var &\text{ an infinite set} \\
+    \Addr &\text{ an infinite set}
+  \end{align*}
+  \caption{The CESK semantic spaces}
+\label{fig:cesk-spaces}
+\end{figure}
+
+\begin{figure}
+  \centering
+  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
+  \begin{tabular}{r|l}
+    \hline
+% variable lookup
+    $\tpl{(\svar\mvar, \menv), \mstore, \mkont}$
+    &
+    $\tpl{\mval,\mstore,\mkont}$ if $\mval \in \mstore(\menv(\mvar))$
+    \\
+% application
+    $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont}$
+    &
+    $\tpl{(\mexpri0, \menv), \mstore, \kar{\mexpri,\menv,\mkont}}$
+    \\
+% argument evaluation
+    $\tpl{\mval, \mstore, \kar{\mexpr,\menv,\mkont}}$
+    &
+    $\tpl{(\mexpr, \menv), \mstore, \kfn{\mval, \mkont}}$
+    \\
+% function call
+    $\tpl{\mval,\mstore,\kfn{\slam{\mvar}{\mexpr}, \mkont}}$
+    &
+    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\mval}, \mkont}$
+  \end{tabular}
+  \caption{The CESK machine}
+  \label{fig:base-semantics}
+\end{figure}
+
+\begin{figure}
+  \begin{align*}
+    \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \times \KTab \times \Memo \\
+    \mkont \in \Kont &::= \kmt \mid \kar{\mexpr,\menv,\mkont} \mid \kfn{\mval,\mkont} \mid \krt{(\mexpr,\menv),\mstore}\\
+    \KTab &= (\Expr \times \Env)\times \Store \to \wp(\Kont) \\
+    \Memo &= (\Expr \times \Env)\times \Store \to \wp(\Value)
+  \end{align*}
+  \caption{Extended semantic spaces}
+\label{fig:cesk-ext-spaces}
+\end{figure}
+
+\begin{figure}
+  \centering
+  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
+  \begin{tabular}{r|l}
+    \hline
+% variable lookup
+    $\tpl{(\svar\mvar, \menv), \mstore, \mkont, \mktab, \mmemo}$
+    &
+    $\tpl{\mval,\mstore,\mkont, \mktab, \mmemo}$ if $\mval \in \mstore(\menv(\mvar))$
+    \\
+% application
+    $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpri0, \menv), \mstore, \kar{\mexpri1,\menv,\mkont}, \mktab, \mmemo}$
+    \\
+% argument evaluation
+    $\tpl{\mval, \mstore, \kar{\mexpr,\menv,\mkont}, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr, \menv), \mstore, \kfn{\mval, \mkont}, \mktab, \mmemo}$
+    \\
+% function call
+    $\tpl{\mval,\mstore,\kfn{\slam{\mvar}{\mexpr}, \mkont},\mktab,\mmemo}$
+    & % actually do call,
+    $\tpl{\mpoint,
+          \mstore',
+          \krt{\mpoint, \mstore'},
+          \joinone{\mktab}{(\mpoint, \mstore')}{\mkont},
+          \mmemo}$
+    \\ % or, lookup in table
+%    & $\tpl{\mval, \mstore', \mkont,\mktab,\mmemo}$ if $\mval \in \mmemo(\mpoint, \mstore')$ \\    
+    where & $\mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr})$ \\
+          & $\mstore' = \joinone{\mstore}{\maddr}{\mval}$
+    \\
+% return
+    $\tpl{\mval, \mstore, \krt{\mpoint,\mstore'}, \mktab, \mmemo}$
+    &
+    $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{\mval}}$
+    if $\mkont \in \mktab(\mpoint, \mstore')$
+  \end{tabular}
+  \caption{The summarizing tabular stack machine}
+  \label{fig:summary-semantics}
+\end{figure}
+
+To turn this semantics into PDCFA's algorithm for constructing a Dyck
+state graph, we apply a widening operator to make $\mstore$, $\mkont$,
+and $\mmemo$ shared amongst all states. Further techniques for a
+performant implementation can be found in
+\citet{ianjohnson:oaam:2013}.
+
+\begin{align*}
+  \mastate \in \sa{State} &= \Expr \times \Env \times \Kont \\
+  \System &= \wp(\sa{State} \times \Store) \times \wp{\sa{State}} \times \Store \times (\Expr \times \Env \times \Store \to \wp(\Kont)) \times (\Expr \times \Env \times \Store \to \wp{\Value}) \\
+  {\mathcal F} : \System \to \System \\
+  {\mathcal F}(S, F, \mstore, \mktab, \mmemo) &= (S \cup S', F', \mstore', \mktab', \mmemo') \\
+  \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mstore,\mktab,\mmemo) \stepto \mstate'} \\
+                S' &= \setbuild{(\mastate',\mstore')}{\wn(\mastate', \mstore', \_, \_) \in I} \\
+                F' &= \setbuild{\mastate'}{(\mastate',\_) \in S' \setminus S} \\
+                \mstore' &= \bigsqcup\setbuild{\mstore'}{\wn(\_, \mstore',\_,\_) \in I} \\
+                \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \_, \mktab', \_) \in I} \\
+                \mmemo' &=  \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \_, \mmemo') \in I} \\
+                \wn(\tpl{\mexpr,\menv,\mkont},\mstore,\mktab,\mmemo) &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
+\end{align*}
 
@@
 
+CFA2 is the first published analysis of a higher-order programming
+language that could properly match calls and returns. Vardoulakis and
+Shivers had a clear goal of harnessing the extra information a
+pushdown model provides to produce a high-precision analysis that
+works well in practice. This resulted in more than just the
+call/return matching of the previous section, which is why we are
+showing the two separately. There are two extra features of the
+semantics:
+\begin{enumerate}
+\item{stack allocation for some bindings}
+\item{strong updates on stack frames for resolved nondeterminism}
+\end{enumerate}
+The first of these is an addition to the stackless context. There is a
+conservative pre-analysis that checks locally whether a binding will
+never escape, and classifies references as able to use the stack frame
+or not. Their criteria for a binding never escaping is that it is
+never referenced in a function that is not its binder. This can be
+extended in a language with more linguistic features; see Kranz's
+thesis about register-allocatable bindings in the Orbit Scheme
+compiler \citep{ianjohnson:kranz:thesis:1988}. A stackable reference
+is one that appears in the body of the binding function, by which we
+mean not within the body of nested function. They use the information
+that a binding never escapes to not bother allocating it in the
+heap. This has the advantage of not changing the heap, and thus leads
+to less propagation. The addition of these stack frames makes the
+analysis exponential in theory, though in practice they have proved to
+decrease running time in most cases.
+
+The second of these is to ameliorate a problem they call ``fake
+rebinding.'' That is, since bindings in the abstract represent several
+values, we don't want to reference a variable \texttt{x} in two
+different places and have it resolve to two different values. In AAM,
+a variable reference non-deterministically steps to all possible
+values associated with that variable. Here we want to say that once
+\texttt{x} is considered to stand for value $v$, then all subsequent
+references of \texttt{x} should be $v$. If they aren't, it looks as if
+\texttt{x} was rebound; it hasn't, and thus it is a ``fake
+rebinding.'' CFA2 does not step to all values on variable reference,
+but instead carries all its values around in superposition until they
+need to be observed at, say, a function call. We give a simplified
+semantics that is more along the AAM style, but CFA2's approach can
+easily be recovered from it.
+
+We show only the significantly modified rules of the semantics in
+figure \ref{fig:frame-semantics}. The other rules simply carry along
+the extra $\mframe$ component untouched.
+\begin{figure}
+  \centering
+  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
+  \begin{tabular}{r|l}
+    \hline
+% variable lookup
+    $\tpl{(\svar[\mlab]{\mvar}, \menv), \mstore, \mframe, \mkont}$
+    &
+    $\tpl{\mval,\mstore,\mframe',\mkont}$ if $(\mframe', \mval) \in \lookup(\mstore,\mframe,\menv,\mvar,\mlab)$
+    \\
+% % application
+%     $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mframe, \mkont}$
+%     &
+%     $\tpl{(\mexpri0, \menv), \mstore, \mframe, \kar{\mexpri1,\menv,\mkont}}$
+%     \\
+% % argument evaluation
+%     $\tpl{\mval, \mstore, \mframe, \kar{\mexpr,\menv,\mkont}}$
+%     &
+%     $\tpl{(\mexpr, \menv), \mstore, \mframe, \kfn{\mval, \mkont}}$
+%     \\
+% function call
+    $\tpl{\mval,\mstore,\mframe,\kfn{\slam{\mvar}{\mexpr}, \mkont}}$
+    &
+    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore', \mframe', \mkont}$
+    \\ where & $(\mstore',\mframe') = \bind(\mstore,\mframe,\maddr,\mvar,\mval)$
+  \end{tabular}
+  \caption{The CES$\xi$K machine}
+  \label{fig:frame-semantics}
+\end{figure}
+
+\begin{align*}
+  \bind(\mstore,\mframe,\maddr,\mvar,\mval) &=
+   \left\{\begin{array}{ll}
+            (\mstore, \snglm{\maddr}{\mval}) & \text{if } \mvar \text{ never escapes} \\
+            (\joinone{\mstore}{\maddr}{\mval}, \snglm{\maddr}{\mval}) & \text{otherwise } \\
+          \end{array}\right. \\
+  \lookup(\mstore,\mframe,\menv,\mvar,\mlab) &=
+    \left\{\begin{array}{ll}
+          \setbuild{(\extm{\mframe}{\menv(\mvar)}{\set{\mval}}, \mval)}
+                   {\mval \in \mframe(\menv(\mvar))}
+                   & \text{if } \mlab \text{ non-escaping} \\
+          \setbuild{(\mframe,\mval)}{\mval \in \mstore(\menv(\mvar))} & \text{otherwise}
+           \end{array}\right.
+\end{align*}
+
 \section{Analysis of the PLT machine}
 \label{sec:plt}
+
+There is contention among programming language researchers whether
+\rackett{call/cc} should be a language primitive, since it captures
+the entire stack, leading to space leaks
+\citep{ianjohnson:kiselyov:against-callcc}. Alternative control
+operators have been proposed that delimit how much of the stack to
+capture, such as Felleisen's $\%$ (read ``prompt'') and capture
+operator ${\mathcal F}$ (read ``control'')
+\citep{ianjohnson:felleisen:control:1988}, or Danvy and Filinski's
+\texttt{reset} (equivalent to $\%$) and \texttt{shift}
+\citep{ianjohnson:danvy:filinski:delim:1990}. However, the stacks
+captured by these operators always extend the stack when invoked,
+rather than replace it like those captured with
+\rackett{call/cc}. Continuations that have this extension behavior are
+called ``composable continuations.'' Stack replacement is simple to
+model in a regular analysis using the AAM approach, and Vardoulakis
+and Shivers showed it can be done in a pushdown approach (although it
+breaks the pushdown model). Stack extension, however, poses a new
+challenge for pushdown analysis, since one application of a composable
+continuation means pushing an unbounded amount of frames onto the
+stack. Vardoulakis' and Shivers' approach does not immediately apply
+in this situation, since their technique drops all knowledge of the
+stack at a continuation's invocation site; extension, however, must
+preserve it.
+
+Composable and non-composable continuations both have their uses, as
+it has been shown the two are not co-expressive when it comes to space
+complexity ~\citet{ianjohnson:dybvig:control:2007}. There is still
+merit in delimiting the stack in either case, as is argued by
+\citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}. Their paper gives a
+semantics for the coexistence of composable and non-composable
+continuations in the presence of first-class prompts,
+\rackett{dynamic-wind} (protects entry/exit of a continuation with
+pre- and post-processors), and continuation marks. We will show how to
+apply pushdown analysis techniques in the presence of all this
+complexity.
+
+The way we have been splitting continuations at function calls has
+similarities to the meta-continuation approach to modeling delimited
+control ~\citep{ianjohnson:Biernacki2006274}. We could view each
+function call as inserting a prompt, and returns as aborting to the
+nearest prompt. This view contends with tail calls, but we can
+identify tail calls easily - any call with a prompt as its
+continuation is a tail call. Tail calls are important in language
+implementations for space complexity reasons
+\citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these
+concerns are less important. The repeated popping of prompts inserted
+by what otherwise were tail-calls is synonymous with CFA2's
+``transitive summaries.''
+
+We first give a CESK semantics of the PLT machine with the allocation
+procedure as a parameter in order to first discuss the complexities
+any computable abstract semantics will have. We then present the key
+ideas to making it a table-based semantics, and how we can abstract it
+to a computable approximation.
+
+\subsection{PLT's CESK semantics}
+
+The treatment of prompts as first-class values and
+\rackett{dynamic-wind}'s interaction with non-composable continuations
+requires equality tests of dynamically generated values. Because of
+this, once we relax the freshness condition for $\alloc$, we need a
+semantics that can handle the possibility that two values aren't just
+equal or disequal - they \emph{may} be equal. This complication is
+separate from the pushdown approach we are presenting, so we give it a
+separate presentation in this section to subsequently amend with pushdown
+techniques.
+
+...TODO...
+
+\subsection{Tabular PLT semantics}
+
+Since continuations can now appear in the store, and the store is in
+the $\krt{}$ continuation, we introduced a circularity in the data
+represented in our machine - this is precisely what the AAM approach
+avoids in order to have a finite model. There is currently no known
+model of computation that allows stack capture and
+replacement/composition that has decidable reachability, so we choose
+to approximate in this case. We represent only a bounded amount of a
+continuation, after which we say that any continuation matching that
+prefix is acceptable. The amount to represent is a tunable parameter
+that we simply set to be one of the local continuations we already
+have in our model.
+
+\section{Related Work}
+
+AAM, PDCFA, CFA2, CFA2+, Reps
+
+\section{Conclusion}
 

### 2013-05-06 paper.tex
Made System definition prettier.
Fixed bad definition of S'
Slight correction to F'

@@
 To turn this semantics into PDCFA's algorithm for constructing a Dyck
-state graph, we apply a widening operator to make $\mstore$, $\mkont$,
+state graph, we apply a widening operator ${\mathcal F}$ to make $\mstore$, $\mkont$,
 and $\mmemo$ shared amongst all states. Further techniques for a
@@
   \mastate \in \sa{State} &= \Expr \times \Env \times \Kont \\
-  \System &= \wp(\sa{State} \times \Store) \times \wp{\sa{State}} \times \Store \times (\Expr \times \Env \times \Store \to \wp(\Kont)) \times (\Expr \times \Env \times \Store \to \wp{\Value}) \\
-  {\mathcal F} : \System \to \System \\
+  \System &= \wp(\sa{State} \times \Store) \times \wp(\sa{State}) \times \Store \times \KTab \times \Memo \\
+  {\mathcal F} &: \System \to \System \\
   {\mathcal F}(S, F, \mstore, \mktab, \mmemo) &= (S \cup S', F', \mstore', \mktab', \mmemo') \\
   \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mstore,\mktab,\mmemo) \stepto \mstate'} \\
-                S' &= \setbuild{(\mastate',\mstore')}{\wn(\mastate', \mstore', \_, \_) \in I} \\
-                F' &= \setbuild{\mastate'}{(\mastate',\_) \in S' \setminus S} \\
                 \mstore' &= \bigsqcup\setbuild{\mstore'}{\wn(\_, \mstore',\_,\_) \in I} \\
@@
                 \mmemo' &=  \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \_, \mmemo') \in I} \\
-                \wn(\tpl{\mexpr,\menv,\mkont},\mstore,\mktab,\mmemo) &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
+                S' &= \setbuild{(\mastate',\mstore')}{\wn(\mastate', \_, \_, \_) \in I} \\
+                F' &= \setbuild{\mastate'}{(\mastate',\_) \in S', (\mastate',\mstore') \notin S} \\
+                \wn(\tpl{(\mexpr,\menv),\mkont},\mstore,\mktab,\mmemo) &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
 \end{align*}

### 2013-05-09 paper.tex
Nixed prompt tags and just went with shift/reset.

@@
 
+\section{Introduction}
+
+Programs in higher-order languages heavily use function calls and
+method dispatch as part of their control flow. Until recently, flow
+analyses could not handle return flow precisely
+\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:earl2010pdcfa}, which
+leads to several spurious paths (and thus false positives) due to
+their pervasiveness. These works, called CFA2 and PDCFA respectively,
+use pushdown automata as their approximation's target model of
+computation. They are hence called ``pushdown analyses.''\footnote{We
+  will refer to the classic finite model analyses as ``regular
+  analyses'' after regular languages.} The precision and performance
+gains of these analyses are so beneficial that we believe new analyses
+going forward should be using pushdown techniques. CFA2 and PDCFA have
+difficult details to easily apply to an off-the-shelf semantics ---
+especially if they feature non-local control transfer that breaks the
+pushdown model.
+
+...
+
 \section{What to expect}
+
+CFA2 uses a technique called ``summarization'' from
+\citet[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the
+$\epsilon$-closure graph that PDCFA constructs. Summarization
+algorithms need not be restricted to languages with well-bracketed
+calls and returns. We can adopt the technique for higher precision in
+the common case but still handle difficult cases such as first-class
+control. This was shown for the
+\rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc})
+operator in ~\citet{ianjohnson:Vardoulakis2011Pushdown}. This
+impressive work illuminated the fact that we can harness the enhanced
+technology of pushdown analyses in non-pushdown models of
+computation. Doing this sacrifices call/return matching in the general
+case, but in practice the precision is much better than the
+alternative regular model that, say, AAM would provide.
+
 This paper lays out a common framework to talk about pushdown analysis
@@
 
-Summarization algorithms need not be restricted to languages with
-well-bracketed calls and returns. We can adopt the technique for
-higher precision in the common case but still handle difficult cases
-such as first-class control. This was shown for the
-\rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc})
-operator in ~\citet{ianjohnson:Vardoulakis2011Pushdown}. This
-impressive work illuminated the fact that we can harness the enhanced
-technology of pushdown analyses in non-pushdown models of
-computation. Doing this sacrifices call/return matching in the general
-case, but in practice the precision is much better than the
-alternative regular\footnote{Regular as in regular language.} model
-that, say, AAM would provide. A downside of the work is that it was an
-algorithmic change to the already complicated CFA2 - there was no
-recipe for how to do this for one's favorite control operator. Instead
-of deriving a ``pushdown'' analysis for a language with
-\rackett{call/cc}, we will show a new analysis for a language with the
-all the control operators in
-\citet{ianjohnson:Flatt:2007:ADC:1291151.1291178} (call it the PLT
-machine) in order to demonstrate the applicability of this viewpoint
-even in the context of complex control operators.
+A downside of the work providing \rackett{call/cc} is that it was an
+algorithmic change to the already complicated CFA2 --- there was no
+recipe for how to do this for one's favorite control operator. This
+paper seeks to do just that with an operational view of what
+summarization is, in essence. In other words, we give a concrete
+semantics to the tricks that the analyses in the literature use in the
+abstract. We also give intuitive analogies to well-established
+ideas/techniques so that you too can write a pushdown analysis for
+your language. In order to demonstrate the applicability of this
+viewpoint, we show a new analysis for a language with composable
+control. All of the semantics modeled in this paper are implemented in
+full detail in PLT Redex and available at
+
+\begin{center}
+  \url{http://github.com/ianj/concrete-summaries}
+\end{center}
 
@@
 \item{section \ref{sec:cfa2}: we make additions to the previous semantics to get a direct-style CFA2 without first-class control}
-\item{section \ref{sec:plt}: we give a novel analysis of the PLT machine, a core calculus of Racket's control operations.}
+\item{section \ref{sec:sr}: we give a novel analysis of delimited and composable first-class control.}
 \end{itemize}
@@
 \label{sec:pdcfa}
-PDCFA is a simpler analysis than CFA2, and thus enjoys a polynomial
-complexity in the monovariant case. Its final ${\mathcal O}(n^6)$ form
-unfortunately is not a complete abstraction of an unbounded stack
-model. In this section, we show how to derive a ${\mathcal O}(n^9)$
-complete abstraction\footnote{Interestingly, the available
-  \emph{implementations} of PDCFA incorporate the fix and are
-  themselves ${\mathcal O}(n^9)$.}. These bounds are absolute worst
-case that would never ever happen in practice. The complete
-abstraction restricts value flow to fewer spurious paths, often
-leading to a faster analysis than a typical 0CFA.
 
-The magic of the method is in keeping a table of (local) continuations
-for each function $\times$ store pair. This closely mirrors the
-technique of store-allocating continuations used in AAM. Instead of
-deferring to the table (store) for the remainder of a continuation for
-each frame, we only have indirections at function call
-boundaries. Note that since AAM is about finitizing the state space,
-this step itself fits within the AAM method, since continuations
-within functions truncated at call boundaries are bounded by the
-nesting depth of those functions, thereby making the continuation
-space finite. In the monovariant case, the number of continuations is
-still linear in the size of the program.
+PDCFA does not have some orthogonal semantic components that CFA2
+features to improve precision, so we start by deriving PDCFA from an
+operational semantics for the untyped lambda calculus (figure
+\ref{fig:base-semantics}). The semantic spaces for the machine follow.
 
-The tail of a continuation in the context of a function will contain
-the stackless context in which the function was called, in order to
-link up with the proper callsite(s). The context includes the function
-(or unique label of the function), the environment, and the store. If
-doing a flow analysis that takes advantage of the pushdown model, the
-context would also have a lattice element.
-
-Once we finitize the address space of the store, this computes a
-characteristic finite state machine that has only reachable states of
-the pushdown system the unbounded stack model embodies.
-
-A ``summary edge'' that is in the literature, in terms of pushdown
-systems, is an edge from the source of a push edge to the target of
-the matching pop edge. ``Matching'' here means there is a path through
-machine edges that don't change the stack, or through summary
-edges. There is an analogy to something more operational: summary
-edges embody memoization. Instead of following an entire path through
-a call to return a value, we simply jump from the call to the return
-(the result of the call). We are very conservative with how much of
-the context is necessary to memoize, so we use the entire heap. This
-conservativeness isn't useful for skipping any work in computing the
-flow analysis (so much so that we don't even bother consulting the
-memo table in the semantics), but it is useful for interpreting the
-results to talk about the behavior of the body of a function without
-detours to other functions. We will return to this analogy and what it
-means in the presence of first-class control in section \ref{sec:plt}.
-
-\begin{figure}
 \begin{align*}
@@
   \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \\
-    \mval \in \Value &::= (\slam{\mvar}{\mexpr},\ \menv) \\
+  \mval \in \Value &::= \sclo{\slam{\mvar}{\mexpr}}{\menv} \\
   \mkont \in \Kont &::= \kmt \mid \kar{\mexpr,\menv,\mkont} \mid \kfn{\mval,\mkont} \\
@@
   \mstore \in \Store &= \Addr \to \wp(\Value) \\
-    \Var &\text{ an infinite set} \\
-    \Addr &\text{ an infinite set}
+  \mvar \in \Var &\text{ an infinite set} \\
+  \maddr \in \Addr &\text{ an infinite set}
 \end{align*}
-  \caption{The CESK semantic spaces}
-\label{fig:cesk-spaces}
-\end{figure}
+
+The $\Addr$ space is what controls the precision of the model. For a
+concrete semantics, we require the allocation metafunction $\alloc :
+\State \to \Addr$ to return fresh addresses ($\maddr \notin
+\dom(\mstore)$), but any choice of address is sound.
 
@@
 % function call
-    $\tpl{\mval,\mstore,\kfn{\slam{\mvar}{\mexpr}, \mkont}}$
+    $\tpl{\mval,\mstore,\kfn{\sclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont}}$
     &
@@
 
-\begin{figure}
+Since PDCFA is a simpler analysis than CFA2, it enjoys a polynomial
+complexity in the monovariant case. Its final ${\mathcal O}(n^6)$ form
+is a naive ``step all states'' widening, which yields poor
+results. The common widening uses a worklist to only step states that
+need stepping --- the fact that the store is shared amongst all states
+makes this a semantic difference, and not just an implementation
+difference. Doing this bumps the complexity up to ${\mathcal
+  O}(n^9)$\footnote{The available \emph{implementations} of PDCFA
+  indeed do this and are themselves ${\mathcal O}(n^9)$.}. These
+bounds are (rough) absolute worst case that is highly unlikely, if
+ever, happen in practice. The worklist abstraction restricts value
+flow to fewer spurious paths, often leading to a faster analysis than
+a typical 0CFA.
+
+\paragraph{Tracking return points:} the magic of the method is in
+keeping a table of (local) continuations for each function and store
+pair. This closely mirrors the technique of store-allocating
+continuations used in AAM. Instead of deferring to the table (store)
+for the remainder of a continuation for each frame, we only have
+indirections at function call boundaries. Note that since AAM is about
+finitizing the state space, this step itself fits within the AAM
+method, since continuations within functions truncated at call
+boundaries are bounded by the nesting depth of those functions,
+thereby making the continuation space finite. In the monovariant case,
+the number of continuations is still linear in the size of the
+program.
+
+The tail of a continuation in the context of a function will contain
+the stackless context in which the function was called, in order to
+link up with the proper call-site(s). The context includes the
+function (or unique label of the function), the environment, and the
+store. Notice that this choice can be viewed as a special allocation
+strategy for continuations in the AAM viewpoint; the store is the
+important ingredient to maintaining enough precision to keep a
+pushdown abstraction. We separate the concepts in this paper for the
+exposition. The semantic spaces for the machine are modified thusly:
+
 \begin{align*}
   \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \times \KTab \times \Memo \\
-    \mkont \in \Kont &::= \kmt \mid \kar{\mexpr,\menv,\mkont} \mid \kfn{\mval,\mkont} \mid \krt{(\mexpr,\menv),\mstore}\\
-    \KTab &= (\Expr \times \Env)\times \Store \to \wp(\Kont) \\
-    \Memo &= (\Expr \times \Env)\times \Store \to \wp(\Value)
+  \mkont \in \Kont &::= \kmt
+                     \mid \kar{\mexpr,\menv,\mkont} 
+                     \mid \kfn{\mval,\mkont}
+                     \mid \krt{(\mexpr,\menv),\mstore}\\
+  \mktab \in \KTab &= (\Expr \times \Env)\times \Store \to \wp(\Kont) \\
+  \mmemo \in \Memo &= (\Expr \times \Env)\times \Store \to \wp(\Value)
 \end{align*}
-  \caption{Extended semantic spaces}
-\label{fig:cesk-ext-spaces}
-\end{figure}
+
+\paragraph{The role of summaries:} notice the additional $\Memo$
+component. A ``summary edge'' in CFA2 or equivalently an
+$\epsilon$-edge in PDCFA is an edge from the source of a push edge to
+the target of the matching pop edge. ``Matching'' here means there is
+a path through machine reductions that don't change the stack, or
+through summary edges. In our model, we only ``push'' when we call a
+function, and we only ``pop'' when we return with a value. There is an
+analogy to something more operational: summary edges embody
+memoization. Instead of following an entire path through a call to
+return a value, we simply jump from the call to the return (the result
+of the call). We will return to this analogy and what it
+means in the presence of first-class control in section \ref{sec:sr}.
 
@@
 % function call
-    $\tpl{\mval,\mstore,\kfn{\slam{\mvar}{\mexpr}, \mkont},\mktab,\mmemo}$
+    $\tpl{\mval,\mstore,\kfn{\sclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont},\mktab,\mmemo}$
     & % actually do call,
@@
           \mmemo}$
+\\
+    & $\tpl{\mval, \mstore', \mkont, \mktab, \mmemo} \text{ if } \mval \in \mmemo(\mpoint,\mstore')$
     \\ % or, lookup in table
@@
 
-To turn this semantics into PDCFA's algorithm for constructing a Dyck
+Once we finitize the address space of the store, this computes a
+characteristic finite state machine that has only reachable states of
+the pushdown system the unbounded stack model embodies.
+To turn this semantics into PDCFA's (implemented) algorithm for constructing a Dyck
 state graph, we apply a widening operator ${\mathcal F}$ to make $\mstore$, $\mkont$,
-and $\mmemo$ shared amongst all states. Further techniques for a
-performant implementation can be found in
-\citet{ianjohnson:oaam:2013}.
+and $\mmemo$ shared amongst all states.
 
 \begin{align*}
-  \mastate \in \sa{State} &= \Expr \times \Env \times \Kont \\
+  \mastate \in \sa{State} &= (\Expr \times \Env) \times \Kont \\
   \System &= \wp(\sa{State} \times \Store) \times \wp(\sa{State}) \times \Store \times \KTab \times \Memo \\
@@
                 F' &= \setbuild{\mastate'}{(\mastate',\_) \in S', (\mastate',\mstore') \notin S} \\
-                \wn(\tpl{(\mexpr,\menv),\mkont},\mstore,\mktab,\mmemo) &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
+                \wn(\tpl{(\mexpr,\menv),\mkont},\mstore,\mktab,\mmemo)
+                   &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
 \end{align*}
+
+A $\System$ embodies the states seen and at which store, $S$, the
+frontier set of states yet to analyze, $F$, the shared store for the
+frontier, $\mstore$, the continuation table $\mktab$ and the memo
+table $\mmemo$. All the states in the frontier are stepped with the
+current store, after which the next store is the least upper bound of
+all the resulting stores. The next frontier contains only states
+resulting from stepping the previous frontier that we haven't seen at
+this next store. Systematic techniques for a performant implementation
+can be found in \citet{ianjohnson:oaam:2013}.
 
@@
 CFA2 is the first published analysis of a higher-order programming
-language that could properly match calls and returns. Vardoulakis and
-Shivers had a clear goal of harnessing the extra information a
-pushdown model provides to produce a high-precision analysis that
-works well in practice. This resulted in more than just the
-call/return matching of the previous section, which is why we are
-showing the two separately. There are two extra features of the
-semantics:
+language that could properly match calls and returns. We will show
+that it fits well into the same presentation we gave for
+PDCFA. Vardoulakis and Shivers had a clear goal of harnessing the
+extra information a pushdown model provides to produce a
+high-precision analysis that works well in practice. This resulted in
+more than just the call/return matching of the previous section, which
+is why we are showing the two separately. There are two orthogonal
+features of the semantics:
 \begin{enumerate}
@@
 to less propagation. The addition of these stack frames makes the
-analysis exponential in theory, though in practice they have proved to
-decrease running time in most cases.
+analysis exponential in theory, though in practice they have been
+observed to decrease running time in most cases.
 
@@
 semantics that is more along the AAM style, but CFA2's approach can
-easily be recovered from it.
+easily be recovered from it\footnote{Our Redex model implements the fake rebinding fix after function return}.
 
@@
 % function call
-    $\tpl{\mval,\mstore,\mframe,\kfn{\slam{\mvar}{\mexpr}, \mkont}}$
+    $\tpl{\mval,\mstore,\mframe,\kfn{\sclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont}}$
     &
@@
 \end{figure}
+
+The metafunctions the semantics uses to entend and consult the heap
+and stack use implicit information from the preanalysis we described
+above:
 
@@
 
-\section{Analysis of the PLT machine}
-\label{sec:plt}
+Sidestepping fake rebinding does not need to be restricted to
+stackable references, but that is what CFA2 does. Indeed, as soon as
+the nondeterminism has been determined, we could extend the stack
+frame so any subsequent reference means what it meant previously in
+the function. Lookup would then always try the stack frame before
+falling back on the heap.
+
+The widening we applied to PDCFA also applies here in order to get
+CFA2. Both PDCFA and CFA2 use $\epsilon$-edges and summaries to
+seemingly skip work when new stacks reach a function, but the summary
+is not likely to be there, so we just drop that whole bit of
+complexity from the semantics. CFA2 also included what they called
+``transitive summaries'' to deal with tail calls. We insert an $\rt$
+frame at every function call.  This view contends with tail calls, but
+we can identify tail calls easily --- any call with an $\rt$ as its
+continuation is a tail call. Tail calls are important in language
+implementations for space complexity reasons
+\citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these
+concerns are less important. The repeated popping of $\rt$ inserted
+by what otherwise were tail-calls is synonymous with CFA2's transitive
+summaries.
+
+\section{Analysis of delimited, composable control}
+\label{sec:sr}
 
@@
 
-Composable and non-composable continuations both have their uses, as
-it has been shown the two are not co-expressive when it comes to space
-complexity ~\citet{ianjohnson:dybvig:control:2007}. There is still
-merit in delimiting the stack in either case, as is argued by
-\citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}. Their paper gives a
-semantics for the coexistence of composable and non-composable
-continuations in the presence of first-class prompts,
-\rackett{dynamic-wind} (protects entry/exit of a continuation with
-pre- and post-processors), and continuation marks. We will show how to
-apply pushdown analysis techniques in the presence of all this
-complexity.
-
 The way we have been splitting continuations at function calls has
 similarities to the meta-continuation approach to modeling delimited
-control ~\citep{ianjohnson:Biernacki2006274}. We could view each
-function call as inserting a prompt, and returns as aborting to the
-nearest prompt. This view contends with tail calls, but we can
-identify tail calls easily - any call with a prompt as its
-continuation is a tail call. Tail calls are important in language
-implementations for space complexity reasons
-\citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these
-concerns are less important. The repeated popping of prompts inserted
-by what otherwise were tail-calls is synonymous with CFA2's
-``transitive summaries.''
+control, given in figure \ref{fig:shift-reset} (adapted from
+~\citep{ianjohnson:Biernacki2006274}). We could view each function call
+as inserting a prompt, and returns as aborting to the nearest prompt.
 
-We first give a CESK semantics of the PLT machine with the allocation
-procedure as a parameter in order to first discuss the complexities
-any computable abstract semantics will have. We then present the key
-ideas to making it a table-based semantics, and how we can abstract it
-to a computable approximation.
+\begin{figure}
+  \centering
+  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
+  \begin{tabular}{r|l}
+    \hline
+% Reset
+    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont}$
+    &
+    $\tpl{(\mexpr, \menv), \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
+    \\
+% Pop prompt
+    $\tpl{\mval, \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
+    &
+    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}}$
+    \\
+% Shift
+    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont}$
+    &
+    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\mkont},\kmt,\mmkont}$
+    \\
+% continuation call
+    $\tpl{\mval,\mstore,\kfn{\mkont', \mkont}, \mmkont}$
+    &
+    $\tpl{\mval, \mstore, \mkont', \mkapp{\mkont}{\mmkont}}$
+  \end{tabular}  
+  \caption{Machine semantics for shift/reset}
+  \label{fig:shift-reset}
+\end{figure}
 
-\subsection{PLT's CESK semantics}
-
-The treatment of prompts as first-class values and
-\rackett{dynamic-wind}'s interaction with non-composable continuations
-requires equality tests of dynamically generated values. Because of
-this, once we relax the freshness condition for $\alloc$, we need a
-semantics that can handle the possibility that two values aren't just
-equal or disequal - they \emph{may} be equal. This complication is
-separate from the pushdown approach we are presenting, so we give it a
-separate presentation in this section to subsequently amend with pushdown
-techniques.
-
-...TODO...
-
-\subsection{Tabular PLT semantics}
-
-Since continuations can now appear in the store, and the store is in
-the $\krt{}$ continuation, we introduced a circularity in the data
-represented in our machine - this is precisely what the AAM approach
-avoids in order to have a finite model. There is currently no known
-model of computation that allows stack capture and
-replacement/composition that has decidable reachability, so we choose
-to approximate in this case. We represent only a bounded amount of a
-continuation, after which we say that any continuation matching that
-prefix is acceptable. The amount to represent is a tunable parameter
-that we simply set to be one of the local continuations we already
-have in our model.
+% Since continuations can now appear in the store, and the store is in
+% the $\rt$ continuation, we introduced a circularity in the data
+% represented in our machine, which can lead to a non-terminating
+% analysis --- this is precisely what the AAM approach avoids. There is
+% currently no known model of computation that allows stack capture and
+% replacement/composition that has decidable reachability, so we choose
+% to approximate in this case. We represent only a bounded amount of a
+% continuation, possibly with some gaps, after which we say that any
+% continuation matching that ``outline'' is acceptable. The amount to
+% represent is a tunable parameter; we can represent the continuation up
+% to boundedly-many function calls, or have innermost prompt tags
+% matching $\many\mtag$ for boundedly-long $\many\mtag$ (or a mix of
+% $dw$ frames and prompt tags), etc.
 
@@
 
+Tables, truncation and prompts are the solution to all life's ills.
+
 \bibliographystyle{chicago}

### 2013-05-09 paper.tex
Shift-reset semantics in paper done. Added memo tables to cfa2 and pdcfa semantics.

@@
   \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \\
-  \mval \in \Value &::= \sclo{\slam{\mvar}{\mexpr}}{\menv} \\
+  \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \\
   \mkont \in \Kont &::= \kmt \mid \kar{\mexpr,\menv,\mkont} \mid \kfn{\mval,\mkont} \\
@@
 % function call
-    $\tpl{\mval,\mstore,\kfn{\sclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont}}$
+    $\tpl{\mval,\mstore,\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont}}$
     &
@@
 % function call
-    $\tpl{\mval,\mstore,\kfn{\sclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont},\mktab,\mmemo}$
+    $\tpl{\mval,\mstore,\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont},\mktab,\mmemo}$
     & % actually do call,
@@
 % function call
-    $\tpl{\mval,\mstore,\mframe,\kfn{\sclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont}}$
+    $\tpl{\mval,\mstore,\mframe,\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont}}$
     &
@@
 
+Turning this into a table-based semantics involves making prompts a
+point of indirection, just like function calls. Memoization also gets
+a new context to consider, calling a continuation, because composable
+continuations act like functions. Figure \ref{fig:shift-reset-table0}
+has what one might naturally write using just this
+information. Unfortunately, since $\rt$ frames contain a store, and
+continuations can now appear in the store, this introduces a
+circularity that could cause the analysis to never terminate. For
+example, a loop that continually captures continuations would
+introduce unboundedly many continuation values due to the ever-growing
+store. This means the store is not finite height, and there may be no
+fixed point.
+
+\paragraph{Possible fixes to non-termination:} The simplest fix to the
+circularity would be to strip stores out of the context of the $\rt$
+frame in a captured continuation, and additionally consider contexts
+with a stripped store an abstraction of the same context with any
+store. This is a fairly brutal approximation of a captured
+continuation, so an alternative is to make this tunable with our
+precision-tuning friend, $\alloc$. Then, instead of entirely removing
+the store component of the $\rt$ context, we can replace it with an
+address of possible stores that it would approximate. The first
+brute-force approach is an instantiation of this, where the allocated
+address is always the same.
+
+\begin{figure}
+  \centering
+  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
+  \begin{tabular}{r|l}
+    \hline
+% Reset
+    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
+    &
+    $\tpl{\mpoint, \mstore, \kmt, \kprompt{\mctx}, \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}, \mmemo}$
+    \\
+    where & $\mpoint = (\mexpr, \menv)$, $\mctx = (\mpoint, \mstore)$.
+    \\
+% Pop prompt
+    $\tpl{\mval, \mstore, \kmt, \kprompt{\mctx}, \mktab, \mmemo}$
+    &
+    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}, \mktab, \joinone{\mmemo}{\mctx}{\mval}}$
+    if $(\mkont,\mmkont) \in \mktab(\mctx)$
+    \\
+% Shift
+    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont}},\kmt,\mmkont,\mktab,\mmemo}$
+    \\
+% continuation call
+    $\tpl{\mval,\mstore,\kfn{\vcomp{\mkont'}, \mkont}, \mmkont, \mktab, \mmemo}$
+    &
+    $\tpl{\mval, \mstore, \mkont', \kprompt{\mctx}, \mktab',\mmemo}$
+    \\
+    where & $\mctx = (\mkont, \mval, \mstore')$ \\
+          & $\mktab' = \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}$
+    \\
+% Memoized continuation call
+% continuation call
+    $\tpl{\mval,\mstore,\kfn{\vcomp{\mkont'}, \mkont}, \mmkont, \mktab, \mmemo}$
+    &
+    $\tpl{\mval', \mstore, \mkont, \mmkont, \mktab,\mmemo}$ if $\mval' \in \mmemo(\mkont',\mval,\mstore)$
+  \end{tabular}  
+  \caption{Faulty table-based semantics for shift/reset}
+  \label{fig:shift-reset-table0}
+\end{figure}
+
+The new indirection possibility changes $\Context$ to also include
+$\maddr$ where there was previously a $\mstore$, and $\KTab$ now
+additionally maps $\Addr$ to a set of stores. We introduce three new
+metafunctions to facilitate the new approximation:
+
+\newcommand{\replacectx}{\mathit{replace}\mstore}
+\newcommand{\addstore}{\mathit{add}\mstore}
+\begin{align*}
+  \returns(\mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
+  \returns(\mktab, ((\mexpr,\menv), \maddr)) &=
+    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mktab(\maddr)}
+  \\[2pt]
+  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \mstore), \mval) &=
+    \joinone{\mmemo}{((\mexpr,\menv), \mstore)}{\mval} \\
+  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \maddr), \mval) &=
+    \mmemo \sqcup \bigsqcup\limits_{\mstore \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore)}{\mval}
+  \\[2pt]
+  \approximate(\mkont, \mktab, \maddr) &= (\replacectx(\mkont), \addstore{\mkont}) \\
+  \text{where }
+   \replacectx(\krt{((\mexpr,\menv),\_)}) &= \krt{((\mexpr,\menv),\maddr)} \\
+   \replacectx(\kar{\mexpr,\menv,\mkont}) &= \kar{\mexpr,\menv,\replacectx(\mkont)} \\
+   \replacectx(\kfn{\mval, \mkont}) &= \kfn{\mval,\replacectx(\mkont)}
+  \\[2pt]
+   \addstore(\krt{((\mexpr,\menv),\mstore)}) &= \joinone{\mktab}{\maddr}{\mstore} \\
+   \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mktab}{\maddr}{\mktab(\maddr')} \\
+   \addstore(\kar{\mexpr,\menv,\mkont}) &= \addstore(\mkont) \\
+   \addstore(\kfn{\mval, \mkont}) &= \addstore(\mkont)
+\end{align*}
+
+\begin{figure}
+  \centering
+% Shift
+  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
+  \begin{tabular}{r|l}
+    \hline
+    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont'}},\kmt,\mmkont,\mktab',\mmemo}$
+    \\
+    where & $\maddr = \alloc(\varsigma)$, $(\mkont',\mktab') = \approximate(\mkont, \mktab, \maddr)$
+\\
+% Return
+   $\tpl{\mval, \mstore, \krt{\mctx}, \mmkont, \mktab, \mmemo}$
+   &
+   $\tpl{\mval, \mstore, \mkont, \mmkont, \mktab, \mmemo'}$
+   if $\mkont \in \returns(\mktab, \mctx)$
+   \\
+   where & $\mmemo' = \memoize(\mmemo, \mktab, \mctx, \mval)$
+  \end{tabular}
+  \caption{Fixed table-based semantics for shift/reset}
+  \label{fig:shift-reset-table1}
+\end{figure}
+
 % Since continuations can now appear in the store, and the store is in

### 2013-05-09 paper.tex
Took a stab at some related work and a conclusion.

@@
                 S' &= \setbuild{(\mastate',\mstore')}{\wn(\mastate', \_, \_, \_) \in I} \\
-                F' &= \setbuild{\mastate'}{(\mastate',\_) \in S', (\mastate',\mstore') \notin S} \\
+                F' &= \setbuild{\mastate'}{(\mastate',\_) \in S' \setminus S} \\
                 \wn(\tpl{(\mexpr,\menv),\mkont},\mstore,\mktab,\mmemo)
@@
 control, given in figure \ref{fig:shift-reset} (adapted from
-~\citep{ianjohnson:Biernacki2006274}). We could view each function call
-as inserting a prompt, and returns as aborting to the nearest prompt.
+~\citep{ianjohnson:Biernacki2006274}). We could view each function
+call as inserting a prompt, and returns as aborting to the nearest
+prompt. The changed semantic spaces for the shift/reset semantics are as
+follows:
+
+\begin{align*}
+  \mstate \in \State &::= \tpl{\mpoint, \mstore, \mkont, \mmkont} \\
+  \mpoint \in \Point &::= (\mexpr, \menv) \mid \mval \\
+  \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \mid \vcomp{\mkont} \\
+  \mmkont \in \MKont &::= \kmt \mid \mkapp{\mkont}{\mmkont}
+\end{align*}
 
@@
 a new context to consider, calling a continuation, because composable
-continuations act like functions. Figure \ref{fig:shift-reset-table0}
-has what one might naturally write using just this
-information. Unfortunately, since $\rt$ frames contain a store, and
-continuations can now appear in the store, this introduces a
-circularity that could cause the analysis to never terminate. For
-example, a loop that continually captures continuations would
-introduce unboundedly many continuation values due to the ever-growing
-store. This means the store is not finite height, and there may be no
-fixed point.
+continuations act like functions. The new semantic spaces are then
+
+\begin{align*}
+  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid (\vcomp{\mkont},\mval,\mstore) \\
+  \mkont \in \Kont &= \kmt \mid \krt{\mctx} \mid \kar{\mexpr,\menv,\mkont} \mid \kfn{\mval,\mkont} \\
+  \mmkont \in \MKont &::= \kmt \mid \kprompt{\mctx} \\
+  \mktab \in \KTab &= \Context \to (\wp(\Kont \times \MKont) \cup \Kont) \\
+  \mmemo \in \Memo &= \Context \to \wp(\Value)
+\end{align*}
+
+Figure \ref{fig:shift-reset-table0} has what one might naturally write
+using just this information. Unfortunately, since $\rt$ frames contain
+a store, and continuations can now appear in the store, this
+introduces a circularity that could cause the analysis to never
+terminate. For example, a loop that continually captures continuations
+would introduce unboundedly many continuation values due to the
+ever-growing store. This means the store is not finite height, and
+there may be no fixed point.
 
@@
 the store component of the $\rt$ context, we can replace it with an
-address of possible stores that it would approximate. The first
-brute-force approach is an instantiation of this, where the allocated
-address is always the same.
+address of possible stores that it would approximate.
 
@@
 The new indirection possibility changes $\Context$ to also include
-$\maddr$ where there was previously a $\mstore$, and $\KTab$ now
-additionally maps $\Addr$ to a set of stores. We introduce three new
-metafunctions to facilitate the new approximation:
+$\maddr$ where there was previously a $\mstore$ (though contexts for
+continuation calls remain unchanged), and $\KTab$ now additionally
+maps $\Addr$ to a set of stores. The rules that change are presented in figure \ref{fig:shift-reset-table1}.
 
-\newcommand{\replacectx}{\mathit{replace}\mstore}
-\newcommand{\addstore}{\mathit{add}\mstore}
 \begin{align*}
-  \returns(\mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
-  \returns(\mktab, ((\mexpr,\menv), \maddr)) &=
-    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mktab(\maddr)}
-  \\[2pt]
-  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \mstore), \mval) &=
-    \joinone{\mmemo}{((\mexpr,\menv), \mstore)}{\mval} \\
-  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \maddr), \mval) &=
-    \mmemo \sqcup \bigsqcup\limits_{\mstore \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore)}{\mval}
-  \\[2pt]
-  \approximate(\mkont, \mktab, \maddr) &= (\replacectx(\mkont), \addstore{\mkont}) \\
-  \text{where }
-   \replacectx(\krt{((\mexpr,\menv),\_)}) &= \krt{((\mexpr,\menv),\maddr)} \\
-   \replacectx(\kar{\mexpr,\menv,\mkont}) &= \kar{\mexpr,\menv,\replacectx(\mkont)} \\
-   \replacectx(\kfn{\mval, \mkont}) &= \kfn{\mval,\replacectx(\mkont)}
-  \\[2pt]
-   \addstore(\krt{((\mexpr,\menv),\mstore)}) &= \joinone{\mktab}{\maddr}{\mstore} \\
-   \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mktab}{\maddr}{\mktab(\maddr')} \\
-   \addstore(\kar{\mexpr,\menv,\mkont}) &= \addstore(\mkont) \\
-   \addstore(\kfn{\mval, \mkont}) &= \addstore(\mkont)
+  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid ((\mexpr,\menv),\maddr) \mid (\vcomp{\mkont},\mval,\mstore) \\
+  \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
+  \mskont \in \SKont &= \krt{\msctx} \mid \ldots \\
+  \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \mid \vcomp{\mskont} \\
+  \mmkont \in \MKont &::= \kmt \mid \kprompt{\mctx} \\
+  \mktab \in \KTab &= (\Context \to (\wp(\Kont \times \MKont) \cup \Kont)) \\
+                   &\cup (\Addr \to \wp(\Store))
 \end{align*}
@@
 
-% Since continuations can now appear in the store, and the store is in
-% the $\rt$ continuation, we introduced a circularity in the data
-% represented in our machine, which can lead to a non-terminating
-% analysis --- this is precisely what the AAM approach avoids. There is
-% currently no known model of computation that allows stack capture and
-% replacement/composition that has decidable reachability, so we choose
-% to approximate in this case. We represent only a bounded amount of a
-% continuation, possibly with some gaps, after which we say that any
-% continuation matching that ``outline'' is acceptable. The amount to
-% represent is a tunable parameter; we can represent the continuation up
-% to boundedly-many function calls, or have innermost prompt tags
-% matching $\many\mtag$ for boundedly-long $\many\mtag$ (or a mix of
-% $dw$ frames and prompt tags), etc.
+The metafunctions mentioned in the fixed semantics all deal with the
+addition of $\maddr$ to contexts. If the context in an $\rt$ frame is
+approximate, we must return to all the continuations known for all the
+contexts it approximates ($\returns$). At return boundaries, the memo
+table must add the result to all the represented contexts
+($\memoize$). At capture time, we strip the store in the $\rt$ frame
+if there is one, and replace it with an address.
+
+\newcommand{\replacectx}{\mathit{replace}\mstore}
+\newcommand{\addstore}{\mathit{add}\mstore}
+\begin{align*}
+  \returns(\mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
+  \returns(\mktab, ((\mexpr,\menv), \maddr)) &=
+    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mktab(\maddr)}
+  \\[2pt]
+  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \mstore), \mval) &=
+    \joinone{\mmemo}{((\mexpr,\menv), \mstore)}{\mval} \\
+  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \maddr), \mval) &=
+    \mmemo \sqcup \bigsqcup\limits_{\mstore \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore)}{\mval}
+  \\[2pt]
+  \approximate(\mkont, \mktab, \maddr) &= (\replacectx(\mkont), \addstore{\mkont}) \\
+  \text{where }
+   \replacectx(\kmt) &= \kmt \\
+   \replacectx(\krt{((\mexpr,\menv),\_)}) &= \krt{((\mexpr,\menv),\maddr)} \\
+   \replacectx(\kar{\mexpr,\menv,\mkont}) &= \kar{\mexpr,\menv,\replacectx(\mkont)} \\
+   \replacectx(\kfn{\mval, \mkont}) &= \kfn{\mval,\replacectx(\mkont)}
+  \\[2pt]
+   \addstore(\kmt) &= \mktab \\
+   \addstore(\krt{((\mexpr,\menv),\mstore)}) &= \joinone{\mktab}{\maddr}{\mstore} \\
+   \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mktab}{\maddr}{\mktab(\maddr')} \\
+   \addstore(\kar{\mexpr,\menv,\mkont}) &= \addstore(\mkont) \\
+   \addstore(\kfn{\mval, \mkont}) &= \addstore(\mkont)
+\end{align*}
+
+The result of this viewpoint is a sound, precise, and computable
+semantics for a ``pushdown approach'' to analyzing delimited,
+composable control. To handle non-composable control, we can simply
+remove the prompts and keep $\approximate$.
 
@@
 
-AAM, PDCFA, CFA2, CFA2+, Reps
+The immediately related work is that of PDCFA
+\citet{ianjohnson:earl2010pdcfa},
+CFA2~\citet{ianjohnson:vardoulakis-lmcs11,
+  ianjohnson:Vardoulakis2011Pushdown}, and AAM
+~\citet{dvanhorn:VanHorn2010Abstracting}, the first two of which we
+recreated in full detail. The version of CFA2 that handles
+\rackett{call/cc} does not handle composable control, and is dependent
+on a restricted CPS representation. They also had no way to tune the
+precision of their first class continuation approximation. Since
+\rackett{call/cc} can be simulated with shift/reset, this work
+supercedes theirs.
+
+Pushdown models have existed in the first-order static analysis
+literature ~\citet[Chapter
+7]{local:muchnick:jones:flow-analysis:1981}\citet{ianjohnson:reps:pushdown:1995},
+and the first-order model checking literature
+\citet{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time. The
+higher-order setting imposes additional challenges that make their
+methods difficult to adapt. The most important constraint is that we
+can't know all call-sites of a function/method before the analysis
+begins, which their methods heavily rely on.
 
@@
 
-Tables, truncation and prompts are the solution to all life's ills.
+As the programming world continues to embrace behavioral values, it
+becomes more important to import the powerful techniques pioneered by
+the first-order analysis and model checking communities. CFA2 and
+PDCFA paved the way, and in large part inspired this work. It is our
+view that systematic approaches to applying the techniques are pivotal
+to scaling them to ``real languages.'' We believe that the recipe that
+this paper set forth is a step in that direction. That is, make
+continuation tables keyed with enough context, and memoize at the
+introduced indirection points. The result in a language with
+well-bracketed control is a ``pushdown analysis'' using
+summarization. In a language without well-bracketed control, we are
+not chained to a pushdown automaton as the target of the
+approximation, so the techniques still apply and give better precision
+than regular methods.
 

### 2013-05-09 paper.tex
Added an example.

@@
 
-\usepackage{amsmath,stmaryrd,natbib}
+\usepackage{amsmath,stmaryrd,natbib,listings}
 \usepackage[usenames,dvipsnames]{color}
@@
 % Conclusion
+
+\abstract{GAH}
 
@@
 
+\paragraph{Example:} Let's consider a monovariant allocation strategy
+(the address for each binding is the variable name itself) to run the
+following example:
+\begin{lstlisting}[mathescape]
+  (let* ([id ($\lambda$ (x) x)]
+         [y (id 0)]
+         [z (id 1)])
+    ($\le$ y z))
+\end{lstlisting}
+We should expect that a pushdown analysis would predict this evaluates
+to true, and there are no loops in the program. 0CFA claims there is a
+loop from the second call of \texttt{id} to the first, and thusly
+predicts this program evaluates to true or false. PDCFA claims there
+is no loop, and depending on the implementation, that the result is
+true or false (paper), or just true (implemented). Let's take a look
+at the evaluation after binding \texttt{id} ($\mstore_0 = [\texttt{id} \mapsto \set{\vclo{\slam{\texttt{x}}{\texttt{x}}}{\bot}}]$):
+
+\begin{enumerate}
+\item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
+    \mstore_0[\texttt{x} \mapsto \set{0}]$ and
+    (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
+         $\mkont_1$ = \texttt{(letk y (let ([z (id 1)]) ($\le$ y z)) $[\texttt{id}\mapsto{\texttt{id}}]$)})
+    $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
+\item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
+\item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
+\item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \texttt{(letk z ($\le$ y z) $[\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]$)} (call this $\mkont_2$).}
+\item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
+      (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
+\item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
+\item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
+\end{enumerate}
+
+We maintained enough context to distinguish the return points of
+\texttt{id} to not rebind \texttt{y} to 1. When determining control
+flow through the expression, we consult $\mmemo$ to continue past
+function calls, so there is no confusion about a back edge from the
+second call to \texttt{id}.
+
 \section{Deriving CFA2}
@@
 summaries.
+
+\paragraph{Example:} if we use this semantics to run through the
+example of the previous section, we find that with the addition of
+stack frames, there is no confusion about \texttt{z}'s value either,
+so if the operator weren't $\le$ but instead $<$ or $+$, we could
+still constant fold that away.
 
@@
 The immediately related work is that of PDCFA
-\citet{ianjohnson:earl2010pdcfa},
-CFA2~\citet{ianjohnson:vardoulakis-lmcs11,
+\citep{ianjohnson:earl2010pdcfa},
+CFA2~\citep{ianjohnson:vardoulakis-lmcs11,
   ianjohnson:Vardoulakis2011Pushdown}, and AAM
-~\citet{dvanhorn:VanHorn2010Abstracting}, the first two of which we
+~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we
 recreated in full detail. The version of CFA2 that handles
@@
 Pushdown models have existed in the first-order static analysis
-literature ~\citet[Chapter
-7]{local:muchnick:jones:flow-analysis:1981}\citet{ianjohnson:reps:pushdown:1995},
+literature ~\citep[Chapter
+7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995},
 and the first-order model checking literature
-\citet{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time. The
+\citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time. The
 higher-order setting imposes additional challenges that make their
@@
 
-\section{Conclusion}
+\section{Conclusion and future work}
 
@@
 
+Our goal in the future is to show that this technique is even more
+widely applicable than shift and reset, but indeed to the most
+intricate control operators in production languages such as in
+\citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}. The control
+structures there are difficult to model even with AAM's techniques,
+but once we can tackle AAM, there should be a straightforward route to
+extending it with the pushdown techniques of this paper.
+
+There is also the question of clients of these analyses. Most
+obviously we would want to know where we can implement first-class
+control more efficiently. On top of optimizations, there are security
+analyses. Greater control of the stack gives us the ability to drill
+deep into context-sensitive security properties and make precise
+predictions.
+
 \bibliographystyle{chicago}

### 2013-05-09 paper.tex
Removed rundandant sentences and changed cfa2 model to fix fake rebinding at function call, not return.

@@
 
-\abstract{GAH}
+\abstract{Lorem Ipsum}
 
@@
 method dispatch as part of their control flow. Until recently, flow
-analyses could not handle return flow precisely
+analyses for higher-order languages could not handle return flow precisely
 \citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:earl2010pdcfa}, which
@@
 especially if they feature non-local control transfer that breaks the
-pushdown model.
+pushdown model. \\
 
-...
+This paper gives a common, simple framework to discuss both CFA2 and
+PDCFA along with a recipe to apply pushdown analysis techniques to an
+off-the-shelf semantics in a manner similar to
+~\citet{dvanhorn:VanHorn2010Abstracting} (a technique called
+abstracting abstract machines, or AAM). We then extend the state of
+the art by giving a novel analysis for delimited, composable control. \\
+
+Testing new ideas in analysis for improving precision or increasing
+performance can be a difficult venture. We contend that the machinery
+that we employ in the abstract should have a concrete counterpart that
+maintains the meaning of the language so that we can see their effect
+without introducing approximations which might mask correctness
+issues. Abstraction should be a simple process that is ``obviously
+correct.'' The framework we present in this paper is applicable in the
+concrete such that a pointwise abstraction leads to the pushdown
+analyses in the literature.
 
@@
 
-This paper lays out a common framework to talk about pushdown analysis
-for higher-order languages. In particular, we extract the ``essence''
-of the existing techniques into a quality of the machine semantics
-that specify a language's meaning, much in the same way as
-~\citet{dvanhorn:VanHorn2012Systematic} (a method we will refer to as
-AAM: abstracting abstract machines). Once the machine semantics is in
-this form, simple pointwise abstraction leads to the summarization
-algorithms that we see in the literature. In effect, we give a
-\emph{concrete} semantics to pushdown analysis.
-
 A downside of the work providing \rackett{call/cc} is that it was an
@@
 semantics to the tricks that the analyses in the literature use in the
-abstract. We also give intuitive analogies to well-established
-ideas/techniques so that you too can write a pushdown analysis for
-your language. In order to demonstrate the applicability of this
-viewpoint, we show a new analysis for a language with composable
-control. All of the semantics modeled in this paper are implemented in
-full detail in PLT Redex and available at
+abstract, and maintain the original meaning of the language. We also
+give intuitive analogies to well-established ideas/techniques so that
+you too can write a pushdown analysis for your language. In order to
+demonstrate the applicability of this viewpoint, we show a new
+analysis for a language with composable control. All of the semantics
+modeled in this paper are implemented in full detail in PLT Redex and
+available at
 
@@
 \end{center}
+
+There are common underpinnings of PDCFA and CFA2 that can be embodied
+as concrete machinery in the programming language semantics: breaking
+the recursive structure of continuations by indirecting them through a
+table with appropriately precise keys, and memoization.  Once the
+machine semantics is in this form, simple pointwise abstraction leads
+to the summarization algorithms that we see in the literature. In
+effect, we give a \emph{concrete} semantics to pushdown analysis.
 
@@
 PDCFA does not have some orthogonal semantic components that CFA2
-features to improve precision, so we start by deriving PDCFA from an
-operational semantics for the untyped lambda calculus (figure
-\ref{fig:base-semantics}). The semantic spaces for the machine follow.
+features to improve precision and is thus the simpler of the two. We
+start by deriving PDCFA from an operational semantics for the untyped
+lambda calculus (figure \ref{fig:base-semantics}). The semantic spaces
+for the machine follow.
 
@@
 difference. Doing this bumps the complexity up to ${\mathcal
-  O}(n^9)$\footnote{The available \emph{implementations} of PDCFA
-  indeed do this and are themselves ${\mathcal O}(n^9)$.}. These
-bounds are (rough) absolute worst case that is highly unlikely, if
-ever, happen in practice. The worklist abstraction restricts value
-flow to fewer spurious paths, often leading to a faster analysis than
-a typical 0CFA.
+  O}(n^9)$. The available \emph{implementations} of PDCFA indeed do
+this and are themselves ${\mathcal O}(n^9)$. These bounds are (rough)
+absolute worst case that is highly unlikely, if ever, happen in
+practice. The worklist abstraction restricts value flow to fewer
+spurious paths, often leading to a faster analysis than a typical
+${\mathcal O}(n^3)$ 0CFA.
 
@@
 is one that appears in the body of the binding function, by which we
-mean not within the body of nested function. They use the information
+mean not within the body of a nested function. They use the information
 that a binding never escapes to not bother allocating it in the
@@
 semantics that is more along the AAM style, but CFA2's approach can
-easily be recovered from it\footnote{Our Redex model implements the fake rebinding fix after function return}.
+easily be recovered from it\footnote{Our Redex model implements the fake rebinding strategy CFA2 itself employs}.
 
@@
 
-The widening we applied to PDCFA also applies here in order to get
-CFA2. Both PDCFA and CFA2 use $\epsilon$-edges and summaries to
-seemingly skip work when new stacks reach a function, but the summary
-is not likely to be there, so we just drop that whole bit of
-complexity from the semantics. CFA2 also included what they called
-``transitive summaries'' to deal with tail calls. We insert an $\rt$
-frame at every function call.  This view contends with tail calls, but
-we can identify tail calls easily --- any call with an $\rt$ as its
-continuation is a tail call. Tail calls are important in language
-implementations for space complexity reasons
-\citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these
-concerns are less important. The repeated popping of $\rt$ inserted
-by what otherwise were tail-calls is synonymous with CFA2's transitive
-summaries.
+CFA2 also includes what they called ``transitive summaries'' to deal
+with tail calls. We insert an $\rt$ frame at every function call.
+This view contends with tail calls, but we can identify tail calls
+easily --- any call with an $\rt$ as its continuation is a tail
+call. Tail calls are important in language implementations for space
+complexity reasons \citep{ianjohnson:clinger:tail-calls:1998}, but in
+an analysis, these concerns are less important. The repeated popping
+of $\rt$ inserted by what otherwise were tail-calls is synonymous with
+CFA2's transitive summaries.
 
@@
     \\
-    where & $\maddr = \alloc(\varsigma)$, $(\mkont',\mktab') = \approximate(\mkont, \mktab, \maddr)$
+    where & $(\mkont',\mktab') = \approximate(\mkont, \mktab, \maddr)$
 \\

### 2013-05-10 paper.tex
Incorporated some of Stephen's comments.

@@
 
-\abstract{Lorem Ipsum}
+\begin{abstract}
+  Lorem Ipsum
+\end{abstract}
 
@@
 method dispatch as part of their control flow. Until recently, flow
-analyses for higher-order languages could not handle return flow precisely
-\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:earl2010pdcfa}, which
-leads to several spurious paths (and thus false positives) due to
-their pervasiveness. These works, called CFA2 and PDCFA respectively,
-use pushdown automata as their approximation's target model of
-computation. They are hence called ``pushdown analyses.''\footnote{We
-  will refer to the classic finite model analyses as ``regular
-  analyses'' after regular languages.} The precision and performance
-gains of these analyses are so beneficial that we believe new analyses
-going forward should be using pushdown techniques. CFA2 and PDCFA have
-difficult details to easily apply to an off-the-shelf semantics ---
-especially if they feature non-local control transfer that breaks the
-pushdown model. \\
+analyses for higher-order languages could not handle return flow
+precisely \citep{ianjohnson:vardoulakis-lmcs11,
+  ianjohnson:earl2010pdcfa}, which leads to several spurious paths
+(and thus false positives) due to the pervasiveness of function/method
+calls and subsequent returns. These works, called CFA2 and PDCFA
+respectively, use pushdown automata as their approximation's target
+model of computation. They are hence called ``pushdown
+analyses.''\footnote{We will refer to the classic finite model
+  analyses as ``regular analyses'' after regular languages.} CFA2 and
+PDCFA have difficult details to easily apply to an off-the-shelf
+semantics --- especially if they feature non-local control transfer
+that breaks the pushdown model.
 
-This paper gives a common, simple framework to discuss both CFA2 and
-PDCFA along with a recipe to apply pushdown analysis techniques to an
-off-the-shelf semantics in a manner similar to
-~\citet{dvanhorn:VanHorn2010Abstracting} (a technique called
-abstracting abstract machines, or AAM). We then extend the state of
-the art by giving a novel analysis for delimited, composable control. \\
+There is a systematic process for transforming off-the-shelf
+programming language semantics into a form amenable to \emph{regular}
+analysis that has been widely applied with great success to production
+programming languages ~\citet{dvanhorn:VanHorn2010Abstracting} (a
+technique called abstracting abstract machines, or AAM). A
+contribution of this paper is a systematic process to construct
+\emph{pushdown} analyses of programming languages, due to the
+precision (and often performance) benefits.
 
@@
 analyses in the literature.
+
+This paper gives a common, simple framework to derive both CFA2 and
+PDCFA using a recipe to apply pushdown analysis techniques to
+arbitrary semantics in a manner similar to AAM. To exercise the recipe further, 
+we give a novel analysis for delimited, composable control. \\
 
@@
 give intuitive analogies to well-established ideas/techniques so that
-you too can write a pushdown analysis for your language. In order to
-demonstrate the applicability of this viewpoint, we show a new
-analysis for a language with composable control. All of the semantics
-modeled in this paper are implemented in full detail in PLT Redex and
-available at
-
-\begin{center}
-  \url{http://github.com/ianj/concrete-summaries}
-\end{center}
+the working semanticist can write a pushdown analysis for her
+language. In order to demonstrate the applicability of this viewpoint,
+we show a new analysis for a language with composable control. All of
+the semantics modeled in this paper are implemented in full detail in
+PLT Redex~\citep{ianjohnson:Felleisen:2009:SEP:1795772} and available
+online\footnote{\url{http://github.com/ianj/concrete-summaries}}.
 
 There are common underpinnings of PDCFA and CFA2 that can be embodied
-as concrete machinery in the programming language semantics: breaking
-the recursive structure of continuations by indirecting them through a
-table with appropriately precise keys, and memoization.  Once the
-machine semantics is in this form, simple pointwise abstraction leads
-to the summarization algorithms that we see in the literature. In
-effect, we give a \emph{concrete} semantics to pushdown analysis.
+as concrete machinery in the programming language semantics: 
+\begin{enumerate}
+\item{breaking the recursive structure of continuations by indirecting them through a
+table with appropriately precise keys, and}
+\item{memoization.}
+\end{enumerate}
+Once the machine semantics is in this form, simple pointwise
+abstraction leads to the summarization algorithms that we see in the
+literature.
 
@@
 \State \to \Addr$ to return fresh addresses ($\maddr \notin
-\dom(\mstore)$), but any choice of address is sound.
+\dom(\mstore)$), but any choice of address is sound. This metafunction
+along with the commitment to having no recursive datastructures is the
+key to the AAM technique. All recursion can be expressed by
+indirecting through addresses into a store --- tying Landin's knot. We
+borrow this technique on top of our own.
 
@@
 for the remainder of a continuation for each frame, we only have
-indirections at function call boundaries. Note that since AAM is about
-finitizing the state space, this step itself fits within the AAM
-method, since continuations within functions truncated at call
-boundaries are bounded by the nesting depth of those functions,
-thereby making the continuation space finite. In the monovariant case,
-the number of continuations is still linear in the size of the
-program.
+indirections at function call boundaries (see rules 4 and 5 in figure
+\ref{fig:summary-semantics} for saving and restoring
+continuations). Note that since AAM is about finitizing the state
+space, this step itself fits within the AAM method, since
+continuations within functions truncated at call boundaries are
+bounded by the nesting depth of those functions, thereby making the
+continuation space finite. In the monovariant case, the number of
+continuations is still linear in the size of the program.
 
@@
 analogy to something more operational: summary edges embody
-memoization. Instead of following an entire path through a call to
-return a value, we simply jump from the call to the return (the result
-of the call). We will return to this analogy and what it
-means in the presence of first-class control in section \ref{sec:sr}.
+memoization (see the last rule in figure
+\ref{fig:summary-semantics}). Instead of following an entire path
+through a call to return a value, we simply jump from the call to the
+return with the result of the call (the second case of rule 4). We
+will return to this analogy and what it means in the presence of
+first-class control in section \ref{sec:sr}.
 
@@
 \\
-    & $\tpl{\mval, \mstore', \mkont, \mktab, \mmemo} \text{ if } \mval \in \mmemo(\mpoint,\mstore')$
+    & or \\
+    & $\tpl{\mval_\mathit{result}, \mstore', \mkont, \mktab, \mmemo} \text{ if } \mval_\mathit{result} \in \mmemo(\mpoint,\mstore')$
     \\ % or, lookup in table
@@
 \rackett{call/cc} can be simulated with shift/reset, this work
-supercedes theirs.
+supercedes theirs. The extended version of PDCFA that does garbage
+collection \citep{dvanhorn:Earl2012Introspective} also fits into our
+model, although we did not explicitly show it. The addresses that the
+stack keeps alive can be accumulated by ``reading through'' the
+continuation table, building up the set of addresses in each portion
+of the stack that we come across.
 
@@
 begins, which their methods heavily rely on.
+
+A new temporal logic that itself understands well-bracketing of call
+and returns for stating and validating propositions, NWTL
+~\citep{ianjohnson:alur:nwtl:2007}, applies to visibly pushdown
+systems, which can precisely model progamming languages that only have
+well-bracketed call and returns. The propositional form of the logic
+has decidable satisfiability checking, and control-flow queries are
+shallow propositions to pose in the logic. However, the satisfiability
+problem is \textsc{2Exptime}-hard and the algorithm is not given in
+the on-the-fly form that a higher-order language would need to
+effectively answer control-flow queries without already knowing the
+answers. The logic itself is an exciting new frontier for expressing
+program correctness properties, and a higher-order version would be a
+welcome addition to the functional programmer's toolbelt.
+
+The trend of deriving static analyses from abstract machines does not
+stop at flow analyses. The model-checking community showed how to
+check temporal logic queries for collapsible pushdown automata (CPDA),
+or equivalently, higher-order recursion schemes, by deriving the
+checking algorithm from the Krivine
+machine~\citep{ianjohnson:Salvati:2011:KMH:2027223.2027239}. The
+expressiveness of CPDAs outweighs that of PDAs, but it is unclear how
+to adapt higher-order recursion schemes to model arbitrary programming
+language features. The method is strongly tied to the simply-typed
+call-by-name lambda calculus and depends on finite sized base-types.
 
@@
 Our goal in the future is to show that this technique is even more
-widely applicable than shift and reset, but indeed to the most
-intricate control operators in production languages such as in
+widely applicable than shift and reset. We conjecture that the same
+recipe will apply to the most intricate control operators in
+production languages such as in
 \citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}. The control
-structures there are difficult to model even with AAM's techniques,
+structures there are difficult to model even with AAM's techniques due
+to the ability to capture and compose continuations with arbitrarily many prompts,
 but once we can tackle AAM, there should be a straightforward route to
@@
 obviously we would want to know where we can implement first-class
-control more efficiently. On top of optimizations, there are security
-analyses. Greater control of the stack gives us the ability to drill
-deep into context-sensitive security properties and make precise
-predictions.
+control more efficiently. In particular, we would want to have an
+escape analysis to find first-class continuations that don't need to
+be heap-allocated, and single target abort operations that can be
+turned into long jumps after a computed stack unraveling. On top of
+optimizations, there are security analyses. Greater control of the
+stack gives us the ability to drill deep into context-sensitive
+security properties and make precise predictions such as, ``is it
+possible to send image data on the network outside the context of the
+user's permission?'' This way a user can use an image taking/sharing
+application without the application surrepticiously taking and sending
+images the user doesn't know about.
 

### 2013-05-10 paper.tex
Added short overviews of what PDCFA and CFA2 actually do before going into our own presentation.

@@
 
+\pagestyle{plain}
 \usepackage{amsmath,stmaryrd,natbib,listings}
@@
 correct.'' The framework we present in this paper is applicable in the
-concrete such that a pointwise abstraction leads to the pushdown
+concrete such that a point-wise abstraction leads to the pushdown
 analyses in the literature.
@@
 \end{enumerate}
-Once the machine semantics is in this form, simple pointwise
+Once the machine semantics is in this form, simple point-wise
 abstraction leads to the summarization algorithms that we see in the
@@
 PDCFA does not have some orthogonal semantic components that CFA2
-features to improve precision and is thus the simpler of the two. We
-start by deriving PDCFA from an operational semantics for the untyped
-lambda calculus (figure \ref{fig:base-semantics}). The semantic spaces
-for the machine follow.
+features to improve precision and is thus the simpler of the two. The
+key to their method is to notice in languages without stack-capturing
+features, the continuation is only modified a frame at a time. If the
+state space without the stack is finite (and it can be made finite
+with an approximation ala AAM), then the model falls directly into the
+realm of pushdown systems. They thus recast the problem in terms of
+pushdown systems. The whole of their machinery is then computing the
+pushdown system on-the-fly, only considering states reachable from the
+root (initial) state. To match nodes that push frames with later nodes
+following the pop of that frame, they additionally compute an
+$\epsilon$-closure graph, which keeps edges between nodes that have a
+net-zero stack change path between them. We skip the detour to
+pushdown systems altogether and show an operational understanding of
+the resulting analysis.
+
+We start by deriving PDCFA from an operational semantics for the
+untyped lambda calculus (figure \ref{fig:base-semantics}). The
+semantic spaces for the machine follow.
 
@@
 The $\Addr$ space is what controls the precision of the model. For a
-concrete semantics, we require the allocation metafunction $\alloc :
+concrete semantics, we require the allocation meta-function $\alloc :
 \State \to \Addr$ to return fresh addresses ($\maddr \notin
-\dom(\mstore)$), but any choice of address is sound. This metafunction
-along with the commitment to having no recursive datastructures is the
-key to the AAM technique. All recursion can be expressed by
-indirecting through addresses into a store --- tying Landin's knot. We
-borrow this technique on top of our own.
+\dom(\mstore)$), but any choice of address is sound. If $\alloc$ only
+uses addresses from a finite subset of $\Addr$, then the state space
+without continuations is finite, and the space of contination frames
+is finite (thus the pushdown system interpretation is apt). This
+meta-function along with the commitment to having no recursive
+data-structures is the key to the AAM technique. All recursion can be
+expressed by indirecting through addresses into a store --- tying
+Landin's knot. We borrow this technique on top of our own.
 
@@
     &
-    $\tpl{(\mexpri0, \menv), \mstore, \kar{\mexpri,\menv,\mkont}}$
+    $\tpl{(\mexpri0, \menv), \mstore, \kar{\mexpri1,\menv,\mkont}}$
     \\
@@
 
-Since PDCFA is a simpler analysis than CFA2, it enjoys a polynomial
-complexity in the monovariant case. Its final ${\mathcal O}(n^6)$ form
-is a naive ``step all states'' widening, which yields poor
-results. The common widening uses a worklist to only step states that
-need stepping --- the fact that the store is shared amongst all states
-makes this a semantic difference, and not just an implementation
-difference. Doing this bumps the complexity up to ${\mathcal
-  O}(n^9)$. The available \emph{implementations} of PDCFA indeed do
-this and are themselves ${\mathcal O}(n^9)$. These bounds are (rough)
-absolute worst case that is highly unlikely, if ever, happen in
-practice. The worklist abstraction restricts value flow to fewer
-spurious paths, often leading to a faster analysis than a typical
-${\mathcal O}(n^3)$ 0CFA.
-
 \paragraph{Tracking return points:} the magic of the method is in
-keeping a table of (local) continuations for each function and store
-pair. This closely mirrors the technique of store-allocating
-continuations used in AAM. Instead of deferring to the table (store)
-for the remainder of a continuation for each frame, we only have
-indirections at function call boundaries (see rules 4 and 5 in figure
+keeping a table of continuation segments for each function and store
+pair, where the continuation is segmented at function boundaries. This
+closely mirrors the technique of store-allocating continuations used
+in AAM. Instead of deferring to the table (store) for the remainder of
+a continuation for each frame, we only have indirections at function
+call boundaries (see rules 4 and 5 in figure
 \ref{fig:summary-semantics} for saving and restoring
@@
 The tail of a continuation in the context of a function will contain
-the stackless context in which the function was called, in order to
+the stack-less context in which the function was called, in order to
 link up with the proper call-site(s). The context includes the
@@
 store. Notice that this choice can be viewed as a special allocation
-strategy for continuations in the AAM viewpoint; the store is the
-important ingredient to maintaining enough precision to keep a
-pushdown abstraction. We separate the concepts in this paper for the
-exposition. The semantic spaces for the machine are modified thusly:
+strategy for continuations in the AAM viewpoint, but we separate the
+table of continuations from the store since continuations themselves
+contain stores in $\rt$ frames --- this leads to a recursive
+data-structure that we are trying to avoid. The store is an important
+ingredient to maintaining enough precision to keep a pushdown
+abstraction. The semantic spaces for the machine are modified thusly:
 
@@
 
+\paragraph{PDCFA paper versus implementation:} since PDCFA is a
+simpler analysis than CFA2, it enjoys a polynomial complexity in the
+monovariant case. Its final ${\mathcal O}(n^6)$ form is a naive ``step
+all states'' widening, which yields poor results as evident in the
+above example. The common widening uses a work-list to only step
+states that need stepping like the frontier-based widening we use ---
+the fact that the store is shared amongst all states makes this a
+semantic difference, and not just an implementation difference. Doing
+this bumps the complexity up to ${\mathcal O}(n^9)$. The available
+\emph{implementations} of PDCFA indeed do this and are themselves
+${\mathcal O}(n^9)$. These bounds are (rough) absolute worst case that
+is highly unlikely, if ever, to happen in practice. The work-list
+abstraction restricts value flow to fewer spurious paths, often
+leading to a faster analysis than a typical ${\mathcal O}(n^3)$ 0CFA
+~\citep{ianjohnson:vardoulakis-lmcs11,
+  dvanhorn:Earl2012Introspective}.
+
 \section{Deriving CFA2}
@@
 \item{stack allocation for some bindings}
-\item{strong updates on stack frames for resolved nondeterminism}
+\item{strong updates on stack frames for resolved non-determinism}
 \end{enumerate}
-The first of these is an addition to the stackless context. There is a
+The first of these is an addition to the stack-less context. There is a
 conservative pre-analysis that checks locally whether a binding will
@@
 semantics that is more along the AAM style, but CFA2's approach can
-easily be recovered from it\footnote{Our Redex model implements the fake rebinding strategy CFA2 itself employs}.
+easily be recovered from it\footnote{Our Redex model implements the
+  fake rebinding strategy CFA2 itself employs}.
+
+CFA2 uses a ``local semantics'' that is similar to our segmenting
+continuations at function boundaries, but gets stuck when it gets to a
+point where it would need to ``return.'' It instead appeals to an
+external, imperative algorithm for summarization to sew the function
+calls and returns together. What we show here is almost wholly the
+same, only embodied still in terms of a machine semantics and thus
+more easily reasoned about. The analogy to memoization as
+summarization guide us in the construction of the analysis in the
+coming section.
 
@@
 
-The metafunctions the semantics uses to entend and consult the heap
-and stack use implicit information from the preanalysis we described
+The meta-functions the semantics uses to extend and consult the heap
+and stack use implicit information from the pre-analysis we described
 above:
@@
 stackable references, but that is what CFA2 does. Indeed, as soon as
-the nondeterminism has been determined, we could extend the stack
+the non-determinism has been determined, we could extend the stack
 frame so any subsequent reference means what it meant previously in
-the function. Lookup would then always try the stack frame before
+the function. Look-up would then always try the stack frame before
 falling back on the heap.
@@
 
-The metafunctions mentioned in the fixed semantics all deal with the
+The meta-functions mentioned in the fixed semantics all deal with the
 addition of $\maddr$ to contexts. If the context in an $\rt$ frame is
@@
 ~\citep{ianjohnson:alur:nwtl:2007}, applies to visibly pushdown
-systems, which can precisely model progamming languages that only have
+systems, which can precisely model programming languages that only have
 well-bracketed call and returns. The propositional form of the logic
@@
 program correctness properties, and a higher-order version would be a
-welcome addition to the functional programmer's toolbelt.
+welcome addition to the functional programmer's tool belt.
 
@@
 user's permission?'' This way a user can use an image taking/sharing
-application without the application surrepticiously taking and sending
+application without the application surreptitiously taking and sending
 images the user doesn't know about.

### 2013-05-10 paper.tex
Attempted an abstract. Removed the rambling about complexity. The extra rendered space is brutal.

@@
 \begin{abstract}
-  Lorem Ipsum
+Pushdown analysis is better than finite-state analysis in precision and
+performance. Why then have we not seen total widespread adoption of these
+techniques? For one, the known techniques are technically burdened and
+difficult to understand or extend. Control structure of the programming language
+gets pulled into the model of computation, which makes extensions to
+non-pushdown control structures, such as \rackett{call/cc} or \rackett{shift} and
+\rackett{reset}, non-trivial.
+
+We show a derivational approach to abstract interpretation that yields
+transparently sound static analyses that can precisely match calls and
+returns when applied to well-known abstract machines. We show that
+adding memoization and segmenting the continuation into bounded pieces
+leads to machines that abstract to static analyses for context-free
+reachability by simply bounding the stores. This technique allows us
+to derive existing, more technically involved analyses, and a novel
+pushdown analysis for delimited, composable control.
 \end{abstract}
@@
 characteristic finite state machine that has only reachable states of
-the pushdown system the unbounded stack model embodies.
-To turn this semantics into PDCFA's (implemented) algorithm for constructing a Dyck
-state graph, we apply a widening operator ${\mathcal F}$ to make $\mstore$, $\mkont$,
-and $\mmemo$ shared amongst all states.
+the pushdown system the unbounded stack model embodies.  To turn this
+semantics into PDCFA's algorithm for constructing a Dyck state graph,
+we apply a widening operator ${\mathcal F}$ to make $\mstore$,
+$\mkont$, and $\mmemo$ shared amongst all states.
 
@@
 is no loop, and depending on the implementation, that the result is
-true or false (paper), or just true (implemented). Let's take a look
-at the evaluation after binding \texttt{id} ($\mstore_0 = [\texttt{id} \mapsto \set{\vclo{\slam{\texttt{x}}{\texttt{x}}}{\bot}}]$):
+true or false (paper), or just true (implemented)\footnote{This is
+  because the paper steps every seen state with the current store
+  every iteration, but the implementation only steps states that need stepping.}. Let's take a look at the evaluation after binding
+\texttt{id} ($\mstore_0 = [\texttt{id} \mapsto
+\set{\vclo{\slam{\texttt{x}}{\texttt{x}}}{\bot}}]$):
 
@@
 second call to \texttt{id}.
-
-\paragraph{PDCFA paper versus implementation:} since PDCFA is a
-simpler analysis than CFA2, it enjoys a polynomial complexity in the
-monovariant case. Its final ${\mathcal O}(n^6)$ form is a naive ``step
-all states'' widening, which yields poor results as evident in the
-above example. The common widening uses a work-list to only step
-states that need stepping like the frontier-based widening we use ---
-the fact that the store is shared amongst all states makes this a
-semantic difference, and not just an implementation difference. Doing
-this bumps the complexity up to ${\mathcal O}(n^9)$. The available
-\emph{implementations} of PDCFA indeed do this and are themselves
-${\mathcal O}(n^9)$. These bounds are (rough) absolute worst case that
-is highly unlikely, if ever, to happen in practice. The work-list
-abstraction restricts value flow to fewer spurious paths, often
-leading to a faster analysis than a typical ${\mathcal O}(n^3)$ 0CFA
-~\citep{ianjohnson:vardoulakis-lmcs11,
-  dvanhorn:Earl2012Introspective}.
 
@@
 external, imperative algorithm for summarization to sew the function
-calls and returns together. What we show here is almost wholly the
-same, only embodied still in terms of a machine semantics and thus
-more easily reasoned about. The analogy to memoization as
-summarization guide us in the construction of the analysis in the
-coming section.
+calls and returns together by pattern matching on the states that the
+local semantics produces. What we show here is almost wholly the same
+in character, only embodied still in terms of a machine semantics and
+thus more easily reasoned about, and can be run in the concrete.
 
@@
 call as inserting a prompt, and returns as aborting to the nearest
-prompt. The changed semantic spaces for the shift/reset semantics are as
-follows:
+prompt. Resets then insert a second-tier prompt. The changed semantic
+spaces for the shift/reset semantics are as follows:
 

### 2013-07-10 paper.tex
Reformatted for sigplan.

@@
-\documentclass{llncs}
+\documentclass{sigplanconf}
 
@@
 \title{Concrete Semantics for Pushdown Analysis:\\The Essence of Summarization}
-\author{J. Ian Johnson and David Van Horn}
-\institute{Northeastern University \\
-           \email{\{ianj,dvanhorn\}@ccs.neu.edu}}
+\authorinfo{J. Ian Johnson and David Van Horn}{Northeastern University}{\{ianj,dvanhorn\}@ccs.neu.edu}
 
@@
 pushdown system on-the-fly, only considering states reachable from the
-root (initial) state. To match nodes that push frames with later nodes
+root (initial) state. They call the main data-structure for this a
+Dyck state graph. To match nodes that push frames with later nodes
 following the pop of that frame, they additionally compute an
@@
 through a call to return a value, we simply jump from the call to the
-return with the result of the call (the second case of rule 4). We
-will return to this analogy and what it means in the presence of
-first-class control in section \ref{sec:sr}.
+return with the result of the call (the second case of rule 4).
 
@@
           \krt{\mpoint, \mstore'},
-          \joinone{\mktab}{(\mpoint, \mstore')}{\mkont},
+          \mktab',
           \mmemo}$
@@
     & or \\
-    & $\tpl{\mval_\mathit{result}, \mstore', \mkont, \mktab, \mmemo} \text{ if } \mval_\mathit{result} \in \mmemo(\mpoint,\mstore')$
+    & $\tpl{\mval_\mathit{result},
+            \mstore',
+            \mkont,
+            \mktab',
+            \mmemo}$
+    \\ & \quad if $\mval_\mathit{result} \in \mmemo(\mpoint,\mstore')$
     \\ % or, lookup in table
@@
     where & $\mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr})$ \\
-          & $\mstore' = \joinone{\mstore}{\maddr}{\mval}$
+          & $\mstore' = \joinone{\mstore}{\maddr}{\mval}$ \\
+          & $\mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{\mkont}$
     \\
@@
     $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{\mval}}$
-    if $\mkont \in \mktab(\mpoint, \mstore')$
+    \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
   \end{tabular}
@@
 
-Once we finitize the address space of the store, this computes a
-characteristic finite state machine that has only reachable states of
-the pushdown system the unbounded stack model embodies.  To turn this
-semantics into PDCFA's algorithm for constructing a Dyck state graph,
-we apply a widening operator ${\mathcal F}$ to make $\mstore$,
-$\mkont$, and $\mmemo$ shared amongst all states.
+To turn this semantics into PDCFA's algorithm for constructing a Dyck
+state graph, given the appropriate $\alloc$, we apply a widening
+operator to make $\mstore$, $\mkont$, and $\mmemo$ shared amongst all
+states (meaning states then become represented without these
+compontents), in what is then called a $\System$.  Since these
+components are shared, they are always the least upper bound of any
+respective component that the semantics produces (the intermediate set
+$I$) in order to stay sound. We use a meta-function $\wn$, ``wide to
+narrow'', to shift between the two representations of states.
 
@@
   \mastate \in \sa{State} &= (\Expr \times \Env) \times \Kont \\
-  \System &= \wp(\sa{State} \times \Store) \times \wp(\sa{State}) \times \Store \times \KTab \times \Memo \\
+  \System &= \wp(\sa{State} \times \Store) \times \wp(\sa{State}) \\
+          &\phantom{=} \times \Store \times \KTab \times \Memo \\
   {\mathcal F} &: \System \to \System \\
@@
 this next store. Systematic techniques for a performant implementation
-can be found in \citet{ianjohnson:oaam:2013}.
+can be found in \citet{DBLP:journals/corr/abs-1211-3722}.
 
@@
 \end{lstlisting}
+Suppose we extend our semantics to allow numbers, numeric primitives
+and \texttt{let}. The continuation frame for a \texttt{let} contains
+the identifier to bind to the resulting value, along with the body of
+the \texttt{let} with its environment. Call the constructor of this
+frame $\mathbf{lt}$.
+
 We should expect that a pushdown analysis would predict this evaluates
-to true, and there are no loops in the program. 0CFA claims there is a
+to true, and there are no loops in the program. 0CFA ~\citep{ianjohnson:Shivers:1991:CFA} claims there is a
 loop from the second call of \texttt{id} to the first, and thusly
@@
     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
-         $\mkont_1$ = \texttt{(letk y (let ([z (id 1)]) ($\le$ y z)) $[\texttt{id}\mapsto{\texttt{id}}]$)})
+         $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
@@
 \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
-\item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \texttt{(letk z ($\le$ y z) $[\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]$)} (call this $\mkont_2$).}
+\item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
+      $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
+      (call this $\mkont_2$).}
 \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
@@
 \begin{enumerate}
-\item{stack allocation for some bindings}
+\item{stack allocation for some bindings in an additional $\mframe \in \Store$}
 \item{strong updates on stack frames for resolved non-determinism}
 \end{enumerate}
-The first of these is an addition to the stack-less context. There is a
-conservative pre-analysis that checks locally whether a binding will
-never escape, and classifies references as able to use the stack frame
-or not. Their criteria for a binding never escaping is that it is
-never referenced in a function that is not its binder. This can be
-extended in a language with more linguistic features; see Kranz's
-thesis about register-allocatable bindings in the Orbit Scheme
-compiler \citep{ianjohnson:kranz:thesis:1988}. A stackable reference
-is one that appears in the body of the binding function, by which we
-mean not within the body of a nested function. They use the information
-that a binding never escapes to not bother allocating it in the
-heap. This has the advantage of not changing the heap, and thus leads
-to less propagation. The addition of these stack frames makes the
-analysis exponential in theory, though in practice they have been
-observed to decrease running time in most cases.
+The first of these is an addition to the stack-less context. There is
+a conservative pre-analysis that checks locally whether a binding will
+never escape, and classifies references (labeled with a distinguishing
+$\mlab$ from an arbitrary space of labels) as able to use the stack frame or not. Their criteria for a
+binding never escaping is that it is never referenced in a function
+that is not its binder. This can be extended in a language with more
+linguistic features; see Kranz's thesis about register-allocatable
+bindings in the Orbit Scheme compiler
+\citep{ianjohnson:kranz:thesis:1988}. A stackable reference is one
+that appears in the body of the binding function, by which we mean not
+within the body of a nested function. They use the information that a
+binding never escapes to not bother allocating it in the heap. This
+has the advantage of not changing the heap, and thus leads to less
+propagation. The addition of these stack frames makes the analysis
+exponential in theory, though in practice they have been observed to
+decrease running time in most cases.
 
@@
   \bind(\mstore,\mframe,\maddr,\mvar,\mval) &=
-   \left\{\begin{array}{ll}
-            (\mstore, \snglm{\maddr}{\mval}) & \text{if } \mvar \text{ never escapes} \\
-            (\joinone{\mstore}{\maddr}{\mval}, \snglm{\maddr}{\mval}) & \text{otherwise } \\
+   \left\{\begin{array}{l}
+            (\mstore, \snglm{\maddr}{\mval}) \text{ if } \mvar \text{ never escapes} \\
+            (\joinone{\mstore}{\maddr}{\mval}, \snglm{\maddr}{\mval}) \text{ otherwise} \\
           \end{array}\right. \\
   \lookup(\mstore,\mframe,\menv,\mvar,\mlab) &=
-    \left\{\begin{array}{ll}
+    \left\{\begin{array}{l}
           \setbuild{(\extm{\mframe}{\menv(\mvar)}{\set{\mval}}, \mval)}
                    {\mval \in \mframe(\menv(\mvar))}
-                   & \text{if } \mlab \text{ non-escaping} \\
-          \setbuild{(\mframe,\mval)}{\mval \in \mstore(\menv(\mvar))} & \text{otherwise}
+                   \\ \qquad \text{if } \mlab \text{ non-escaping} \\
+          \setbuild{(\mframe,\mval)}{\mval \in \mstore(\menv(\mvar))} \text{ otherwise}
            \end{array}\right.
@@
 operators have been proposed that delimit how much of the stack to
-capture, such as Felleisen's $\%$ (read ``prompt'') and capture
-operator ${\mathcal F}$ (read ``control'')
-\citep{ianjohnson:felleisen:control:1988}, or Danvy and Filinski's
-\texttt{reset} (equivalent to $\%$) and \texttt{shift}
+capture, such as $\%$ (read ``prompt'') and capture operator
+${\mathcal F}$ (read ``control'')
+\citep{ianjohnson:felleisen:control:1988}, or \texttt{reset}
+(equivalent to $\%$) and \texttt{shift}
 \citep{ianjohnson:danvy:filinski:delim:1990}. However, the stacks
@@
 \rackett{call/cc}. Continuations that have this extension behavior are
-called ``composable continuations.'' Stack replacement is simple to
-model in a regular analysis using the AAM approach, and Vardoulakis
+called ``composable continuations.'' Stack replacement is easily
+modeled in a regular analysis using the AAM approach, and Vardoulakis
 and Shivers showed it can be done in a pushdown approach (although it
@@
     &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\mkont},\kmt,\mmkont}$
+    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore',\kmt,\mmkont}$
+    \\ & where $\mstore' = \joinone{\mstore}{\maddr}{\mkont}$
     \\
@@
   \mmkont \in \MKont &::= \kmt \mid \kprompt{\mctx} \\
-  \mktab \in \KTab &= \Context \to (\wp(\Kont \times \MKont) \cup \Kont) \\
+  \mathit{KTabRange} &= \wp(\Kont \times \MKont) \cup \wp(\Kont) \\
+  \mktab \in \KTab &= \Context \to \mathit{KTabRange} \\
   \mmemo \in \Memo &= \Context \to \wp(\Value)
@@
 
-\begin{figure}
+\begin{figure*}
   \centering
@@
     \\
-    where & $\mctx = (\mkont, \mval, \mstore')$ \\
+    where & $\mctx = (\mkont', \mval, \mstore)$ \\
           & $\mktab' = \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}$
@@
   \label{fig:shift-reset-table0}
-\end{figure}
+\end{figure*}
 
@@
 \begin{align*}
-  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid ((\mexpr,\menv),\maddr) \mid (\vcomp{\mkont},\mval,\mstore) \\
+  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid ((\mexpr,\menv),\maddr)
+                     \\ & \mid (\vcomp{\mkont},\mval,\mstore) \\
   \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
@@
   \mmkont \in \MKont &::= \kmt \mid \kprompt{\mctx} \\
-  \mktab \in \KTab &= (\Context \to (\wp(\Kont \times \MKont) \cup \Kont)) \\
+  \mktab \in \KTab &= (\Context \to \mathit{KTabRange}) \\
                    &\cup (\Addr \to \wp(\Store))
@@
   \label{fig:shift-reset-table1}
-\end{figure}
+\end{figure*}
 
@@
 approximate, we must return to all the continuations known for all the
-contexts it approximates ($\returns$). At return boundaries, the memo
-table must add the result to all the represented contexts
-($\memoize$). At capture time, we strip the store in the $\rt$ frame
-if there is one, and replace it with an address.
+contexts it approximates:
+\begin{align*}
+  \returns(\mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
+  \returns(\mktab, ((\mexpr,\menv), \maddr)) &=
+    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mktab(\maddr)}
+\end{align*}
+
+At return boundaries, the memo
+table must add the result to all the represented contexts:=
+\begin{align*}
+  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \mstore), \mval) &=
+  \joinone{\mmemo}{((\mexpr,\menv), \mstore)}{\mval} \\
+  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \maddr), \mval) &=
+  \mmemo \sqcup \bigsqcup\limits_{\mstore \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore)}{\mval}
+\end{align*}
+At capture time, we strip the store in the $\rt$ frame
+if there is one, and replace it with an address:
 
@@
 \begin{align*}
-  \returns(\mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
-  \returns(\mktab, ((\mexpr,\menv), \maddr)) &=
-    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mktab(\maddr)}
-  \\[2pt]
-  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \mstore), \mval) &=
-    \joinone{\mmemo}{((\mexpr,\menv), \mstore)}{\mval} \\
-  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \maddr), \mval) &=
-    \mmemo \sqcup \bigsqcup\limits_{\mstore \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore)}{\mval}
-  \\[2pt]
-  \approximate(\mkont, \mktab, \maddr) &= (\replacectx(\mkont), \addstore{\mkont}) \\
+  \approximate(\mkont, \mktab, \maddr) &= (\replacectx(\mkont), \addstore(\mkont)) \\
   \text{where }
@@
    \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mktab}{\maddr}{\mktab(\maddr')} \\
-   \addstore(\kar{\mexpr,\menv,\mkont}) &= \addstore(\mkont) \\
-   \addstore(\kfn{\mval, \mkont}) &= \addstore(\mkont)
+   \addstore(\kar{\mexpr,\menv,\mkont}) &= \addstore(\kfn{\mval, \mkont}) = \addstore(\mkont)
 \end{align*}
@@
 stack gives us the ability to drill deep into context-sensitive
-security properties and make precise predictions such as, ``is it
-possible to send image data on the network outside the context of the
-user's permission?'' This way a user can use an image taking/sharing
-application without the application surreptitiously taking and sending
-images the user doesn't know about.
+security properties and make precise predictions.
 
-\bibliographystyle{chicago}
+\bibliographystyle{plainnat}
 \bibliography{bibliography}

### 2013-09-30 abstract.tex
Cleaned up repo. Added semantic fix and conjectures for correctness.

--- 2013-05-10 paper.tex
@@
-\documentclass{llncs}
-
-\pagestyle{plain}
-\usepackage{amsmath,stmaryrd,natbib,listings}
-\usepackage[usenames,dvipsnames]{color}
-
-\newcommand{\Scribtexttt}[1]{{\texttt{#1}}}
-\newcommand{\SColorize}[2]{\color{#1}{#2}}
-\newcommand{\inColor}[2]{{\Scribtexttt{\SColorize{#1}{#2}}}}
-\definecolor{PaleBlue}{rgb}{0.90,0.90,1.0}
-\newcommand{\rackett}[1]{\inColor{black}{#1}}
-\input{local-macros}
-
-\title{Concrete Semantics for Pushdown Analysis:\\The Essence of Summarization}
-\author{J. Ian Johnson and David Van Horn}
-\institute{Northeastern University \\
-           \email{\{ianj,dvanhorn\}@ccs.neu.edu}}
-
-\usepackage{hyperref}
-\begin{document}
-\maketitle
-
-% Outline:
-% Introduction.
-% High level
-% PDCFA
-% - Concrete
-% - Abstract
-% CFA2
-% - Concrete
-% - Abstract
-% WCM machine
-% - Concrete
-% - Abstract
-% Related Work
-% Future Work
-% Conclusion
-
-\begin{abstract}
 Pushdown analysis is better than finite-state analysis in precision and
@@
 pushdown analysis for delimited, composable control.
-\end{abstract}
-
-\section{Introduction}
-
-Programs in higher-order languages heavily use function calls and
-method dispatch as part of their control flow. Until recently, flow
-analyses for higher-order languages could not handle return flow
-precisely \citep{ianjohnson:vardoulakis-lmcs11,
-  ianjohnson:earl2010pdcfa}, which leads to several spurious paths
-(and thus false positives) due to the pervasiveness of function/method
-calls and subsequent returns. These works, called CFA2 and PDCFA
-respectively, use pushdown automata as their approximation's target
-model of computation. They are hence called ``pushdown
-analyses.''\footnote{We will refer to the classic finite model
-  analyses as ``regular analyses'' after regular languages.} CFA2 and
-PDCFA have difficult details to easily apply to an off-the-shelf
-semantics --- especially if they feature non-local control transfer
-that breaks the pushdown model.
-
-There is a systematic process for transforming off-the-shelf
-programming language semantics into a form amenable to \emph{regular}
-analysis that has been widely applied with great success to production
-programming languages ~\citet{dvanhorn:VanHorn2010Abstracting} (a
-technique called abstracting abstract machines, or AAM). A
-contribution of this paper is a systematic process to construct
-\emph{pushdown} analyses of programming languages, due to the
-precision (and often performance) benefits.
-
-Testing new ideas in analysis for improving precision or increasing
-performance can be a difficult venture. We contend that the machinery
-that we employ in the abstract should have a concrete counterpart that
-maintains the meaning of the language so that we can see their effect
-without introducing approximations which might mask correctness
-issues. Abstraction should be a simple process that is ``obviously
-correct.'' The framework we present in this paper is applicable in the
-concrete such that a point-wise abstraction leads to the pushdown
-analyses in the literature.
-
-This paper gives a common, simple framework to derive both CFA2 and
-PDCFA using a recipe to apply pushdown analysis techniques to
-arbitrary semantics in a manner similar to AAM. To exercise the recipe further, 
-we give a novel analysis for delimited, composable control. \\
-
-\section{What to expect}
-
-CFA2 uses a technique called ``summarization'' from
-\citet[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the
-$\epsilon$-closure graph that PDCFA constructs. Summarization
-algorithms need not be restricted to languages with well-bracketed
-calls and returns. We can adopt the technique for higher precision in
-the common case but still handle difficult cases such as first-class
-control. This was shown for the
-\rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc})
-operator in ~\citet{ianjohnson:Vardoulakis2011Pushdown}. This
-impressive work illuminated the fact that we can harness the enhanced
-technology of pushdown analyses in non-pushdown models of
-computation. Doing this sacrifices call/return matching in the general
-case, but in practice the precision is much better than the
-alternative regular model that, say, AAM would provide.
-
-A downside of the work providing \rackett{call/cc} is that it was an
-algorithmic change to the already complicated CFA2 --- there was no
-recipe for how to do this for one's favorite control operator. This
-paper seeks to do just that with an operational view of what
-summarization is, in essence. In other words, we give a concrete
-semantics to the tricks that the analyses in the literature use in the
-abstract, and maintain the original meaning of the language. We also
-give intuitive analogies to well-established ideas/techniques so that
-the working semanticist can write a pushdown analysis for her
-language. In order to demonstrate the applicability of this viewpoint,
-we show a new analysis for a language with composable control. All of
-the semantics modeled in this paper are implemented in full detail in
-PLT Redex~\citep{ianjohnson:Felleisen:2009:SEP:1795772} and available
-online\footnote{\url{http://github.com/ianj/concrete-summaries}}.
-
-There are common underpinnings of PDCFA and CFA2 that can be embodied
-as concrete machinery in the programming language semantics: 
-\begin{enumerate}
-\item{breaking the recursive structure of continuations by indirecting them through a
-table with appropriately precise keys, and}
-\item{memoization.}
-\end{enumerate}
-Once the machine semantics is in this form, simple point-wise
-abstraction leads to the summarization algorithms that we see in the
-literature.
-
-The remaining sections of the paper are
-\begin{itemize}
-\item{section \ref{sec:pdcfa}: we derive a cousin of PDCFA}
-\item{section \ref{sec:cfa2}: we make additions to the previous semantics to get a direct-style CFA2 without first-class control}
-\item{section \ref{sec:sr}: we give a novel analysis of delimited and composable first-class control.}
-\end{itemize}
-
-\section{Deriving PDCFA}
-\label{sec:pdcfa}
-
-PDCFA does not have some orthogonal semantic components that CFA2
-features to improve precision and is thus the simpler of the two. The
-key to their method is to notice in languages without stack-capturing
-features, the continuation is only modified a frame at a time. If the
-state space without the stack is finite (and it can be made finite
-with an approximation ala AAM), then the model falls directly into the
-realm of pushdown systems. They thus recast the problem in terms of
-pushdown systems. The whole of their machinery is then computing the
-pushdown system on-the-fly, only considering states reachable from the
-root (initial) state. To match nodes that push frames with later nodes
-following the pop of that frame, they additionally compute an
-$\epsilon$-closure graph, which keeps edges between nodes that have a
-net-zero stack change path between them. We skip the detour to
-pushdown systems altogether and show an operational understanding of
-the resulting analysis.
-
-We start by deriving PDCFA from an operational semantics for the
-untyped lambda calculus (figure \ref{fig:base-semantics}). The
-semantic spaces for the machine follow.
-
-\begin{align*}
-  \mexpr \in \Expr &= \svar{\mvar} \mid \sapp{\mexpr}{\mexpr} \mid \slam{\mvar}{\mexpr} \\
-  \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \\
-  \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \\
-  \mkont \in \Kont &::= \kmt \mid \kar{\mexpr,\menv,\mkont} \mid \kfn{\mval,\mkont} \\
-  \menv \in \Env &= \Var \to \Addr \\
-  \mstore \in \Store &= \Addr \to \wp(\Value) \\
-  \mvar \in \Var &\text{ an infinite set} \\
-  \maddr \in \Addr &\text{ an infinite set}
-\end{align*}
-
-The $\Addr$ space is what controls the precision of the model. For a
-concrete semantics, we require the allocation meta-function $\alloc :
-\State \to \Addr$ to return fresh addresses ($\maddr \notin
-\dom(\mstore)$), but any choice of address is sound. If $\alloc$ only
-uses addresses from a finite subset of $\Addr$, then the state space
-without continuations is finite, and the space of contination frames
-is finite (thus the pushdown system interpretation is apt). This
-meta-function along with the commitment to having no recursive
-data-structures is the key to the AAM technique. All recursion can be
-expressed by indirecting through addresses into a store --- tying
-Landin's knot. We borrow this technique on top of our own.
-
-\begin{figure}
-  \centering
-  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
-  \begin{tabular}{r|l}
-    \hline
-% variable lookup
-    $\tpl{(\svar\mvar, \menv), \mstore, \mkont}$
-    &
-    $\tpl{\mval,\mstore,\mkont}$ if $\mval \in \mstore(\menv(\mvar))$
-    \\
-% application
-    $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont}$
-    &
-    $\tpl{(\mexpri0, \menv), \mstore, \kar{\mexpri1,\menv,\mkont}}$
-    \\
-% argument evaluation
-    $\tpl{\mval, \mstore, \kar{\mexpr,\menv,\mkont}}$
-    &
-    $\tpl{(\mexpr, \menv), \mstore, \kfn{\mval, \mkont}}$
-    \\
-% function call
-    $\tpl{\mval,\mstore,\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont}}$
-    &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\mval}, \mkont}$
-  \end{tabular}
-  \caption{The CESK machine}
-  \label{fig:base-semantics}
-\end{figure}
-
-\paragraph{Tracking return points:} the magic of the method is in
-keeping a table of continuation segments for each function and store
-pair, where the continuation is segmented at function boundaries. This
-closely mirrors the technique of store-allocating continuations used
-in AAM. Instead of deferring to the table (store) for the remainder of
-a continuation for each frame, we only have indirections at function
-call boundaries (see rules 4 and 5 in figure
-\ref{fig:summary-semantics} for saving and restoring
-continuations). Note that since AAM is about finitizing the state
-space, this step itself fits within the AAM method, since
-continuations within functions truncated at call boundaries are
-bounded by the nesting depth of those functions, thereby making the
-continuation space finite. In the monovariant case, the number of
-continuations is still linear in the size of the program.
-
-The tail of a continuation in the context of a function will contain
-the stack-less context in which the function was called, in order to
-link up with the proper call-site(s). The context includes the
-function (or unique label of the function), the environment, and the
-store. Notice that this choice can be viewed as a special allocation
-strategy for continuations in the AAM viewpoint, but we separate the
-table of continuations from the store since continuations themselves
-contain stores in $\rt$ frames --- this leads to a recursive
-data-structure that we are trying to avoid. The store is an important
-ingredient to maintaining enough precision to keep a pushdown
-abstraction. The semantic spaces for the machine are modified thusly:
-
-\begin{align*}
-  \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \times \KTab \times \Memo \\
-  \mkont \in \Kont &::= \kmt
-                     \mid \kar{\mexpr,\menv,\mkont} 
-                     \mid \kfn{\mval,\mkont}
-                     \mid \krt{(\mexpr,\menv),\mstore}\\
-  \mktab \in \KTab &= (\Expr \times \Env)\times \Store \to \wp(\Kont) \\
-  \mmemo \in \Memo &= (\Expr \times \Env)\times \Store \to \wp(\Value)
-\end{align*}
-
-\paragraph{The role of summaries:} notice the additional $\Memo$
-component. A ``summary edge'' in CFA2 or equivalently an
-$\epsilon$-edge in PDCFA is an edge from the source of a push edge to
-the target of the matching pop edge. ``Matching'' here means there is
-a path through machine reductions that don't change the stack, or
-through summary edges. In our model, we only ``push'' when we call a
-function, and we only ``pop'' when we return with a value. There is an
-analogy to something more operational: summary edges embody
-memoization (see the last rule in figure
-\ref{fig:summary-semantics}). Instead of following an entire path
-through a call to return a value, we simply jump from the call to the
-return with the result of the call (the second case of rule 4). We
-will return to this analogy and what it means in the presence of
-first-class control in section \ref{sec:sr}.
-
-\begin{figure}
-  \centering
-  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
-  \begin{tabular}{r|l}
-    \hline
-% variable lookup
-    $\tpl{(\svar\mvar, \menv), \mstore, \mkont, \mktab, \mmemo}$
-    &
-    $\tpl{\mval,\mstore,\mkont, \mktab, \mmemo}$ if $\mval \in \mstore(\menv(\mvar))$
-    \\
-% application
-    $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpri0, \menv), \mstore, \kar{\mexpri1,\menv,\mkont}, \mktab, \mmemo}$
-    \\
-% argument evaluation
-    $\tpl{\mval, \mstore, \kar{\mexpr,\menv,\mkont}, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr, \menv), \mstore, \kfn{\mval, \mkont}, \mktab, \mmemo}$
-    \\
-% function call
-    $\tpl{\mval,\mstore,\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont},\mktab,\mmemo}$
-    & % actually do call,
-    $\tpl{\mpoint,
-          \mstore',
-          \krt{\mpoint, \mstore'},
-          \joinone{\mktab}{(\mpoint, \mstore')}{\mkont},
-          \mmemo}$
-\\
-    & or \\
-    & $\tpl{\mval_\mathit{result}, \mstore', \mkont, \mktab, \mmemo} \text{ if } \mval_\mathit{result} \in \mmemo(\mpoint,\mstore')$
-    \\ % or, lookup in table
-%    & $\tpl{\mval, \mstore', \mkont,\mktab,\mmemo}$ if $\mval \in \mmemo(\mpoint, \mstore')$ \\    
-    where & $\mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr})$ \\
-          & $\mstore' = \joinone{\mstore}{\maddr}{\mval}$
-    \\
-% return
-    $\tpl{\mval, \mstore, \krt{\mpoint,\mstore'}, \mktab, \mmemo}$
-    &
-    $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{\mval}}$
-    if $\mkont \in \mktab(\mpoint, \mstore')$
-  \end{tabular}
-  \caption{The summarizing tabular stack machine}
-  \label{fig:summary-semantics}
-\end{figure}
-
-Once we finitize the address space of the store, this computes a
-characteristic finite state machine that has only reachable states of
-the pushdown system the unbounded stack model embodies.  To turn this
-semantics into PDCFA's algorithm for constructing a Dyck state graph,
-we apply a widening operator ${\mathcal F}$ to make $\mstore$,
-$\mkont$, and $\mmemo$ shared amongst all states.
-
-\begin{align*}
-  \mastate \in \sa{State} &= (\Expr \times \Env) \times \Kont \\
-  \System &= \wp(\sa{State} \times \Store) \times \wp(\sa{State}) \times \Store \times \KTab \times \Memo \\
-  {\mathcal F} &: \System \to \System \\
-  {\mathcal F}(S, F, \mstore, \mktab, \mmemo) &= (S \cup S', F', \mstore', \mktab', \mmemo') \\
-  \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mstore,\mktab,\mmemo) \stepto \mstate'} \\
-                \mstore' &= \bigsqcup\setbuild{\mstore'}{\wn(\_, \mstore',\_,\_) \in I} \\
-                \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \_, \mktab', \_) \in I} \\
-                \mmemo' &=  \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \_, \mmemo') \in I} \\
-                S' &= \setbuild{(\mastate',\mstore')}{\wn(\mastate', \_, \_, \_) \in I} \\
-                F' &= \setbuild{\mastate'}{(\mastate',\_) \in S' \setminus S} \\
-                \wn(\tpl{(\mexpr,\menv),\mkont},\mstore,\mktab,\mmemo)
-                   &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
-\end{align*}
-
-A $\System$ embodies the states seen and at which store, $S$, the
-frontier set of states yet to analyze, $F$, the shared store for the
-frontier, $\mstore$, the continuation table $\mktab$ and the memo
-table $\mmemo$. All the states in the frontier are stepped with the
-current store, after which the next store is the least upper bound of
-all the resulting stores. The next frontier contains only states
-resulting from stepping the previous frontier that we haven't seen at
-this next store. Systematic techniques for a performant implementation
-can be found in \citet{ianjohnson:oaam:2013}.
-
-\paragraph{Example:} Let's consider a monovariant allocation strategy
-(the address for each binding is the variable name itself) to run the
-following example:
-\begin{lstlisting}[mathescape]
-  (let* ([id ($\lambda$ (x) x)]
-         [y (id 0)]
-         [z (id 1)])
-    ($\le$ y z))
-\end{lstlisting}
-We should expect that a pushdown analysis would predict this evaluates
-to true, and there are no loops in the program. 0CFA claims there is a
-loop from the second call of \texttt{id} to the first, and thusly
-predicts this program evaluates to true or false. PDCFA claims there
-is no loop, and depending on the implementation, that the result is
-true or false (paper), or just true (implemented)\footnote{This is
-  because the paper steps every seen state with the current store
-  every iteration, but the implementation only steps states that need stepping.}. Let's take a look at the evaluation after binding
-\texttt{id} ($\mstore_0 = [\texttt{id} \mapsto
-\set{\vclo{\slam{\texttt{x}}{\texttt{x}}}{\bot}}]$):
-
-\begin{enumerate}
-\item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
-    \mstore_0[\texttt{x} \mapsto \set{0}]$ and
-    (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
-         $\mkont_1$ = \texttt{(letk y (let ([z (id 1)]) ($\le$ y z)) $[\texttt{id}\mapsto{\texttt{id}}]$)})
-    $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
-\item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
-\item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
-\item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \texttt{(letk z ($\le$ y z) $[\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]$)} (call this $\mkont_2$).}
-\item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
-      (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
-\item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
-\item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
-\end{enumerate}
-
-We maintained enough context to distinguish the return points of
-\texttt{id} to not rebind \texttt{y} to 1. When determining control
-flow through the expression, we consult $\mmemo$ to continue past
-function calls, so there is no confusion about a back edge from the
-second call to \texttt{id}.
-
-\section{Deriving CFA2}
-\label{sec:cfa2}
-
-CFA2 is the first published analysis of a higher-order programming
-language that could properly match calls and returns. We will show
-that it fits well into the same presentation we gave for
-PDCFA. Vardoulakis and Shivers had a clear goal of harnessing the
-extra information a pushdown model provides to produce a
-high-precision analysis that works well in practice. This resulted in
-more than just the call/return matching of the previous section, which
-is why we are showing the two separately. There are two orthogonal
-features of the semantics:
-\begin{enumerate}
-\item{stack allocation for some bindings}
-\item{strong updates on stack frames for resolved non-determinism}
-\end{enumerate}
-The first of these is an addition to the stack-less context. There is a
-conservative pre-analysis that checks locally whether a binding will
-never escape, and classifies references as able to use the stack frame
-or not. Their criteria for a binding never escaping is that it is
-never referenced in a function that is not its binder. This can be
-extended in a language with more linguistic features; see Kranz's
-thesis about register-allocatable bindings in the Orbit Scheme
-compiler \citep{ianjohnson:kranz:thesis:1988}. A stackable reference
-is one that appears in the body of the binding function, by which we
-mean not within the body of a nested function. They use the information
-that a binding never escapes to not bother allocating it in the
-heap. This has the advantage of not changing the heap, and thus leads
-to less propagation. The addition of these stack frames makes the
-analysis exponential in theory, though in practice they have been
-observed to decrease running time in most cases.
-
-The second of these is to ameliorate a problem they call ``fake
-rebinding.'' That is, since bindings in the abstract represent several
-values, we don't want to reference a variable \texttt{x} in two
-different places and have it resolve to two different values. In AAM,
-a variable reference non-deterministically steps to all possible
-values associated with that variable. Here we want to say that once
-\texttt{x} is considered to stand for value $v$, then all subsequent
-references of \texttt{x} should be $v$. If they aren't, it looks as if
-\texttt{x} was rebound; it hasn't, and thus it is a ``fake
-rebinding.'' CFA2 does not step to all values on variable reference,
-but instead carries all its values around in superposition until they
-need to be observed at, say, a function call. We give a simplified
-semantics that is more along the AAM style, but CFA2's approach can
-easily be recovered from it\footnote{Our Redex model implements the
-  fake rebinding strategy CFA2 itself employs}.
-
-CFA2 uses a ``local semantics'' that is similar to our segmenting
-continuations at function boundaries, but gets stuck when it gets to a
-point where it would need to ``return.'' It instead appeals to an
-external, imperative algorithm for summarization to sew the function
-calls and returns together by pattern matching on the states that the
-local semantics produces. What we show here is almost wholly the same
-in character, only embodied still in terms of a machine semantics and
-thus more easily reasoned about, and can be run in the concrete.
-
-We show only the significantly modified rules of the semantics in
-figure \ref{fig:frame-semantics}. The other rules simply carry along
-the extra $\mframe$ component untouched.
-\begin{figure}
-  \centering
-  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
-  \begin{tabular}{r|l}
-    \hline
-% variable lookup
-    $\tpl{(\svar[\mlab]{\mvar}, \menv), \mstore, \mframe, \mkont}$
-    &
-    $\tpl{\mval,\mstore,\mframe',\mkont}$ if $(\mframe', \mval) \in \lookup(\mstore,\mframe,\menv,\mvar,\mlab)$
-    \\
-% % application
-%     $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mframe, \mkont}$
-%     &
-%     $\tpl{(\mexpri0, \menv), \mstore, \mframe, \kar{\mexpri1,\menv,\mkont}}$
-%     \\
-% % argument evaluation
-%     $\tpl{\mval, \mstore, \mframe, \kar{\mexpr,\menv,\mkont}}$
-%     &
-%     $\tpl{(\mexpr, \menv), \mstore, \mframe, \kfn{\mval, \mkont}}$
-%     \\
-% function call
-    $\tpl{\mval,\mstore,\mframe,\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont}}$
-    &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore', \mframe', \mkont}$
-    \\ where & $(\mstore',\mframe') = \bind(\mstore,\mframe,\maddr,\mvar,\mval)$
-  \end{tabular}
-  \caption{The CES$\xi$K machine}
-  \label{fig:frame-semantics}
-\end{figure}
-
-The meta-functions the semantics uses to extend and consult the heap
-and stack use implicit information from the pre-analysis we described
-above:
-
-\begin{align*}
-  \bind(\mstore,\mframe,\maddr,\mvar,\mval) &=
-   \left\{\begin{array}{ll}
-            (\mstore, \snglm{\maddr}{\mval}) & \text{if } \mvar \text{ never escapes} \\
-            (\joinone{\mstore}{\maddr}{\mval}, \snglm{\maddr}{\mval}) & \text{otherwise } \\
-          \end{array}\right. \\
-  \lookup(\mstore,\mframe,\menv,\mvar,\mlab) &=
-    \left\{\begin{array}{ll}
-          \setbuild{(\extm{\mframe}{\menv(\mvar)}{\set{\mval}}, \mval)}
-                   {\mval \in \mframe(\menv(\mvar))}
-                   & \text{if } \mlab \text{ non-escaping} \\
-          \setbuild{(\mframe,\mval)}{\mval \in \mstore(\menv(\mvar))} & \text{otherwise}
-           \end{array}\right.
-\end{align*}
-
-Sidestepping fake rebinding does not need to be restricted to
-stackable references, but that is what CFA2 does. Indeed, as soon as
-the non-determinism has been determined, we could extend the stack
-frame so any subsequent reference means what it meant previously in
-the function. Look-up would then always try the stack frame before
-falling back on the heap.
-
-CFA2 also includes what they called ``transitive summaries'' to deal
-with tail calls. We insert an $\rt$ frame at every function call.
-This view contends with tail calls, but we can identify tail calls
-easily --- any call with an $\rt$ as its continuation is a tail
-call. Tail calls are important in language implementations for space
-complexity reasons \citep{ianjohnson:clinger:tail-calls:1998}, but in
-an analysis, these concerns are less important. The repeated popping
-of $\rt$ inserted by what otherwise were tail-calls is synonymous with
-CFA2's transitive summaries.
-
-\paragraph{Example:} if we use this semantics to run through the
-example of the previous section, we find that with the addition of
-stack frames, there is no confusion about \texttt{z}'s value either,
-so if the operator weren't $\le$ but instead $<$ or $+$, we could
-still constant fold that away.
-
-\section{Analysis of delimited, composable control}
-\label{sec:sr}
-
-There is contention among programming language researchers whether
-\rackett{call/cc} should be a language primitive, since it captures
-the entire stack, leading to space leaks
-\citep{ianjohnson:kiselyov:against-callcc}. Alternative control
-operators have been proposed that delimit how much of the stack to
-capture, such as Felleisen's $\%$ (read ``prompt'') and capture
-operator ${\mathcal F}$ (read ``control'')
-\citep{ianjohnson:felleisen:control:1988}, or Danvy and Filinski's
-\texttt{reset} (equivalent to $\%$) and \texttt{shift}
-\citep{ianjohnson:danvy:filinski:delim:1990}. However, the stacks
-captured by these operators always extend the stack when invoked,
-rather than replace it like those captured with
-\rackett{call/cc}. Continuations that have this extension behavior are
-called ``composable continuations.'' Stack replacement is simple to
-model in a regular analysis using the AAM approach, and Vardoulakis
-and Shivers showed it can be done in a pushdown approach (although it
-breaks the pushdown model). Stack extension, however, poses a new
-challenge for pushdown analysis, since one application of a composable
-continuation means pushing an unbounded amount of frames onto the
-stack. Vardoulakis' and Shivers' approach does not immediately apply
-in this situation, since their technique drops all knowledge of the
-stack at a continuation's invocation site; extension, however, must
-preserve it.
-
-The way we have been splitting continuations at function calls has
-similarities to the meta-continuation approach to modeling delimited
-control, given in figure \ref{fig:shift-reset} (adapted from
-~\citep{ianjohnson:Biernacki2006274}). We could view each function
-call as inserting a prompt, and returns as aborting to the nearest
-prompt. Resets then insert a second-tier prompt. The changed semantic
-spaces for the shift/reset semantics are as follows:
-
-\begin{align*}
-  \mstate \in \State &::= \tpl{\mpoint, \mstore, \mkont, \mmkont} \\
-  \mpoint \in \Point &::= (\mexpr, \menv) \mid \mval \\
-  \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \mid \vcomp{\mkont} \\
-  \mmkont \in \MKont &::= \kmt \mid \mkapp{\mkont}{\mmkont}
-\end{align*}
-
-\begin{figure}
-  \centering
-  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
-  \begin{tabular}{r|l}
-    \hline
-% Reset
-    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont}$
-    &
-    $\tpl{(\mexpr, \menv), \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
-    \\
-% Pop prompt
-    $\tpl{\mval, \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
-    &
-    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}}$
-    \\
-% Shift
-    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont}$
-    &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\mkont},\kmt,\mmkont}$
-    \\
-% continuation call
-    $\tpl{\mval,\mstore,\kfn{\mkont', \mkont}, \mmkont}$
-    &
-    $\tpl{\mval, \mstore, \mkont', \mkapp{\mkont}{\mmkont}}$
-  \end{tabular}  
-  \caption{Machine semantics for shift/reset}
-  \label{fig:shift-reset}
-\end{figure}
-
-Turning this into a table-based semantics involves making prompts a
-point of indirection, just like function calls. Memoization also gets
-a new context to consider, calling a continuation, because composable
-continuations act like functions. The new semantic spaces are then
-
-\begin{align*}
-  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid (\vcomp{\mkont},\mval,\mstore) \\
-  \mkont \in \Kont &= \kmt \mid \krt{\mctx} \mid \kar{\mexpr,\menv,\mkont} \mid \kfn{\mval,\mkont} \\
-  \mmkont \in \MKont &::= \kmt \mid \kprompt{\mctx} \\
-  \mktab \in \KTab &= \Context \to (\wp(\Kont \times \MKont) \cup \Kont) \\
-  \mmemo \in \Memo &= \Context \to \wp(\Value)
-\end{align*}
-
-Figure \ref{fig:shift-reset-table0} has what one might naturally write
-using just this information. Unfortunately, since $\rt$ frames contain
-a store, and continuations can now appear in the store, this
-introduces a circularity that could cause the analysis to never
-terminate. For example, a loop that continually captures continuations
-would introduce unboundedly many continuation values due to the
-ever-growing store. This means the store is not finite height, and
-there may be no fixed point.
-
-\paragraph{Possible fixes to non-termination:} The simplest fix to the
-circularity would be to strip stores out of the context of the $\rt$
-frame in a captured continuation, and additionally consider contexts
-with a stripped store an abstraction of the same context with any
-store. This is a fairly brutal approximation of a captured
-continuation, so an alternative is to make this tunable with our
-precision-tuning friend, $\alloc$. Then, instead of entirely removing
-the store component of the $\rt$ context, we can replace it with an
-address of possible stores that it would approximate.
-
-\begin{figure}
-  \centering
-  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
-  \begin{tabular}{r|l}
-    \hline
-% Reset
-    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
-    &
-    $\tpl{\mpoint, \mstore, \kmt, \kprompt{\mctx}, \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}, \mmemo}$
-    \\
-    where & $\mpoint = (\mexpr, \menv)$, $\mctx = (\mpoint, \mstore)$.
-    \\
-% Pop prompt
-    $\tpl{\mval, \mstore, \kmt, \kprompt{\mctx}, \mktab, \mmemo}$
-    &
-    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}, \mktab, \joinone{\mmemo}{\mctx}{\mval}}$
-    if $(\mkont,\mmkont) \in \mktab(\mctx)$
-    \\
-% Shift
-    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont}},\kmt,\mmkont,\mktab,\mmemo}$
-    \\
-% continuation call
-    $\tpl{\mval,\mstore,\kfn{\vcomp{\mkont'}, \mkont}, \mmkont, \mktab, \mmemo}$
-    &
-    $\tpl{\mval, \mstore, \mkont', \kprompt{\mctx}, \mktab',\mmemo}$
-    \\
-    where & $\mctx = (\mkont, \mval, \mstore')$ \\
-          & $\mktab' = \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}$
-    \\
-% Memoized continuation call
-% continuation call
-    $\tpl{\mval,\mstore,\kfn{\vcomp{\mkont'}, \mkont}, \mmkont, \mktab, \mmemo}$
-    &
-    $\tpl{\mval', \mstore, \mkont, \mmkont, \mktab,\mmemo}$ if $\mval' \in \mmemo(\mkont',\mval,\mstore)$
-  \end{tabular}  
-  \caption{Faulty table-based semantics for shift/reset}
-  \label{fig:shift-reset-table0}
-\end{figure}
-
-The new indirection possibility changes $\Context$ to also include
-$\maddr$ where there was previously a $\mstore$ (though contexts for
-continuation calls remain unchanged), and $\KTab$ now additionally
-maps $\Addr$ to a set of stores. The rules that change are presented in figure \ref{fig:shift-reset-table1}.
-
-\begin{align*}
-  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid ((\mexpr,\menv),\maddr) \mid (\vcomp{\mkont},\mval,\mstore) \\
-  \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
-  \mskont \in \SKont &= \krt{\msctx} \mid \ldots \\
-  \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \mid \vcomp{\mskont} \\
-  \mmkont \in \MKont &::= \kmt \mid \kprompt{\mctx} \\
-  \mktab \in \KTab &= (\Context \to (\wp(\Kont \times \MKont) \cup \Kont)) \\
-                   &\cup (\Addr \to \wp(\Store))
-\end{align*}
-
-\begin{figure}
-  \centering
-% Shift
-  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
-  \begin{tabular}{r|l}
-    \hline
-    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont'}},\kmt,\mmkont,\mktab',\mmemo}$
-    \\
-    where & $(\mkont',\mktab') = \approximate(\mkont, \mktab, \maddr)$
-\\
-% Return
-   $\tpl{\mval, \mstore, \krt{\mctx}, \mmkont, \mktab, \mmemo}$
-   &
-   $\tpl{\mval, \mstore, \mkont, \mmkont, \mktab, \mmemo'}$
-   if $\mkont \in \returns(\mktab, \mctx)$
-   \\
-   where & $\mmemo' = \memoize(\mmemo, \mktab, \mctx, \mval)$
-  \end{tabular}
-  \caption{Fixed table-based semantics for shift/reset}
-  \label{fig:shift-reset-table1}
-\end{figure}
-
-The meta-functions mentioned in the fixed semantics all deal with the
-addition of $\maddr$ to contexts. If the context in an $\rt$ frame is
-approximate, we must return to all the continuations known for all the
-contexts it approximates ($\returns$). At return boundaries, the memo
-table must add the result to all the represented contexts
-($\memoize$). At capture time, we strip the store in the $\rt$ frame
-if there is one, and replace it with an address.
-
-\newcommand{\replacectx}{\mathit{replace}\mstore}
-\newcommand{\addstore}{\mathit{add}\mstore}
-\begin{align*}
-  \returns(\mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
-  \returns(\mktab, ((\mexpr,\menv), \maddr)) &=
-    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mktab(\maddr)}
-  \\[2pt]
-  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \mstore), \mval) &=
-    \joinone{\mmemo}{((\mexpr,\menv), \mstore)}{\mval} \\
-  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \maddr), \mval) &=
-    \mmemo \sqcup \bigsqcup\limits_{\mstore \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore)}{\mval}
-  \\[2pt]
-  \approximate(\mkont, \mktab, \maddr) &= (\replacectx(\mkont), \addstore{\mkont}) \\
-  \text{where }
-   \replacectx(\kmt) &= \kmt \\
-   \replacectx(\krt{((\mexpr,\menv),\_)}) &= \krt{((\mexpr,\menv),\maddr)} \\
-   \replacectx(\kar{\mexpr,\menv,\mkont}) &= \kar{\mexpr,\menv,\replacectx(\mkont)} \\
-   \replacectx(\kfn{\mval, \mkont}) &= \kfn{\mval,\replacectx(\mkont)}
-  \\[2pt]
-   \addstore(\kmt) &= \mktab \\
-   \addstore(\krt{((\mexpr,\menv),\mstore)}) &= \joinone{\mktab}{\maddr}{\mstore} \\
-   \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mktab}{\maddr}{\mktab(\maddr')} \\
-   \addstore(\kar{\mexpr,\menv,\mkont}) &= \addstore(\mkont) \\
-   \addstore(\kfn{\mval, \mkont}) &= \addstore(\mkont)
-\end{align*}
-
-The result of this viewpoint is a sound, precise, and computable
-semantics for a ``pushdown approach'' to analyzing delimited,
-composable control. To handle non-composable control, we can simply
-remove the prompts and keep $\approximate$.
-
-\section{Related Work}
-
-The immediately related work is that of PDCFA
-\citep{ianjohnson:earl2010pdcfa},
-CFA2~\citep{ianjohnson:vardoulakis-lmcs11,
-  ianjohnson:Vardoulakis2011Pushdown}, and AAM
-~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we
-recreated in full detail. The version of CFA2 that handles
-\rackett{call/cc} does not handle composable control, and is dependent
-on a restricted CPS representation. They also had no way to tune the
-precision of their first class continuation approximation. Since
-\rackett{call/cc} can be simulated with shift/reset, this work
-supercedes theirs. The extended version of PDCFA that does garbage
-collection \citep{dvanhorn:Earl2012Introspective} also fits into our
-model, although we did not explicitly show it. The addresses that the
-stack keeps alive can be accumulated by ``reading through'' the
-continuation table, building up the set of addresses in each portion
-of the stack that we come across.
-
-Pushdown models have existed in the first-order static analysis
-literature ~\citep[Chapter
-7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995},
-and the first-order model checking literature
-\citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time. The
-higher-order setting imposes additional challenges that make their
-methods difficult to adapt. The most important constraint is that we
-can't know all call-sites of a function/method before the analysis
-begins, which their methods heavily rely on.
-
-A new temporal logic that itself understands well-bracketing of call
-and returns for stating and validating propositions, NWTL
-~\citep{ianjohnson:alur:nwtl:2007}, applies to visibly pushdown
-systems, which can precisely model programming languages that only have
-well-bracketed call and returns. The propositional form of the logic
-has decidable satisfiability checking, and control-flow queries are
-shallow propositions to pose in the logic. However, the satisfiability
-problem is \textsc{2Exptime}-hard and the algorithm is not given in
-the on-the-fly form that a higher-order language would need to
-effectively answer control-flow queries without already knowing the
-answers. The logic itself is an exciting new frontier for expressing
-program correctness properties, and a higher-order version would be a
-welcome addition to the functional programmer's tool belt.
-
-The trend of deriving static analyses from abstract machines does not
-stop at flow analyses. The model-checking community showed how to
-check temporal logic queries for collapsible pushdown automata (CPDA),
-or equivalently, higher-order recursion schemes, by deriving the
-checking algorithm from the Krivine
-machine~\citep{ianjohnson:Salvati:2011:KMH:2027223.2027239}. The
-expressiveness of CPDAs outweighs that of PDAs, but it is unclear how
-to adapt higher-order recursion schemes to model arbitrary programming
-language features. The method is strongly tied to the simply-typed
-call-by-name lambda calculus and depends on finite sized base-types.
-
-\section{Conclusion and future work}
-
-As the programming world continues to embrace behavioral values, it
-becomes more important to import the powerful techniques pioneered by
-the first-order analysis and model checking communities. CFA2 and
-PDCFA paved the way, and in large part inspired this work. It is our
-view that systematic approaches to applying the techniques are pivotal
-to scaling them to ``real languages.'' We believe that the recipe that
-this paper set forth is a step in that direction. That is, make
-continuation tables keyed with enough context, and memoize at the
-introduced indirection points. The result in a language with
-well-bracketed control is a ``pushdown analysis'' using
-summarization. In a language without well-bracketed control, we are
-not chained to a pushdown automaton as the target of the
-approximation, so the techniques still apply and give better precision
-than regular methods.
-
-Our goal in the future is to show that this technique is even more
-widely applicable than shift and reset. We conjecture that the same
-recipe will apply to the most intricate control operators in
-production languages such as in
-\citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}. The control
-structures there are difficult to model even with AAM's techniques due
-to the ability to capture and compose continuations with arbitrarily many prompts,
-but once we can tackle AAM, there should be a straightforward route to
-extending it with the pushdown techniques of this paper.
-
-There is also the question of clients of these analyses. Most
-obviously we would want to know where we can implement first-class
-control more efficiently. In particular, we would want to have an
-escape analysis to find first-class continuations that don't need to
-be heap-allocated, and single target abort operations that can be
-turned into long jumps after a computed stack unraveling. On top of
-optimizations, there are security analyses. Greater control of the
-stack gives us the ability to drill deep into context-sensitive
-security properties and make precise predictions such as, ``is it
-possible to send image data on the network outside the context of the
-user's permission?'' This way a user can use an image taking/sharing
-application without the application surreptitiously taking and sending
-images the user doesn't know about.
-
-\bibliographystyle{chicago}
-\bibliography{bibliography}
-
-\end{document}

### 2013-09-30 content.tex
Cleaned up repo. Added semantic fix and conjectures for correctness.

--- 2013-07-10 paper.tex
@@
-\documentclass{sigplanconf}
-
-\pagestyle{plain}
-\usepackage{amsmath,stmaryrd,natbib,listings}
-\usepackage[usenames,dvipsnames]{color}
-
-\newcommand{\Scribtexttt}[1]{{\texttt{#1}}}
-\newcommand{\SColorize}[2]{\color{#1}{#2}}
-\newcommand{\inColor}[2]{{\Scribtexttt{\SColorize{#1}{#2}}}}
-\definecolor{PaleBlue}{rgb}{0.90,0.90,1.0}
-\newcommand{\rackett}[1]{\inColor{black}{#1}}
-\input{local-macros}
-
-\title{Concrete Semantics for Pushdown Analysis:\\The Essence of Summarization}
-\authorinfo{J. Ian Johnson and David Van Horn}{Northeastern University}{\{ianj,dvanhorn\}@ccs.neu.edu}
-
-\usepackage{hyperref}
-\begin{document}
-\maketitle
-
-% Outline:
-% Introduction.
-% High level
-% PDCFA
-% - Concrete
-% - Abstract
-% CFA2
-% - Concrete
-% - Abstract
-% WCM machine
-% - Concrete
-% - Abstract
-% Related Work
-% Future Work
-% Conclusion
-
-\begin{abstract}
-Pushdown analysis is better than finite-state analysis in precision and
-performance. Why then have we not seen total widespread adoption of these
-techniques? For one, the known techniques are technically burdened and
-difficult to understand or extend. Control structure of the programming language
-gets pulled into the model of computation, which makes extensions to
-non-pushdown control structures, such as \rackett{call/cc} or \rackett{shift} and
-\rackett{reset}, non-trivial.
-
-We show a derivational approach to abstract interpretation that yields
-transparently sound static analyses that can precisely match calls and
-returns when applied to well-known abstract machines. We show that
-adding memoization and segmenting the continuation into bounded pieces
-leads to machines that abstract to static analyses for context-free
-reachability by simply bounding the stores. This technique allows us
-to derive existing, more technically involved analyses, and a novel
-pushdown analysis for delimited, composable control.
-\end{abstract}
-
 \section{Introduction}
@@
   \mktab \in \KTab &= (\Expr \times \Env)\times \Store \to \wp(\Kont) \\
-  \mmemo \in \Memo &= (\Expr \times \Env)\times \Store \to \wp(\Value)
+  \mmemo \in \Memo &= (\Expr \times \Env)\times \Store \to \wp(\Value \times \Store)
 \end{align*}
@@
 through a call to return a value, we simply jump from the call to the
-return with the result of the call (the second case of rule 4).
+return with the result of the call, with the resulting context changes---in this case only the store (the second case of rule 4).
 
@@
     & $\tpl{\mval_\mathit{result},
-            \mstore',
+            \mstore'',
             \mkont,
@@
             \mmemo}$
-    \\ & \quad if $\mval_\mathit{result} \in \mmemo(\mpoint,\mstore')$
+    \\ & \quad if $(\mval_\mathit{result},\mstore'') \in \mmemo(\mpoint,\mstore')$
     \\ % or, lookup in table
@@
     &
-    $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{\mval}}$
+    $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}$
     \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
@@
 
+%% FIXME: remove stores from memo-table since they will always be unchanged
+%% TODO: Show that store widening and store/memo/KTab widening have a bisimulation.
 \begin{align*}
@@
 frontier, $\mstore$, the continuation table $\mktab$ and the memo
-table $\mmemo$. All the states in the frontier are stepped with the
-current store, after which the next store is the least upper bound of
-all the resulting stores. The next frontier contains only states
-resulting from stepping the previous frontier that we haven't seen at
-this next store. Systematic techniques for a performant implementation
-can be found in \citet{DBLP:journals/corr/abs-1211-3722}.
+table $\mmemo$. The memo table need only map to values, since the
+global store will always subsume the store that would otherwise be
+mapped. All the states in the frontier are stepped with the current
+store, after which the next store is the least upper bound of all the
+resulting stores. The next frontier contains only states resulting
+from stepping the previous frontier that we haven't seen at this next
+store. Systematic techniques for a performant implementation can be
+found in \citet{ianjohnson:oaam:icfp2013}.
 
@@
 
+%% FIXME stores in memo-table
 \begin{enumerate}
@@
 second call to \texttt{id}.
+
+\begin{conjecture}
+  Tabular semantics has a bisimulation with the standard semantics with a sound allocation strategy.
+\end{conjecture}
+\begin{proof}
+
+\end{proof}
+
+\begin{conjecture}
+  Store-widened tabular semantics has a bisimulation with store/memo/k-table-widened semantics.
+\end{conjecture}
 
@@
 the function. Look-up would then always try the stack frame before
-falling back on the heap.
+falling back on the heap (of course this must be limited to immutable addresses).
 
@@
 
+\begin{conjecture}
+  Tabular semantics has a bisimulation with standard semantics.
+\end{conjecture}
+
 \section{Analysis of delimited, composable control}
 \label{sec:sr}
+
+%% TODO (thanks to ccshan): Discuss difference between CPS for shift/reset and this approach,
+%% Discuss how this does not generalize to first class prompts due to append versus cons.
 
@@
   \mktab \in \KTab &= \Context \to \mathit{KTabRange} \\
-  \mmemo \in \Memo &= \Context \to \wp(\Value)
+  \mmemo \in \Memo &= \Context \to \wp(\Value \times \Store)
 \end{align*}
@@
     &
-    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}, \mktab, \joinone{\mmemo}{\mctx}{\mval}}$
+    $\tpl{\mval, \mstore, {\mkont}, {\mmkont}, \mktab, \joinone{\mmemo}{\mctx}{(\mval,\mstore)}}$
     if $(\mkont,\mmkont) \in \mktab(\mctx)$
@@
     &
-    $\tpl{\mval', \mstore, \mkont, \mmkont, \mktab,\mmemo}$ if $\mval' \in \mmemo(\mkont',\mval,\mstore)$
+    $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab,\mmemo}$ if $(\mval',\mstore') \in \mmemo(\mkont',\mval,\mstore)$
   \end{tabular}  
@@
    \\
-   where & $\mmemo' = \memoize(\mmemo, \mktab, \mctx, \mval)$
+   where & $\mmemo' = \memoize(\mstore,\mmemo, \mktab, \mctx, \mval)$
   \end{tabular}
@@
 \begin{align*}
-  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \mstore), \mval) &=
-  \joinone{\mmemo}{((\mexpr,\menv), \mstore)}{\mval} \\
-  \memoize(\mmemo, \mktab, ((\mexpr,\menv), \maddr), \mval) &=
-  \mmemo \sqcup \bigsqcup\limits_{\mstore \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore)}{\mval}
+  \memoize(\mstore,\mmemo, \mktab, ((\mexpr,\menv), \mstore'), \mval) &=
+  \joinone{\mmemo}{((\mexpr,\menv), \mstore')}{(\mval,\mstore')} \\
+  \memoize(\mstore,\mmemo, \mktab, ((\mexpr,\menv), \maddr), \mval) &=
+  \mmemo \sqcup \bigsqcup\limits_{\mstore' \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore')}{(\mval,\mstore)}
 \end{align*}
@@
 remove the prompts and keep $\approximate$.
+
+\begin{conjecture}
+  Tabular semantics has a bisimulation with standard semantics given fresh allocation.
+\end{conjecture}
+
+\begin{conjecture}
+  Tabular semantics simulates standard semantics with abstract allocation.
+\end{conjecture}
+
+%% TODO: complete
+\paragraph{Comparison to CFA2 CPS'd to remove {\tt shift} and {\tt reset}:}{
+  One might ask what this different model buys us over using a CPS transform to compile away {\tt shift} and {\tt reset} forms.
+%
+The short answer is that we get better precision for capture and invocation of continuations.
+%
+In pushdown analysis, we track function return points with entire calling contexts rather than the $k$-CFA way of allocating the continuation in the heap at some lower-precision address, like a variable.
+%
+In our summarizing analysis of delimited composable control, we get the same added precision for {\tt reset}, {\tt shift}, and continuation invocation points.
+%
+Consider the following program and its CPS transform for comparison:
+\begin{lstlisting}
+  mumble
+\end{lstlisting}
+
+PDCFA with a monovariant allocation strategy will predict the CPS'd version returns ???.
+%
+Our specialized analysis for delimited control, also with a monovariant allocation strategy, will predict the non-CPS'd version returns ???.}
 
@@
 security properties and make precise predictions.
-
-\bibliographystyle{plainnat}
-\bibliography{bibliography}
-
-\end{document}

### 2013-10-02 content.tex
Moved proofs to appendix, reformatted sentences to be on single lines for better diffs, added GC section.

@@
 
-Programs in higher-order languages heavily use function calls and
-method dispatch as part of their control flow. Until recently, flow
-analyses for higher-order languages could not handle return flow
-precisely \citep{ianjohnson:vardoulakis-lmcs11,
-  ianjohnson:earl2010pdcfa}, which leads to several spurious paths
-(and thus false positives) due to the pervasiveness of function/method
-calls and subsequent returns. These works, called CFA2 and PDCFA
-respectively, use pushdown automata as their approximation's target
-model of computation. They are hence called ``pushdown
-analyses.''\footnote{We will refer to the classic finite model
-  analyses as ``regular analyses'' after regular languages.} CFA2 and
-PDCFA have difficult details to easily apply to an off-the-shelf
-semantics --- especially if they feature non-local control transfer
-that breaks the pushdown model.
+Programs in higher-order languages heavily use function calls and method dispatch as part of their control flow.
+%
+Until recently, flow analyses for higher-order languages could not handle return flow precisely \citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:earl2010pdcfa}, which leads to several spurious paths (and thus false positives) due to the pervasiveness of function/method calls and subsequent returns.
+%
+These works, called CFA2 and PDCFA respectively, use pushdown automata as their approximation's target model of computation.
+%
+They are hence called ``pushdown analyses.''
+%
+\footnote{We will refer to the classic finite model analyses as ``regular analyses'' after regular languages.}
+%
+CFA2 and PDCFA have difficult details to easily apply to an off-the-shelf semantics --- especially if they feature non-local control transfer that breaks the pushdown model.
+%%
 
-There is a systematic process for transforming off-the-shelf
-programming language semantics into a form amenable to \emph{regular}
-analysis that has been widely applied with great success to production
-programming languages ~\citet{dvanhorn:VanHorn2010Abstracting} (a
-technique called abstracting abstract machines, or AAM). A
-contribution of this paper is a systematic process to construct
-\emph{pushdown} analyses of programming languages, due to the
-precision (and often performance) benefits.
+%%
+There is a systematic process for transforming off-the-shelf programming language semantics into a form amenable to \emph{regular} analysis that has been widely applied with great success to production programming languages ~\citet{dvanhorn:VanHorn2010Abstracting} (a technique called abstracting abstract machines, or AAM).
+%
+A contribution of this paper is a systematic process to construct \emph{pushdown} analyses of programming languages, due to the precision (and often performance) benefits.
+%%
 
-Testing new ideas in analysis for improving precision or increasing
-performance can be a difficult venture. We contend that the machinery
-that we employ in the abstract should have a concrete counterpart that
-maintains the meaning of the language so that we can see their effect
-without introducing approximations which might mask correctness
-issues. Abstraction should be a simple process that is ``obviously
-correct.'' The framework we present in this paper is applicable in the
-concrete such that a point-wise abstraction leads to the pushdown
-analyses in the literature.
+%%
+Testing new ideas in analysis for improving precision or increasing performance can be a difficult venture.
+%
+We contend that the machinery that we employ in the abstract should have a concrete counterpart that maintains the meaning of the language so that we can see their effect without introducing approximations which might mask correctness issues.
+%
+Abstraction should be a simple process that is ``obviously correct.''
+%
+The framework we present in this paper is applicable in the concrete such that a point-wise abstraction leads to the pushdown analyses in the literature.
+%%
 
-This paper gives a common, simple framework to derive both CFA2 and
-PDCFA using a recipe to apply pushdown analysis techniques to
-arbitrary semantics in a manner similar to AAM. To exercise the recipe further, 
-we give a novel analysis for delimited, composable control. \\
+%%
+This paper gives a common, simple framework to derive both CFA2 and PDCFA using a recipe to apply pushdown analysis techniques to arbitrary semantics in a manner similar to AAM.
+%
+To exercise the recipe further, we give a novel analysis for delimited, composable control. \\
 
@@
 
-CFA2 uses a technique called ``summarization'' from
-\citet[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the
-$\epsilon$-closure graph that PDCFA constructs. Summarization
-algorithms need not be restricted to languages with well-bracketed
-calls and returns. We can adopt the technique for higher precision in
-the common case but still handle difficult cases such as first-class
-control. This was shown for the
-\rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc})
-operator in ~\citet{ianjohnson:Vardoulakis2011Pushdown}. This
-impressive work illuminated the fact that we can harness the enhanced
-technology of pushdown analyses in non-pushdown models of
-computation. Doing this sacrifices call/return matching in the general
-case, but in practice the precision is much better than the
-alternative regular model that, say, AAM would provide.
+CFA2 uses a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the $\epsilon$-closure graph that PDCFA constructs.
+%
+Summarization algorithms need not be restricted to languages with well-bracketed calls and returns.
+%
+We can adopt the technique for higher precision in the common case but still handle difficult cases such as first-class control.
+%
+This was shown for the \rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc}) operator in ~\citet{ianjohnson:Vardoulakis2011Pushdown}.
+%
+This impressive work illuminated the fact that we can harness the enhanced technology of pushdown analyses in non-pushdown models of computation.
+%
+Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, say, AAM would provide.
+%%
 
-A downside of the work providing \rackett{call/cc} is that it was an
-algorithmic change to the already complicated CFA2 --- there was no
-recipe for how to do this for one's favorite control operator. This
-paper seeks to do just that with an operational view of what
-summarization is, in essence. In other words, we give a concrete
-semantics to the tricks that the analyses in the literature use in the
-abstract, and maintain the original meaning of the language. We also
-give intuitive analogies to well-established ideas/techniques so that
-the working semanticist can write a pushdown analysis for her
-language. In order to demonstrate the applicability of this viewpoint,
-we show a new analysis for a language with composable control. All of
-the semantics modeled in this paper are implemented in full detail in
-PLT Redex~\citep{ianjohnson:Felleisen:2009:SEP:1795772} and available
-online\footnote{\url{http://github.com/ianj/concrete-summaries}}.
+%%
+A downside of the work providing \rackett{call/cc} is that it was an algorithmic change to the already complicated CFA2 --- there was no recipe for how to do this for one's favorite control operator.
+%
+This paper seeks to do just that with an operational view of what summarization is, in essence.
+%
+In other words, we give a concrete semantics to the tricks that the analyses in the literature use in the abstract, and maintain the original meaning of the language.
+%
+We also give intuitive analogies to well-established ideas/techniques so that the working semanticist can write a pushdown analysis for her language.
+%
+In order to demonstrate the applicability of this viewpoint, we show a new analysis for a language with composable control.
+%
+All of the semantics modeled in this paper are implemented in full detail in PLT Redex~\citep{ianjohnson:Felleisen:2009:SEP:1795772} and available online\footnote{\url{http://github.com/ianj/concrete-summaries}}.
 
-There are common underpinnings of PDCFA and CFA2 that can be embodied
-as concrete machinery in the programming language semantics: 
+There are common underpinnings of PDCFA and CFA2 that can be embodied as concrete machinery in the programming language semantics: 
 \begin{enumerate}
-\item{breaking the recursive structure of continuations by indirecting them through a
-table with appropriately precise keys, and}
+\item{breaking the recursive structure of continuations by indirecting them through a table with appropriately precise keys, and}
 \item{memoization.}
 \end{enumerate}
-Once the machine semantics is in this form, simple point-wise
-abstraction leads to the summarization algorithms that we see in the
-literature.
+Once the machine semantics is in this form, simple point-wise abstraction leads to the summarization algorithms that we see in the literature.
 
@@
 
-PDCFA does not have some orthogonal semantic components that CFA2
-features to improve precision and is thus the simpler of the two. The
-key to their method is to notice in languages without stack-capturing
-features, the continuation is only modified a frame at a time. If the
-state space without the stack is finite (and it can be made finite
-with an approximation ala AAM), then the model falls directly into the
-realm of pushdown systems. They thus recast the problem in terms of
-pushdown systems. The whole of their machinery is then computing the
-pushdown system on-the-fly, only considering states reachable from the
-root (initial) state. They call the main data-structure for this a
-Dyck state graph. To match nodes that push frames with later nodes
-following the pop of that frame, they additionally compute an
-$\epsilon$-closure graph, which keeps edges between nodes that have a
-net-zero stack change path between them. We skip the detour to
-pushdown systems altogether and show an operational understanding of
-the resulting analysis.
+PDCFA does not have some orthogonal semantic components that CFA2 features to improve precision and is thus the simpler of the two.
+%
+The key to their method is to notice in languages without stack-capturing features, the continuation is only modified a frame at a time.
+%
+If the state space without the stack is finite (and it can be made finite with an approximation ala AAM), then the model falls directly into the realm of pushdown systems.
+%
+They thus recast the problem in terms of pushdown systems.
+%
+The whole of their machinery is then computing the pushdown system on-the-fly, only considering states reachable from the root (initial) state.
+%
+They call the main data-structure for this a Dyck state graph.
+%
+To match nodes that push frames with later nodes following the pop of that frame, they additionally compute an $\epsilon$-closure graph, which keeps edges between nodes that have a net-zero stack change path between them.
+%
+We skip the detour to pushdown systems altogether and show an operational understanding of the resulting analysis.
+%%
 
-We start by deriving PDCFA from an operational semantics for the
-untyped lambda calculus (figure \ref{fig:base-semantics}). The
-semantic spaces for the machine follow.
+%%
+We start by deriving PDCFA from an operational semantics for the untyped lambda calculus (figure \ref{fig:base-semantics}).
+%
+The semantic spaces for the machine follow.
 
+\begin{minipage}[b]{.55\linewidth}
   \begin{align*}
@@
   \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \\
-  \mkont \in \Kont &::= \kmt \mid \kar{\mexpr,\menv,\mkont} \mid \kfn{\mval,\mkont} \\
+  \mkont \in \Kont &::= \kmt \mid \kcons{\mkframe}{\mkont} \\
+  \mkframe \in \Frame &::= \kar{\mexpr,\menv} \mid \kfn{\mval}
+  \end{align*}\hrule height 0pt\end{minipage}
+\begin{minipage}[b]{.40\linewidth}
+  \begin{align*}
   \menv \in \Env &= \Var \to \Addr \\
@@
   \maddr \in \Addr &\text{ an infinite set}
-\end{align*}
+  \end{align*}\hrule height 0pt\end{minipage}
 
-The $\Addr$ space is what controls the precision of the model. For a
-concrete semantics, we require the allocation meta-function $\alloc :
-\State \to \Addr$ to return fresh addresses ($\maddr \notin
-\dom(\mstore)$), but any choice of address is sound. If $\alloc$ only
-uses addresses from a finite subset of $\Addr$, then the state space
-without continuations is finite, and the space of contination frames
-is finite (thus the pushdown system interpretation is apt). This
-meta-function along with the commitment to having no recursive
-data-structures is the key to the AAM technique. All recursion can be
-expressed by indirecting through addresses into a store --- tying
-Landin's knot. We borrow this technique on top of our own.
+The $\Addr$ space is what controls the precision of the model.
+%
+For a concrete semantics, we require the allocation meta-function $\alloc :
+\State \to \Addr$ to return fresh addresses ($\maddr \notin \dom(\mstore)$), but any choice of address is sound.
+%
+If $\alloc$ only uses addresses from a finite subset of $\Addr$, then the state space without continuations is finite, and the space of contination frames is finite (thus the pushdown system interpretation is apt).
+%
+This meta-function along with the commitment to having no recursive data-structures is the key to the AAM technique.
+%
+All recursion can be expressed by indirecting through addresses into a store --- tying Landin's knot.
+%
+We borrow this technique on top of our own.
 
@@
     &
-    $\tpl{(\mexpri0, \menv), \mstore, \kar{\mexpri1,\menv,\mkont}}$
+    $\tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}}$
     \\
 % argument evaluation
-    $\tpl{\mval, \mstore, \kar{\mexpr,\menv,\mkont}}$
+    $\tpl{\mval, \mstore, \kcons{\kar{\mexpr,\menv}}{\mkont}}$
     &
-    $\tpl{(\mexpr, \menv), \mstore, \kfn{\mval, \mkont}}$
+    $\tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}}$
     \\
 % function call
-    $\tpl{\mval,\mstore,\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont}}$
+    $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont}}$
     &
@@
 
-\paragraph{Tracking return points:} the magic of the method is in
-keeping a table of continuation segments for each function and store
-pair, where the continuation is segmented at function boundaries. This
-closely mirrors the technique of store-allocating continuations used
-in AAM. Instead of deferring to the table (store) for the remainder of
-a continuation for each frame, we only have indirections at function
-call boundaries (see rules 4 and 5 in figure
-\ref{fig:summary-semantics} for saving and restoring
-continuations). Note that since AAM is about finitizing the state
-space, this step itself fits within the AAM method, since
-continuations within functions truncated at call boundaries are
-bounded by the nesting depth of those functions, thereby making the
-continuation space finite. In the monovariant case, the number of
-continuations is still linear in the size of the program.
+\paragraph{Tracking return points:} the magic of the method is in keeping a table of continuation segments for each function and store pair, where the continuation is segmented at function boundaries.
+%
+This closely mirrors the technique of store-allocating continuations used in AAM.
+%
+Instead of deferring to the table (store) for the remainder of a continuation for each frame, we only have indirections at function call boundaries (see rules 4 and 5 in figure \ref{fig:summary-semantics} for saving and restoring continuations).
+%
+Note that since AAM is about finitizing the state space, this step itself fits within the AAM method, since continuations within functions truncated at call boundaries are bounded by the nesting depth of those functions, thereby making the continuation space finite.
+%
+In the monovariant case, the number of continuations is still linear in the size of the program.
+%%
 
-The tail of a continuation in the context of a function will contain
-the stack-less context in which the function was called, in order to
-link up with the proper call-site(s). The context includes the
-function (or unique label of the function), the environment, and the
-store. Notice that this choice can be viewed as a special allocation
-strategy for continuations in the AAM viewpoint, but we separate the
-table of continuations from the store since continuations themselves
-contain stores in $\rt$ frames --- this leads to a recursive
-data-structure that we are trying to avoid. The store is an important
-ingredient to maintaining enough precision to keep a pushdown
-abstraction. The semantic spaces for the machine are modified thusly:
+%%
+The tail of a continuation in the context of a function will contain the stack-less context in which the function was called, in order to link up with the proper call-site(s).
+%
+The context includes the function (or unique label of the function), the environment, and the store.
+%
+Notice that this choice can be viewed as a special allocation strategy for continuations in the AAM viewpoint, but we separate the table of continuations from the store since continuations themselves contain stores in $\rt$ frames --- this leads to a recursive data-structure that we are trying to avoid.
+%
+The store is an important ingredient to maintaining enough precision to keep a pushdown abstraction.
+%
+The semantic spaces for the machine are modified thusly:
 
@@
   \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \times \KTab \times \Memo \\
-  \mkont \in \Kont &::= \kmt
-                     \mid \kar{\mexpr,\menv,\mkont} 
-                     \mid \kfn{\mval,\mkont}
-                     \mid \krt{(\mexpr,\menv),\mstore}\\
+  \mkont \in \Kont &::= \kmt \mid \krt{(\mexpr,\menv),\mstore} \mid \kcons{\mkframe}{\mkont} \\
   \mktab \in \KTab &= (\Expr \times \Env)\times \Store \to \wp(\Kont) \\
@@
 
-\paragraph{The role of summaries:} notice the additional $\Memo$
-component. A ``summary edge'' in CFA2 or equivalently an
-$\epsilon$-edge in PDCFA is an edge from the source of a push edge to
-the target of the matching pop edge. ``Matching'' here means there is
-a path through machine reductions that don't change the stack, or
-through summary edges. In our model, we only ``push'' when we call a
-function, and we only ``pop'' when we return with a value. There is an
-analogy to something more operational: summary edges embody
-memoization (see the last rule in figure
-\ref{fig:summary-semantics}). Instead of following an entire path
-through a call to return a value, we simply jump from the call to the
-return with the result of the call, with the resulting context changes---in this case only the store (the second case of rule 4).
+\paragraph{The role of summaries:} notice the additional $\Memo$ component.
+%
+A ``summary edge'' in CFA2 or equivalently an $\epsilon$-edge in PDCFA is an edge from the source of a push edge to the target of the matching pop edge.
+%
+``Matching'' here means there is a path through machine reductions that don't change the stack, or through summary edges.
+%
+In our model, we only ``push'' when we call a function, and we only ``pop'' when we return with a value.
+%
+There is an analogy to something more operational: summary edges embody memoization (see the last rule in figure \ref{fig:summary-semantics}).
+%
+Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store (the second case of rule 4).
 
@@
     &
-    $\tpl{(\mexpri0, \menv), \mstore, \kar{\mexpri1,\menv,\mkont}, \mktab, \mmemo}$
+    $\tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}$
     \\
 % argument evaluation
-    $\tpl{\mval, \mstore, \kar{\mexpr,\menv,\mkont}, \mktab, \mmemo}$
+    $\tpl{\mval, \mstore, \kcons{\kar{\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
     &
-    $\tpl{(\mexpr, \menv), \mstore, \kfn{\mval, \mkont}, \mktab, \mmemo}$
+    $\tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}$
     \\
 % function call
-    $\tpl{\mval,\mstore,\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont},\mktab,\mmemo}$
+    $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo}$
     & % actually do call,
@@
 
-To turn this semantics into PDCFA's algorithm for constructing a Dyck
-state graph, given the appropriate $\alloc$, we apply a widening
-operator to make $\mstore$, $\mkont$, and $\mmemo$ shared amongst all
-states (meaning states then become represented without these
-compontents), in what is then called a $\System$.  Since these
-components are shared, they are always the least upper bound of any
-respective component that the semantics produces (the intermediate set
-$I$) in order to stay sound. We use a meta-function $\wn$, ``wide to
-narrow'', to shift between the two representations of states.
+To turn this semantics into PDCFA's algorithm for constructing a Dyck state graph, given the appropriate $\alloc$, we apply a widening operator to make $\mstore$, $\mkont$, and $\mmemo$ shared amongst all states (meaning states then become represented without these compontents), in what is then called a $\System$.
+%
+Since these components are shared, they are always the least upper bound of any respective component that the semantics produces (the intermediate set $I$) in order to stay sound.
+%
+We use a meta-function $\wn$, ``wide to narrow'', to shift between the two representations of states.
 
@@
   \mastate \in \sa{State} &= (\Expr \times \Env) \times \Kont \\
-  \System &= \wp(\sa{State} \times \Store) \times \wp(\sa{State}) \\
-          &\phantom{=} \times \Store \times \KTab \times \Memo \\
+  \System &= \wp(\sa{State} \times \Store) \times \wp(\sa{State}) \times \Store \times \KTab \times \Memo \\
   {\mathcal F} &: \System \to \System \\
@@
 
-A $\System$ embodies the states seen and at which store, $S$, the
-frontier set of states yet to analyze, $F$, the shared store for the
-frontier, $\mstore$, the continuation table $\mktab$ and the memo
-table $\mmemo$. The memo table need only map to values, since the
-global store will always subsume the store that would otherwise be
-mapped. All the states in the frontier are stepped with the current
-store, after which the next store is the least upper bound of all the
-resulting stores. The next frontier contains only states resulting
-from stepping the previous frontier that we haven't seen at this next
-store. Systematic techniques for a performant implementation can be
-found in \citet{ianjohnson:oaam:icfp2013}.
+A $\System$ embodies the states seen and at which store, $S$, the frontier set of states yet to analyze, $F$, the shared store for the frontier, $\mstore$, the continuation table $\mktab$ and the memo table $\mmemo$.
+%
+The memo table need only map to values, since the global store will always subsume the store that would otherwise be mapped.
+%
+All the states in the frontier are stepped with the current store, after which the next store is the least upper bound of all the resulting stores.
+%
+The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.
+%
+Systematic techniques for a performant implementation can be found in \citet{ianjohnson:oaam:icfp2013}.
 
-\paragraph{Example:} Let's consider a monovariant allocation strategy
-(the address for each binding is the variable name itself) to run the
-following example:
+\paragraph{Example:} Let's consider a monovariant allocation strategy (the address for each binding is the variable name itself) to run the following example:
 \begin{lstlisting}[mathescape]
@@
 \end{lstlisting}
-Suppose we extend our semantics to allow numbers, numeric primitives
-and \texttt{let}. The continuation frame for a \texttt{let} contains
-the identifier to bind to the resulting value, along with the body of
-the \texttt{let} with its environment. Call the constructor of this
-frame $\mathbf{lt}$.
+Suppose we extend our semantics to allow numbers, numeric primitives and \texttt{let}.
+%
+The continuation frame for a \texttt{let} contains the identifier to bind to the resulting value, along with the body of the \texttt{let} with its environment.
+%
+Call the constructor of this frame $\mathbf{lt}$.
+%%
 
-We should expect that a pushdown analysis would predict this evaluates
-to true, and there are no loops in the program. 0CFA ~\citep{ianjohnson:Shivers:1991:CFA} claims there is a
-loop from the second call of \texttt{id} to the first, and thusly
-predicts this program evaluates to true or false. PDCFA claims there
-is no loop, and depending on the implementation, that the result is
-true or false (paper), or just true (implemented)\footnote{This is
-  because the paper steps every seen state with the current store
-  every iteration, but the implementation only steps states that need stepping.}. Let's take a look at the evaluation after binding
-\texttt{id} ($\mstore_0 = [\texttt{id} \mapsto
-\set{\vclo{\slam{\texttt{x}}{\texttt{x}}}{\bot}}]$):
+%%
+We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
+%
+0CFA ~\citep{ianjohnson:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
+%
+PDCFA claims there is no loop, and depending on the implementation, that the result is true or false (paper), or just true (implemented)
+\footnote{This is because the paper steps every seen state with the current store every iteration, but the implementation only steps states that need stepping.}.
+%
+Let's take a look at the evaluation after binding \texttt{id} ($\mstore_0 = [\texttt{id} \mapsto \set{\vclo{\slam{\texttt{x}}{\texttt{x}}}{\bot}}]$):
 
@@
 
-We maintained enough context to distinguish the return points of
-\texttt{id} to not rebind \texttt{y} to 1. When determining control
-flow through the expression, we consult $\mmemo$ to continue past
-function calls, so there is no confusion about a back edge from the
-second call to \texttt{id}.
+We maintained enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
+%
+When determining control flow through the expression, we consult $\mmemo$ to continue past function calls, so there is no confusion about a back edge from the second call to \texttt{id}.
+%
 
-\begin{conjecture}
-  Tabular semantics has a bisimulation with the standard semantics with a sound allocation strategy.
-\end{conjecture}
-\begin{proof}
+\begin{theorem}
+  Tabular semantics has a bisimulation with the standard semantics with comparable allocation strategies.
+\end{theorem}
 
-\end{proof}
+This follows directly from the invariants we prove of $\mktab$ and $\mmemo$ in the tabular semantics.
+%
+Allocation strategies are comparable if they produce equal addresses regardless of the differences in the state representations.
+%
+More formally, $\alloc$ and $\alloc^*$ are comparable if this implication holds:
+\begin{mathpar}
+  \inferrule{\mkont \in \unroll{\mktab}{\mkont^*}}
+            {\alloc(\tpl{\mpoint,\mstore,\mkont}) = \alloc^*(\tpl{\mpoint,\mstore,\mkont^*,\mktab,\mmemo})}
+\end{mathpar}
 
-\begin{conjecture}
-  Store-widened tabular semantics has a bisimulation with store/memo/k-table-widened semantics.
-\end{conjecture}
+The $\mathit{unroll}$ function interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$:
+\begin{mathpar}
+  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt}} \quad
+  \inferrule{\mkont \in \unroll{\mktab}{\mkont'}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont'}} \quad
+  \inferrule{\mkont^\circ \in \mktab(\mctx) \\
+             \mkont \in \unroll{\mktab}{\mkont^\circ}}
+            {\mkont \in \unroll{\mktab}{\krt{\mctx}}}
+\end{mathpar}
+
+The astute reader may find this definition is dubious, since there is a possibility for unguarded corecursion here.
+%
+A function with ill-founded recursion can lead to a table such that $\krt{\mctx} \in \mktab(\mctx)$.
+%
+The abstractions that we impose on our data might also cause this situation, even though the concrete execution might always terminate.
+%
+Our interpretation of $\stepto$ is in a fixed-point computation, so circularities are not a problem for termination.
+
+\begin{lemma}[Stack irrelevance]\label{lem:irrelevance}
+  $\hastail(\mtrace,\mkont)$ implies $\replacetail(\mtrace,\mkont,\mkont')$ is a valid trace.
+\end{lemma}
+\begin{mathpar}
+  \inferrule{ }{\hastail(\epsilon,\mkont)} \quad
+  \inferrule{ }{\hastail(\tpl{\mpoint,\mstore,\append{\mkont'}{\mkont}},\mkont)} \quad
+  \inferrule{\hastail(\mtrace\mstate,\mkont) \quad
+             \mstate \stepto \mstate' \quad
+             \hastail(\mstate',\mkont)}
+            {\hastail(\mtrace\mstate\mstate',\mkont)}
+\end{mathpar}
+
+\begin{align*}
+  \replacetail(\tpl{\mpoint,\mstore,\append{\mkont'}{\mkont}},\mkont,\mkont'') &= \tpl{\mpoint,\mstore,\append{\mkont'}{\mkont''}} \\
+  \replacetail(\epsilon,\mkont,\mkont'') &= \epsilon \\
+  \replacetail(\mtrace\mstate,\mkont,\mkont') &= \replacetail(\mtrace,\mkont,\mkont')\replacetail(\mstate,\mkont,\mkont') \\
+\end{align*}
+
+The different tables we have encode information about execution history that we prove is invariant.
+\begin{align*}
+  \inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo}) &=
+          \dom(\mmemo) \subseteq \dom(\mktab) \\
+  &\wedge \forall \mctx\equiv(\mpoint',\mstore') \in \dom(\mktab),
+                  \mkont^\circ \in \mktab(\mctx),
+                  \mkont' \in \unroll{\mktab}{\mkont^\circ}. \\
+  &\qquad \exists \mkont^* \in \unroll{\mktab}{\mkont}.
+            \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont^*} \\
+  &\qquad \wedge \forall (\mval, \mstore') \in \mmemo(\mctx).
+            \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore', \mkont'}
+\end{align*}
+
+\begin{lemma}[Table invariants]\label{lem:tab-inv}
+  If $\inv(\mstate)$ and $\mstate \stepto \mstate'$ then $\inv(\mstate')$.
+\end{lemma}
+
+\begin{theorem}[Equivalence with global tables]\label{thm:global}
+  Tabular semantics has a bisimulation with memo/k-table-widened semantics.
+\end{theorem}
+
+\section{Deriving GC from the introspective pushdown model}
+
+Abstract garbage collection~\citep{dvanhorn:Might:2006:GammaCFA} is a powerful technique for regaining precision in an abstract interpretation.
+%
+The followup to PDCFA~\citet{dvanhorn:Earl2012Introspective} that allows inspecting the stack in order to perform sound GC introduces an entirely new class of automata for this one purpose.
+%
+In our view with tabular semantics, the GC story mimics the simplicity of GC in the regular setting, without the detour into automata theory.
+%
+We follow the GC style of \citet{dvanhorn:Might:2006:GammaCFA} with a stop-and-copy collector.
+%
+Address reachability is in terms of a base ``touches'' function, $\touches$, that is different for our continuations.
+%
+Entries in $\mktab$ might be circular, so we compute the continuation frames in a fixed-point before computing touched addresses:
+\begin{align*}
+  \konts(\mkont_0, \mktab)(K) &= \Phi(\mkont_0) \cup K \cup \bigcup\limits_{\krt{\mctx} \in K,\, \mkont' \in \mktab(\mctx)}{\Phi(\mkont')}
+\\
+  \touches(\mkont, \mktab) &= \bigcup\limits_{\mkframe \in \lfp(\konts(\mkont, \mktab))}{\touchesf(\mkframe)}
+\\[2pt]
+  \Phi(\kcons{\mkframe}{\mkont}) &= \set{\mkframe} \cup \Phi(\mkont) \\
+  \Phi(\krt{\mctx}) &= \set{\krt{\mctx}} \\
+  \Phi(\kmt) &= \emptyset 
+\\[2pt]
+  \touchesf(\kar{\mexpr,\menv}) &= \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)}) \\
+  \touchesf(\kfn{(\slam{\mvar}{\mexpr},\menv)}) &= \mathit{rng}(\menv|_{\mathit{fv}(\slam{\mvar}{\mexpr})})
+\end{align*}
+
+Garbage collection ($\Gamma$) amounts to restricting the heap to just the reachable addresses, so a garbage-collecting semantics introduces the following reduction rule:
+\begin{align*}
+  \mstate &\stepto \Gamma(\mstate) \\
+  \Gamma(\tpl{(\mexpr,\menv),\mstore,\mkont, \mktab, \mmemo}) &= \tpl{(\mexpr,\menv),\mstore|_{\reaches(\mathit{root}, \mstore)}, \mkont, \mktab, \mmemo} \\
+\end{align*}
+%
+The ``reaches'' metafunction ($\reaches$) is a transitive closure of ``touches'' as it weaves through the store, starting from some root set of addresses:
+
+\begin{align*}
+  \text{where } \mathit{root} &= \touches(\mkont,\mktab) \cup \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)}) \\
+  \reaches(\mathit{root}, \mstore) &= \setbuild{b}{\maddr \in \mathit{root}, \maddr \rightsquigarrow_\mstore^* b} \\
+   &
+               \inferrule{\mval \in \mstore(\maddr) \\
+                             b \in \touches(\mval)}
+                            {\maddr \rightsquigarrow_\mstore b}
+\end{align*}
 
@@
 
-CFA2 is the first published analysis of a higher-order programming
-language that could properly match calls and returns. We will show
-that it fits well into the same presentation we gave for
-PDCFA. Vardoulakis and Shivers had a clear goal of harnessing the
-extra information a pushdown model provides to produce a
-high-precision analysis that works well in practice. This resulted in
-more than just the call/return matching of the previous section, which
-is why we are showing the two separately. There are two orthogonal
-features of the semantics:
+CFA2 is the first published analysis of a higher-order programming language that could properly match calls and returns.
+%
+We will show that it fits well into the same presentation we gave for PDCFA.
+%
+Vardoulakis and Shivers had a clear goal of harnessing the extra information a pushdown model provides to produce a high-precision analysis that works well in practice.
+%
+This resulted in more than just the call/return matching of the previous section, which is why we are showing the two separately.
+%
+There are two orthogonal features of the semantics:
 \begin{enumerate}
@@
 \end{enumerate}
-The first of these is an addition to the stack-less context. There is
-a conservative pre-analysis that checks locally whether a binding will
-never escape, and classifies references (labeled with a distinguishing
-$\mlab$ from an arbitrary space of labels) as able to use the stack frame or not. Their criteria for a
-binding never escaping is that it is never referenced in a function
-that is not its binder. This can be extended in a language with more
-linguistic features; see Kranz's thesis about register-allocatable
-bindings in the Orbit Scheme compiler
-\citep{ianjohnson:kranz:thesis:1988}. A stackable reference is one
-that appears in the body of the binding function, by which we mean not
-within the body of a nested function. They use the information that a
-binding never escapes to not bother allocating it in the heap. This
-has the advantage of not changing the heap, and thus leads to less
-propagation. The addition of these stack frames makes the analysis
-exponential in theory, though in practice they have been observed to
-decrease running time in most cases.
+The first of these is an addition to the stack-less context.
+%
+There is a conservative pre-analysis that checks locally whether a binding will never escape, and classifies references (labeled with a distinguishing $\mlab$ from an arbitrary space of labels) as able to use the stack frame or not.
+%
+Their criteria for a binding never escaping is that it is never referenced in a function that is not its binder.
+%
+This can be extended in a language with more linguistic features; see Kranz's thesis about register-allocatable bindings in the Orbit Scheme compiler \citep{ianjohnson:kranz:thesis:1988}.
+%
+A stackable reference is one that appears in the body of the binding function, by which we mean not within the body of a nested function.
+%
+They use the information that a binding never escapes to not bother allocating it in the heap.
+%
+This has the advantage of not changing the heap, and thus leads to less propagation.
+%
+The addition of these stack frames makes the analysis exponential in theory, though in practice they have been observed to decrease running time in most cases.
+%%
 
-The second of these is to ameliorate a problem they call ``fake
-rebinding.'' That is, since bindings in the abstract represent several
-values, we don't want to reference a variable \texttt{x} in two
-different places and have it resolve to two different values. In AAM,
-a variable reference non-deterministically steps to all possible
-values associated with that variable. Here we want to say that once
-\texttt{x} is considered to stand for value $v$, then all subsequent
-references of \texttt{x} should be $v$. If they aren't, it looks as if
-\texttt{x} was rebound; it hasn't, and thus it is a ``fake
-rebinding.'' CFA2 does not step to all values on variable reference,
-but instead carries all its values around in superposition until they
-need to be observed at, say, a function call. We give a simplified
-semantics that is more along the AAM style, but CFA2's approach can
-easily be recovered from it\footnote{Our Redex model implements the
-  fake rebinding strategy CFA2 itself employs}.
+%%
+The second of these is to ameliorate a problem they call ``fake rebinding.''
+%
+That is, since bindings in the abstract represent several values, we don't want to reference a variable \texttt{x} in two different places and have it resolve to two different values.
+%
+In AAM, a variable reference non-deterministically steps to all possible values associated with that variable.
+%
+Here we want to say that once \texttt{x} is considered to stand for value $v$, then all subsequent references of \texttt{x} should be $v$.
+%
+If they aren't, it looks as if \texttt{x} was rebound; it hasn't, and thus it is a ``fake rebinding.''
+%
+CFA2 does not step to all values on variable reference, but instead carries all its values around in superposition until they need to be observed at, say, a function call.
+%
+We give a simplified semantics that is more along the AAM style, but CFA2's approach can easily be recovered from it\footnote{Our Redex model implements the fake rebinding strategy CFA2 itself employs}.
+%%
 
-CFA2 uses a ``local semantics'' that is similar to our segmenting
-continuations at function boundaries, but gets stuck when it gets to a
-point where it would need to ``return.'' It instead appeals to an
-external, imperative algorithm for summarization to sew the function
-calls and returns together by pattern matching on the states that the
-local semantics produces. What we show here is almost wholly the same
-in character, only embodied still in terms of a machine semantics and
-thus more easily reasoned about, and can be run in the concrete.
+%%
+CFA2 uses a ``local semantics'' that is similar to our segmenting continuations at function boundaries, but gets stuck when it gets to a point where it would need to ``return.''
+%
+It instead appeals to an external, imperative algorithm for summarization to sew the function calls and returns together by pattern matching on the states that the local semantics produces.
+%
+What we show here is almost wholly the same in character, only embodied still in terms of a machine semantics and thus more easily reasoned about, and can be run in the concrete.
+%%
 
-We show only the significantly modified rules of the semantics in
-figure \ref{fig:frame-semantics}. The other rules simply carry along
-the extra $\mframe$ component untouched.
+%%
+We show only the significantly modified rules of the semantics in figure \ref{fig:frame-semantics}.
+%
+The other rules simply carry along the extra $\mframe$ component untouched.
 \begin{figure}
@@
 % function call
-    $\tpl{\mval,\mstore,\mframe,\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}, \mkont}}$
+    $\tpl{\mval,\mstore,\mframe,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont}}$
     &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore', \mframe', \mkont}$
+    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore', \mframe', \kcons{\kpush{\mframe}}{\mkont}}$
     \\ where & $(\mstore',\mframe') = \bind(\mstore,\mframe,\maddr,\mvar,\mval)$
+    \\
+    $\tpl{\mval,\mstore,\mframe,\kcons{\kpush{\mframe'}}{\mkont}}$
+    &
+    $\tpl{\mval,\mstore,\mframe',\mkont}$
   \end{tabular}
@@
 
-Sidestepping fake rebinding does not need to be restricted to
-stackable references, but that is what CFA2 does. Indeed, as soon as
-the non-determinism has been determined, we could extend the stack
-frame so any subsequent reference means what it meant previously in
-the function. Look-up would then always try the stack frame before
-falling back on the heap (of course this must be limited to immutable addresses).
+Sidestepping fake rebinding does not need to be restricted to stackable references, but that is what CFA2 does.
+%
+Indeed, as soon as the non-determinism has been determined, we could extend the stack frame so any subsequent reference means what it meant previously in the function (reminiscent of how equality information of unknowns is tracked in \citet{ianjohnson:DBLP:journals/cacm/DilligDA10}).
+%
+Look-up would then always try the stack frame before falling back on the heap (of course this must be limited to immutable addresses).
+%%
 
-CFA2 also includes what they called ``transitive summaries'' to deal
-with tail calls. We insert an $\rt$ frame at every function call.
-This view contends with tail calls, but we can identify tail calls
-easily --- any call with an $\rt$ as its continuation is a tail
-call. Tail calls are important in language implementations for space
-complexity reasons \citep{ianjohnson:clinger:tail-calls:1998}, but in
-an analysis, these concerns are less important. The repeated popping
-of $\rt$ inserted by what otherwise were tail-calls is synonymous with
-CFA2's transitive summaries.
+%%
+CFA2 also includes what they called ``transitive summaries'' to deal with tail calls.
+%
+We insert an $\rt$ frame at every function call.
+%
+This view contends with tail calls, but we can identify tail calls easily --- any call with an $\rt$ as its continuation is a tail call.
+%
+Tail calls are important in language implementations for space complexity reasons \citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these concerns are less important.
+%
+The repeated popping of $\rt$ inserted by what otherwise were tail-calls is synonymous with CFA2's transitive summaries.
 
-\paragraph{Example:} if we use this semantics to run through the
-example of the previous section, we find that with the addition of
-stack frames, there is no confusion about \texttt{z}'s value either,
-so if the operator weren't $\le$ but instead $<$ or $+$, we could
-still constant fold that away.
+\paragraph{Example:} if we use this semantics to run through the example of the previous section, we find that with the addition of stack frames, there is no confusion about \texttt{z}'s value either, so if the operator weren't $\le$ but instead $<$ or $+$, we could still constant fold that away.
+%%
 
-\begin{conjecture}
+%%
+The stack frame introduces a refinement over the original CESK machine, so we first prove it correct with respect to the concrete semantics.
+%
+We follow up with a similar proof to that of PDCFA, which is that the CES$\xi$K machine has a bisimulation with its tabular semantics, with comparable allocation strategies.
+\begin{theorem}[Correctness of stack frame refinement]\label{thm:refinement}
+  The CES$\xi$K machine has a bisimulation with the CESK machine with fresh allocation.
+\end{theorem}
+This follows from the invariant that if a stack frame maps $\maddr$ to $\mval$, then the store maps $\maddr$ to $\mval$, meaning that the special lookup function is equivalent to lookup directly in the store (exploits immutability).
+%
+That is, if we always allocated bindings in the heap.
+%
+The additional escape information that allows us to elide allocation in the heap introduces a second level of reasoning, where the burden of proof falls on the escape analysis.
+%
+The additional escape analysis adds information to use in the proof, that if $\mvar$ never escapes, then every reference to allocations of $\mvar$'s binding will always be able to use $\mframe$ for the lookup.
+%
+With this proposition, we can prove that the CES$\xi$K machine that always allocates in the store has a bisimulation with the refined machine that exploits escape analysis to avoid unnecessary changes to the store.
+%
+
+\begin{theorem}[Correctness of CFA2]\label{thm:cfa2}
   Tabular semantics has a bisimulation with standard semantics.
-\end{conjecture}
+\end{theorem}
+This proof has the same structure as that of PDCFA.
 
@@
 
-There is contention among programming language researchers whether
-\rackett{call/cc} should be a language primitive, since it captures
-the entire stack, leading to space leaks
-\citep{ianjohnson:kiselyov:against-callcc}. Alternative control
-operators have been proposed that delimit how much of the stack to
-capture, such as $\%$ (read ``prompt'') and capture operator
-${\mathcal F}$ (read ``control'')
-\citep{ianjohnson:felleisen:control:1988}, or \texttt{reset}
-(equivalent to $\%$) and \texttt{shift}
-\citep{ianjohnson:danvy:filinski:delim:1990}. However, the stacks
-captured by these operators always extend the stack when invoked,
-rather than replace it like those captured with
-\rackett{call/cc}. Continuations that have this extension behavior are
-called ``composable continuations.'' Stack replacement is easily
-modeled in a regular analysis using the AAM approach, and Vardoulakis
-and Shivers showed it can be done in a pushdown approach (although it
-breaks the pushdown model). Stack extension, however, poses a new
-challenge for pushdown analysis, since one application of a composable
-continuation means pushing an unbounded amount of frames onto the
-stack. Vardoulakis' and Shivers' approach does not immediately apply
-in this situation, since their technique drops all knowledge of the
-stack at a continuation's invocation site; extension, however, must
-preserve it.
+There is contention among programming language researchers whether \rackett{call/cc} should be a language primitive, since it captures the entire stack, leading to space leaks \citep{ianjohnson:kiselyov:against-callcc}.
+%
+Alternative control operators have been proposed that delimit how much of the stack to capture, such as $\%$ (read ``prompt'') and capture operator ${\mathcal F}$ (read ``control'') \citep{ianjohnson:felleisen:control:1988}, or \texttt{reset} (equivalent to $\%$) and \texttt{shift} \citep{ianjohnson:danvy:filinski:delim:1990}.
+%
+However, the stacks captured by these operators always extend the stack when invoked, rather than replace it like those captured with \rackett{call/cc}.
+%
+Continuations that have this extension behavior are called ``composable continuations.''
+%
+Stack replacement is easily modeled in a regular analysis using the AAM approach, and Vardoulakis and Shivers showed it can be done in a pushdown approach (although it breaks the pushdown model).
+%
+Stack extension, however, poses a new challenge for pushdown analysis, since one application of a composable continuation means pushing an unbounded amount of frames onto the stack.
+%
+Vardoulakis' and Shivers' approach does not immediately apply in this situation, since their technique drops all knowledge of the stack at a continuation's invocation site; extension, however, must preserve it.
+%%
 
-The way we have been splitting continuations at function calls has
-similarities to the meta-continuation approach to modeling delimited
-control, given in figure \ref{fig:shift-reset} (adapted from
-~\citep{ianjohnson:Biernacki2006274}). We could view each function
-call as inserting a prompt, and returns as aborting to the nearest
-prompt. Resets then insert a second-tier prompt. The changed semantic
-spaces for the shift/reset semantics are as follows:
+%%
+The way we have been splitting continuations at function calls has similarities to the meta-continuation approach to modeling delimited control, given in figure \ref{fig:shift-reset} (adapted from~\citep{ianjohnson:Biernacki2006274}).
+%
+We could view each function call as inserting a prompt, and returns as aborting to the nearest prompt.
+%
+Resets then insert a second-tier prompt.
+%
+The changed semantic spaces for the shift/reset semantics are as follows:
 
@@
     $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore',\kmt,\mmkont}$
-    \\ & where $\mstore' = \joinone{\mstore}{\maddr}{\mkont}$
+    \\ & where $\mstore' = \joinone{\mstore}{\maddr}{\vcomp{\mkont}}$
     \\
 % continuation call
-    $\tpl{\mval,\mstore,\kfn{\mkont', \mkont}, \mmkont}$
+    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont}$
     &
@@
 
-Turning this into a table-based semantics involves making prompts a
-point of indirection, just like function calls. Memoization also gets
-a new context to consider, calling a continuation, because composable
-continuations act like functions. The new semantic spaces are then
+Turning this into a table-based semantics involves making prompts a point of indirection, just like function calls.
+%
+Memoization also gets a new context to consider, calling a continuation, because composable continuations act like functions.
+%
+The new semantic spaces are then
 
@@
   \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid (\vcomp{\mkont},\mval,\mstore) \\
-  \mkont \in \Kont &= \kmt \mid \krt{\mctx} \mid \kar{\mexpr,\menv,\mkont} \mid \kfn{\mval,\mkont} \\
+  \mkont \in \Kont &= \kmt \mid \krt{\mctx} \mid \kcons{\mkframe}{\mkont} \\
   \mmkont \in \MKont &::= \kmt \mid \kprompt{\mctx} \\
@@
 
-Figure \ref{fig:shift-reset-table0} has what one might naturally write
-using just this information. Unfortunately, since $\rt$ frames contain
-a store, and continuations can now appear in the store, this
-introduces a circularity that could cause the analysis to never
-terminate. For example, a loop that continually captures continuations
-would introduce unboundedly many continuation values due to the
-ever-growing store. This means the store is not finite height, and
-there may be no fixed point.
+Figure \ref{fig:shift-reset-table0} has what one might naturally write using just this information.
+%
+Unfortunately, since $\rt$ frames contain a store, and continuations can now appear in the store, this introduces a circularity that could cause the analysis to never terminate.
+%
+For example, a loop that continually captures continuations would introduce unboundedly many continuation values due to the ever-growing store.
+%
+This means the store is not finite height, and there may be no fixed point.
 
-\paragraph{Possible fixes to non-termination:} The simplest fix to the
-circularity would be to strip stores out of the context of the $\rt$
-frame in a captured continuation, and additionally consider contexts
-with a stripped store an abstraction of the same context with any
-store. This is a fairly brutal approximation of a captured
-continuation, so an alternative is to make this tunable with our
-precision-tuning friend, $\alloc$. Then, instead of entirely removing
-the store component of the $\rt$ context, we can replace it with an
-address of possible stores that it would approximate.
+\paragraph{Possible fixes to non-termination:} The simplest fix to the circularity would be to strip stores out of the context of the $\rt$ frame in a captured continuation, and additionally consider contexts with a stripped store an abstraction of the same context with any store.
+%
+This is a fairly brutal approximation of a captured continuation, so an alternative is to make this tunable with our precision-tuning friend, $\alloc$.
+%
+Then, instead of entirely removing the store component of the $\rt$ context, we can replace it with an address of possible stores that it would approximate.
 
@@
     \\
-    where & $\mctx = (\mkont', \mval, \mstore)$ \\
+    & or \\
+    & $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab',\mmemo}$
+    \\ & \quad if $(\mval',\mstore') \in \mmemo(\mctx)$
+    \\
+    where & $\mctx = (\vcomp{\mkont'}, \mval, \mstore)$ \\
           & $\mktab' = \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}$
@@
 % continuation call
-    $\tpl{\mval,\mstore,\kfn{\vcomp{\mkont'}, \mkont}, \mmkont, \mktab, \mmemo}$
+    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont, \mktab, \mmemo}$
     &
@@
 
-The new indirection possibility changes $\Context$ to also include
-$\maddr$ where there was previously a $\mstore$ (though contexts for
-continuation calls remain unchanged), and $\KTab$ now additionally
-maps $\Addr$ to a set of stores. The rules that change are presented in figure \ref{fig:shift-reset-table1}.
+The new indirection possibility changes $\Context$ to also include $\maddr$ where there was previously a $\mstore$ (though contexts for continuation calls remain unchanged), and $\KTab$ now additionally maps $\Addr$ to a set of stores.
+%
+The rules that change are presented in figure \ref{fig:shift-reset-table1}.
 
 \begin{align*}
-  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid ((\mexpr,\menv),\maddr)
-                     \\ & \mid (\vcomp{\mkont},\mval,\mstore) \\
+  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid (\vcomp{\mkont},\mval,\mstore) \mid \msctx \\
   \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
@@
    \\
-   where & $\mmemo' = \memoize(\mstore,\mmemo, \mktab, \mctx, \mval)$
+   where & $\mmemo' = \mmemo \sqcup \memoize(\mktab, \mctx, (\mval, \mstore))$
   \end{tabular}
@@
 
-The meta-functions mentioned in the fixed semantics all deal with the
-addition of $\maddr$ to contexts. If the context in an $\rt$ frame is
-approximate, we must return to all the continuations known for all the
-contexts it approximates:
+The meta-functions mentioned in the fixed semantics all deal with the addition of $\maddr$ to contexts.
+%
+If the context in an $\rt$ frame is approximate, we must return to all the continuations known for all the contexts it approximates:
 \begin{align*}
@@
 
-At return boundaries, the memo
-table must add the result to all the represented contexts:=
+At return boundaries, the memo table must add the result to all the represented contexts:=
 \begin{align*}
-  \memoize(\mstore,\mmemo, \mktab, ((\mexpr,\menv), \mstore'), \mval) &=
-  \joinone{\mmemo}{((\mexpr,\menv), \mstore')}{(\mval,\mstore')} \\
-  \memoize(\mstore,\mmemo, \mktab, ((\mexpr,\menv), \maddr), \mval) &=
-  \mmemo \sqcup \bigsqcup\limits_{\mstore' \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore')}{(\mval,\mstore)}
+  \memoize(\mktab, ((\mexpr,\menv), \mstore'), \mathit{v\sigma}) &=
+  [((\mexpr,\menv), \mstore') \mapsto \set{{\mathit{v\sigma}}}] \\
+  \memoize(\mktab, ((\mexpr,\menv), \maddr), \mathit{v\sigma}) &=
+  \bigsqcup\limits_{\mstore' \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore')}{\mathit{v\sigma}}
 \end{align*}
-At capture time, we strip the store in the $\rt$ frame
-if there is one, and replace it with an address:
+At capture time, we strip the store in the $\rt$ frame if there is one, and replace it with an address:
 
@@
    \replacectx(\krt{((\mexpr,\menv),\_)}) &= \krt{((\mexpr,\menv),\maddr)} \\
-   \replacectx(\kar{\mexpr,\menv,\mkont}) &= \kar{\mexpr,\menv,\replacectx(\mkont)} \\
-   \replacectx(\kfn{\mval, \mkont}) &= \kfn{\mval,\replacectx(\mkont)}
+   \replacectx(\kcons{\mkframe}{\mkont}) &= \kcons{\mkframe}{\replacectx(\mkont)}
   \\[2pt]
@@
    \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mktab}{\maddr}{\mktab(\maddr')} \\
-   \addstore(\kar{\mexpr,\menv,\mkont}) &= \addstore(\kfn{\mval, \mkont}) = \addstore(\mkont)
+   \addstore(\kcons{\mkframe}{\mkont}) &= \addstore(\mkont)
 \end{align*}
 
-The result of this viewpoint is a sound, precise, and computable
-semantics for a ``pushdown approach'' to analyzing delimited,
-composable control. To handle non-composable control, we can simply
-remove the prompts and keep $\approximate$.
+The result of this viewpoint is a sound, precise, and computable semantics for a ``pushdown approach'' to analyzing delimited, composable control.
+%
+To handle non-composable control, we can simply remove the prompts and keep $\approximate$.
 
-\begin{conjecture}
+\begin{conjecture}\label{thm:concrete-sr}
   Tabular semantics has a bisimulation with standard semantics given fresh allocation.
@@
 
-\begin{conjecture}
+We prove this with the invariant that the continuation and meta-continuation have unique, well-founded unrollings in addition to the table invariants from \autoref{sec:pdcfa}.
+
+\begin{conjecture}\label{thm:sound-sr}
   Tabular semantics simulates standard semantics with abstract allocation.
@@
 \begin{lstlisting}
-  mumble
+???
+\end{lstlisting}
+
+\begin{lstlisting}
+???
 \end{lstlisting}
@@
 
-The immediately related work is that of PDCFA
-\citep{ianjohnson:earl2010pdcfa},
-CFA2~\citep{ianjohnson:vardoulakis-lmcs11,
-  ianjohnson:Vardoulakis2011Pushdown}, and AAM
-~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we
-recreated in full detail. The version of CFA2 that handles
-\rackett{call/cc} does not handle composable control, and is dependent
-on a restricted CPS representation. They also had no way to tune the
-precision of their first class continuation approximation. Since
-\rackett{call/cc} can be simulated with shift/reset, this work
-supercedes theirs. The extended version of PDCFA that does garbage
-collection \citep{dvanhorn:Earl2012Introspective} also fits into our
-model, although we did not explicitly show it. The addresses that the
-stack keeps alive can be accumulated by ``reading through'' the
-continuation table, building up the set of addresses in each portion
-of the stack that we come across.
+The immediately related work is that of PDCFA \citep{ianjohnson:earl2010pdcfa}, CFA2~\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:Vardoulakis2011Pushdown}, and AAM ~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we recreated in full detail.
+%
+The version of CFA2 that handles \rackett{call/cc} does not handle composable control, and is dependent on a restricted CPS representation.
+%
+They also had no way to tune the precision of their first class continuation approximation.
+%
+Since \rackett{call/cc} can be simulated with shift/reset, this work supercedes theirs.
+%
+The extended version of PDCFA that does garbage collection~\citep{dvanhorn:Earl2012Introspective} also fits into our model;
+the addresses that the stack keeps alive can be accumulated by ``reading through'' the continuation table, building up the set of addresses in each portion of the stack that we come across.
+%%
 
-Pushdown models have existed in the first-order static analysis
-literature ~\citep[Chapter
-7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995},
-and the first-order model checking literature
-\citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time. The
-higher-order setting imposes additional challenges that make their
-methods difficult to adapt. The most important constraint is that we
-can't know all call-sites of a function/method before the analysis
-begins, which their methods heavily rely on.
+%%
+Pushdown models have existed in the first-order static analysis literature ~\citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995}, and the first-order model checking literature \citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time.
+%
+The higher-order setting imposes additional challenges that make their methods difficult to adapt.
+%
+The most important constraint is that we can't know all call-sites of a function/method before the analysis begins, which their methods heavily rely on.
+%%
 
-A new temporal logic that itself understands well-bracketing of call
-and returns for stating and validating propositions, NWTL
-~\citep{ianjohnson:alur:nwtl:2007}, applies to visibly pushdown
-systems, which can precisely model programming languages that only have
-well-bracketed call and returns. The propositional form of the logic
-has decidable satisfiability checking, and control-flow queries are
-shallow propositions to pose in the logic. However, the satisfiability
-problem is \textsc{2Exptime}-hard and the algorithm is not given in
-the on-the-fly form that a higher-order language would need to
-effectively answer control-flow queries without already knowing the
-answers. The logic itself is an exciting new frontier for expressing
-program correctness properties, and a higher-order version would be a
-welcome addition to the functional programmer's tool belt.
+%%
+A new temporal logic that itself understands well-bracketing of call and returns for stating and validating propositions, NWTL ~\citep{ianjohnson:alur:nwtl:2007}, applies to visibly pushdown systems, which can precisely model programming languages that only have well-bracketed call and returns.
+%
+The propositional form of the logic has decidable satisfiability checking, and control-flow queries are shallow propositions to pose in the logic.
+%
+However, the satisfiability problem is \textsc{2Exptime}-hard and the algorithm is not given in the on-the-fly form that a higher-order language would need to effectively answer control-flow queries without already knowing the answers.
+%
+The logic itself is an exciting new frontier for expressing program correctness properties, and a higher-order version would be a welcome addition to the functional programmer's tool belt.
+%%
 
-The trend of deriving static analyses from abstract machines does not
-stop at flow analyses. The model-checking community showed how to
-check temporal logic queries for collapsible pushdown automata (CPDA),
-or equivalently, higher-order recursion schemes, by deriving the
-checking algorithm from the Krivine
-machine~\citep{ianjohnson:Salvati:2011:KMH:2027223.2027239}. The
-expressiveness of CPDAs outweighs that of PDAs, but it is unclear how
-to adapt higher-order recursion schemes to model arbitrary programming
-language features. The method is strongly tied to the simply-typed
-call-by-name lambda calculus and depends on finite sized base-types.
+%%
+The trend of deriving static analyses from abstract machines does not stop at flow analyses.
+%
+The model-checking community showed how to check temporal logic queries for collapsible pushdown automata (CPDA), or equivalently, higher-order recursion schemes, by deriving the checking algorithm from the Krivine machine~\citep{ianjohnson:Salvati:2011:KMH:2027223.027239}.
+%
+The expressiveness of CPDAs outweighs that of PDAs, but it is unclear how to adapt higher-order recursion schemes to model arbitrary programming language features.
+%
+The method is strongly tied to the simply-typed call-by-name lambda calculus and depends on finite sized base-types.
 
@@
 
-As the programming world continues to embrace behavioral values, it
-becomes more important to import the powerful techniques pioneered by
-the first-order analysis and model checking communities. CFA2 and
-PDCFA paved the way, and in large part inspired this work. It is our
-view that systematic approaches to applying the techniques are pivotal
-to scaling them to ``real languages.'' We believe that the recipe that
-this paper set forth is a step in that direction. That is, make
-continuation tables keyed with enough context, and memoize at the
-introduced indirection points. The result in a language with
-well-bracketed control is a ``pushdown analysis'' using
-summarization. In a language without well-bracketed control, we are
-not chained to a pushdown automaton as the target of the
-approximation, so the techniques still apply and give better precision
-than regular methods.
+As the programming world continues to embrace behavioral values, it becomes more important to import the powerful techniques pioneered by the first-order analysis and model checking communities.
+%
+CFA2 and PDCFA paved the way, and in large part inspired this work.
+%
+It is our view that systematic approaches to applying the techniques are pivotal to scaling them to ``real languages.''
+%
+We believe that the recipe that this paper set forth is a step in that direction.
+%
+That is, make continuation tables keyed with enough context, and memoize at the introduced indirection points.
+%
+The result in a language with well-bracketed control is a ``pushdown analysis'' using summarization.
+%
+In a language without well-bracketed control, we are not chained to a pushdown automaton as the target of the approximation, so the techniques still apply and give better precision than regular methods.
+%%
 
-Our goal in the future is to show that this technique is even more
-widely applicable than shift and reset. We conjecture that the same
-recipe will apply to the most intricate control operators in
-production languages such as in
-\citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}. The control
-structures there are difficult to model even with AAM's techniques due
-to the ability to capture and compose continuations with arbitrarily many prompts,
-but once we can tackle AAM, there should be a straightforward route to
-extending it with the pushdown techniques of this paper.
+%%
+Our goal in the future is to show that this technique is even more widely applicable than shift and reset.
+%
+We conjecture that the same recipe will apply to the most intricate control operators in production languages such as in \citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}.
+%
+The control structures there are difficult to model even with AAM's techniques due to the ability to capture and compose continuations with arbitrarily many prompts, but once we can tackle AAM, there should be a straightforward route to extending it with the pushdown techniques of this paper.
+%%
 
-There is also the question of clients of these analyses. Most
-obviously we would want to know where we can implement first-class
-control more efficiently. In particular, we would want to have an
-escape analysis to find first-class continuations that don't need to
-be heap-allocated, and single target abort operations that can be
-turned into long jumps after a computed stack unraveling. On top of
-optimizations, there are security analyses. Greater control of the
-stack gives us the ability to drill deep into context-sensitive
-security properties and make precise predictions.
+%%
+There is also the question of clients of these analyses.
+%
+Most obviously we would want to know where we can implement first-class control more efficiently.
+%
+In particular, we would want to have an escape analysis to find first-class continuations that don't need to be heap-allocated, and single target abort operations that can be turned into long jumps after a computed stack unraveling.
+%
+On top of optimizations, there are security analyses.
+%
+Greater control of the stack gives us the ability to drill deep into context-sensitive security properties and make precise predictions.

### 2013-10-02 content.tex
Fixed the circularity issue with unroll after talking with Mitch.

@@
 %%
-There is a systematic process for transforming off-the-shelf programming language semantics into a form amenable to \emph{regular} analysis that has been widely applied with great success to production programming languages ~\citet{dvanhorn:VanHorn2010Abstracting} (a technique called abstracting abstract machines, or AAM).
+There is a systematic process for transforming off-the-shelf programming language semantics into a form amenable to \emph{regular} analysis that has been widely applied with great success to production programming languages~\citet{dvanhorn:VanHorn2010Abstracting} (a technique called abstracting abstract machines, or AAM).
 %
@@
 \begin{itemize}
-\item{section \ref{sec:pdcfa}: we derive a cousin of PDCFA}
-\item{section \ref{sec:cfa2}: we make additions to the previous semantics to get a direct-style CFA2 without first-class control}
-\item{section \ref{sec:sr}: we give a novel analysis of delimited and composable first-class control.}
+\item{\autoref{sec:pdcfa}: we derive a cousin of PDCFA}
+\item{\autoref{sec:gc}: we show how abstract garbage collection ``falls out'' of this approach.}
+\item{\autoref{sec:cfa2}: we make additions to the PDCFA semantics to get a direct-style CFA2 without first-class control}
+\item{\autoref{sec:sr}: we give a novel analysis of delimited and composable first-class control.}
 \end{itemize}
@@
 %
-Notice that this choice can be viewed as a special allocation strategy for continuations in the AAM viewpoint, but we separate the table of continuations from the store since continuations themselves contain stores in $\rt$ frames --- this leads to a recursive data-structure that we are trying to avoid.
+Notice that this choice can be viewed as a special allocation strategy for continuations in the AAM viewpoint, but we separate the table of continuations from the store since continuations themselves contain stores in $\rt$ continuations --- this leads to a recursive data-structure that we are trying to avoid.
 %
@@
 
-The $\mathit{unroll}$ function interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$:
+The $\mathit{unroll}$ function interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, defined as the greatest fixed point of the following rules:
 \begin{mathpar}
@@
 %
-A function with ill-founded recursion can lead to a table such that $\krt{\mctx} \in \mktab(\mctx)$.
+Interpreting a function with ill-founded recursion can lead to a table such that $\krt{\mctx} \in \mktab(\mctx)$.
 %
@@
 %
-Our interpretation of $\stepto$ is in a fixed-point computation, so circularities are not a problem for termination.
+Our interpretation of $\stepto$ is in a fixed-point computation, so circularities are not a problem for termination, but we augment the definition here to guard against circularities in our proofs:
+
+\begin{mathpar}
+  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt, G}} \quad
+  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \bot}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont', G}} \\
+  \inferrule{G \neq \mctx \\
+             \mkont^\circ \in \mktab(\mctx) \\
+             \mkont \in \unroll{\mktab}{\mkont^\circ, \mctx}}
+            {\mkont \in \unroll{\mktab}{\krt{\mctx}, G}}  
+\end{mathpar}
+Consider later uses of $\mathit{unroll}$ to implicitly give $\bot$ for $G$.
 
@@
 
-\section{Deriving GC from the introspective pushdown model}
+\section{Deriving GC from the introspective pushdown model}\label{sec:gc}
 
@@
 
+This semantics of course enjoys the soundness properties that GC does not introduce dangling pointers, and (in the concrete) any two reduction sequences of the same length lead to states equal up to GC.
+%
+Unconstrained allocation functions
+
 \section{Deriving CFA2}
@@
 %
-There are two orthogonal features of the semantics:
+There are two orthogonal features of the semantics, embodied by our semantics' two metafunctions ($\bind$ and $\lookup$, respectively in \autoref{fig:frame-semantics}):
 \begin{enumerate}
@@
 The first of these is an addition to the stack-less context.
+
+\begin{figure}
+  \centering
+  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
+  \begin{tabular}{r|l}
+    \hline
+% variable lookup
+    $\tpl{(\svar[\mlab]{\mvar}, \menv), \mstore, \mframe, \mkont}$
+    &
+    $\tpl{\mval,\mstore,\mframe',\mkont}$ if $(\mframe', \mval) \in \lookup(\mstore,\mframe,\menv,\mvar,\mlab)$
+    \\
+% % application
+%     $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mframe, \mkont}$
+%     &
+%     $\tpl{(\mexpri0, \menv), \mstore, \mframe, \kar{\mexpri1,\menv,\mkont}}$
+%     \\
+% % argument evaluation
+%     $\tpl{\mval, \mstore, \mframe, \kar{\mexpr,\menv,\mkont}}$
+%     &
+%     $\tpl{(\mexpr, \menv), \mstore, \mframe, \kfn{\mval, \mkont}}$
+%     \\
+% function call
+    $\tpl{\mval,\mstore,\mframe,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont}}$
+    &
+    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore', \mframe', \kcons{\kpush{\mframe}}{\mkont}}$
+    \\ where & $(\mstore',\mframe') = \bind(\mstore,\maddr,\mvar,\mval)$
+    \\
+    $\tpl{\mval,\mstore,\mframe,\kcons{\kpush{\mframe'}}{\mkont}}$
+    &
+    $\tpl{\mval,\mstore,\mframe',\mkont}$
+  \end{tabular}
+  \caption{The CES$\xi$K machine}
+  \label{fig:frame-semantics}
+\end{figure}
+
 %
@@
 The other rules simply carry along the extra $\mframe$ component untouched.
-\begin{figure}
-  \centering
-  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
-  \begin{tabular}{r|l}
-    \hline
-% variable lookup
-    $\tpl{(\svar[\mlab]{\mvar}, \menv), \mstore, \mframe, \mkont}$
-    &
-    $\tpl{\mval,\mstore,\mframe',\mkont}$ if $(\mframe', \mval) \in \lookup(\mstore,\mframe,\menv,\mvar,\mlab)$
-    \\
-% % application
-%     $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mframe, \mkont}$
-%     &
-%     $\tpl{(\mexpri0, \menv), \mstore, \mframe, \kar{\mexpri1,\menv,\mkont}}$
-%     \\
-% % argument evaluation
-%     $\tpl{\mval, \mstore, \mframe, \kar{\mexpr,\menv,\mkont}}$
-%     &
-%     $\tpl{(\mexpr, \menv), \mstore, \mframe, \kfn{\mval, \mkont}}$
-%     \\
-% function call
-    $\tpl{\mval,\mstore,\mframe,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont}}$
-    &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore', \mframe', \kcons{\kpush{\mframe}}{\mkont}}$
-    \\ where & $(\mstore',\mframe') = \bind(\mstore,\mframe,\maddr,\mvar,\mval)$
-    \\
-    $\tpl{\mval,\mstore,\mframe,\kcons{\kpush{\mframe'}}{\mkont}}$
-    &
-    $\tpl{\mval,\mstore,\mframe',\mkont}$
-  \end{tabular}
-  \caption{The CES$\xi$K machine}
-  \label{fig:frame-semantics}
-\end{figure}
-
-The meta-functions the semantics uses to extend and consult the heap
-and stack use implicit information from the pre-analysis we described
-above:
+%
+The meta-functions the semantics uses to extend and consult the heap and stack use implicit information from the pre-analysis we described above:
 
 \begin{align*}
-  \bind(\mstore,\mframe,\maddr,\mvar,\mval) &=
+  \bind(\mstore,\maddr,\mvar,\mval) &=
    \left\{\begin{array}{l}
-            (\mstore, \snglm{\maddr}{\mval}) \text{ if } \mvar \text{ never escapes} \\
+            (\mstore, \snglm{\maddr}{\mval}) \text{ if } \mvar \text{ allocations never escape} \\
             (\joinone{\mstore}{\maddr}{\mval}, \snglm{\maddr}{\mval}) \text{ otherwise} \\
@@
 
+A tabular semantics for the CES$\xi$K machine follows the previous section almost entirely, except that $\mktab$ now stores $(\mkont, \mframe)$ pairs, and returns reinstate the saved $\mframe$:
+\begin{align*}
+  \tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} &\stepto
+  \tpl{\mpoint,
+    \mstore',
+    \mframe',
+    \krt{\mpoint, \mstore'},
+    \mktab',
+    \mmemo} \\
+  \text{where }
+  & \mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr}) \\
+  & (\mstore',\mframe') = \bind(\mstore,\maddr,\mvar,\mval) \\
+  & \mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{(\mkont,\mframe)}
+\\[2pt]
+ \tpl{\mval, \mstore, \mframe, \krt{\mpoint,\mstore'}, \mktab, \mmemo} &\stepto
+ \tpl{\mval, \mstore, \mframe', \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}
+  \\ & \quad\text{if } (\mkont,\mframe') \in \mktab(\mpoint, \mstore')
+\end{align*}
+
 %%
@@
 %
-We insert an $\rt$ frame at every function call.
+We create an $\rt$ continuation at every function call.
 %
@@
 %
-Unfortunately, since $\rt$ frames contain a store, and continuations can now appear in the store, this introduces a circularity that could cause the analysis to never terminate.
+Unfortunately, since $\rt$ continuations contain a store, and continuations can now appear in the store, this introduces a circularity that could cause the analysis to never terminate.
 %
@@
 
-\paragraph{Possible fixes to non-termination:} The simplest fix to the circularity would be to strip stores out of the context of the $\rt$ frame in a captured continuation, and additionally consider contexts with a stripped store an abstraction of the same context with any store.
+\paragraph{Possible fixes to non-termination:} The simplest fix to the circularity would be to strip stores out of the context of the $\rt$ in a captured continuation, and additionally consider contexts with a stripped store an abstraction of the same context with any store.
 %
@@
 %
-If the context in an $\rt$ frame is approximate, we must return to all the continuations known for all the contexts it approximates:
+If the context in an $\rt$ continuation is approximate, we must return to all the continuations known for all the contexts it approximates:
 \begin{align*}
@@
 \end{align*}
-At capture time, we strip the store in the $\rt$ frame if there is one, and replace it with an address:
+At capture time, we strip the store in the $\rt$ continuation if there is one, and replace it with an address:
 

### 2013-10-02 content.tex
Added code to describe the math of touches-kont. Fixed circularity fix.

@@
   \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt, G}} \quad
-  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \bot}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont', G}} \\
-  \inferrule{G \neq \mctx \\
+  \inferrule{\mkont \in \unroll{\mktab}{\mkont', G}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont', G}} \\
+  \inferrule{\mctx \notin G  \\
              \mkont^\circ \in \mktab(\mctx) \\
-             \mkont \in \unroll{\mktab}{\mkont^\circ, \mctx}}
+             \mkont \in \unroll{\mktab}{\mkont^\circ, \set{\mctx}\cup G}}
             {\mkont \in \unroll{\mktab}{\krt{\mctx}, G}}  
 \end{mathpar}
-Consider later uses of $\mathit{unroll}$ to implicitly give $\bot$ for $G$.
+Consider later uses of $\mathit{unroll}$ to implicitly give $\emptyset$ for $G$.
 
@@
 \end{align*}
+
+The first is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
+%
+The second (the existential clause) states that all the calling contexts in the continuation table reach some unrolling of the current state.
+%
+The final invariant states that the memo table is correct: all calling contexts reach the memoized result, with well-bracketing.
 
@@
 \end{align*}
+
+An implementation of the same functionality would use a mutable set to track which contexts have been traversed to build up a set of addresses:
+% This code is typeset below by Scribble
+% \begin{lstlisting}[mathescape]
+% (define (touch-kont $\mkont$ $\mktab$)
+%   (define seen (mutable-seteq))
+%   (let loop ([$\mkont$ $\mkont$])
+%     (match $\mkont$
+%       [(mt) $\emptyset$]
+%       [(cons $\mkframe$ $\mkont$) ($\cup$ $\touchesf(\mkframe)$ (loop $\mkont$))]
+%       [(rt $\mctx$)
+%        (cond
+%          [(set-member? seen $\mctx$) $\emptyset$]
+%          [else
+%           (set-add! seen $\mctx$)
+%           (for/union ([$\mkont$ (in-set $\mktab(\mctx)$)]) (loop $\mkont$))])])))           
+% \end{lstlisting}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Begin scary Scribble output
+\newcommand{\SCodePreSkip}{\vskip\abovedisplayskip}
+\newcommand{\SCodePostSkip}{\vskip\belowdisplayskip}
+\newenvironment{SCodeFlow}{\SCodePreSkip\begin{list}{}{\topsep=0pt\partopsep=0pt%
+\listparindent=0pt\itemindent=0pt\labelwidth=0pt\leftmargin=2ex\rightmargin=2ex%
+\itemsep=0pt\parsep=0pt}\item}{\end{list}\SCodePostSkip}
+\newenvironment{SingleColumn}{\begin{list}{}{\topsep=0pt\partopsep=0pt%
+\listparindent=0pt\itemindent=0pt\labelwidth=0pt\leftmargin=0pt\rightmargin=0pt%
+\itemsep=0pt\parsep=0pt}\item}{\end{list}}
+\newenvironment{RktBlk}{}{}
+\definecolor{IdentifierColor}{rgb}{0.15,0.15,0.50}
+\definecolor{ParenColor}{rgb}{0.52,0.24,0.14}
+\newcommand{\RktSym}[1]{\inColor{IdentifierColor}{#1}}
+\newcommand{\RktPn}[1]{\inColor{ParenColor}{#1}}
+
+\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touch{-}kont}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mktab$}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{seen}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{mutable{-}seteq}\RktPn{)}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{let}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{loop}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{$\mkont$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{]}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xxxx}}}\RktPn{(}\RktSym{match}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{mt}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\emptyset$}\RktPn{]}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{cons}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkframe$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\cup$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touches{-}frame}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkframe$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{loop}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{)}\RktPn{)}\RktPn{]}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{rt}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{ctx}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{(}\RktSym{cond}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{set{-}member{\hbox{\texttt{?}}}}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{seen}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{ctx}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\emptyset$}\RktPn{]}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxxx}}}\RktPn{[}\RktSym{else}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{set{-}add{\hbox{\texttt{!}}}}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{seen}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{ctx}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxxxxxxxxx}}}\RktPn{(}\RktSym{for/union}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{$\mkont$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{in{-}set}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{dict{-}ref}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mktab$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{ctx}\RktPn{)}\RktPn{)}\RktPn{]}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxxxxxxxxxxx}}}\RktPn{(}\RktSym{loop}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{)}\RktPn{)}\RktPn{]}\RktPn{)}\RktPn{]}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% End of scary Scribble output
 

### 2013-10-03 abstract.tex
Theorems for GC and updated abstract.

@@
-Pushdown analysis is better than finite-state analysis in precision and
-performance. Why then have we not seen total widespread adoption of these
-techniques? For one, the known techniques are technically burdened and
-difficult to understand or extend. Control structure of the programming language
-gets pulled into the model of computation, which makes extensions to
-non-pushdown control structures, such as \rackett{call/cc} or \rackett{shift} and
-\rackett{reset}, non-trivial.
+Pushdown analysis improves upon finite-state analysis in precision and performance since it can rule out spurious paths caused by improperly matched calls and returns.
+%
+Despite this, we have not seen total widespread adoption of these techniques for analyzing higher-order languages.
+%
+A possible explanation is that the known techniques are technically burdened and difficult to understand or extend.
+%
+Control structure of the programming language gets pulled into the model of computation, which makes extensions to non-pushdown control structures, such as first-class control operators, non-trivial.
+%%
 
-We show a derivational approach to abstract interpretation that yields
-transparently sound static analyses that can precisely match calls and
-returns when applied to well-known abstract machines. We show that
-adding memoization and segmenting the continuation into bounded pieces
-leads to machines that abstract to static analyses for context-free
-reachability by simply bounding the stores. This technique allows us
-to derive existing, more technically involved analyses, and a novel
-pushdown analysis for delimited, composable control.
+%%
+We show a derivational approach to abstract interpretation that yields transparently sound static analyses that can precisely match calls and returns when applied to well-known abstract machines.
+%
+We show that the abstraction techniques in the literature have an intuitive operational meaning, even in the concrete setting, that build off well-known programming techniques such as memoization.
+%
+This approach allows us to derive the existing, more technically involved analyses, and a novel pushdown analysis for delimited, composable control.

### 2013-10-03 content.tex
Theorems for GC and updated abstract.

@@
 For a concrete semantics, we require the allocation meta-function $\alloc :
-\State \to \Addr$ to return fresh addresses ($\maddr \notin \dom(\mstore)$), but any choice of address is sound.
+\State \to \Addr$ to return fresh addresses ($\alloc(\mstate) \notin \dom(\mstate.\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
 %
@@
 
-To turn this semantics into PDCFA's algorithm for constructing a Dyck state graph, given the appropriate $\alloc$, we apply a widening operator to make $\mstore$, $\mkont$, and $\mmemo$ shared amongst all states (meaning states then become represented without these compontents), in what is then called a $\System$.
+To turn this semantics into PDCFA's algorithm for constructing a Dyck state graph, given the appropriate $\alloc$, we apply a widening operator to make $\mstore$, $\mkont$, and $\mmemo$ shared amongst all states (meaning states then become represented without these components), in what is then called a $\System$.
 %
@@
   \System &= \wp(\sa{State} \times \Store) \times \wp(\sa{State}) \times \Store \times \KTab \times \Memo \\
-  {\mathcal F} &: \System \to \System \\
-  {\mathcal F}(S, F, \mstore, \mktab, \mmemo) &= (S \cup S', F', \mstore', \mktab', \mmemo') \\
+  {\mathcal F} &: \sa{\State} \to \System \to \System \\
+  {\mathcal F}(\mstate_0)(S, F, \mstore, \mktab, \mmemo) &= (\set{(\mstate_0,\bot)} \cup S \cup S', F', \mstore', \mktab', \mmemo') \\
   \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mstore,\mktab,\mmemo) \stepto \mstate'} \\
@@
           \dom(\mmemo) \subseteq \dom(\mktab) \\
-  &\wedge \forall \mctx\equiv(\mpoint',\mstore') \in \dom(\mktab),
-                  \mkont^\circ \in \mktab(\mctx),
+  &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab),
+                  \mkont^\circ \in \mktab(\mpoint',\mstore'),
                   \mkont' \in \unroll{\mktab}{\mkont^\circ}. \\   
-  &\qquad \exists \mkont^* \in \unroll{\mktab}{\mkont}.
+  &\qquad \phi_{\reachable} \wedge \phi_{\memo} \\
+  \text{where } \phi_{\reachable} &=
+  \exists \mkont^* \in \unroll{\mktab}{\mkont}.
     \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont^*} \\
-  &\qquad \wedge \forall (\mval, \mstore') \in \mmemo(\mctx).
-            \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore', \mkont'}
+  \phi_{\memo} &=
+ \forall (\mval, \mstore'') \in \mmemo(\mpoint',\mstore'). \\
+  &\qquad \exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore'', \mkont'}.
+     \hastail(\mtrace,\mkont') \\
 \end{align*}
@@
 %
-The second (the existential clause) states that all the calling contexts in the continuation table reach some unrolling of the current state.
-%
-The final invariant states that the memo table is correct: all calling contexts reach the memoized result, with well-bracketing.
+The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
+%
+The final invariant, $\phi_{\memo}$, states that the memo table is correct: all calling contexts reach the memoized result, with the calling context's continuation in the tail (to justify using stack-irrelevance).
 
@@
 
-This semantics of course enjoys the soundness properties that GC does not introduce dangling pointers, and (in the concrete) any two reduction sequences of the same length lead to states equal up to GC.
+This semantics of course enjoys the soundness properties that GC does not introduce dangling pointers (\autoref{thm:gc-pointers}), equal length traces are equal up to GC (in the concrete) (\autoref{thm:gc-concrete}), and abstractions are sound with respect to collected states (\autoref{thm:gc-sound}).
+%
+\begin{theorem}\label{thm:gc-pointers}
+  All addresses in $\Gamma(\mstate)$ are in the domain of $\Gamma(\mstate).\mstore$.
+\end{theorem}
+
+\begin{theorem}\label{thm:gc-concrete}
+  If $\mstate \stepto^n \mstate'$ and $\mstate \stepto^n \mstate''$ then $\Gamma(\mstate') = \Gamma(\mstate'')$.
+\end{theorem}
+
+\begin{theorem}\label{thm:gc-sound}
+  If $\mstate \stepto \mstate'$ and $\alpha(\Gamma(\mstate)) \sqsubseteq \mastate$ then there is a $\mastate'$ such that $\mastate \astepto \mastate'$ and $\alpha(\Gamma(\mstate')) \sqsubseteq \mastate'$.
+\end{theorem}
 %
@@
 \begin{align*}
-  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid (\vcomp{\mkont},\mval,\mstore) \mid \msctx \\
   \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
+  \mctx \in \Context &::= \msctx \mid \ldots \\
   \mskont \in \SKont &= \krt{\msctx} \mid \ldots \\
   \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \mid \vcomp{\mskont} \\
-  \mmkont \in \MKont &::= \kmt \mid \kprompt{\mctx} \\
   \mktab \in \KTab &= (\Context \to \mathit{KTabRange}) \\
@@
 
-At return boundaries, the memo table must add the result to all the represented contexts:=
+At return boundaries, the memo table must add the result to all the represented contexts:
 \begin{align*}
@@
 %
-Since \rackett{call/cc} can be simulated with shift/reset, this work supercedes theirs.
+Since \rackett{call/cc} can be simulated with shift/reset\footnote{\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{call/cc}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{f}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{shift}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{k}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{k}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{f}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{x}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{shift}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{c}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{k}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{x}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}}, this work supercedes theirs (the space complexity difference imposed by the simulation is non-issue in the abstract).
 %

### 2013-10-03 content.tex
Added example that CPS isn't as precise and tested in the redex model.

@@
 %%
-This paper gives a common, simple framework to derive both CFA2 and PDCFA using a recipe to apply pushdown analysis techniques to arbitrary semantics in a manner similar to AAM.
+This paper gives a common, simple framework to derive CFA2, PDCFA and its garbage-collecting extension using a recipe to apply pushdown analysis techniques to arbitrary semantics in a manner similar to AAM.
 %
@@
 %% TODO: complete
-\paragraph{Comparison to CFA2 CPS'd to remove {\tt shift} and {\tt reset}:}{
+\paragraph{Comparison to PDCFA CPS'd to remove {\tt shift} and {\tt reset}:}{
   One might ask what this different model buys us over using a CPS transform to compile away {\tt shift} and {\tt reset} forms.
@@
 Consider the following program and its CPS transform for comparison:
-\begin{lstlisting}
-???
+\begin{lstlisting}[mathescape]
+(let* ([id ($\lambda$ (x) x)]
+       [f ($\lambda$ (y) (shift k (k (k y))))]
+       [g ($\lambda$ (z) (reset (id (f z))))])
+  (<= (g 0) (g 1)))
 \end{lstlisting}
 
-\begin{lstlisting}
-???
+\begin{lstlisting}[mathescape]
+(let* ([id ($\lambda$ (x k) (k x))]
+       [f ($\lambda$ (y j) (j (j y)))]
+       [g ($\lambda$ (z h) (h (f z ($\lambda$ (fv) (id fv ($\lambda$ (i) i))))))])
+  (g 0 ($\lambda$ (g0v) (g 1 ($\lambda$ (g1v) (<= g0v g1v))))))
 \end{lstlisting}
 
-PDCFA with a monovariant allocation strategy will predict the CPS'd version returns ???.
-%
-Our specialized analysis for delimited control, also with a monovariant allocation strategy, will predict the non-CPS'd version returns ???.}
+PDCFA with a monovariant allocation strategy will predict the CPS'd version returns true or false.
+%
+Our specialized analysis for delimited control, also with a monovariant allocation strategy, will predict the non-CPS'd version returns true.}
 

### 2013-10-03 content.tex
Better formatting for shift/reset and CPS example.

@@
 Consider the following program and its CPS transform for comparison:
-\begin{lstlisting}[mathescape]
-(let* ([id ($\lambda$ (x) x)]
-       [f ($\lambda$ (y) (shift k (k (k y))))]
-       [g ($\lambda$ (z) (reset (id (f z))))])
-  (<= (g 0) (g 1)))
-\end{lstlisting}
-
-\begin{lstlisting}[mathescape]
-(let* ([id ($\lambda$ (x k) (k x))]
-       [f ($\lambda$ (y j) (j (j y)))]
-       [g ($\lambda$ (z h) (h (f z ($\lambda$ (fv) (id fv ($\lambda$ (i) i))))))])
-  (g 0 ($\lambda$ (g0v) (g 1 ($\lambda$ (g1v) (<= g0v g1v))))))
-\end{lstlisting}
+% \begin{lstlisting}[mathescape]
+% (let* ([id ($\lambda$ (x) x)]
+%        [f ($\lambda$ (y) (shift k (k (k y))))]
+%        [g ($\lambda$ (z) (reset (id (f z))))])
+%   (<= (g 0) (g 1)))
+% \end{lstlisting}
+\newcommand{\Stttextless}{{\fontencoding{T1}\selectfont<}}
+\definecolor{ValueColor}{rgb}{0.13,0.55,0.13}
+\newcommand{\RktVal}[1]{\inColor{ValueColor}{#1}}
+
+\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{let*}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{x}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{x}\RktPn{)}\RktPn{]}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{[}\RktSym{f}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{y}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{shift}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{k}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{k}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{k}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{y}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{]}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{[}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{z}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{reset}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{f}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{z}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{]}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{{\Stttextless}=}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
+
+% \begin{lstlisting}[mathescape]
+% (let* ([id ($\lambda$ (x k) (k x))]
+%        [f ($\lambda$ (y j) (j (j y)))]
+%        [g ($\lambda$ (z h) (h (f z ($\lambda$ (fv) (id fv ($\lambda$ (i) i))))))])
+%   (g 0 ($\lambda$ (g0v) (g 1 ($\lambda$ (g1v) (<= g0v g1v))))))
+% \end{lstlisting}
+\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{let*}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{x}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{k}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{k}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{x}\RktPn{)}\RktPn{)}\RktPn{]}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{[}\RktSym{f}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{y}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{j}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{j}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{j}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{y}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{]}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{[}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{z}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{h}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{h}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{f}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{z}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{fv}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{fv}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{i}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{i}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{]}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g0v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g1v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{{\Stttextless}=}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g0v}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g1v}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
 

### 2013-10-03 content.tex
More code formatting stuff.

@@
 %% TODO: Show that store widening and store/memo/KTab widening have a bisimulation.
+\setlength{\abovedisplayskip}{0pt}
+\setlength{\belowdisplayskip}{4pt}
+\setlength{\abovedisplayshortskip}{0pt}
+\setlength{\belowdisplayshortskip}{8pt}
 \begin{align*}
@@
   \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mstore,\mktab,\mmemo) \stepto \mstate'} \\
+                \wn(\tpl{(\mexpr,\menv),\mkont},\mstore,\mktab,\mmemo)
+                   &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
+\end{align*}
+\begin{minipage}{.55\linewidth}
+  \begin{align*}
     \mstore' &= \bigsqcup\setbuild{\mstore'}{\wn(\_, \mstore',\_,\_) \in I} \\
     \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \_, \mktab', \_) \in I} \\
-                \mmemo' &=  \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \_, \mmemo') \in I} \\
+    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \_, \mmemo') \in I}
+  \end{align*}
+\end{minipage}
+\begin{minipage}{.40\linewidth}
+  \begin{align*}
     S' &= \setbuild{(\mastate',\mstore')}{\wn(\mastate', \_, \_, \_) \in I} \\
     F' &= \setbuild{\mastate'}{(\mastate',\_) \in S' \setminus S} \\
-                \wn(\tpl{(\mexpr,\menv),\mkont},\mstore,\mktab,\mmemo)
-                   &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
-\end{align*}
+  \end{align*}
+\end{minipage}
 
@@
 \paragraph{Example:} Let's consider a monovariant allocation strategy (the address for each binding is the variable name itself) to run the following example:
-\begin{lstlisting}[mathescape]
-  (let* ([id ($\lambda$ (x) x)]
-         [y (id 0)]
-         [z (id 1)])
-    ($\le$ y z))
-\end{lstlisting}
+% \begin{lstlisting}[mathescape]
+%   (let* ([id ($\lambda$ (x) x)]
+%          [y (id 0)]
+%          [z (id 1)])
+%     ($\le$ y z))
+% \end{lstlisting}
+\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{let*}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{x}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{x}\RktPn{)}\RktPn{]}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{[}\RktSym{y}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\RktPn{)}\RktPn{]}
+
+\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{[}\RktSym{z}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\RktPn{)}\RktPn{]}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{$\leq$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{y}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{z}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
+
 Suppose we extend our semantics to allow numbers, numeric primitives and \texttt{let}.
@@
 %% Begin scary Scribble output
-\newcommand{\SCodePreSkip}{\vskip\abovedisplayskip}
-\newcommand{\SCodePostSkip}{\vskip\belowdisplayskip}
-\newenvironment{SCodeFlow}{\SCodePreSkip\begin{list}{}{\topsep=0pt\partopsep=0pt%
-\listparindent=0pt\itemindent=0pt\labelwidth=0pt\leftmargin=2ex\rightmargin=2ex%
-\itemsep=0pt\parsep=0pt}\item}{\end{list}\SCodePostSkip}
-\newenvironment{SingleColumn}{\begin{list}{}{\topsep=0pt\partopsep=0pt%
-\listparindent=0pt\itemindent=0pt\labelwidth=0pt\leftmargin=0pt\rightmargin=0pt%
-\itemsep=0pt\parsep=0pt}\item}{\end{list}}
-\newenvironment{RktBlk}{}{}
-\definecolor{IdentifierColor}{rgb}{0.15,0.15,0.50}
-\definecolor{ParenColor}{rgb}{0.52,0.24,0.14}
-\newcommand{\RktSym}[1]{\inColor{IdentifierColor}{#1}}
-\newcommand{\RktPn}[1]{\inColor{ParenColor}{#1}}
-
 \begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touch{-}kont}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mktab$}\RktPn{)}
@@
 %
-Unconstrained allocation functions
+\textbf{TODO: wrap up}
 
@@
 % \end{lstlisting}
-\newcommand{\Stttextless}{{\fontencoding{T1}\selectfont<}}
-\definecolor{ValueColor}{rgb}{0.13,0.55,0.13}
-\newcommand{\RktVal}[1]{\inColor{ValueColor}{#1}}
-
 \begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{let*}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{x}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{x}\RktPn{)}\RktPn{]}
@@
 
-\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{{\Stttextless}=}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
+\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{$\le$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
 
@@
 
-\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g0v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g1v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{{\Stttextless}=}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g0v}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g1v}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
+\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g0v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g1v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\le$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g0v}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g1v}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
 
@@
 %
-The model-checking community showed how to check temporal logic queries for collapsible pushdown automata (CPDA), or equivalently, higher-order recursion schemes, by deriving the checking algorithm from the Krivine machine~\citep{ianjohnson:Salvati:2011:KMH:2027223.027239}.
+The model-checking community showed how to check temporal logic queries for collapsible pushdown automata (CPDA), or equivalently, higher-order recursion schemes, by deriving the checking algorithm from the Krivine machine~\citep{ianjohnson:Salvati:2011:KMH:2027223.2027239}.
 %

### 2013-10-04 abstract.tex
More abstract changes.

--- 2013-09-30 abstract.tex
@@
-Pushdown analysis is better than finite-state analysis in precision and
-performance. Why then have we not seen total widespread adoption of these
-techniques? For one, the known techniques are technically burdened and
-difficult to understand or extend. Control structure of the programming language
-gets pulled into the model of computation, which makes extensions to
-non-pushdown control structures, such as \rackett{call/cc} or \rackett{shift} and
-\rackett{reset}, non-trivial.
-
-We show a derivational approach to abstract interpretation that yields
-transparently sound static analyses that can precisely match calls and
-returns when applied to well-known abstract machines. We show that
-adding memoization and segmenting the continuation into bounded pieces
-leads to machines that abstract to static analyses for context-free
-reachability by simply bounding the stores. This technique allows us
-to derive existing, more technically involved analyses, and a novel
-pushdown analysis for delimited, composable control.
+Pushdown analysis improves upon finite-state analysis in precision and performance by not exploring spurious paths caused by improperly matched calls and returns.
+%
+Prior literature on pushdown higher-order flow analysis depends on complicated machinery that is difficult to understand, extend, and prove correct.
+%
+The machinery has no concrete counterpart --- there is little to no operational intuition as to why it works.
+%
+We show that these pushdown analyses have a similar core that can be explained with well-known techniques such as memoization and Abstracting Abstract Machines (AAM).
+%
+We use this interpretation to derive a novel pushdown analysis for delimited, composable control.

### 2013-10-04 content.tex
Started a section on with-continuation-mark, but had to punt due to extension vs copy adding more complexity than I could fit.

@@
 %
-Until recently, flow analyses for higher-order languages could not handle return flow precisely \citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:earl2010pdcfa}, which leads to several spurious paths (and thus false positives) due to the pervasiveness of function/method calls and subsequent returns.
+Until recently, flow analyses for higher-order languages could not handle return flow precisely \citep{ianjohnson:vardoulakis-lmcs11, dvanhorn:Earl2010Pushdown}, which leads to several spurious paths (and thus false positives) due to the pervasiveness of function/method calls and subsequent returns.
 %
@@
 A contribution of this paper is a systematic process to construct \emph{pushdown} analyses of programming languages, due to the precision (and often performance) benefits.
+%
+
 %%
@@
 \end{enumerate}
+
+%%
+
 Once the machine semantics is in this form, simple point-wise abstraction leads to the summarization algorithms that we see in the literature.
-
-The remaining sections of the paper are
-\begin{itemize}
-\item{\autoref{sec:pdcfa}: we derive a cousin of PDCFA}
-\item{\autoref{sec:gc}: we show how abstract garbage collection ``falls out'' of this approach.}
-\item{\autoref{sec:cfa2}: we make additions to the PDCFA semantics to get a direct-style CFA2 without first-class control}
-\item{\autoref{sec:sr}: we give a novel analysis of delimited and composable first-class control.}
-\end{itemize}
+%
+In \autoref{sec:pdcfa} we derive a cousin of PDCFA.
+%
+The following section shows how to extend PDCFA with abstract garbage collection. % \autoref{sec:gc}: we show how abstract garbage collection ``falls out'' of this approach.
+%
+Then in \autoref{sec:cfa2} we show the fundamental difference between PDCFA and CFA2 without first-class control.
+% : we make additions to the PDCFA semantics to get a direct-style CFA2 without first-class control}
+%
+\ifwcm{We show in \autoref{sec:wcm} that semantics with stack-inspection (beyond GC) also fit within this interpretation.}
+%
+Finally, we show in \autoref{sec:sr} that even though first-class control operators don't fit in the pushdown model, we can gracefully loosen the abstraction to analyze delimited, composable first-class control.
 
@@
 %%
-We start by deriving PDCFA from an operational semantics for the untyped lambda calculus (figure \ref{fig:base-semantics}).
+We start by deriving PDCFA from an operational semantics for the untyped lambda calculus (\autoref{fig:base-semantics}).
 %
@@
 
+{
+\setlength{\abovedisplayskip}{0pt}
+\setlength{\belowdisplayskip}{4pt}
+\setlength{\abovedisplayshortskip}{0pt}
+\setlength{\belowdisplayshortskip}{8pt}
+\begin{align*}
+  \mexpr \in \Expr = \svar{\mvar} \alt \sapp{\mexpr}{\mexpr} \alt \slam{\mvar}{\mexpr}
+\end{align*}
 \begin{minipage}[b]{.55\linewidth}
   \begin{align*}
-  \mexpr \in \Expr &= \svar{\mvar} \mid \sapp{\mexpr}{\mexpr} \mid \slam{\mvar}{\mexpr} \\
   \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \\
   \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \\
-  \mkont \in \Kont &::= \kmt \mid \kcons{\mkframe}{\mkont} \\
-  \mkframe \in \Frame &::= \kar{\mexpr,\menv} \mid \kfn{\mval}
+  \mkont \in \Kont &::= \kmt \alt \kcons{\mkframe}{\mkont} \\
+  \mkframe \in \Frame &::= \kar{\mexpr,\menv} \alt \kfn{\mval}
   \end{align*}\hrule height 0pt\end{minipage}
@@
   \end{align*}\hrule height 0pt\end{minipage}
-
+}
+%%
+
+%%
 The $\Addr$ space is what controls the precision of the model.
@@
   \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \times \KTab \times \Memo \\
-  \mkont \in \Kont &::= \kmt \mid \krt{(\mexpr,\menv),\mstore} \mid \kcons{\mkframe}{\mkont} \\
+  \mkont \in \Kont &::= \kmt \alt \krt{(\mexpr,\menv),\mstore} \alt \kcons{\mkframe}{\mkont} \\
   \mktab \in \KTab &= (\Expr \times \Env)\times \Store \to \wp(\Kont) \\
@@
 %
+We show only the memo and continuation table widening, since store widening would require rewriting newly-created contexts to use the global store following the step.
+%
+This rewriting is unnecessary with the store-counting per-wide-step technique presented in \citet{ianjohnson:oaam:icfp2013}.
+%
 Since these components are shared, they are always the least upper bound of any respective component that the semantics produces (the intermediate set $I$) in order to stay sound.
@@
 %% TODO: Show that store widening and store/memo/KTab widening have a bisimulation.
+{
 \setlength{\abovedisplayskip}{0pt}
@@
 \begin{align*}
-  \mastate \in \sa{State} &= (\Expr \times \Env) \times \Kont \\
-  \System &= \wp(\sa{State} \times \Store) \times \wp(\sa{State}) \times \Store \times \KTab \times \Memo \\
+  \mastate \in \sa{State} &= (\Expr \times \Env) \times \Store \times \Kont \\
+  \System &= \wp(\sa{State}) \times \wp(\sa{State}) \times \KTab \times \Memo \\
   {\mathcal F} &: \sa{\State} \to \System \to \System \\
-  {\mathcal F}(\mstate_0)(S, F, \mstore, \mktab, \mmemo) &= (\set{(\mstate_0,\bot)} \cup S \cup S', F', \mstore', \mktab', \mmemo') \\
-  \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mstore,\mktab,\mmemo) \stepto \mstate'} \\
-                \wn(\tpl{(\mexpr,\menv),\mkont},\mstore,\mktab,\mmemo)
+  {\mathcal F}(\mstate_0)(S, F, \mktab, \mmemo) &= (S \cup S', F' \cup (\set{\mstate_0} \setminus S), \mktab', \mmemo') \\
+  \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
+                \wn(\tpl{(\mexpr,\menv),\mstore,\mkont},\mktab,\mmemo)
                    &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
@@
   \begin{align*}
-    \mstore' &= \bigsqcup\setbuild{\mstore'}{\wn(\_, \mstore',\_,\_) \in I} \\
     \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \_, \mktab', \_) \in I} \\
@@
   \begin{align*}
-    S' &= \setbuild{(\mastate',\mstore')}{\wn(\mastate', \_, \_, \_) \in I} \\
-    F' &= \setbuild{\mastate'}{(\mastate',\_) \in S' \setminus S} \\
+    S' &= \setbuild{\mastate'}{\wn(\mastate', \_, \_, \_) \in I} \\
+    F' &= S' \setminus S
   \end{align*}
 \end{minipage}
-
+}
+%%
+
+%%
 A $\System$ embodies the states seen and at which store, $S$, the frontier set of states yet to analyze, $F$, the shared store for the frontier, $\mstore$, the continuation table $\mktab$ and the memo table $\mmemo$.
@@
 %
-The followup to PDCFA~\citet{dvanhorn:Earl2012Introspective} that allows inspecting the stack in order to perform sound GC introduces an entirely new class of automata for this one purpose.
+The followup to PDCFA~\citep{dvanhorn:Earl2012Introspective} that allows inspecting the stack in order to perform sound GC introduces an entirely new class of automata for this one purpose.
 %
@@
 \begin{align*}
-  \konts(\mkont_0, \mktab)(K) &= \Phi(\mkont_0) \cup K \cup \bigcup\limits_{\krt{\mctx} \in K,\, \mkont' \in \mktab(\mctx)}{\Phi(\mkont')}
-\\
-  \touches(\mkont, \mktab) &= \bigcup\limits_{\mkframe \in \lfp(\konts(\mkont, \mktab))}{\touchesf(\mkframe)}
-\\[2pt]
-  \Phi(\kcons{\mkframe}{\mkont}) &= \set{\mkframe} \cup \Phi(\mkont) \\
-  \Phi(\krt{\mctx}) &= \set{\krt{\mctx}} \\
-  \Phi(\kmt) &= \emptyset 
+  \touchesk{\mkont}(\mktab) &= \setbuild{\maddr}{\mkont \rightsquigarrow_\mktab^* \maddr}
 \\[2pt]
   \touchesf(\kar{\mexpr,\menv}) &= \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)}) \\
-  \touchesf(\kfn{(\slam{\mvar}{\mexpr},\menv)}) &= \mathit{rng}(\menv|_{\mathit{fv}(\slam{\mvar}{\mexpr})})
-\end{align*}
+  \touchesf(\kfn{\mval}) &= \touches(\mval)
+\end{align*}
+\begin{mathpar}
+  \inferrule{ }{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont}
+  \quad
+  \inferrule{\maddr \in \touchesf(\mkframe)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
+  \quad
+  \inferrule{\mkont \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}
+  \quad
+  \inferrule{\mkont \in \mktab(\mctx)}{\krt{\mctx} \rightsquigarrow_\mktab \mkont}
+\end{mathpar}
 
@@
   \Gamma(\tpl{(\mexpr,\menv),\mstore,\mkont, \mktab, \mmemo}) &= \tpl{(\mexpr,\menv),\mstore|_{\reaches(\mathit{root}, \mstore)}, \mkont, \mktab, \mmemo} \\
+  \text{where } \mathit{root} &= \touchesk{\mkont}(\mktab) \cup \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)})
 \end{align*}
@@
 \begin{align*}
-  \text{where } \mathit{root} &= \touches(\mkont,\mktab) \cup \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)}) \\
   \reaches(\mathit{root}, \mstore) &= \setbuild{b}{\maddr \in \mathit{root}, \maddr \rightsquigarrow_\mstore^* b} \\
@@
     $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore', \mframe', \kcons{\kpush{\mframe}}{\mkont}}$
-    \\ where & $(\mstore',\mframe') = \bind(\mstore,\maddr,\mvar,\mval)$
+    \\ & where $(\mstore',\mframe') = \bind(\mstore,\maddr,\mvar,\mval)$
     \\
@@
   \end{tabular}
-  \caption{The CES$\xi$K machine}
+  \caption{The CES$\xi$K machine significant rules}
   \label{fig:frame-semantics}
@@
 
+\ifwcm{
+\section{Stack inspection with continuation marks}
+\label{sec:wcm}
+
+We saw in \autoref{sec:gc} that a semantics can inspect the continuation table to compute a property of the ``whole stack,'' which is outside the model of pushdown systems.
+%
+Here we show that Clements and Felleisen's WCM machine (a generalization of the CM machine) is directly analyzable in this setting, though we give up some precision in order to account for all calling contexts instead of the single one a function is currently in.
+%
+The machine can extend the marks of the continuation, reify the current set of marks, and query the set for all the values given to a particular mark key.
+%
+The mark set of the continuation is then part of the ``context'' for function calls, since functions can inspect the marks.
+%
+Mark sets can have unbounded length chains of values for each mark key, so we use $\alloc$ to tune the precision of context comparisons.
+%
+The effect is that we reify the current mark set for each function call to keep as part of the context (one can devise caching strategies for better performance).
+%
+We use a private store to represent the chains of values, in order to not affect the precision of reifications that happen within the analyzed program.
+%
+\begin{align*}
+  \mexpr \in \Expr &::= \swcm{\mexpr}{\mexpr}{\mexpr} \alt \sccms \alt \scmsl{\mexpr}{\mexpr} \alt \ldots \\
+  \mctx \in \Context &::= ((\mexpr, \menv), \mstore, \mmarkset) \\
+  \mkframe \in \Frame &::= \kwcmk{\mexpr, \mexpr, \menv} \alt
+                           \kwcmv{\mval, \mexpr, \menv} \alt
+                           \kcmsls{\mexpr,\menv} \alt
+                           \kcmslk{\mval} \alt \ldots \\ 
+  \mkont \in \Kont &::= \kmt^\mmarks \alt \krt{\mctx} \alt \kcons{\mkframe^\mwmarks}{\mkont} \\
+  \mmarks \in \Marks &= \Value \to \wp(\Value) \\
+  \mwmarks \in \QualifiedMarks &::= \scopy{\mmarks} \alt \sext{\mmarks} \\
+  \mmarkset \in \Markset &= \Marks \times \Store \\
+  \mval \in \Value &::= \texttt{nil} \alt \cons{\maddr}{\maddr} \alt \smarkset{\mmarks} \alt \ldots
+\end{align*}
+%% TODO: distinguish between copied and extended marks so that domarkset only builds
+%% a list of mark extensions
+\begin{figure}
+  \centering
+  $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
+  \begin{tabular}{r|l}
+    \hline
+% wcm eval key
+    $\tpl{(\swcm{\mexpr_k}{\mexpr_v}{\mexpr}, \menv), \mstore, \mkont, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr_k, \menv), \mstore, \kcons{\kwcmk{\mexpr_v,\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
+    \\
+% wcm eval value
+    $\tpl{\mval_k, \mstore, \kcons{\kwcmk{\mexpr_v,\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr_v, \menv), \mstore, \kcons{\kwcmv{\mval_k,\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
+    \\    
+% wcm all done.
+    $\tpl{\mval_v, \mstore, \kcons{\kwcmv{\mval_k, \mexpr, \menv}}{\mkont}, \mktab, \mmemo}$
+    &
+    $\tpl{\mexpr,\menv,\mstore,\mkont',\mktab', \mmemo}$ \\
+    where & $(\mkont',\mktab') = \domark(\mkont,\mktab,\mval_k,\mval_v)$
+    \\
+% function call
+    $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont}}$
+    &
+    $\tpl{(\mexpr, \menv'), \mstore', \krt{\mctx},\mktab',\mmemo}$
+    \\ where & $\menv' = \extm{\menv}{\mvar}{\maddr}$ \\
+    & $\mstore' = \joinone{\mstore}{\maddr}{\mval}$ \\
+    & $\mctx = ((\mexpr,\menv'),\mstore', \domarkset(\mstate))$ \\
+    & $\mktab' = \joinone{\mktab}{\mctx}{\mkont}$
+    \\
+% Current continuation marks
+    $\tpl{(\sccms,\menv),\mstore,\mkont,\mktab,\mmemo}$
+    &
+    $\tpl{\smarkset{\mmarks},\mstore\sqcup\mstore',\mkont,\mktab,\mmemo}$
+    \\ where & $(\mmarks,\mstore') = \domarkset(\mstate)$
+    \\
+% cms->list mark set
+    $\tpl{(\scmsl{\mexpr_\mmarks}{\mexpr_k}, \menv), \mstore, \mkont, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr_\mmarks, \menv), \mstore, \kcons{\kcmsls{\mexpr_k,\menv}}{\mkont}, \mktab, \mmemo}$
+    \\
+% cms->list key
+    $\tpl{\mval, \mstore, \kcons{\kcmsls{\mexpr_k,\menv}}{\mkont}, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr_k, \menv), \mstore, \kcons{\kcmslk{\mval}}{\mkont}, \mktab, \mmemo}$
+    \\
+% cms->list done
+    $\tpl{\mval_k, \mstore, \kcons{\kcmslk{\mmarks}}{\mkont}, \mktab, \mmemo}$
+    &
+    $\tpl{\mval, \mstore, \mktab, \mmemo}$ if $\mval \in \mmarks(\mval_k)$
+  \end{tabular}  
+  \caption{The Tabular WCM machine (list rules not shown)}
+  \label{fig:wcm}
+\end{figure}
+
+The two metafunctions, $\domark$ and $\domarkset$ are what do the heavy-lifting of adding marks to a continuation and reifying the mark set into a map of keys to lists.
+%
+Adding a mark to a continuation has an implicit equality check; if $\mval_k = \mval_k'$ concretely, then {\tt wcm} should do a strong update on $\mmarks$, but otherwise it should join since they could be two different values.
+%
+Concrete equality analysis is outside the scope of this paper, but doable (\eg, with abstract counting~\citep{dvanhorn:Might:2006:GammaCFA}), so we elide the details and call this strong-update-or-join operation $\sqcup_{!{}}$.
+%
+Marking a return continuation means we need to crawl through the table, which again can be circular, so we compute the least fixed point of ${\mathcal M}$ to get an extended table holding the marked continuations.
+%
+\newcommand{\rewritectx}{\mathit{rw}}
+\newcommand{\domarkstep}{\mathit{mark}_0}
+\begin{align*}
+  \domark(\mkont, \mktab, \mval_k, \mval_v) &= (\mkont', \mktab) \text{ if } (\mkont',\bot) = \domarkstep(\mkont) \\
+  \domark(\mkont, \mktab, \mval_k, \mval_v) &= (\mkont', \mathbf{snd}(\lfp({\mathcal M}(\mctx)))) \text{ if } (\mkont',\mctx) = \domarkstep(\mkont) \\
+  \text{where } {\mathcal M}(\mctx_0)(C, \mktab) &= (C \cup C' \cup \set{\mctx_0}, \mktab \sqcup \mktab') \\
+    I &= \setbuild{(\rewritectx(\mctx), \setbuild{\domarkstep(\mkont')}{\mkont' \in \mktab(\mctx)})}{\mctx \in C} \\
+    C' &= \setbuild{\mctx}{(\_, S) \in I, (\_, \mctx) \in S} \\
+    \mktab' &= \bigsqcup\setbuild{[\mctx \mapsto \setbuild{\mkont}{(\mctx, S) \in I, (\mkont, \_) \in S}]}{(\mctx,\_) \in I}
+\\[2pt]
+  \rewritectx((\mexpr,\menv),\mstore,(\mmarks, \mstore')) &= ((\mexpr,\menv),\mstore,(\mmarks\sqcup_{!{}}[\mval_k \mapsto \mval_v], \mstore'))
+\\[2pt]
+  \domarkstep(\krt{\mctx}) &= (\krt{\rewritectx(\mctx)}, \mctx) \\
+  \domarkstep(\kcons{\mkframe^\mmarks}{\mkont}) &= (\kcons{\mkframe^{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}}{\mkont}, \bot) \\
+  \domarkstep(\kmt^\mmarks) &= (\kmt^{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}, \bot)
+\end{align*}
+
+Second, the $\domarkset$ metafunction builds a map of keys to lists, and builds a private store for the list constructions.
+%
+For this, we extend $\alloc$'s specification to allow a second argument, the private store, and produce a pair of addresses.
+%
+%% FIXME: need to crawl through qualified continuations for all extensions to build a list for each found key.
+%% Has another implicit equality check
+\begin{align*}
+  \domarkset(\mstate\equiv\tpl{\mpoint,\mstore,\mkont,\mktab,\mmemo}) &=
+    (\maddr_a, \maddr_d) = \alloc(\mstate,\mstore')
+\end{align*}
+}
+
 \section{Analysis of delimited, composable control}
@@
   \mstate \in \State &::= \tpl{\mpoint, \mstore, \mkont, \mmkont} \\
-  \mpoint \in \Point &::= (\mexpr, \menv) \mid \mval \\
-  \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \mid \vcomp{\mkont} \\
-  \mmkont \in \MKont &::= \kmt \mid \mkapp{\mkont}{\mmkont}
+  \mpoint \in \Point &::= (\mexpr, \menv) \alt \mval \\
+  \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt \vcomp{\mkont} \\
+  \mmkont \in \MKont &::= \kmt \alt \mkapp{\mkont}{\mmkont}
 \end{align*}
@@
 \begin{align*}
-  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \mid (\vcomp{\mkont},\mval,\mstore) \\
-  \mkont \in \Kont &= \kmt \mid \krt{\mctx} \mid \kcons{\mkframe}{\mkont} \\
-  \mmkont \in \MKont &::= \kmt \mid \kprompt{\mctx} \\
-  \mathit{KTabRange} &= \wp(\Kont \times \MKont) \cup \wp(\Kont) \\
-  \mktab \in \KTab &= \Context \to \mathit{KTabRange} \\
+  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \alt (\vcomp{\mkont},\mval,\mstore) \\
+  \mkont \in \Kont &= \kmt \alt \krt{\mctx} \alt \kcons{\mkframe}{\mkont} \\
+  \mmkont \in \MKont &::= \kmt \alt \kprompt{\mctx} \\
+  \mathit{KRange} &= \wp(\Kont \times \MKont) \cup \wp(\Kont) \\
+  \mktab \in \KTab &= \Context \to \mathit{KRange} \\
   \mmemo \in \Memo &= \Context \to \wp(\Value \times \Store)
@@
 
+{
+\setlength{\abovedisplayskip}{0pt}
+\setlength{\belowdisplayskip}{4pt}
+\setlength{\abovedisplayshortskip}{0pt}
+\setlength{\belowdisplayshortskip}{8pt}
+\begin{minipage}{0.45\linewidth}
   \begin{align*}
     \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
-  \mctx \in \Context &::= \msctx \mid \ldots \\
-  \mskont \in \SKont &= \krt{\msctx} \mid \ldots \\
-  \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \mid \vcomp{\mskont} \\
-  \mktab \in \KTab &= (\Context \to \mathit{KTabRange}) \\
+    \mctx \in \Context &::= \msctx \alt \ldots \\
+    \mskont \in \SKont &= \krt{\msctx} \alt \ldots
+  \end{align*}
+\end{minipage}
+\begin{minipage}{0.50\linewidth}
+  \begin{align*}
+    \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt \vcomp{\mskont} \\
+    \mktab \in \KTab &= (\Context \to \mathit{KRange}) \\
     &\cup (\Addr \to \wp(\Store))
   \end{align*}
+\end{minipage}
+
+}
 
@@
 
-The immediately related work is that of PDCFA \citep{ianjohnson:earl2010pdcfa}, CFA2~\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:Vardoulakis2011Pushdown}, and AAM ~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we recreated in full detail.
+The immediately related work is that of PDCFA \citep{dvanhorn:Earl2010Pushdown}, CFA2~\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:Vardoulakis2011Pushdown}, and AAM ~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we recreated in full detail.
 %

### 2013-10-04 content.tex
Filled out wcm semantics more. I think markset is unsound in the circular case. Unsure how to fix.

@@
                            \kcmsls{\mexpr,\menv} \alt
-                           \kcmslk{\mval} \alt \ldots \\ 
-  \mkont \in \Kont &::= \kmt^\mmarks \alt \krt{\mctx} \alt \kcons{\mkframe^\mwmarks}{\mkont} \\
+                           \kcmslk{\mval} \alt \kwcm{\mmarks} \alt \ldots \\
+  \mkont \in \Kont &::= \kmt^\bot \alt \krt{\mctx}^\mmarks \alt \kcons{\mkframe^\mmarks}{\mkont} \\
   \mmarks \in \Marks &= \Value \to \wp(\Value) \\
-  \mwmarks \in \QualifiedMarks &::= \scopy{\mmarks} \alt \sext{\mmarks} \\
   \mmarkset \in \Markset &= \Marks \times \Store \\
@@
 \end{align*}
+
 %% TODO: distinguish between copied and extended marks so that domarkset only builds
@@
 % wcm eval key
-    $\tpl{(\swcm{\mexpr_k}{\mexpr_v}{\mexpr}, \menv), \mstore, \mkont, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr_k, \menv), \mstore, \kcons{\kwcmk{\mexpr_v,\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
+    $\tpl{(\swcm{\mexpr_k}{\mexpr_v}{\mexpr}, \menv), \mstore, \mkont^\mmarks, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr_k, \menv), \mstore, \kcons{\kwcmk{\mexpr_v,\mexpr,\menv}^\mmarks}{\mkont}, \mktab, \mmemo}$
     \\
 % wcm eval value
-    $\tpl{\mval_k, \mstore, \kcons{\kwcmk{\mexpr_v,\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr_v, \menv), \mstore, \kcons{\kwcmv{\mval_k,\mexpr,\menv}}{\mkont}, \mktab, \mmemo}$
+    $\tpl{\mval_k, \mstore, \kcons{\kwcmk{\mexpr_v,\mexpr,\menv}}{\mkont^\mmarks}, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr_v, \menv), \mstore, \kcons{\kwcmv{\mval_k,\mexpr,\menv}^\mmarks}{\mkont}, \mktab, \mmemo}$
     \\    
@@
     \\
+% Pop marks
+   $\tpl{\mval,\mstore,\kcons{\kwcm{\mmarks}^{\mmarks'}}{\mkont}, \mktab, \mmemo}$
+   &
+   $\tpl{\mval,\mstore,\mkont,\mktab,\mmemo}$
+   \\
 % function call
-    $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont}}$
-    &
-    $\tpl{(\mexpr, \menv'), \mstore', \krt{\mctx},\mktab',\mmemo}$
+    $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}^\mmarks}{\mkont}}$
+    &
+    $\tpl{(\mexpr, \menv'), \mstore', \krt{\mctx}^\mmarks,\mktab',\mmemo}$
     \\ where & $\menv' = \extm{\menv}{\mvar}{\maddr}$ \\
@@
 % cms->list mark set
-    $\tpl{(\scmsl{\mexpr_\mmarks}{\mexpr_k}, \menv), \mstore, \mkont, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr_\mmarks, \menv), \mstore, \kcons{\kcmsls{\mexpr_k,\menv}}{\mkont}, \mktab, \mmemo}$
+    $\tpl{(\scmsl{\mexpr_\mmarks}{\mexpr_k}, \menv), \mstore, \mkont^\mmarks, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr_\mmarks, \menv), \mstore, \kcons{\kcmsls{\mexpr_k,\menv}^\mmarks}{\mkont}, \mktab, \mmemo}$
     \\
 % cms->list key
-    $\tpl{\mval, \mstore, \kcons{\kcmsls{\mexpr_k,\menv}}{\mkont}, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr_k, \menv), \mstore, \kcons{\kcmslk{\mval}}{\mkont}, \mktab, \mmemo}$
+    $\tpl{\mval, \mstore, \kcons{\kcmsls{\mexpr_k,\menv}}{\mkont^\mmarks}, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr_k, \menv), \mstore, \kcons{\kcmslk{\mval}^\mmarks}{\mkont}, \mktab, \mmemo}$
     \\
@@
 \\[2pt]
-  \domarkstep(\krt{\mctx}) &= (\krt{\rewritectx(\mctx)}, \mctx) \\
-  \domarkstep(\kcons{\mkframe^\mmarks}{\mkont}) &= (\kcons{\mkframe^{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}}{\mkont}, \bot) \\
-  \domarkstep(\kmt^\mmarks) &= (\kmt^{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}, \bot)
+  \domarkstep(\krt{\mctx}^\mmarks) &= (\krt{\rewritectx(\mctx)}^{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}, \mctx) \\
+  \domarkstep(\kcons{\kwcm{\mmarks}^{\mmarks'}}{\mkont}) &= (\kcons{\kwcm{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}^{\mmarks' \sqcup_{!{}} [\mval_k \mapsto \mval_v]}}{\mkont}, \bot) \\
+  \domarkstep(\mkont^\mmarks) &= (\kcons{\kwcm{[\mval_k \mapsto \mval_v]}^{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}}{\mkont}, \bot)
 \end{align*}
@@
 %
-For this, we extend $\alloc$'s specification to allow a second argument, the private store, and produce a pair of addresses.
+For this, we extend $\alloc$'s specification to allow more arguments --- the private store, and a set of keys that need cons cells allocated for them --- and produce a map from keys to pairs of addresses.
 %
@@
 %% Has another implicit equality check
-\begin{align*}
-  \domarkset(\mstate\equiv\tpl{\mpoint,\mstore,\mkont,\mktab,\mmemo}) &=
-    (\maddr_a, \maddr_d) = \alloc(\mstate,\mstore')
+\newcommand{\markaux}{\mathit{marks}}
+\begin{align*}
+  \domarkset(\mstate\equiv\tpl{\mpoint,\mstore,\mkont,\mktab,\mmemo}) &= \markaux(\mkont,\emptyset) \\
+  \markaux(\kmt^\bot, G) &= (\bot, \bot) \\
+  \markaux(\krt{\mctx}^\mmarks, G) &= (\bot, \bot) \text{ if } \mctx \in G \\ %% FIXME: unsound
+  \markaux(\krt{\mctx}^\mmarks, G) &= \bigsqcup\limits_{\mkont\in \mktab(\mctx)}{\markaux(\mkont,G \cup \set{\mctx})} \text{ otherwise} \\
+  \markaux(\kcons{\kwcm{\mmarks}^{\mmarks'}}{\mkont}, G) &= (\mmarks''[\mval_k \mapsto \set{\cons{a}{d}} : \mval_k \in \dom(\mmarks), (a, d) = A(\mval_k)], \\
+    &\phantom{= (}
+      \mstore\sqcup[a \mapsto \mmarks(\mval_k) : \mval_k \in \dom(\mmarks), (a,\_) = A(\mval_k)] \\
+    &\phantom{= (\mstore}\sqcup[d \mapsto \mmarks''(\mval_k) : \mval_k \in \dom(\mmarks), (\_,d) = A(\mval_k)]
+   \\ \text{where } (\mmarks'', \mstore) &= \markaux(\mkont, G)
+   \\ A &= \alloc(\mstate, \mstore, \dom(\mmarks))
 \end{align*}

### 2013-10-07 content.tex
Better illustration for the example.

@@
 %
-The whole of their machinery is then computing the pushdown system on-the-fly, only considering states reachable from the root (initial) state.
+The whole of their machinery is then computing the pushdown system on-the-fly, only considering states reachable from the initial state.
 %
@@
 %
-To match nodes that push frames with later nodes following the pop of that frame, they additionally compute an $\epsilon$-closure graph, which keeps edges between nodes that have a net-zero stack change path between them.
+They additionally compute an $\epsilon$-closure graph to match nodes that push frames with later nodes following the pop of that frame, which keeps edges between nodes that have a net-zero stack change path between them.
 %
@@
 %
-If $\alloc$ only uses addresses from a finite subset of $\Addr$, then the state space without continuations is finite, and the space of contination frames is finite (thus the pushdown system interpretation is apt).
+If $\alloc$ only uses addresses from a finite subset of $\Addr$, then the state space without continuations is finite, and the space of contination frames is finite --- thus the pushdown system interpretation is apt.
 %
@@
   \end{tabular}
-  \caption{The CESK machine}
+  \caption{The CESK machine with allocation tuning}
   \label{fig:base-semantics}
@@
 Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store (the second case of rule 4).
+%%
+
+%%
+Memoization with respect to the entire store is safe, if not a little wasteful; there are portions of the store that can change and be irrelevant to the result of a function call.
+%
+There are techniques to ignore differences in the heap when they are irrelevant, dubbed ``sparseness analysis'' that have been successful for large-scale first-order analysis~\citep{ianjohnson:DBLP:conf/pldi/OhHLLY12}, but they rely on a pre-analysis that knows the control-flow graph.
+%
+We have promising preliminary results in this direction that don't require a pre-analysis, but we leave further investigation of sparse techniques to future work.
 
@@
 %
-This rewriting is unnecessary with the store-counting per-wide-step technique presented in \citet{ianjohnson:oaam:icfp2013}.
+This rewriting is unnecessary with the store-counting per-wide-step technique presented in \citet{ianjohnson:oaam:icfp2013}, or the less precise timestamp approximation from \citet{dvanhorn:Shivers:1991:CFA}.
 %
@@
 %%
-A $\System$ embodies the states seen and at which store, $S$, the frontier set of states yet to analyze, $F$, the shared store for the frontier, $\mstore$, the continuation table $\mktab$ and the memo table $\mmemo$.
-%
-The memo table need only map to values, since the global store will always subsume the store that would otherwise be mapped.
-%
-All the states in the frontier are stepped with the current store, after which the next store is the least upper bound of all the resulting stores.
-%
-The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.
+A $\System$ embodies the states seen and at which store, $S$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
+% Note: Only if widened store.{
+%
+%The memo table need only map to values, since the global store will always subsume the store that would otherwise be mapped.
+%
+%All the states in the frontier are stepped with the current store, after which the next store is the least upper bound of all the resulting stores.
+%
+%The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.}
 %
@@
 
-\paragraph{Example:} Let's consider a monovariant allocation strategy (the address for each binding is the variable name itself) to run the following example:
-% \begin{lstlisting}[mathescape]
-%   (let* ([id ($\lambda$ (x) x)]
-%          [y (id 0)]
-%          [z (id 1)])
-%     ($\le$ y z))
-% \end{lstlisting}
-\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{let*}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{x}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{x}\RktPn{)}\RktPn{]}
-
-\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{[}\RktSym{y}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\RktPn{)}\RktPn{]}
-
-\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{[}\RktSym{z}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\RktPn{)}\RktPn{]}\RktPn{)}
-
-\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{$\leq$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{y}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{z}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
+\paragraph{Example:} Let's consider a monovariant allocation strategy (the address for each binding is the variable name itself, so $\menv$ is always the identity function $\iota$) to run the following example:
+
+\begin{center}
+  \includegraphics[scale=0.5]{example0}
+\end{center}
+  % \item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
+%     \mstore_0[\texttt{x} \mapsto \set{0}]$ and
+%     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
+%          $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
+%     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
+% \item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
+% \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
+% \item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
+%       $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
+%       (call this $\mkont_2$).}
+% \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
+%       (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
+% \item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
+% \item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
 
@@
 %
-0CFA ~\citep{ianjohnson:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
+0CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
 %
@@
 %
-Let's take a look at the evaluation after binding \texttt{id} ($\mstore_0 = [\texttt{id} \mapsto \set{\vclo{\slam{\texttt{x}}{\texttt{x}}}{\bot}}]$):
-
-%% FIXME stores in memo-table
-\begin{enumerate}
-\item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
-    \mstore_0[\texttt{x} \mapsto \set{0}]$ and
-    (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
-         $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
-    $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
-\item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
-\item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
-\item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
-      $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
-      (call this $\mkont_2$).}
-\item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
-      (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
-\item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
-\item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
-\end{enumerate}
+Let's take a look at the evaluation after binding \texttt{id}:
+
+\begin{center}
+  $\begin{array}{l}\includegraphics[scale=0.4195]{example1}\end{array}$
+  $\Rightarrow$
+  $\begin{array}{l}\includegraphics[scale=0.4195]{example2}\end{array}$
+  $\Rightarrow$
+  $\begin{array}{l}\includegraphics[scale=0.4195]{example3}\end{array}$
+\end{center}
+
+{\small
+\begin{tabular}{|l|c|c|c|c|c|}
+\hline
+\# & $\mstore$ & $\mctx$ & $\mkont$ & $\mktab$ & $\mmemo$ \\
+\hline
+0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\bot)}]$ & $(({\tt x}, \iota), \mstore_1)$ & $\mkont_0$ & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
+1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $(({\tt x}, \iota), \mstore_3)$ & $\krt{\mctx_0}$ & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
+2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & - & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)})$ & - & - \\
+3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & - & $\krt{\mctx_1}$ & - & - \\
+4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & - & $\kmt$ & - & - \\
+\hline
+\end{tabular}}
+Where $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$
+
+% %% FIXME stores in memo-table
+% \begin{enumerate}
+% \item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
+%     \mstore_0[\texttt{x} \mapsto \set{0}]$ and
+%     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
+%          $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
+%     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
+% \item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
+% \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
+% \item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
+%       $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
+%       (call this $\mkont_2$).}
+% \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
+%       (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
+% \item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
+% \item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
+% \end{enumerate}
 
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:pdcfa-tabular}
   Tabular semantics has a bisimulation with the standard semantics with comparable allocation strategies.
@@
   \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt, G}} \quad
-  \inferrule{\mkont \in \unroll{\mktab}{\mkont', G}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont', G}} \\
+  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \emptyset}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont', G}} \\
   \inferrule{\mctx \notin G  \\
@@
 %
-\textbf{TODO: wrap up}
+%TODO: wrap up
 
@@
 
-Figure \ref{fig:shift-reset-table0} has what one might naturally write using just this information.
+\autoref{fig:shift-reset-table0} has what one might naturally write using just this information.
 %
@@
 
-\begin{conjecture}\label{thm:concrete-sr}
+\begin{theorem}\label{thm:concrete-sr}
   Tabular semantics has a bisimulation with standard semantics given fresh allocation.
-\end{conjecture}
-
-We prove this with the invariant that the continuation and meta-continuation have unique, well-founded unrollings in addition to the table invariants from \autoref{sec:pdcfa}.
-
-\begin{conjecture}\label{thm:sound-sr}
+\end{theorem}
+
+We prove this with the invariant that the continuation and meta-continuation have unique, well-founded unrollings in addition to the table invariants from \autoref{sec:pdcfa}, since entries in $\mktab$ are always fresh.
+
+\begin{theorem}\label{thm:sound-sr}
   Tabular semantics simulates standard semantics with abstract allocation.
-\end{conjecture}
+\end{theorem}
 
@@
 The control structures there are difficult to model even with AAM's techniques due to the ability to capture and compose continuations with arbitrarily many prompts, but once we can tackle AAM, there should be a straightforward route to extending it with the pushdown techniques of this paper.
+%
+We have an unproven prototype of the stack inspection capabilities provided by the {\tt with-continuation-mark}, {\tt current-continuation-marks} and {\tt continuation-mark-set->list} primitives, where a ``context'' then also includes what is the (approximated) result of {\tt current-continuation-marks}, since functions can inspect only this amount of the stack.
 %%

### 2013-10-07 content.tex
Example is more readable now.

@@
 \begin{center}
-  $\begin{array}{l}\includegraphics[scale=0.4195]{example1}\end{array}$
-  $\Rightarrow$
-  $\begin{array}{l}\includegraphics[scale=0.4195]{example2}\end{array}$
-  $\Rightarrow$
-  $\begin{array}{l}\includegraphics[scale=0.4195]{example3}\end{array}$
+  \begin{minipage}{0.55\linewidth}
+    \begin{center}
+      $\begin{array}{l}\includegraphics[scale=0.65]{example1}\end{array}$ \\
+      $\Downarrow$ \\
+      $\begin{array}{l}\includegraphics[scale=0.65]{example2}\end{array}$ \\
+      $\Downarrow$ \\
+      $\begin{array}{l}\includegraphics[scale=0.65]{example3}\end{array}$
     \end{center}
-
+  \end{minipage}
+  \begin{minipage}{0.43\linewidth}
+    \begin{center} {\small
+        \begin{tabular}{|l|c|c|}
+          \hline
+          \# & $\mstore$ & $\mkont$ \\
+          \hline
+          0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\bot)}]$ & $\mkont_0$ \\
+          1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
+          2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)})$ \\
+          3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
+          4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
+          \hline
+        \end{tabular}} \\[10pt]
       {\small
-\begin{tabular}{|l|c|c|c|c|c|}
-\hline
-\# & $\mstore$ & $\mctx$ & $\mkont$ & $\mktab$ & $\mmemo$ \\
-\hline
-0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\bot)}]$ & $(({\tt x}, \iota), \mstore_1)$ & $\mkont_0$ & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
-1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $(({\tt x}, \iota), \mstore_3)$ & $\krt{\mctx_0}$ & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
-2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & - & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)})$ & - & - \\
-3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & - & $\krt{\mctx_1}$ & - & - \\
-4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & - & $\kmt$ & - & - \\
+        \begin{tabular}{|l|c|}
+          \hline
+          \# & $\mctx$ \\
+          \hline
+          0  & $(({\tt x}, \iota), \mstore_1)$ \\
+          1  & $(({\tt x}, \iota), \mstore_3)$ \\
           \hline
         \end{tabular}}
+      \\[10pt]
+      {\small
+        \begin{tabular}{|l|c|c|c|}
+          \hline
+          \# & $\mktab$ & $\mmemo$ \\
+          \hline
+          0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
+          1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
+          \hline
+        \end{tabular}}    
+    \end{center}
+  \end{minipage}
+\end{center}
 Where $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$
-
 % %% FIXME stores in memo-table
@@
 
-The astute reader may find this definition is dubious, since there is a possibility for unguarded corecursion here.
+The astute reader may find this definition dubious, since there is a possibility for unguarded corecursion here.
 %
@@
 \end{mathpar}
-Consider later uses of $\mathit{unroll}$ to implicitly give $\emptyset$ for $G$.
+
+%%
+The different tables we have encode information about execution history that we prove is invariant.
+\begin{align*}
+  \inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo}) &=
+          \dom(\mmemo) \subseteq \dom(\mktab) \\
+  &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab),
+                  \mkont^\circ \in \mktab(\mpoint',\mstore'),
+                  \mkont' \in \unroll{\mktab}{\mkont^\circ, \emptyset}. \\   
+  &\qquad \phi_{\reachable} \wedge \phi_{\memo} \\
+  \text{where } \phi_{\reachable} &=
+  \exists \mkont^* \in \unroll{\mktab}{\mkont,\emptyset}.
+    \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont^*} \\
+  \phi_{\memo} &=
+ \forall (\mval, \mstore'') \in \mmemo(\mpoint',\mstore'). \\
+  &\qquad \exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore'', \mkont'}.
+     \hastail(\mtrace,\mkont') \\
+\end{align*}
+
+The first is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
+%
+The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
+%
+The final invariant, $\phi_{\memo}$, states that the memo table is correct: all calling contexts reach the memoized result, with the calling context's continuation in the tail (to justify using stack-irrelevance).
+
+\begin{lemma}[Table invariants]\label{lem:tab-inv}
+  If $\inv(\mstate)$ and $\mstate \stepto \mstate'$ then $\inv(\mstate')$.
+\end{lemma}
+
+In order to prove this, we need a lemma that allows us to plug in any continuation for memoized results.
 
@@
 
-The different tables we have encode information about execution history that we prove is invariant.
-\begin{align*}
-  \inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo}) &=
-          \dom(\mmemo) \subseteq \dom(\mktab) \\
-  &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab),
-                  \mkont^\circ \in \mktab(\mpoint',\mstore'),
-                  \mkont' \in \unroll{\mktab}{\mkont^\circ}. \\   
-  &\qquad \phi_{\reachable} \wedge \phi_{\memo} \\
-  \text{where } \phi_{\reachable} &=
-  \exists \mkont^* \in \unroll{\mktab}{\mkont}.
-    \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont^*} \\
-  \phi_{\memo} &=
- \forall (\mval, \mstore'') \in \mmemo(\mpoint',\mstore'). \\
-  &\qquad \exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore'', \mkont'}.
-     \hastail(\mtrace,\mkont') \\
-\end{align*}
-
-The first is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
-%
-The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
-%
-The final invariant, $\phi_{\memo}$, states that the memo table is correct: all calling contexts reach the memoized result, with the calling context's continuation in the tail (to justify using stack-irrelevance).
-
-\begin{lemma}[Table invariants]\label{lem:tab-inv}
-  If $\inv(\mstate)$ and $\mstate \stepto \mstate'$ then $\inv(\mstate')$.
-\end{lemma}
-
+Finally, we can show that even if the tables are shared amongst all states, they are precise enough to not interfere with the results.
 \begin{theorem}[Equivalence with global tables]\label{thm:global}
@@
 %
+Technically since this semantics inspects the touched addresses of the continuation, the set of touched addresses should be considered part of the ``context'' in the $\mathbf{rt}$ continuations.
+%
+This extra context can be overlooked, since that means the computed touched addresses will be an overapproximation of what is already sound.
+%
+More touched addresses means a less effective GC, but this is an acceptable tradeoff for a possibly smaller state space.
+%
 \begin{theorem}\label{thm:gc-pointers}
@@
 %
-If they aren't, it looks as if \texttt{x} was rebound; it hasn't, and thus it is a ``fake rebinding.''
+If they are not all considered $v$, it looks as if \texttt{x} was rebound; it hasn't, and thus it is a ``fake rebinding.''
 %
@@
 %
-It is our view that systematic approaches to applying the techniques are pivotal to scaling them to ``real languages.''
+It is our view that systematic approaches to applying the techniques are pivotal to scaling them to production programming languages.
 %
@@
 %
-We have an unproven prototype of the stack inspection capabilities provided by the {\tt with-continuation-mark}, {\tt current-continuation-marks} and {\tt continuation-mark-set->list} primitives, where a ``context'' then also includes what is the (approximated) result of {\tt current-continuation-marks}, since functions can inspect only this amount of the stack.
+We have an unproven prototype of the stack inspection capabilities provided by the {\tt with-continuation-mark}, {\tt current-continuation-marks} and {\tt continuation-mark-set->list} primitives, where a ``context'' then also includes what is the result of {\tt current-continuation-marks}, since functions can inspect only this amount of the stack.
 %%

### 2013-10-07 content.tex
Slashes -> and/or. Better discussion of call/cc.

@@
 %
-Until recently, flow analyses for higher-order languages could not handle return flow precisely \citep{ianjohnson:vardoulakis-lmcs11, dvanhorn:Earl2010Pushdown}, which leads to several spurious paths (and thus false positives) due to the pervasiveness of function/method calls and subsequent returns.
+Until recently, flow analyses for higher-order languages could not handle return flow precisely \citep{ianjohnson:vardoulakis-lmcs11, dvanhorn:Earl2010Pushdown}, which leads to several spurious paths (and thus false positives) due to the pervasiveness of function or method calls and subsequent returns.
 %
@@
 %
-We also give intuitive analogies to well-established ideas/techniques so that the working semanticist can write a pushdown analysis for her language.
+We also give intuitive analogies to well-established ideas and techniques so that the working semanticist can write a pushdown analysis for her language.
 %
@@
 To turn this semantics into PDCFA's algorithm for constructing a Dyck state graph, given the appropriate $\alloc$, we apply a widening operator to make $\mstore$, $\mkont$, and $\mmemo$ shared amongst all states (meaning states then become represented without these components), in what is then called a $\System$.
-%
+We use a meta-function $\wn$, ``wide to narrow,'' to shift between the two representations of states.
+%%
+
+%%
 We show only the memo and continuation table widening, since store widening would require rewriting newly-created contexts to use the global store following the step.
@@
 Since these components are shared, they are always the least upper bound of any respective component that the semantics produces (the intermediate set $I$) in order to stay sound.
-%
-We use a meta-function $\wn$, ``wide to narrow'', to shift between the two representations of states.
 
@@
 \begin{theorem}[Equivalence with global tables]\label{thm:global}
-  Tabular semantics has a bisimulation with memo/k-table-widened semantics.
+  Tabular semantics has a bisimulation with memo-and-k-table-widened semantics.
 \end{theorem}
@@
 The result of this viewpoint is a sound, precise, and computable semantics for a ``pushdown approach'' to analyzing delimited, composable control.
-%
-To handle non-composable control, we can simply remove the prompts and keep $\approximate$.
 
@@
 
-We prove this with the invariant that the continuation and meta-continuation have unique, well-founded unrollings in addition to the table invariants from \autoref{sec:pdcfa}, since entries in $\mktab$ are always fresh.
+We prove this with the invariant that the continuation and meta-continuation have unique unrollings in addition to the table invariants from \autoref{sec:pdcfa}, since entries in $\mktab$ are always fresh.
 
@@
 %
-Since \rackett{call/cc} can be simulated with shift/reset\footnote{\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{call/cc}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{f}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{shift}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{k}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{k}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{f}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{x}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{shift}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{c}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{k}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{x}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}}, this work supercedes theirs (the space complexity difference imposed by the simulation is non-issue in the abstract).
+Our semantics adapts to \rackett{call/cc} by removing the meta-continuation operations, and thus this work supercedes theirs.
 %
@@
 %
-The most important constraint is that we can't know all call-sites of a function/method before the analysis begins, which their methods heavily rely on.
+The most important constraint is that we can't know all call-sites of a function or method before the analysis begins, which their methods heavily rely on.
 %%

### 2013-10-08 content.tex
Changed page size to the actual page size used. Updated related work.

@@
 
-\section{What to expect}
+\section{The essence of summarization}
 
@@
 %
-Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, say, AAM would provide.
+Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, say, 0-CFA would provide.
 %%
@@
 
-There are common underpinnings of PDCFA and CFA2 that can be embodied as concrete machinery in the programming language semantics: 
+There are common underpinnings of PDCFA and CFA2 that can be embodied as concrete ``summarization'' machinery in the programming language semantics: 
 \begin{enumerate}
-\item{breaking the recursive structure of continuations by indirecting them through a table with appropriately precise keys, and}
-\item{memoization.}
+\item{bound continuations with with delimiters that specify the exact context in which any continuation replacement would lead to the same result, and}
+\item{memoize at these delimiters}
 \end{enumerate}
@@
 %%
-
+The first of these two might seem unfamiliar, but it is just identifying what is needed for sound memoization, and we delimit the continuations in order to mark where to memoize results.
+%
+In the case of a CESK machine, where the store is relevant to the evaluation of functions, a context is just the CES portion since stacks don't matter.
+%
+For stack inspecting analyses, a context is CES plus the inspected portion of K.
+%
+As soon as we have a value that such a context evaluates to, all subsequent visits to that context can jump straight to the memoized result(s).
+%%
+
+%%
 Once the machine semantics is in this form, simple point-wise abstraction leads to the summarization algorithms that we see in the literature.
@@
   {\mathcal F} &: \sa{\State} \to \System \to \System \\
-  {\mathcal F}(\mstate_0)(S, F, \mktab, \mmemo) &= (S \cup S', F' \cup (\set{\mstate_0} \setminus S), \mktab', \mmemo') \\
+  {\mathcal F}(\mastate_0)(S, F, \mktab, \mmemo) &= (S \cup S', F', \mktab', \mmemo') \\
   \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
@@
   \begin{align*}
-    S' &= \setbuild{\mastate'}{\wn(\mastate', \_, \_, \_) \in I} \\
+    S' &= \setbuild{\mastate'}{\wn(\mastate', \_, \_, \_) \in I} \cup \set{\mastate_0}\\
     F' &= S' \setminus S
@@
 %The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.}
+%
+The least fixed-point of ${\mathcal F}(\tpl{(\mexpr,\bot),\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
 %
@@
 %
-Let's take a look at the evaluation after binding \texttt{id}:
-
-\begin{center}
+The example evaluation is in \autoref{fig:ex-eval}.
+%
+We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
+%
+When determining control flow through the expression, we consult $\mmemo$ to continue past function calls, so there is no confusion about a back edge from the second call to \texttt{id}.
+%
+
+\begin{figure}
+  \centering
     \begin{minipage}{0.55\linewidth}
@@
   \end{minipage}
-\end{center}
 Where $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$
+\caption{Example analysis evaluation}
+\label{fig:ex-eval}
+\end{figure}
 % %% FIXME stores in memo-table
@@
 
-We maintained enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
-%
-When determining control flow through the expression, we consult $\mmemo$ to continue past function calls, so there is no confusion about a back edge from the second call to \texttt{id}.
-%
-
 \begin{theorem}\label{thm:pdcfa-tabular}
@@
 The $\mathit{unroll}$ function interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, defined as the greatest fixed point of the following rules:
-\begin{mathpar}
-  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt}} \quad
-  \inferrule{\mkont \in \unroll{\mktab}{\mkont'}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont'}} \quad
-  \inferrule{\mkont^\circ \in \mktab(\mctx) \\
-             \mkont \in \unroll{\mktab}{\mkont^\circ}}
-            {\mkont \in \unroll{\mktab}{\krt{\mctx}}}
-\end{mathpar}
-
-The astute reader may find this definition dubious, since there is a possibility for unguarded corecursion here.
-%
-Interpreting a function with ill-founded recursion can lead to a table such that $\krt{\mctx} \in \mktab(\mctx)$.
-%
-The abstractions that we impose on our data might also cause this situation, even though the concrete execution might always terminate.
-%
-Our interpretation of $\stepto$ is in a fixed-point computation, so circularities are not a problem for termination, but we augment the definition here to guard against circularities in our proofs:
-
 \begin{mathpar}
@@
 
-%%
-The different tables we have encode information about execution history that we prove is invariant.
+We add $G$ to protect against unguarded corecursion in order for the definition to be well-defined.
+%
+Interpreting a function with ill-founded recursion can lead to a table such that $\krt{\mctx} \in \mktab(\mctx)$.
+%
+The abstractions that we impose on our data might also cause this situation, even though the concrete execution might always terminate.
+
+%%
+The different tables we have encode information about execution history that we prove is invariant (\autoref{fig:inv}).
+%
+The first is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
+%
+The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
+%
+The final invariant, $\phi_{\memo}$, states that the memo table is correct: all calling contexts reach the memoized result, with the calling context's continuation in the tail (to justify using stack-irrelevance).
+
+\begin{figure}
+  \centering
   \begin{align*}
@@
     \dom(\mmemo) \subseteq \dom(\mktab) \\
-  &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab),
-                  \mkont^\circ \in \mktab(\mpoint',\mstore'),
+    &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab), \mkont^\circ
+    \in \mktab(\mpoint',\mstore'),
     \mkont' \in \unroll{\mktab}{\mkont^\circ, \emptyset}. \\
     &\qquad \phi_{\reachable} \wedge \phi_{\memo} \\
-  \text{where } \phi_{\reachable} &=
-  \exists \mkont^* \in \unroll{\mktab}{\mkont,\emptyset}.
+    \text{where } \phi_{\reachable} &= \exists \mkont^* \in
+    \unroll{\mktab}{\mkont,\emptyset}.
     \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont^*} \\
@@
     \forall (\mval, \mstore'') \in \mmemo(\mpoint',\mstore'). \\
-  &\qquad \exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore'', \mkont'}.
+    &\qquad \exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'}
+    \stepto^* \tpl{\mval,\mstore'', \mkont'}.
     \hastail(\mtrace,\mkont') \\
   \end{align*}
-
-The first is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
-%
-The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
-%
-The final invariant, $\phi_{\memo}$, states that the memo table is correct: all calling contexts reach the memoized result, with the calling context's continuation in the tail (to justify using stack-irrelevance).
-
+  \caption{Table invariants}
+\label{fig:inv}
+\end{figure}
 \begin{lemma}[Table invariants]\label{lem:tab-inv}
@@
   \quad
-  \inferrule{\mkont \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}
+  \inferrule{\mkont \rightsquigarrow_\mktab \mkont'}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont'}
+  \quad
+  \inferrule{\mkont \rightsquigarrow_\mktab \maddr}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
+%  \inferrule{\mkont \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}
   \quad
@@
 %% Begin scary Scribble output
-\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touch{-}kont}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mktab$}\RktPn{)}
+\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touch{-}kont}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\Xi$}\RktPn{)}
 
@@
 
-\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{let}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{loop}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{$\mkont$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{]}\RktPn{)}
-
-\mbox{\hphantom{\Scribtexttt{xxxx}}}\RktPn{(}\RktSym{match}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}
+\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{fix{-}touch}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xxxx}}}\RktPn{(}\RktSym{match}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}
 
@@
 
-\mbox{\hphantom{\Scribtexttt{xxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{cons}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkframe$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\cup$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touches{-}frame}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkframe$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{loop}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{)}\RktPn{)}\RktPn{]}
+\mbox{\hphantom{\Scribtexttt{xxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{cons}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\phi$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\cup$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touches{-}frame}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\phi$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{fix{-}touch}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}\RktPn{)}\RktPn{]}
 
@@
 
-\mbox{\hphantom{\Scribtexttt{xxxxxxxxxxxxxx}}}\RktPn{(}\RktSym{for/union}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{$\mkont$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{in{-}set}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{dict{-}ref}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mktab$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{ctx}\RktPn{)}\RktPn{)}\RktPn{]}\RktPn{)}
-
-\mbox{\hphantom{\Scribtexttt{xxxxxxxxxxxxxxxx}}}\RktPn{(}\RktSym{loop}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{)}\RktPn{)}\RktPn{]}\RktPn{)}\RktPn{]}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
+\mbox{\hphantom{\Scribtexttt{xxxxxxxxxxxxxx}}}\RktPn{(}\RktSym{union{-}map}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{fix{-}touch}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{ref}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\Xi$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{ctx}\RktPn{)}\RktPn{)}\RktPn{]}\RktPn{)}\RktPn{]}\RktPn{)}\RktPn{)}
+
+\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{fix{-}touch}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %
-\begin{theorem}\label{thm:gc-pointers}
+\begin{theorem}[No dangling pointers]\label{thm:gc-pointers}
   All addresses in $\Gamma(\mstate)$ are in the domain of $\Gamma(\mstate).\mstore$.
@@
 
-\begin{theorem}\label{thm:gc-concrete}
+\begin{theorem}[Garbage is irrelevant]\label{thm:gc-concrete}
   If $\mstate \stepto^n \mstate'$ and $\mstate \stepto^n \mstate''$ then $\Gamma(\mstate') = \Gamma(\mstate'')$.
@@
 
-\begin{theorem}\label{thm:gc-sound}
+\begin{theorem}[Abstract GC is sound for garbage-free states]\label{thm:gc-sound}
   If $\mstate \stepto \mstate'$ and $\alpha(\Gamma(\mstate)) \sqsubseteq \mastate$ then there is a $\mastate'$ such that $\mastate \astepto \mastate'$ and $\alpha(\Gamma(\mstate')) \sqsubseteq \mastate'$.
 \end{theorem}
-%
-%TODO: wrap up
-
+
+This final theorem is not in the original work on abstract GC; they prove soundness with respect to a concrete semantics that performs GC after every step.
+%
+The moral correctness of the abstraction is that it never under-approximates \emph{relevant} portions of the state space --- in this case, the reachable subset of the store.
+%
+The definitions of $\alpha$, $\astepto$ and $\sqsubseteq$ are the simple pointwise abstractions of addresses through states and the reduction relation, and finally a lifted subset relation between states and their components, respectively.
+%
+Both $\alpha$ and $\sqsubseteq$ are parameterized on the choice of $\Addr$, for which the following soundness criterion should hold:
+\begin{mathpar}
+  \inferrule*[right={[Sound allocation]}]{\alpha(\mstate) \sqsubseteq \mastate}{\alpha(\alloc(\mstate)) \sqsubseteq \widehat{\alloc}(\mastate)}
+\end{mathpar}
 \section{Deriving CFA2}
@@
 \end{align*}
-At capture time, we strip the store in the $\rt$ continuation if there is one, and replace it with an address:
+At capture time, we strip the store in the $\rt$ continuation if there is one, and replace it with an address to make the context storeable:
 
@@
 \paragraph{Comparison to PDCFA CPS'd to remove {\tt shift} and {\tt reset}:}{
-  One might ask what this different model buys us over using a CPS transform to compile away {\tt shift} and {\tt reset} forms.
-%
-The short answer is that we get better precision for capture and invocation of continuations.
+We lose precision if we use a CPS transform to compile away {\tt shift} and {\tt reset} forms, because variables are treated less precisely than continuations.
 %
@@
 
-The immediately related work is that of PDCFA \citep{dvanhorn:Earl2010Pushdown}, CFA2~\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:Vardoulakis2011Pushdown}, and AAM ~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we recreated in full detail.
-%
-The version of CFA2 that handles \rackett{call/cc} does not handle composable control, and is dependent on a restricted CPS representation.
-%
-They also had no way to tune the precision of their first class continuation approximation.
+The immediately related work is that of PDCFA \citep{dvanhorn:Earl2010Pushdown, dvanhorn:Earl2012Introspective}, CFA2~\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:Vardoulakis2011Pushdown}, and AAM ~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we recreated in full detail.
+%
+The version of CFA2 that handles \rackett{call/cc} does not handle composable control, is dependent on a restricted CPS representation, and has untunable precision for first-class continuations.
 %
@@
 %
-The extended version of PDCFA that does garbage collection~\citep{dvanhorn:Earl2012Introspective} also fits into our model;
+The extended version of PDCFA that inspects the stack to do garbage collection~\citep{dvanhorn:Earl2012Introspective} also fits into our model;
 the addresses that the stack keeps alive can be accumulated by ``reading through'' the continuation table, building up the set of addresses in each portion of the stack that we come across.
@@
 %%
+Stack inspecting flow analyses also exist, but operate on pre-constructed regular control-flow graphs~\citep{ianjohnson:bartoletti2004stack}, so the CFGs cannot be trimmed due to the extra information at construction time, leading to less precision.
+%
+Backward analyses for stack inspection also exist, with the same prerequisite~\citep{ianjohnson:DBLP:journals/sigplan/Chang06}.
+%%
+
+%%
 Pushdown models have existed in the first-order static analysis literature ~\citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995}, and the first-order model checking literature \citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time.
@@
 %%
-A new temporal logic that itself understands well-bracketing of call and returns for stating and validating propositions, NWTL ~\citep{ianjohnson:alur:nwtl:2007}, applies to visibly pushdown systems, which can precisely model programming languages that only have well-bracketed call and returns.
-%
-The propositional form of the logic has decidable satisfiability checking, and control-flow queries are shallow propositions to pose in the logic.
-%
-However, the satisfiability problem is \textsc{2Exptime}-hard and the algorithm is not given in the on-the-fly form that a higher-order language would need to effectively answer control-flow queries without already knowing the answers.
-%
-The logic itself is an exciting new frontier for expressing program correctness properties, and a higher-order version would be a welcome addition to the functional programmer's tool belt.
-%%
-
-%%
 The trend of deriving static analyses from abstract machines does not stop at flow analyses.
@@
 %%
-There is also the question of clients of these analyses.
+There is also the question of novel clients of these analyses.
 %

### 2013-10-08 content.tex
Fixed soundness issue in shift/reset

@@
 
-CFA2 uses a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the $\epsilon$-closure graph that PDCFA constructs.
+CFA2 is an online version of a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the $\epsilon$-closure graph that PDCFA constructs.
+%
+The online quality of the algorithm is what makes it suitable for higher-order languages.
 %
@@
 %
-This was shown for the \rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc}) operator in ~\citet{ianjohnson:Vardoulakis2011Pushdown}.
-%
-This impressive work illuminated the fact that we can harness the enhanced technology of pushdown analyses in non-pushdown models of computation.
+This was shown for the \rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc}) operator in~\citet{ianjohnson:Vardoulakis2011Pushdown}.
+%
+This work illuminated the fact that we can harness the enhanced technology of pushdown analyses in non-pushdown models of computation.
 %
@@
 %
-The key to their method is to notice in languages without stack-capturing features, the continuation is only modified a frame at a time.
+The key insight of PDCFA is to notice in languages without stack-capturing features, the continuation is only modified a frame at a time.
 %
@@
 %
+
+\subsection{Concrete semantics and summarization}
 We skip the detour to pushdown systems altogether and show an operational understanding of the resulting analysis.
-%%
-
-%%
+%
 We start by deriving PDCFA from an operational semantics for the untyped lambda calculus (\autoref{fig:base-semantics}).
@@
 For a concrete semantics, we require the allocation meta-function $\alloc :
-\State \to \Addr$ to return fresh addresses ($\alloc(\mstate) \notin \dom(\mstate.\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
-%
-If $\alloc$ only uses addresses from a finite subset of $\Addr$, then the state space without continuations is finite, and the space of contination frames is finite --- thus the pushdown system interpretation is apt.
-%
-This meta-function along with the commitment to having no recursive data-structures is the key to the AAM technique.
-%
-All recursion can be expressed by indirecting through addresses into a store --- tying Landin's knot.
+\State \to \Addr$ to return fresh addresses ($\alloc(\tpl{(\mexpr, \menv), \mstore, \mkont}) \notin \dom(\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
+%
+If $\alloc$ only uses addresses from a finite subset of $\Addr$, then the state space without continuations is finite, and the space of continuation frames is finite.
+%
+A finite allocation strategy makes the continuation the only unbounded data structure in the machine, which follows a stack discipline with a finite ``stack alphabet'' --- thus the pushdown system interpretation is apt.
+%
+The $\alloc$ meta-function along with the commitment to having no recursive data-structures is the key to the AAM technique.
+%
+All recursion can be expressed by indirecting through addresses into a store.
 %
@@
 %
-Instead of deferring to the table (store) for the remainder of a continuation for each frame, we only have indirections at function call boundaries (see rules 4 and 5 in figure \ref{fig:summary-semantics} for saving and restoring continuations).
+Instead of deferring to the table (store) for the remainder of a continuation for each frame, we only have indirections at function call boundaries (see last 2 rules in figure \ref{fig:summary-semantics} for saving and restoring continuations).
 %
@@
 %%
-The tail of a continuation in the context of a function will contain the stack-less context in which the function was called, in order to link up with the proper call-site(s).
-%
-The context includes the function (or unique label of the function), the environment, and the store.
-%
-Notice that this choice can be viewed as a special allocation strategy for continuations in the AAM viewpoint, but we separate the table of continuations from the store since continuations themselves contain stores in $\rt$ continuations --- this leads to a recursive data-structure that we are trying to avoid.
+The tail of a continuation within a function is some $\krt{\mctx}$, which instructs the semantics to memoize the result of the function evaluation at $\mctx$.
+%
+A context, $\mctx$, must contain enough information for memoization to be sound, \ie, any extra components not in $\mctx$ must be irrelevant to the evaluation of the function.
+%
+In this case, the context includes the function, the environment, and the store.
+%
+All the extra components at a context $\mctx$ get stored in a separate table, $\mktab$, so that updates to the memo table will propagate to all callers that depend on the result of $\mctx$.
+%%
+
+%%
+Continuations count as ``extra'' in this semantics since functions cannot inspect the stack, so $\krt{\mctx}$ might be considered a special allocation strategy for continuations in the AAM viewpoint.
+%
+We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores --- this leads to a recursive data-structure that we are trying to avoid.
 %
@@
           \mmemo}$
+     \\ & \quad if $(\mpoint, \mstore') \notin \dom(\mmemo)$
 \\
@@
     $\tpl{\mval, \mstore, \mkont', \kprompt{\mctx}, \mktab',\mmemo}$
+    \\ & \quad if $\mctx \notin \dom(\mmemo)$
     \\
@@
 
-The new indirection possibility changes $\Context$ to also include $\maddr$ where there was previously a $\mstore$ (though contexts for continuation calls remain unchanged), and $\KTab$ now additionally maps $\Addr$ to a set of stores.
-%
-The rules that change are presented in figure \ref{fig:shift-reset-table1}.
+The new indirection possibility changes $\Context$ to also include $\maddr$ where there was previously a $\mstore$, and the approximations of continuations live in the store is relevant to contexts now.
+%
+We add a new component to states and contexts, $\mmktab \in \MKTab : \Addr \to \wp(\Store)$, that is also subject to GC for enhanced precision if included.
+%
+Continuations in the store are further closed by $\mmktab$, so that becomes part of the memoized content in addition to the value and store.
+%
+The rules that change significantly are presented in figure \ref{fig:shift-reset-table1}.
 
@@
 \setlength{\belowdisplayshortskip}{8pt}
-\begin{minipage}{0.45\linewidth}
 \begin{align*}
   \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
-    \mctx \in \Context &::= \msctx \alt \ldots \\
-    \mskont \in \SKont &= \krt{\msctx} \alt \ldots
-  \end{align*}
-\end{minipage}
-\begin{minipage}{0.50\linewidth}
-  \begin{align*}
-    \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt \vcomp{\mskont} \\
-    \mktab \in \KTab &= (\Context \to \mathit{KRange}) \\
-    &\cup (\Addr \to \wp(\Store))
-  \end{align*}
-\end{minipage}
-
+  \mctx \in \Context &::= \msctx \alt ((\mexpr,\menv),\mstore,\mmktab) \alt (\vcomp{\mskont}, \mval, \mstore, \mmktab) \\
+  \mskont \in \SKont &= \kmt \alt \kcons{\mkframe}{\mskont} \alt \krt{\msctx} \\
+  \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt \vcomp{\mskont}
+  \end{align*}
 }
@@
     \hline
-    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont'}},\kmt,\mmkont,\mktab',\mmemo}$
-    \\
-    where & $(\mkont',\mktab') = \approximate(\mkont, \mktab, \maddr)$
+    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mmktab, \mktab, \mmemo}$
+    &
+    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore',\kmt,\mmkont, \mmktab', \mktab,\mmemo}$
+    \\
+    where & $\mstore' = \joinone{\mstore}{\maddr}{\vcomp{\mkont'}}$ \\
+    & $(\mkont',\mmktab') = \approximate(\mkont, \mmktab, \maddr)$
 \\
 % Return
-   $\tpl{\mval, \mstore, \krt{\mctx}, \mmkont, \mktab, \mmemo}$
-   &
-   $\tpl{\mval, \mstore, \mkont, \mmkont, \mktab, \mmemo'}$
-   if $\mkont \in \returns(\mktab, \mctx)$
-   \\
-   where & $\mmemo' = \mmemo \sqcup \memoize(\mktab, \mctx, (\mval, \mstore))$
+   $\tpl{\mval, \mstore, \krt{\mctx}, \mmkont, \mmktab, \mktab, \mmemo}$
+   &
+   $\tpl{\mval, \mstore, \mkont, \mmkont, \mmktab, \mktab, \mmemo'}$
+   if $\mkont \in \returns(\mmktab, \mktab, \mctx)$
+   \\
+   where & $\mmemo' = \mmemo \sqcup \memoize(\mmktab, \mctx, (\mval, \mstore, \mmktab))$
   \end{tabular}
@@
 \begin{align*}
-  \returns(\mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
-  \returns(\mktab, ((\mexpr,\menv), \maddr)) &=
-    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mktab(\maddr)}
+  \returns(\mmktab, \mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
+  \returns(\mmktab, \mktab, ((\mexpr,\menv), \maddr)) &=
+    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mmktab(\maddr)}
 \end{align*}
@@
 \begin{align*}
-  \memoize(\mktab, ((\mexpr,\menv), \mstore'), \mathit{v\sigma}) &=
-  [((\mexpr,\menv), \mstore') \mapsto \set{{\mathit{v\sigma}}}] \\
-  \memoize(\mktab, ((\mexpr,\menv), \maddr), \mathit{v\sigma}) &=
-  \bigsqcup\limits_{\mstore' \in \mktab(\maddr)}\snglm{((\mexpr,\menv), \mstore')}{\mathit{v\sigma}}
+  \memoize(\mmktab, ((\mexpr,\menv), \mstore'), \mathit{out}) &=
+  [((\mexpr,\menv), \mstore') \mapsto \set{{\mathit{out}}}] \\
+  \memoize(\mmktab, ((\mexpr,\menv), \maddr), \mathit{out}) &=
+  \bigsqcup\limits_{\mstore' \in \mmktab(\maddr)}\snglm{((\mexpr,\menv), \mstore')}{\mathit{out}}
 \end{align*}
@@
 \begin{align*}
-  \approximate(\mkont, \mktab, \maddr) &= (\replacectx(\mkont), \addstore(\mkont)) \\
+  \approximate(\mkont, \mmktab, \maddr) &= (\replacectx(\mkont), \addstore(\mkont)) \\
   \text{where }
@@
   \\[2pt]
-   \addstore(\kmt) &= \mktab \\
-   \addstore(\krt{((\mexpr,\menv),\mstore)}) &= \joinone{\mktab}{\maddr}{\mstore} \\
-   \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mktab}{\maddr}{\mktab(\maddr')} \\
+   \addstore(\kmt) &= \mmktab \\
+   \addstore(\krt{((\mexpr,\menv),\mstore)}) &= \joinone{\mmktab}{\maddr}{\mstore} \\
+   \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mmktab}{\maddr}{\mmktab(\maddr')} \\
    \addstore(\kcons{\mkframe}{\mkont}) &= \addstore(\mkont)
@@
 
-The immediately related work is that of PDCFA \citep{dvanhorn:Earl2010Pushdown, dvanhorn:Earl2012Introspective}, CFA2~\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:Vardoulakis2011Pushdown}, and AAM ~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we recreated in full detail.
+The immediately related work is that of PDCFA \citep{dvanhorn:Earl2010Pushdown, dvanhorn:Earl2012Introspective}, CFA2~\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:Vardoulakis2011Pushdown}, and AAM~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we recreated in full detail.
 %
@@
 %%
-Pushdown models have existed in the first-order static analysis literature ~\citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995}, and the first-order model checking literature \citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time.
+Pushdown models have existed in the first-order static analysis literature~\citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995}, and the first-order model checking literature \citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time.
 %

### 2013-10-08 content.tex
Okay, better fix for soundness issue.

@@
 %%
-
-%%
-Memoization with respect to the entire store is safe, if not a little wasteful; there are portions of the store that can change and be irrelevant to the result of a function call.
-%
-There are techniques to ignore differences in the heap when they are irrelevant, dubbed ``sparseness analysis'' that have been successful for large-scale first-order analysis~\citep{ianjohnson:DBLP:conf/pldi/OhHLLY12}, but they rely on a pre-analysis that knows the control-flow graph.
-%
-We have promising preliminary results in this direction that don't require a pre-analysis, but we leave further investigation of sparse techniques to future work.
 
@@
 
-%% FIXME: remove stores from memo-table since they will always be unchanged
-%% TODO: Show that store widening and store/memo/KTab widening have a bisimulation.
 {
@@
 
-%% TODO (thanks to ccshan): Discuss difference between CPS for shift/reset and this approach,
+%% (thanks to ccshan): Discuss difference between CPS for shift/reset and this approach,
 %% Discuss how this does not generalize to first class prompts due to append versus cons.
@@
    \\
-   where & $\mmemo' = \mmemo \sqcup \memoize(\mmktab, \mctx, (\mval, \mstore, \mmktab))$
+   where & $\mmemo' = \mmemo \sqcup \memoize(\mmktab, \mktab, \mctx, (\mval, \mstore, \mmktab))$
   \end{tabular}
@@
 
-The meta-functions mentioned in the fixed semantics all deal with the addition of $\maddr$ to contexts.
-%
-If the context in an $\rt$ continuation is approximate, we must return to all the continuations known for all the contexts it approximates:
-\begin{align*}
-  \returns(\mmktab, \mktab, ((\mexpr,\menv), \mstore)) &= \mktab((\mexpr,\menv), \mstore) \\
-  \returns(\mmktab, \mktab, ((\mexpr,\menv), \maddr)) &=
-    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore)}{\mstore \in \mmktab(\maddr)}
-\end{align*}
-
-At return boundaries, the memo table must add the result to all the represented contexts:
-\begin{align*}
-  \memoize(\mmktab, ((\mexpr,\menv), \mstore'), \mathit{out}) &=
-  [((\mexpr,\menv), \mstore') \mapsto \set{{\mathit{out}}}] \\
-  \memoize(\mmktab, ((\mexpr,\menv), \maddr), \mathit{out}) &=
-  \bigsqcup\limits_{\mstore' \in \mmktab(\maddr)}\snglm{((\mexpr,\menv), \mstore')}{\mathit{out}}
-\end{align*}
-At capture time, we strip the store in the $\rt$ continuation if there is one, and replace it with an address to make the context storeable:
+The meta-functions mentioned in the fixed semantics deal with the addition of $\maddr$ to contexts.
+%
+At capture time, we strip the store and stored continuations closure in the $\rt$ continuation if there is one, and replace them with an address to make the context storeable.
+%
+We can't recursively map addresses to stores and continuation closures, so we flatten the context's closure into the current one.
+%
+This operation makes lookups more costly since we have to find closures bounded above by the current closure, rather than one exact closure.
 
@@
    \replacectx(\kmt) &= \kmt \\
-   \replacectx(\krt{((\mexpr,\menv),\_)}) &= \krt{((\mexpr,\menv),\maddr)} \\
+   \replacectx(\krt{((\mexpr,\menv),\ldots)}) &= \krt{((\mexpr,\menv),\maddr)} \\
    \replacectx(\kcons{\mkframe}{\mkont}) &= \kcons{\mkframe}{\replacectx(\mkont)}
@@
    \addstore(\kmt) &= \mmktab \\
-   \addstore(\krt{((\mexpr,\menv),\mstore)}) &= \joinone{\mmktab}{\maddr}{\mstore} \\
+   \addstore(\krt{((\mexpr,\menv),\mstore,\mmktab')}) &= \joinone{\mmktab}{\maddr}{\mstore}\sqcup \mmktab' \\
    \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mmktab}{\maddr}{\mmktab(\maddr')} \\
    \addstore(\kcons{\mkframe}{\mkont}) &= \addstore(\mkont)
+\end{align*}
+%
+If the context in an $\rt$ continuation is approximate, we must return to all the continuations known for all the contexts it approximates:
+\begin{align*}
+  \returns(\mmktab, \mktab, ((\mexpr,\menv), \mstore, \mmktab')) &= \mktab((\mexpr,\menv), \mstore, \mmktab') \\
+  \returns(\mmktab, \mktab, ((\mexpr,\menv), \maddr)) &=
+    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore,\mmktab')}{(\mstore \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
+\end{align*}
+
+At return boundaries, the memo table must add the result to all the represented contexts:
+\begin{align*}
+  \memoize(\mmktab, \mktab, ((\mexpr,\menv), \mstore', \mmktab'), \mathit{out}) &=
+  [((\mexpr,\menv), \mstore',\mmktab') \mapsto \set{{\mathit{out}}}] \\
+  \memoize(\mmktab, \mktab, ((\mexpr,\menv), \maddr), \mathit{out}) &= [\mctx \mapsto \set{\mathit{out}} : \mctx \in \dom(\mktab) \cap C] \\
+    \text{where } C &= \setbuild{((\mexpr,\menv), \mstore', \mmktab')}{\mstore' \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
 \end{align*}
@@
 
-We prove this with the invariant that the continuation and meta-continuation have unique unrollings in addition to the table invariants from \autoref{sec:pdcfa}, since entries in $\mktab$ are always fresh.
+We prove this with the invariant that the continuation and meta-continuation have unique unrollings in addition to the table invariants from \autoref{sec:pdcfa}, since the distinct $\mstore$ at push and capture time makes entries in $\mktab$ unique, regardless of $\mmktab$.
 
@@
 
-%% TODO: complete
 \paragraph{Comparison to PDCFA CPS'd to remove {\tt shift} and {\tt reset}:}{
@@
 %%
+Memoization with respect to the entire store is safe, if not a little wasteful; there are portions of the store that can change and be irrelevant to the result of a function call.
+%
+There are techniques to ignore differences in the heap when they are irrelevant, dubbed ``sparseness analysis'' that have been successful for large-scale first-order analysis~\citep{ianjohnson:DBLP:conf/pldi/OhHLLY12}, but they rely on a pre-analysis that knows the control-flow graph.
+%
+We have promising preliminary results in this direction that don't require a pre-analysis, but we leave further investigation of sparse techniques to future work.
+
+%%
 There is also the question of novel clients of these analyses.

### 2013-10-09 content.tex
Incorporating DVH's comments.

@@
 %
-All of the semantics modeled in this paper are implemented in full detail in PLT Redex~\citep{ianjohnson:Felleisen:2009:SEP:1795772} and available online\footnote{\url{http://github.com/ianj/concrete-summaries}}.
+All of the semantics modeled in this paper are implemented in full detail in PLT Redex~\citep{ianjohnson:Felleisen:2009:SEP:1795772} and available online.\footnote{\url{http://github.com/ianj/concrete-summaries}}
 
@@
 We use a meta-function $\wn$, ``wide to narrow,'' to shift between the two representations of states.
-%%
-
-%%
-We show only the memo and continuation table widening, since store widening would require rewriting newly-created contexts to use the global store following the step.
-%
-This rewriting is unnecessary with the store-counting per-wide-step technique presented in \citet{ianjohnson:oaam:icfp2013}, or the less precise timestamp approximation from \citet{dvanhorn:Shivers:1991:CFA}.
-%
-Since these components are shared, they are always the least upper bound of any respective component that the semantics produces (the intermediate set $I$) in order to stay sound.
+%
+We show only the memo and continuation table widening, since store widening is standard.
+%
 
@@
 %
-PDCFA claims there is no loop, and depending on the implementation, that the result is true or false (paper), or just true (implemented)
-\footnote{This is because the paper steps every seen state with the current store every iteration, but the implementation only steps states that need stepping.}.
+PDCFA claims there is no loop, and depending on the implementation, that the result is true or false (paper), or just true (implemented).
+\footnote{This is because the paper steps every seen state with the current store every iteration, but the implementation only steps states that need stepping.}
 %
@@
 
-\begin{theorem}\label{thm:pdcfa-tabular}
-  Tabular semantics has a bisimulation with the standard semantics with comparable allocation strategies.
-\end{theorem}
-
-This follows directly from the invariants we prove of $\mktab$ and $\mmemo$ in the tabular semantics.
+\subsection{Metatheory}
+
+Correctness follows directly from the invariants we prove of $\mktab$ and $\mmemo$ in the tabular semantics.
 %
@@
 \begin{mathpar}
-  \inferrule{\mkont \in \unroll{\mktab}{\mkont^*}}
-            {\alloc(\tpl{\mpoint,\mstore,\mkont}) = \alloc^*(\tpl{\mpoint,\mstore,\mkont^*,\mktab,\mmemo})}
+  \inferrule{\mkont \in \unroll{\mktab}{\mkont'}}
+            {\alloc(\tpl{\mpoint,\mstore,\mkont}) = \alloc^*(\tpl{\mpoint,\mstore,\mkont',\mktab,\mmemo})}
 \end{mathpar}
@@
   \inferrule{\mctx \notin G  \\
-             \mkont^\circ \in \mktab(\mctx) \\
-             \mkont \in \unroll{\mktab}{\mkont^\circ, \set{\mctx}\cup G}}
+             \mkont'' \in \mktab(\mctx) \\
+             \mkont \in \unroll{\mktab}{\mkont'', \set{\mctx}\cup G}}
             {\mkont \in \unroll{\mktab}{\krt{\mctx}, G}}  
@@
     \dom(\mmemo) \subseteq \dom(\mktab) \\
-    &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab), \mkont^\circ
+    &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab), \mkont''
     \in \mktab(\mpoint',\mstore'),
-    \mkont' \in \unroll{\mktab}{\mkont^\circ, \emptyset}. \\
+    \mkont' \in \unroll{\mktab}{\mkont'', \emptyset}. \\
     &\qquad \phi_{\reachable} \wedge \phi_{\memo} \\
-    \text{where } \phi_{\reachable} &= \exists \mkont^* \in
+    \text{where } \phi_{\reachable} &= \exists \mkont'' \in
     \unroll{\mktab}{\mkont,\emptyset}.
-    \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont^*} \\
+    \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont''} \\
     \phi_{\memo} &=
@@
 In order to prove this, we need a lemma that allows us to plug in any continuation for memoized results.
-
-\begin{lemma}[Stack irrelevance]\label{lem:irrelevance}
+%
+When traces (sequences of states such that adjacent states are related by $\stepto$) have a common tail, we can replace it with anything.
+
+\begin{lemma}[Stack irrelevance in CESK]\label{lem:irrelevance}
   $\hastail(\mtrace,\mkont)$ implies $\replacetail(\mtrace,\mkont,\mkont')$ is a valid trace.
@@
 
+\begin{claim}(Correctness of PDCFA)\label{thm:pdcfa-tabular}
+  Tabular semantics has a bisimulation with the standard semantics with comparable allocation strategies.
+\end{claim}
+
 Finally, we can show that even if the tables are shared amongst all states, they are precise enough to not interfere with the results.
-\begin{theorem}[Equivalence with global tables]\label{thm:global}
+\begin{claim}[Equivalence with global tables]\label{thm:global}
   Tabular semantics has a bisimulation with memo-and-k-table-widened semantics.
-\end{theorem}
+\end{claim}
 
@@
 
-An implementation of the same functionality would use a mutable set to track which contexts have been traversed to build up a set of addresses:
-% This code is typeset below by Scribble
-% \begin{lstlisting}[mathescape]
-% (define (touch-kont $\mkont$ $\mktab$)
-%   (define seen (mutable-seteq))
-%   (let loop ([$\mkont$ $\mkont$])
-%     (match $\mkont$
-%       [(mt) $\emptyset$]
-%       [(cons $\mkframe$ $\mkont$) ($\cup$ $\touchesf(\mkframe)$ (loop $\mkont$))]
-%       [(rt $\mctx$)
-%        (cond
-%          [(set-member? seen $\mctx$) $\emptyset$]
-%          [else
-%           (set-add! seen $\mctx$)
-%           (for/union ([$\mkont$ (in-set $\mktab(\mctx)$)]) (loop $\mkont$))])])))           
-% \end{lstlisting}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Begin scary Scribble output
-\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touch{-}kont}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\Xi$}\RktPn{)}
-
-\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{seen}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{mutable{-}seteq}\RktPn{)}\RktPn{)}
-
-\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{fix{-}touch}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}
-
-\mbox{\hphantom{\Scribtexttt{xxxx}}}\RktPn{(}\RktSym{match}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}
-
-\mbox{\hphantom{\Scribtexttt{xxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{mt}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\emptyset$}\RktPn{]}
-
-\mbox{\hphantom{\Scribtexttt{xxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{cons}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\phi$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\cup$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touches{-}frame}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\phi$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{fix{-}touch}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}\RktPn{)}\RktPn{]}
-
-\mbox{\hphantom{\Scribtexttt{xxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{rt}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{ctx}\RktPn{)}
-
-\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{(}\RktSym{cond}
-
-\mbox{\hphantom{\Scribtexttt{xxxxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{set{-}member{\hbox{\texttt{?}}}}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{seen}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{ctx}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\emptyset$}\RktPn{]}
-
-\mbox{\hphantom{\Scribtexttt{xxxxxxxx}}}\RktPn{[}\RktSym{else}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{set{-}add{\hbox{\texttt{!}}}}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{seen}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{ctx}\RktPn{)}
-
-\mbox{\hphantom{\Scribtexttt{xxxxxxxxxxxxxx}}}\RktPn{(}\RktSym{union{-}map}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{fix{-}touch}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{ref}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\Xi$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{ctx}\RktPn{)}\RktPn{)}\RktPn{]}\RktPn{)}\RktPn{]}\RktPn{)}\RktPn{)}
-
-\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{fix{-}touch}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% End of scary Scribble output
-
 Garbage collection ($\Gamma$) amounts to restricting the heap to just the reachable addresses, so a garbage-collecting semantics introduces the following reduction rule:
@@
 %
-\begin{theorem}[No dangling pointers]\label{thm:gc-pointers}
+\begin{claim}[No dangling pointers]\label{thm:gc-pointers}
   All addresses in $\Gamma(\mstate)$ are in the domain of $\Gamma(\mstate).\mstore$.
-\end{theorem}
-
-\begin{theorem}[Garbage is irrelevant]\label{thm:gc-concrete}
+\end{claim}
+
+\begin{claim}[Garbage is irrelevant]\label{thm:gc-concrete}
   If $\mstate \stepto^n \mstate'$ and $\mstate \stepto^n \mstate''$ then $\Gamma(\mstate') = \Gamma(\mstate'')$.
-\end{theorem}
-
-\begin{theorem}[Abstract GC is sound for garbage-free states]\label{thm:gc-sound}
+\end{claim}
+
+\begin{claim}[Abstract GC is sound for garbage-free states]\label{thm:gc-sound}
   If $\mstate \stepto \mstate'$ and $\alpha(\Gamma(\mstate)) \sqsubseteq \mastate$ then there is a $\mastate'$ such that $\mastate \astepto \mastate'$ and $\alpha(\Gamma(\mstate')) \sqsubseteq \mastate'$.
-\end{theorem}
-
-This final theorem is not in the original work on abstract GC; they prove soundness with respect to a concrete semantics that performs GC after every step.
+\end{claim}
+
+This final proposition is not in the original work on abstract GC; they prove soundness with respect to a concrete semantics that performs GC after every step.
 %
@@
 %
-There are two orthogonal features of the semantics, embodied by our semantics' two metafunctions ($\bind$ and $\lookup$, respectively in \autoref{fig:frame-semantics}):
+
 \begin{enumerate}
@@
 \end{enumerate}
+There are two orthogonal features of the semantics, embodied by our semantics' two metafunctions ($\bind$ and $\lookup$, respectively in \autoref{fig:frame-semantics}).
+%
 The first of these is an addition to the stack-less context.
@@
 %
-We give a simplified semantics that is more along the AAM style, but CFA2's approach can easily be recovered from it\footnote{Our Redex model implements the fake rebinding strategy CFA2 itself employs}.
+We give a simplified semantics that is more along the AAM style, but CFA2's approach can easily be recovered from it.\footnote{Our Redex model implements the fake rebinding strategy CFA2 itself employs.}
 %%
@@
   \bind(\mstore,\maddr,\mvar,\mval) &=
-   \left\{\begin{array}{l}
-            (\mstore, \snglm{\maddr}{\mval}) \text{ if } \mvar \text{ allocations never escape} \\
-            (\joinone{\mstore}{\maddr}{\mval}, \snglm{\maddr}{\mval}) \text{ otherwise} \\
+   \left\{\begin{array}{ll}
+            (\mstore, \snglm{\maddr}{\mval}) &\text{ if } \mvar \text{ allocations never escape} \\
+            (\joinone{\mstore}{\maddr}{\mval}, \snglm{\maddr}{\mval}) &\text{ otherwise} \\
           \end{array}\right. \\
   \lookup(\mstore,\mframe,\menv,\mvar,\mlab) &=
-    \left\{\begin{array}{l}
+    \left\{\begin{array}{ll}
           \setbuild{(\extm{\mframe}{\menv(\mvar)}{\set{\mval}}, \mval)}
-                   {\mval \in \mframe(\menv(\mvar))}
-                   \\ \qquad \text{if } \mlab \text{ non-escaping} \\
-          \setbuild{(\mframe,\mval)}{\mval \in \mstore(\menv(\mvar))} \text{ otherwise}
+                   {\mval \in \mframe(\menv(\mvar))} & \text{if } \mlab \text{ non-escaping} \\
+          \setbuild{(\mframe,\mval)}{\mval \in \mstore(\menv(\mvar))} & \text{ otherwise}
            \end{array}\right.
@@
 We follow up with a similar proof to that of PDCFA, which is that the CES$\xi$K machine has a bisimulation with its tabular semantics, with comparable allocation strategies.
-\begin{theorem}[Correctness of stack frame refinement]\label{thm:refinement}
+\begin{claim}[Correctness of stack frame refinement]\label{thm:refinement}
   The CES$\xi$K machine has a bisimulation with the CESK machine with fresh allocation.
-\end{theorem}
+\end{claim}
 This follows from the invariant that if a stack frame maps $\maddr$ to $\mval$, then the store maps $\maddr$ to $\mval$, meaning that the special lookup function is equivalent to lookup directly in the store (exploits immutability).
@@
 
-\begin{theorem}[Correctness of CFA2]\label{thm:cfa2}
+\begin{claim}[Correctness of CFA2]\label{thm:cfa2}
   Tabular semantics has a bisimulation with standard semantics.
-\end{theorem}
+\end{claim}
 This proof has the same structure as that of PDCFA.
@@
 
-\begin{theorem}\label{thm:concrete-sr}
+\begin{claim}\label{thm:concrete-sr}
   Tabular semantics has a bisimulation with standard semantics given fresh allocation.
-\end{theorem}
+\end{claim}
 
@@
 
-\begin{theorem}\label{thm:sound-sr}
+\begin{claim}\label{thm:sound-sr}
   Tabular semantics simulates standard semantics with abstract allocation.
-\end{theorem}
+\end{claim}
 

### 2013-10-11 content.tex
Fixed claims of correctness w.r.t. concerns about lost paths given memoization.

@@
 
+This semantics is only correct for fresh allocation strategies, which in turn make the reduction relation deterministic.
+%
+As it is, the CESK machine simulates the tabular machine, but not the other way around;
+%
+if we use a finite allocation strategy, then we don't explore some non-deterministic paths that lead to different results for a function call.
+%
+We thus explore all paths and share the k-table and memo tables between all paths.
+
 To turn this semantics into PDCFA's algorithm for constructing a Dyck state graph, given the appropriate $\alloc$, we apply a widening operator to make $\mstore$, $\mkont$, and $\mmemo$ shared amongst all states (meaning states then become represented without these components), in what is then called a $\System$.
@@
   {\mathcal F} &: \sa{\State} \to \System \to \System \\
-  {\mathcal F}(\mastate_0)(S, F, \mktab, \mmemo) &= (S \cup S', F', \mktab', \mmemo') \\
-  \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
+  {\mathcal F}(\mastate_0)(S, \emptyset, \mktab, \mmemo) &= (S, \emptyset, \mktab, \mmemo) \\
+  {\mathcal F}(\mastate_0)(S, F\cup\set{\mastate}, \mktab, \mmemo) &= (S \cup S', F\cup F', \mktab', \mmemo') \\
+  \text{where } I &= \setbuild{\mstate'}{\wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
                 \wn(\tpl{(\mexpr,\menv),\mstore,\mkont},\mktab,\mmemo)
@@
   \begin{align*}
-    \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \_, \mktab', \_) \in I} \\
-    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \_, \mmemo') \in I}
+    \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \mktab', \_) \in I} \\
+    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \mmemo') \in I}
   \end{align*}
@@
   \begin{align*}
-    S' &= \setbuild{\mastate'}{\wn(\mastate', \_, \_, \_) \in I} \cup \set{\mastate_0}\\
+    S' &= \setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \cup \set{\mastate_0}\\
     F' &= S' \setminus S
@@
 %
-PDCFA claims there is no loop, and depending on the implementation, that the result is true or false (paper), or just true (implemented).
-\footnote{This is because the paper steps every seen state with the current store every iteration, but the implementation only steps states that need stepping.}
+PDCFA claims there is no loop, and depending on the implementation, that the result is true.
 %
@@
 %
-When determining control flow through the expression, we consult $\mmemo$ to continue past function calls, so there is no confusion about a back edge from the second call to \texttt{id}.
+When determining control flow through the expression, we consult $\mktab$ to continue past function returns, so there is no confusion about a back edge from the second call to \texttt{id}.
 %
@@
       \begin{center}
-        $\begin{array}{l}\includegraphics[scale=0.65]{example1}\end{array}$ \\
+        $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$ \\
         $\Downarrow$ \\
-        $\begin{array}{l}\includegraphics[scale=0.65]{example2}\end{array}$ \\
+        $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$ \\
         $\Downarrow$ \\
-        $\begin{array}{l}\includegraphics[scale=0.65]{example3}\end{array}$
+        $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
       \end{center}
@@
           \hline
-          0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\bot)}]$ & $\mkont_0$ \\
+          0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\iota)}]$ & $\mkont_0$ \\
           1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
-          2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)})$ \\
+          2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)}, \iota)$ \\
           3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
@@
 \begin{mathpar}
-  \inferrule{\mkont \in \unroll{\mktab}{\mkont'}}
+  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \emptyset}}
             {\alloc(\tpl{\mpoint,\mstore,\mkont}) = \alloc^*(\tpl{\mpoint,\mstore,\mkont',\mktab,\mmemo})}
@@
 
-We add $G$ to protect against unguarded corecursion in order for the definition to be well-defined.
+We add $G$ to protect against unguarded corecursion in order for $\mathit{unroll}$ to be well-defined.
 %
@@
 %
-The final invariant, $\phi_{\memo}$, states that the memo table is correct: all calling contexts reach the memoized result, with the calling context's continuation in the tail (to justify using stack-irrelevance).
+The final invariant, $\phi_{\memo}$, states that all paths starting at a function call either reach the memoized result, or if the path does not include a return, there is an extension that will.
+%
+The portion of these paths between call and return hal the calling context's continuation in the tail to justify using stack-irrelevance.
 
@@
     \forall (\mval, \mstore'') \in \mmemo(\mpoint',\mstore'). \\
-    &\qquad \exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'}
-    \stepto^* \tpl{\mval,\mstore'', \mkont'}.
-    \hastail(\mtrace,\mkont') \\
+    &\qquad\exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore'',\mkont'}. \hastail(\mtrace,\mkont')
   \end{align*}
@@
 \end{figure}
-\begin{lemma}[Table invariants]\label{lem:tab-inv}
+\begin{lemma}[Table invariants in CESK$\mathit{\Xi{}M}$]\label{lem:tab-inv}
   If $\inv(\mstate)$ and $\mstate \stepto \mstate'$ then $\inv(\mstate')$.
@@
 
-\begin{claim}(Correctness of PDCFA)\label{thm:pdcfa-tabular}
-  Tabular semantics has a bisimulation with the standard semantics with comparable allocation strategies.
+\begin{claim}(Correctness of Concrete Summarization)\label{thm:concrete-tabular}
+  The standard semantics has a bisimulation with the tabular semantics given comparable fresh allocation strategies.
 \end{claim}
 
-Finally, we can show that even if the tables are shared amongst all states, they are precise enough to not interfere with the results.
-\begin{claim}[Equivalence with global tables]\label{thm:global}
-  Tabular semantics has a bisimulation with memo-and-k-table-widened semantics.
+Finally, we can show that the reductions we find between states paired with the shared tables define a sound and complete abstraction of the CESK machine with a comparable allocation strategy.
+\begin{claim}[Correctness of PDCFA]\label{thm:global}
+  Widened tabular semantics reifies a relation that has a bisimulation with the standard semantics.
 \end{claim}
@@
 \begin{align*}
-  \reaches(\mathit{root}, \mstore) &= \setbuild{b}{\maddr \in \mathit{root}, \maddr \rightsquigarrow_\mstore^* b} \\
-   &
+  \reaches(\mathit{root}, \mstore) = \setbuild{b}{\maddr \in \mathit{root}, \maddr \rightsquigarrow_\mstore^* b}
+   \qquad
                \inferrule{\mval \in \mstore(\maddr) \\
@@
 
-This semantics of course enjoys the soundness properties that GC does not introduce dangling pointers (\autoref{thm:gc-pointers}), equal length traces are equal up to GC (in the concrete) (\autoref{thm:gc-concrete}), and abstractions are sound with respect to collected states (\autoref{thm:gc-sound}).
+This semantics of course enjoys the soundness properties that GC does not introduce dangling pointers, equal length traces are equal up to GC (in the concrete), and abstractions are sound with respect to collected states.
 %
@@
 \item{stack allocation for some bindings in an additional $\mframe \in \Store$}
-\item{strong updates on stack frames for resolved non-determinism}
+\item{strong updates on stack frames for resolved addresses}
 \end{enumerate}
@@
 %
-In AAM, a variable reference non-deterministically steps to all possible values associated with that variable.
+In AAM, a variable reference non-deterministically steps to all possible values associated with that variable, because there are several possible values in the store for allocated bindings.
 %
@@
 \begin{claim}[Correctness of CFA2]\label{thm:cfa2}
-  Tabular semantics has a bisimulation with standard semantics.
+  The widened tabular semantics reifies a relation that has a bisimulation with standard semantics.
 \end{claim}
@@
 \begin{claim}\label{thm:concrete-sr}
-  Tabular semantics has a bisimulation with standard semantics given fresh allocation.
+  Tabular semantics simulates the standard semantics given fresh allocation.
 \end{claim}
@@
 \begin{claim}\label{thm:sound-sr}
-  Tabular semantics simulates standard semantics with abstract allocation.
+  Widened tabular semantics reifies a relation that simulates standard semantics.
 \end{claim}

### 2013-12-09 content.tex
Tightening wording and split out some text to 'tech report' version using an iftr macro.

@@
 %%
-There is a systematic process for transforming off-the-shelf programming language semantics into a form amenable to \emph{regular} analysis that has been widely applied with great success to production programming languages~\citet{dvanhorn:VanHorn2010Abstracting} (a technique called abstracting abstract machines, or AAM).
+There is a systematic process for transforming off-the-shelf programming language semantics into a form amenable to \emph{regular} analysis that has been widely applied with great success to production programming languages (a technique called abstracting abstract machines, or AAM~\citep{dvanhorn:VanHorn2010Abstracting}).
 %
@@
 %
-This was shown for the \rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc}) operator in~\citet{ianjohnson:Vardoulakis2011Pushdown}.
+This was shown for the \rackett{call-with-current-continuation} (or \rackett{call/cc}) operator in~\citet{ianjohnson:Vardoulakis2011Pushdown}.
 %
@@
 %%
-A downside of the work providing \rackett{call/cc} is that it was an algorithmic change to the already complicated CFA2 --- there was no recipe for how to do this for one's favorite control operator.
-%
-This paper seeks to do just that with an operational view of what summarization is, in essence.
-%
-In other words, we give a concrete semantics to the tricks that the analyses in the literature use in the abstract, and maintain the original meaning of the language.
-%
-We also give intuitive analogies to well-established ideas and techniques so that the working semanticist can write a pushdown analysis for her language.
+This paper extends the work providing \rackett{call/cc} by giving an operational view of what summarization is, in essence.
+%
+In other words, we give a concrete semantics to the algorithmic techniques that the analyses in the literature use in the abstract, and maintain the original meaning of the language.
+%
+We also give intuitive analogies to well-established ideas and techniques so that the working semanticist can write a pushdown analysis for their language.
 %
@@
 \begin{enumerate}
-\item{bound continuations with with delimiters that specify the exact context in which any continuation replacement would lead to the same result, and}
-\item{memoize at these delimiters}
+\item{memoize calls with a ``return'' stack frame that tracks calling context, and}
+\item{store stacks that reach calls in order to continue to them after memoizing}
 \end{enumerate}
@@
 %%
-The first of these two might seem unfamiliar, but it is just identifying what is needed for sound memoization, and we delimit the continuations in order to mark where to memoize results.
-%
-In the case of a CESK machine, where the store is relevant to the evaluation of functions, a context is just the CES portion since stacks don't matter.
-%
-For stack inspecting analyses, a context is CES plus the inspected portion of K.
-%
-As soon as we have a value that such a context evaluates to, all subsequent visits to that context can jump straight to the memoized result(s).
+The first of these is a semantic approach to memoizing that is a straight-forward analog to the programmatic approach.
+%
+A memoizing program will call a function in a non-tail context in order to obtain the result, store the result to an ``enough context to obtain the same result'' key in a table (in a pure setting, just the function and arguments), and continue with the result.
+%
+The semantics we use is functionally equivalent, only we have a special stack frame that acts as the non-tail context that additionally stores results in a table that is part of the machine state.
+%
+In the case of a baseline abstract machine for AAM, a CESK machine~\citep{dvanhorn:Felleisen1987Calculus}, the store is relevant to the evaluation of functions.
+%
+As such, the CES portion of a state suffices as ``enough context.''
+%
+For a semantics that inspects the stack, ``enough context'' is CES plus the inspected portion of K.
+%%
+
+%%
+The second piece of machinery serves two purposes:
+%
+(1) the state space contains all function callsites to project out a control-flow graph, and %
+(2) the worklist semantics of execution makes it possible to appeal to an incomplete memo table, so the saved stacks allow new results to flow to calls that used incomplete memo information.
+%
+Consider a non-deterministic semantics where a function call reaches a result and memoizes, a second call uses the memo table (storing its stack), and then the original call reaches another result (since the worklist allows for such parallel reduction).
+%
+The second result had better also flow to the second call to maintain soundness!
 %%
@@
 %
-If the state space without the stack is finite (and it can be made finite with an approximation ala AAM), then the model falls directly into the realm of pushdown systems.
-%
-They thus recast the problem in terms of pushdown systems.
+If the state space without the stack is finite (and it can be made finite with an approximation ala AAM), then the model is naturally recast into the realm of pushdown systems.
 %
@@
 %
-They call the main data-structure for this a Dyck state graph.
-%
-They additionally compute an $\epsilon$-closure graph to match nodes that push frames with later nodes following the pop of that frame, which keeps edges between nodes that have a net-zero stack change path between them.
+A key component is computing the $\epsilon$-closure graph to match nodes that push frames with later nodes following the pop of that frame.
+%
+This stores all edges between nodes that have a net-zero stack change path between them.
 %
@@
 For a concrete semantics, we require the allocation meta-function $\alloc :
-\State \to \Addr$ to return fresh addresses ($\alloc(\tpl{(\mexpr, \menv), \mstore, \mkont}) \notin \dom(\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
+\State \to \Addr$ to return fresh addresses ($\alloc(\mstate) \notin \dom(\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
 %
@@
 %
-There is an analogy to something more operational: summary edges embody memoization (see the last rule in figure \ref{fig:summary-semantics}).
-%
-Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store (the second case of rule 4).
+There is an analogy to something more operational: summary edges embody memoization.
+%
+Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store [memo lookup].
 %%
@@
   $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
-  \begin{tabular}{r|l}
+  \begin{tabular}{r|ll}
     \hline
@@
     $\tpl{\mval,\mstore,\mkont, \mktab, \mmemo}$ if $\mval \in \mstore(\menv(\mvar))$
+    &
+    [variable lookup]
     \\
@@
     $\tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}$
+    &
+    [function eval]
     \\
@@
     $\tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}$
+    &
+    [argument eval]
     \\
@@
           \mmemo}$
+    &
+    [un-memoized call] 
      \\ & \quad if $(\mpoint, \mstore') \notin \dom(\mmemo)$
@@
             \mmemo}$
+    & [memo lookup]
     \\ & \quad if $(\mval_\mathit{result},\mstore'') \in \mmemo(\mpoint,\mstore')$
@@
     $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}$
+    &
+    [memoize/return]
     \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
@@
 %
-We thus explore all paths and share the k-table and memo tables between all paths.
-
-To turn this semantics into PDCFA's algorithm for constructing a Dyck state graph, given the appropriate $\alloc$, we apply a widening operator to make $\mstore$, $\mkont$, and $\mmemo$ shared amongst all states (meaning states then become represented without these components), in what is then called a $\System$.
+We thus explore all paths and share the KTable and memo tables between all paths.
+
+To turn this semantics into PDCFA's algorithm for constructing a pushdown system with only reachable states, given the appropriate $\alloc$, we apply a widening operator to make $\mkont$, and $\mmemo$ shared amongst all states (meaning states then become represented without these components), in what is then called a $\System$.
 We use a meta-function $\wn$, ``wide to narrow,'' to shift between the two representations of states.
-%
-We show only the memo and continuation table widening, since store widening is standard.
 %
@@
   {\mathcal F}(\mastate_0)(S, \emptyset, \mktab, \mmemo) &= (S, \emptyset, \mktab, \mmemo) \\
-  {\mathcal F}(\mastate_0)(S, F\cup\set{\mastate}, \mktab, \mmemo) &= (S \cup S', F\cup F', \mktab', \mmemo') \\
+  {\mathcal F}(\mastate_0)(S, F\uplus\set{\mastate}, \mktab, \mmemo) &= (S \cup S', F\cup F', \mktab', \mmemo') \\
   \text{where } I &= \setbuild{\mstate'}{\wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
@@
 %
-0CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
-%
-PDCFA claims there is no loop, and depending on the implementation, that the result is true.
+0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
+%
+PDCFA claims there is no loop and that the result is true.
 %
@@
 \begin{figure}
-  \centering
-    \begin{minipage}{0.55\linewidth}
+      \begin{tabular}{rcl}
+        $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$
+        & $\Rightarrow$ &
+        $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$ \\
+        & & \multicolumn{1}{c}{$\Downarrow$} \\
+        \begin{minipage}{0.5\linewidth}
           \begin{center}
-        $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$ \\
-        $\Downarrow$ \\
-        $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$ \\
-        $\Downarrow$ \\
+            {\small
+              \begin{tabular}{|l|c|c|c|}
+                \hline
+                \# & $\mktab$ & $\mmemo$ \\
+                \hline
+                0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
+                1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
+                \hline
+              \end{tabular}} \\[8pt]
+            {\small
+              \begin{tabular}{|l|c|}
+                \hline
+                \# & $\mctx$ \\
+                \hline
+                0  & $(({\tt x}, \iota), \mstore_1)$ \\
+                1  & $(({\tt x}, \iota), \mstore_3)$ \\
+                \hline
+              \end{tabular}}
+          \end{center}
+
+      \end{minipage} & &
         $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
-      \end{center}
-    \end{minipage}
-    \begin{minipage}{0.43\linewidth}
+      \end{tabular}
     \begin{center} {\small
@@
         \end{tabular}} \\[10pt]
-      {\small
-        \begin{tabular}{|l|c|}
-          \hline
-          \# & $\mctx$ \\
-          \hline
-          0  & $(({\tt x}, \iota), \mstore_1)$ \\
-          1  & $(({\tt x}, \iota), \mstore_3)$ \\
-          \hline
-        \end{tabular}}
-      \\[10pt]
-      {\small
-        \begin{tabular}{|l|c|c|c|}
-          \hline
-          \# & $\mktab$ & $\mmemo$ \\
-          \hline
-          0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
-          1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
-          \hline
-        \end{tabular}}    
+ Where $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$
     \end{center}
-  \end{minipage}
-Where $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$
 \caption{Example analysis evaluation}
@@
 
+\iftr{%
 \subsection{Metatheory}
@@
 
-\begin{claim}(Correctness of Concrete Summarization)\label{thm:concrete-tabular}
-  The standard semantics has a bisimulation with the tabular semantics given comparable fresh allocation strategies.
-\end{claim}
+\begin{theorem}[Correctness of Concrete Summarization]\label{thm:concrete-tabular}
+  The standard semantics has a skipping bisimulation with the tabular semantics given comparable fresh allocation strategies.
+\end{theorem}
+\begin{proof}[Sketch]
+  The tabular semantics skipping simulates that standard semantics by appealing to the table invariants to show the long ``recomputation'' path.
+  Crucially, since the memo table is only set once before being used to bypass to the result, there must be only one possible result in the standard semantics.
+Otherwise, the tabular semantics would miss out on the other results for subsequent calls with the same context.
+  The fresh allocation strategy ensures that relation is deterministic and thus function calls only reduce to one result.
+
+  The standard semantics skipping simulates the tabular semantics because ``recomputing'' paths always have the calling stack as its tail, so we can step forward on the invariant's inductively built path until the result.
+\end{proof}
 
 Finally, we can show that the reductions we find between states paired with the shared tables define a sound and complete abstraction of the CESK machine with a comparable allocation strategy.
-\begin{claim}[Correctness of PDCFA]\label{thm:global}
-  Widened tabular semantics reifies a relation that has a bisimulation with the standard semantics.
-\end{claim}
+}
+\begin{theorem}[Correctness of PDCFA]\label{thm:global}
+  With a fair state selection from the frontier, the widened tabular semantics reflects a relation that has a bisimulation with the standard semantics.
+\end{theorem}
+\iftr{%
+\begin{proof}[Sketch]
+  The invariants on the tables still hold. 
+  The widening to share $\mmemo$ and $\mktab$ amongst all explored states reifies the search state into a $\sa{System}$. We can add a set of edges simply.
+  If we add $\mmemo$ and $\mktab$ to all states in these edges, we get a reduction relation comparable to the unwidened tabular semantics.
+  We prove that the infinite unrolling of the standard reduction relation stuttering bisimulates the relation reified from an infinite stepping of the widened tabular semantics.
+  This only holds if the state selection from the frontier is fair, since unfair choices can leave some function calls unresolved.
+  Formally,
+  \begin{equation*}
+    \setbuild{\mtrace}{\mtrace : i \le \omega, \mstate_0 \longmapsto^i \mstate} \simeq {\mathcal F}(\mstate_0)^\omega
+  \end{equation*}
+  where $\simeq$ is TODO
+by coinduction and appeals to the table invariants.
+  If a new function result is found after a call already used the memo table, the context of the memo user is used after memoizing, so it looks like that result was part of the original memo table lookup.
+  The fair selection guarantees that all function results will be eventually discovered, so the memo table appeals eventually catch up to the found results.
+\end{proof}
+}
 
@@
 Entries in $\mktab$ might be circular, so we compute the continuation frames in a fixed-point before computing touched addresses:
-\begin{align*}
-  \touchesk{\mkont}(\mktab) &= \setbuild{\maddr}{\mkont \rightsquigarrow_\mktab^* \maddr}
-\\[2pt]
-  \touchesf(\kar{\mexpr,\menv}) &= \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)}) \\
-  \touchesf(\kfn{\mval}) &= \touches(\mval)
-\end{align*}
+\begin{center}
+  \,
+  \begin{minipage}{0.32\linewidth}
+    $\touchesk{\mkont}(\mktab) = \setbuild{\maddr}{\mkont \rightsquigarrow_\mktab^*\maddr}$
+  \end{minipage}
+  \begin{minipage}{0.36\linewidth}
+    $\touchesf(\kar{\mexpr,\menv}) = \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)})$
+  \end{minipage}
+  \begin{minipage}{0.3\linewidth}
+    $\touchesf(\kfn{\mval}) = \touches(\mval)$
+  \end{minipage}
+\end{center}
 \begin{mathpar}
@@
 Garbage collection ($\Gamma$) amounts to restricting the heap to just the reachable addresses, so a garbage-collecting semantics introduces the following reduction rule:
-\begin{align*}
-  \mstate &\stepto \Gamma(\mstate) \\
+\begin{center}
+  \begin{minipage}{0.3\linewidth}
+    \begin{align*}
+      \frac{\mstate \stepto \mstate'}{\mstate \stepto \Gamma(\mstate')}
+    \end{align*}
+  \end{minipage}
+  \begin{minipage}{0.65\linewidth}
+    \begin{align*}
       \Gamma(\tpl{(\mexpr,\menv),\mstore,\mkont, \mktab, \mmemo}) &= \tpl{(\mexpr,\menv),\mstore|_{\reaches(\mathit{root}, \mstore)}, \mkont, \mktab, \mmemo} \\
-  \text{where } \mathit{root} &= \touchesk{\mkont}(\mktab) \cup \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)})
-\end{align*}
+      \text{where } \mathit{root} &= \touchesk{\mkont}(\mktab) \cup
+      \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)})
+    \end{align*}
+  \end{minipage}
+\end{center}
 %
@@
 %
+The proofs are all similar to the original abstract GC paper's.
+%
 Technically since this semantics inspects the touched addresses of the continuation, the set of touched addresses should be considered part of the ``context'' in the $\mathbf{rt}$ continuations.
 %
-This extra context can be overlooked, since that means the computed touched addresses will be an overapproximation of what is already sound.
-%
-More touched addresses means a less effective GC, but this is an acceptable tradeoff for a possibly smaller state space.
-%
-\begin{claim}[No dangling pointers]\label{thm:gc-pointers}
+This extra context can be overlooked, but it leads to larger live address sets.
+%
+More live addresses means a less effective GC, but this is an acceptable tradeoff for a possibly smaller state space.
+%
+Indeed, the introspective pushdown analysis work chooses this more approximate method.
+%
+\iftr{%
+\begin{theorem}[No dangling pointers]\label{thm:gc-pointers}
   All addresses in $\Gamma(\mstate)$ are in the domain of $\Gamma(\mstate).\mstore$.
-\end{claim}
-
-\begin{claim}[Garbage is irrelevant]\label{thm:gc-concrete}
+\end{theorem}
+
+\begin{theorem}[Garbage is irrelevant]\label{thm:gc-concrete}
   If $\mstate \stepto^n \mstate'$ and $\mstate \stepto^n \mstate''$ then $\Gamma(\mstate') = \Gamma(\mstate'')$.
-\end{claim}
-
-\begin{claim}[Abstract GC is sound for garbage-free states]\label{thm:gc-sound}
+\end{theorem}
+
+\begin{theorem}[Abstract GC is sound for garbage-free states]\label{thm:gc-sound}
   If $\mstate \stepto \mstate'$ and $\alpha(\Gamma(\mstate)) \sqsubseteq \mastate$ then there is a $\mastate'$ such that $\mastate \astepto \mastate'$ and $\alpha(\Gamma(\mstate')) \sqsubseteq \mastate'$.
-\end{claim}
+\end{theorem}
 
@@
 \end{mathpar}
+}
 \section{Deriving CFA2}
@@
 
-CFA2 is the first published analysis of a higher-order programming language that could properly match calls and returns.
-%
-We will show that it fits well into the same presentation we gave for PDCFA.
+CFA2 is the first published analysis of an untyped higher-order programming language that could properly match calls and returns.
 %
@@
     &
-    $\tpl{\mval,\mstore,\mframe',\mkont}$ if $(\mframe', \mval) \in \lookup(\mstore,\mframe,\menv,\mvar,\mlab)$
+    $\tpl{\mval,\mstore,\mframe',\mkont}$ if $(\mframe', \mval) \in \lookup(\mstore,\mframe,\menv(\mvar),\mlab)$
     \\
@@
 %
-Their criteria for a binding never escaping is that it is never referenced in a function that is not its binder.
-%
-This can be extended in a language with more linguistic features; see Kranz's thesis about register-allocatable bindings in the Orbit Scheme compiler \citep{ianjohnson:kranz:thesis:1988}.
-%
-A stackable reference is one that appears in the body of the binding function, by which we mean not within the body of a nested function.
+Their criteria for a binding never escaping is like the criteria of register-allocatable bindings in the Orbit Scheme compiler~\citep{ianjohnson:kranz:thesis:1988}.
 %
@@
           \end{array}\right. \\
-  \lookup(\mstore,\mframe,\menv,\mvar,\mlab) &=
+  \lookup(\mstore,\mframe,\maddr,\mlab) &=
     \left\{\begin{array}{ll}
-          \setbuild{(\extm{\mframe}{\menv(\mvar)}{\set{\mval}}, \mval)}
-                   {\mval \in \mframe(\menv(\mvar))} & \text{if } \mlab \text{ non-escaping} \\
-          \setbuild{(\mframe,\mval)}{\mval \in \mstore(\menv(\mvar))} & \text{ otherwise}
+          \setbuild{(\extm{\mframe}{\maddr}{\set{\mval}}, \mval)}
+                   {\mval \in \mframe(\maddr)} & \text{if } \mlab \text{ non-escaping} \\
+          \setbuild{(\mframe,\mval)}{\mval \in \mstore(\maddr)} & \text{ otherwise}
            \end{array}\right.
@@
 %
-Tail calls are important in language implementations for space complexity reasons \citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these concerns are less important.
+Tail calls are important in language implementations for space complexity reasons \citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these concerns are less important since there is no unbounded recursion.
 %
@@
 We follow up with a similar proof to that of PDCFA, which is that the CES$\xi$K machine has a bisimulation with its tabular semantics, with comparable allocation strategies.
-\begin{claim}[Correctness of stack frame refinement]\label{thm:refinement}
+\begin{theorem}[Correctness of stack frame refinement]\label{thm:refinement}
   The CES$\xi$K machine has a bisimulation with the CESK machine with fresh allocation.
-\end{claim}
+\end{theorem}
 This follows from the invariant that if a stack frame maps $\maddr$ to $\mval$, then the store maps $\maddr$ to $\mval$, meaning that the special lookup function is equivalent to lookup directly in the store (exploits immutability).
 %
-That is, if we always allocated bindings in the heap.
+%That is, if we always allocated bindings in the heap.
 %
@@
 %
-The additional escape analysis adds information to use in the proof, that if $\mvar$ never escapes, then every reference to allocations of $\mvar$'s binding will always be able to use $\mframe$ for the lookup.
+The escape analysis adds information to use in the proof, that if $\mvar$ never escapes, then every reference to allocations of $\mvar$'s binding will always be able to use $\mframe$ for the lookup.
 %
@@
 
-\begin{claim}[Correctness of CFA2]\label{thm:cfa2}
+\begin{theorem}[Correctness of CFA2]\label{thm:cfa2}
   The widened tabular semantics reifies a relation that has a bisimulation with standard semantics.
-\end{claim}
+\end{theorem}
 This proof has the same structure as that of PDCFA.
@@
 %
-Stack replacement is easily modeled in a regular analysis using the AAM approach, and Vardoulakis and Shivers showed it can be done in a pushdown approach (although it breaks the pushdown model).
+Stack replacement is easily modeled in a regular analysis using the AAM approach, and Vardoulakis and Shivers showed it can be done in a pushdown approach.
 %
@@
 
+\begin{center}
+  \begin{minipage}{0.45\linewidth}
     \begin{align*}
       \mstate \in \State &::= \tpl{\mpoint, \mstore, \mkont, \mmkont} \\
-  \mpoint \in \Point &::= (\mexpr, \menv) \alt \mval \\
+      \mpoint \in \Point &::= (\mexpr, \menv) \alt \mval
+    \end{align*}
+  \end{minipage}
+  \begin{minipage}{0.50\linewidth}
+    \begin{align*}
       \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt \vcomp{\mkont} \\
@@
     \end{align*}
-
+  \end{minipage}
+\end{center}
 \begin{figure}
@@
     $\tpl{(\mexpr, \menv), \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
+    &
+    [push prompt]
     \\
@@
     $\tpl{\mval, \mstore, {\mkont}, {\mmkont}}$
+    &
+    [pop prompt]
     \\
@@
     $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore',\kmt,\mmkont}$
+    &
+    [capture continuation]
     \\ & where $\mstore' = \joinone{\mstore}{\maddr}{\vcomp{\mkont}}$
@@
     $\tpl{\mval, \mstore, \mkont', \mkapp{\mkont}{\mmkont}}$
+    &
+    [compose continuation]
   \end{tabular}  
@@
 %
-This is a fairly brutal approximation of a captured continuation, so an alternative is to make this tunable with our precision-tuning friend, $\alloc$.
-%
-Then, instead of entirely removing the store component of the $\rt$ context, we can replace it with an address of possible stores that it would approximate.
+An alternative, more tunable approximation, uses our precision-tuning friend, $\alloc$.
+%
+Instead of entirely removing the store component of the $\rt$ context, we can replace it with an address of possible stores that it would approximate.
 
@@
 
-{
-\setlength{\abovedisplayskip}{0pt}
-\setlength{\belowdisplayskip}{4pt}
-\setlength{\abovedisplayshortskip}{0pt}
-\setlength{\belowdisplayshortskip}{8pt}
+\begin{center}
+  \begin{minipage}{0.50\linewidth}
     \begin{align*}
       \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
-  \mctx \in \Context &::= \msctx \alt ((\mexpr,\menv),\mstore,\mmktab) \alt (\vcomp{\mskont}, \mval, \mstore, \mmktab) \\
+      \mctx \in \Context &::= \msctx \alt
+      ((\mexpr,\menv),\mstore,\mmktab) \\ &\alt (\vcomp{\mskont},
+      \mval, \mstore, \mmktab)
+    \end{align*}
+  \end{minipage}
+  \begin{minipage}{0.45\linewidth}
+    \begin{align*}
       \mskont \in \SKont &= \kmt \alt \kcons{\mkframe}{\mskont} \alt \krt{\msctx} \\
-  \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt \vcomp{\mskont}
-  \end{align*}
-}
+      \mval \in \Value &= \vclo{\slam{\mvar}{\mexpr}}{\menv} \alt
+      \vcomp{\mskont}
+    \end{align*}
+  \end{minipage}
+\end{center}
 
@@
 
-\begin{claim}\label{thm:concrete-sr}
+\iftr{%
+\begin{theorem}\label{thm:concrete-sr}
   Tabular semantics simulates the standard semantics given fresh allocation.
-\end{claim}
-
-We prove this with the invariant that the continuation and meta-continuation have unique unrollings in addition to the table invariants from \autoref{sec:pdcfa}, since the distinct $\mstore$ at push and capture time makes entries in $\mktab$ unique, regardless of $\mmktab$.
-
-\begin{claim}\label{thm:sound-sr}
-  Widened tabular semantics reifies a relation that simulates standard semantics.
-\end{claim}
-
+\end{theorem}
+\begin{proof}[Sketch]
+  We have the invariant that the continuation and meta-continuation have unique unrollings in addition to the table invariants from \autoref{sec:pdcfa},
+  since the distinct $\mstore$ at push and capture time makes entries in $\mktab$ unique, regardless of $\mmktab$.
+\end{proof}
+}
+\begin{theorem}\label{thm:sound-sr}
+  Widened tabular semantics reflects a relation that simulates standard semantics.
+\end{theorem}
+\iftr{%
+\begin{proof}[Sketch]
+  This follows the same line of reasoning as the correctness of PDCFA.
+\end{proof}
+}
 \paragraph{Comparison to PDCFA CPS'd to remove {\tt shift} and {\tt reset}:}{
@@
 %
-The higher-order setting imposes additional challenges that make their methods difficult to adapt.
-%
-The most important constraint is that we can't know all call-sites of a function or method before the analysis begins, which their methods heavily rely on.
+The important difference when we move higher-order is that the model construction to feed these methods is an additional problem --- the one we solve here.
 %%
@@
 %
-The expressiveness of CPDAs outweighs that of PDAs, but it is unclear how to adapt higher-order recursion schemes to model arbitrary programming language features.
+The expressiveness of CPDAs outweighs that of PDAs, but it is unclear how to adapt higher-order recursion schemes (HORS) to model arbitrary programming language features.
 %
 The method is strongly tied to the simply-typed call-by-name lambda calculus and depends on finite sized base-types.
+%
+HORSs are a powerful abstraction target for which monadic second order logic (MSO) is decidable, but systematic constructions of models in HORS from arbitrary programming languages are still out of reach.
 
@@
 
-As the programming world continues to embrace behavioral values, it becomes more important to import the powerful techniques pioneered by the first-order analysis and model checking communities.
+As the programming world continues to embrace behavioral values, it becomes more important to import the powerful techniques pioneered by the first-order analysis and model-checking communities.
 %
@@
 %
-We believe that the recipe that this paper set forth is a step in that direction.
-%
-That is, make continuation tables keyed with enough context, and memoize at the introduced indirection points.
-%
-The result in a language with well-bracketed control is a ``pushdown analysis'' using summarization.
-%
-In a language without well-bracketed control, we are not chained to a pushdown automaton as the target of the approximation, so the techniques still apply and give better precision than regular methods.
+We showed how to systematically construct executable concrete semantics that pointwise abstract to known algorithms for pushdown analyses of higher-order languages.
+%
+We bypass the automata theoretic approach so that we are not chained to a pushdown automaton to model features such as first-class composable control operators.
+%
+The techniques still apply and give better precision than regular methods.
 %%
@@
 %
-The control structures there are difficult to model even with AAM's techniques due to the ability to capture and compose continuations with arbitrarily many prompts, but once we can tackle AAM, there should be a straightforward route to extending it with the pushdown techniques of this paper.
-%
 We have an unproven prototype of the stack inspection capabilities provided by the {\tt with-continuation-mark}, {\tt current-continuation-marks} and {\tt continuation-mark-set->list} primitives, where a ``context'' then also includes what is the result of {\tt current-continuation-marks}, since functions can inspect only this amount of the stack.
@@
 %
-We have promising preliminary results in this direction that don't require a pre-analysis, but we leave further investigation of sparse techniques to future work.
-
-%%
-There is also the question of novel clients of these analyses.
-%
-Most obviously we would want to know where we can implement first-class control more efficiently.
-%
-In particular, we would want to have an escape analysis to find first-class continuations that don't need to be heap-allocated, and single target abort operations that can be turned into long jumps after a computed stack unraveling.
-%
-On top of optimizations, there are security analyses.
-%
-Greater control of the stack gives us the ability to drill deep into context-sensitive security properties and make precise predictions.
+Alternatively, there is an analogous concept in concrete semantics called ``self-adjusting computation'' that has been used for higher-order programs that looks like the right starting point for systematically constructing sparse analyses.
+%
+%% XXX: Don't show this
+%We have promising preliminary results in this direction that don't require a pre-analysis, but we leave further investigation of sparse techniques to future work.
+
+%%
+% There is also the question of novel clients of these analyses.
+% %
+% Most obviously we would want to know where we can implement first-class control more efficiently.
+% %
+% In particular, we would want to have an escape analysis to find first-class continuations that don't need to be heap-allocated, and single target abort operations that can be turned into long jumps after a computed stack unraveling.
+% %
+% On top of optimizations, there are security analyses.
+% %
+% Greater control of the stack gives us the ability to drill deep into context-sensitive security properties and make precise predictions.

### 2013-12-09 content.tex
Shortened some wording. Added cite to self-adjusting computation. Might want to kill that part.

@@
 
-\paragraph{Tracking return points:} the magic of the method is in keeping a table of continuation segments for each function and store pair, where the continuation is segmented at function boundaries.
+\paragraph{Tracking return points:} the crux of the method is the table of continuations for each function and store pair.
 %
@@
 %
-Instead of deferring to the table (store) for the remainder of a continuation for each frame, we only have indirections at function call boundaries (see last 2 rules in figure \ref{fig:summary-semantics} for saving and restoring continuations).
-%
-Note that since AAM is about finitizing the state space, this step itself fits within the AAM method, since continuations within functions truncated at call boundaries are bounded by the nesting depth of those functions, thereby making the continuation space finite.
-%
-In the monovariant case, the number of continuations is still linear in the size of the program.
+The primary differences are that we separate the table from the store, and continuations do not indirect through the table for each continuation frame\footnote{They could, but it is unnecessary and wasteful}.
 %%
@@
 %
-It instead appeals to an external, imperative algorithm for summarization to sew the function calls and returns together by pattern matching on the states that the local semantics produces.
+Returns are un-stuck by an imperative summarization algorithm driving the state exploration.
 %
@@
 %
-The other rules simply carry along the extra $\mframe$ component untouched.
+%The other rules simply carry along the extra $\mframe$ component untouched.
 %
@@
 %
-Indeed, as soon as the non-determinism has been determined, we could extend the stack frame so any subsequent reference means what it meant previously in the function (reminiscent of how equality information of unknowns is tracked in \citet{ianjohnson:DBLP:journals/cacm/DilligDA10}).
-%
-Look-up would then always try the stack frame before falling back on the heap (of course this must be limited to immutable addresses).
+As soon as the non-determinism has been determined, we could extend the stack frame so any subsequent reference means what it meant previously in the function\footnote{This is reminiscent of how equality of unknowns is tracked in \citet{ianjohnson:DBLP:journals/cacm/DilligDA10}}.
+%
+Look-up would then always try the stack frame before falling back on the heap (this must be limited to immutable addresses).
 %%
@@
 
-\paragraph{Possible fixes to non-termination:} The simplest fix to the circularity would be to strip stores out of the context of the $\rt$ in a captured continuation, and additionally consider contexts with a stripped store an abstraction of the same context with any store.
-%
-An alternative, more tunable approximation, uses our precision-tuning friend, $\alloc$.
-%
-Instead of entirely removing the store component of the $\rt$ context, we can replace it with an address of possible stores that it would approximate.
+\paragraph{A fix to non-termination:} we allow for a tunable approximation by using $\alloc$.
+%
+Captured continuations' store component in the $\rt$ context gets replaced with an address of possible stores that it would approximate.
 
@@
 %
-Alternatively, there is an analogous concept in concrete semantics called ``self-adjusting computation'' that has been used for higher-order programs that looks like the right starting point for systematically constructing sparse analyses.
+Alternatively, there is an analogous concept in concrete semantics called ``self-adjusting computation''~\citep{ianjohnson:DBLP:conf/afp/AcarL08} that has been used for higher-order programs that looks like the right starting point for systematically constructing sparse analyses.
 %

### 2013-12-11 content.tex
Trimmed more verbage. Adding supplemental document, though still incomplete.

@@
 %%
-This paper gives a common, simple framework to derive CFA2, PDCFA and its garbage-collecting extension using a recipe to apply pushdown analysis techniques to arbitrary semantics in a manner similar to AAM.
-%
-To exercise the recipe further, we give a novel analysis for delimited, composable control. \\
+This paper gives a common, simple framework to derive CFA2, PDCFA and its garbage-collecting extension in a manner similar to AAM.
+%
+To exercise the framework further, we give a novel analysis for delimited, composable control. \\
 
@@
 
-CFA2 is an online version of a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the $\epsilon$-closure graph that PDCFA constructs.
+CFA2 is an online version of a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with PDCFA's crucial ``$\epsilon$-closure graph''.
 %
@@
 %
-This was shown for the \rackett{call-with-current-continuation} (or \rackett{call/cc}) operator in~\citet{ianjohnson:Vardoulakis2011Pushdown}.
+This was shown for the \rackett{call/cc} operator in~\citet{ianjohnson:Vardoulakis2011Pushdown}.
 %
@@
 For a concrete semantics, we require the allocation meta-function $\alloc :
-\State \to \Addr$ to return fresh addresses ($\alloc(\mstate) \notin \dom(\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
+\State \to \Addr$ to return fresh addresses ($\alloc(\tpl{(\mexpr,\menv),\mstore,\mkont}) \notin \dom(\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
 %
@@
 %
-Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store [memo lookup].
+Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store \textsc{[memo lookup]}.
 %%
@@
     &
-    [variable lookup]
+    \textsc{[variable lookup]}
     \\
@@
     &
-    [function eval]
+    \textsc{[function eval]}
     \\
@@
     &
-    [argument eval]
+    \textsc{[argument eval]}
     \\
@@
     &
-    [un-memoized call] 
+    \textsc{[un-memoized call]}
      \\ & \quad if $(\mpoint, \mstore') \notin \dom(\mmemo)$
@@
             \mmemo}$
-    & [memo lookup]
+    & \textsc{[memo lookup]}
     \\ & \quad if $(\mval_\mathit{result},\mstore'') \in \mmemo(\mpoint,\mstore')$
@@
     &
-    [memoize/return]
+    \textsc{[memoize/return]}
     \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
@@
 %
-We thus explore all paths and share the KTable and memo tables between all paths.
-
-To turn this semantics into PDCFA's algorithm for constructing a pushdown system with only reachable states, given the appropriate $\alloc$, we apply a widening operator to make $\mkont$, and $\mmemo$ shared amongst all states (meaning states then become represented without these components), in what is then called a $\System$.
+We thus explore all paths and share $\mktab$ and $\mmemo$ between all paths.
+
+To turn this semantics into PDCFA's algorithm for constructing a pushdown system with only reachable states, given the appropriate $\alloc$, we apply a widening operator to make $\mktab$, and $\mmemo$ shared amongst all states, in what is then called a $\System$.
 We use a meta-function $\wn$, ``wide to narrow,'' to shift between the two representations of states.
@@
 
-\iftr{%
-\subsection{Metatheory}
-
-Correctness follows directly from the invariants we prove of $\mktab$ and $\mmemo$ in the tabular semantics.
-%
-Allocation strategies are comparable if they produce equal addresses regardless of the differences in the state representations.
-%
-More formally, $\alloc$ and $\alloc^*$ are comparable if this implication holds:
-\begin{mathpar}
-  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \emptyset}}
-            {\alloc(\tpl{\mpoint,\mstore,\mkont}) = \alloc^*(\tpl{\mpoint,\mstore,\mkont',\mktab,\mmemo})}
-\end{mathpar}
-
-The $\mathit{unroll}$ function interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, defined as the greatest fixed point of the following rules:
-\begin{mathpar}
-  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt, G}} \quad
-  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \emptyset}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont', G}} \\
-  \inferrule{\mctx \notin G  \\
-             \mkont'' \in \mktab(\mctx) \\
-             \mkont \in \unroll{\mktab}{\mkont'', \set{\mctx}\cup G}}
-            {\mkont \in \unroll{\mktab}{\krt{\mctx}, G}}  
-\end{mathpar}
-
-We add $G$ to protect against unguarded corecursion in order for $\mathit{unroll}$ to be well-defined.
-%
-Interpreting a function with ill-founded recursion can lead to a table such that $\krt{\mctx} \in \mktab(\mctx)$.
-%
-The abstractions that we impose on our data might also cause this situation, even though the concrete execution might always terminate.
-
-%%
-The different tables we have encode information about execution history that we prove is invariant (\autoref{fig:inv}).
-%
-The first is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
-%
-The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
-%
-The final invariant, $\phi_{\memo}$, states that all paths starting at a function call either reach the memoized result, or if the path does not include a return, there is an extension that will.
-%
-The portion of these paths between call and return hal the calling context's continuation in the tail to justify using stack-irrelevance.
-
-\begin{figure}
-  \centering
-  \begin{align*}
-    \inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo}) &=
-    \dom(\mmemo) \subseteq \dom(\mktab) \\
-    &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab), \mkont''
-    \in \mktab(\mpoint',\mstore'),
-    \mkont' \in \unroll{\mktab}{\mkont'', \emptyset}. \\
-    &\qquad \phi_{\reachable} \wedge \phi_{\memo} \\
-    \text{where } \phi_{\reachable} &= \exists \mkont'' \in
-    \unroll{\mktab}{\mkont,\emptyset}.
-    \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont''} \\
-    \phi_{\memo} &=
-    \forall (\mval, \mstore'') \in \mmemo(\mpoint',\mstore'). \\
-    &\qquad\exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore'',\mkont'}. \hastail(\mtrace,\mkont')
-  \end{align*}
-  \caption{Table invariants}
-\label{fig:inv}
-\end{figure}
-\begin{lemma}[Table invariants in CESK$\mathit{\Xi{}M}$]\label{lem:tab-inv}
-  If $\inv(\mstate)$ and $\mstate \stepto \mstate'$ then $\inv(\mstate')$.
-\end{lemma}
-
-In order to prove this, we need a lemma that allows us to plug in any continuation for memoized results.
-%
-When traces (sequences of states such that adjacent states are related by $\stepto$) have a common tail, we can replace it with anything.
-
-\begin{lemma}[Stack irrelevance in CESK]\label{lem:irrelevance}
-  $\hastail(\mtrace,\mkont)$ implies $\replacetail(\mtrace,\mkont,\mkont')$ is a valid trace.
-\end{lemma}
-\begin{mathpar}
-  \inferrule{ }{\hastail(\epsilon,\mkont)} \quad
-  \inferrule{ }{\hastail(\tpl{\mpoint,\mstore,\append{\mkont'}{\mkont}},\mkont)} \quad
-  \inferrule{\hastail(\mtrace\mstate,\mkont) \quad
-             \mstate \stepto \mstate' \quad
-             \hastail(\mstate',\mkont)}
-            {\hastail(\mtrace\mstate\mstate',\mkont)}
-\end{mathpar}
-
-\begin{align*}
-  \replacetail(\tpl{\mpoint,\mstore,\append{\mkont'}{\mkont}},\mkont,\mkont'') &= \tpl{\mpoint,\mstore,\append{\mkont'}{\mkont''}} \\
-  \replacetail(\epsilon,\mkont,\mkont'') &= \epsilon \\
-  \replacetail(\mtrace\mstate,\mkont,\mkont') &= \replacetail(\mtrace,\mkont,\mkont')\replacetail(\mstate,\mkont,\mkont') \\
-\end{align*}
-
-\begin{theorem}[Correctness of Concrete Summarization]\label{thm:concrete-tabular}
-  The standard semantics has a skipping bisimulation with the tabular semantics given comparable fresh allocation strategies.
-\end{theorem}
-\begin{proof}[Sketch]
-  The tabular semantics skipping simulates that standard semantics by appealing to the table invariants to show the long ``recomputation'' path.
-  Crucially, since the memo table is only set once before being used to bypass to the result, there must be only one possible result in the standard semantics.
-Otherwise, the tabular semantics would miss out on the other results for subsequent calls with the same context.
-  The fresh allocation strategy ensures that relation is deterministic and thus function calls only reduce to one result.
-
-  The standard semantics skipping simulates the tabular semantics because ``recomputing'' paths always have the calling stack as its tail, so we can step forward on the invariant's inductively built path until the result.
-\end{proof}
-
-Finally, we can show that the reductions we find between states paired with the shared tables define a sound and complete abstraction of the CESK machine with a comparable allocation strategy.
-}
+\iftr{\input{metatheory}}
+
 \begin{theorem}[Correctness of PDCFA]\label{thm:global}
@@
 \end{theorem}
-\iftr{%
-\begin{proof}[Sketch]
-  The invariants on the tables still hold. 
-  The widening to share $\mmemo$ and $\mktab$ amongst all explored states reifies the search state into a $\sa{System}$. We can add a set of edges simply.
-  If we add $\mmemo$ and $\mktab$ to all states in these edges, we get a reduction relation comparable to the unwidened tabular semantics.
-  We prove that the infinite unrolling of the standard reduction relation stuttering bisimulates the relation reified from an infinite stepping of the widened tabular semantics.
-  This only holds if the state selection from the frontier is fair, since unfair choices can leave some function calls unresolved.
-  Formally,
-  \begin{equation*}
-    \setbuild{\mtrace}{\mtrace : i \le \omega, \mstate_0 \longmapsto^i \mstate} \simeq {\mathcal F}(\mstate_0)^\omega
-  \end{equation*}
-  where $\simeq$ is TODO
-by coinduction and appeals to the table invariants.
-  If a new function result is found after a call already used the memo table, the context of the memo user is used after memoizing, so it looks like that result was part of the original memo table lookup.
-  The fair selection guarantees that all function results will be eventually discovered, so the memo table appeals eventually catch up to the found results.
-\end{proof}
-}
 
@@
 
-Garbage collection ($\Gamma$) amounts to restricting the heap to just the reachable addresses, so a garbage-collecting semantics introduces the following reduction rule:
+Garbage collection ($\Gamma$) restricts the heap to just the reachable addresses, so a semantics with GC introduces the following reduction rule:
 \begin{center}
@@
 %
-More live addresses means a less effective GC, but this is an acceptable tradeoff for a possibly smaller state space.
-%
-Indeed, the introspective pushdown analysis work chooses this more approximate method.
-%
-\iftr{%
-\begin{theorem}[No dangling pointers]\label{thm:gc-pointers}
-  All addresses in $\Gamma(\mstate)$ are in the domain of $\Gamma(\mstate).\mstore$.
-\end{theorem}
-
-\begin{theorem}[Garbage is irrelevant]\label{thm:gc-concrete}
-  If $\mstate \stepto^n \mstate'$ and $\mstate \stepto^n \mstate''$ then $\Gamma(\mstate') = \Gamma(\mstate'')$.
-\end{theorem}
-
-\begin{theorem}[Abstract GC is sound for garbage-free states]\label{thm:gc-sound}
-  If $\mstate \stepto \mstate'$ and $\alpha(\Gamma(\mstate)) \sqsubseteq \mastate$ then there is a $\mastate'$ such that $\mastate \astepto \mastate'$ and $\alpha(\Gamma(\mstate')) \sqsubseteq \mastate'$.
-\end{theorem}
-
-This final proposition is not in the original work on abstract GC; they prove soundness with respect to a concrete semantics that performs GC after every step.
-%
-The moral correctness of the abstraction is that it never under-approximates \emph{relevant} portions of the state space --- in this case, the reachable subset of the store.
-%
-The definitions of $\alpha$, $\astepto$ and $\sqsubseteq$ are the simple pointwise abstractions of addresses through states and the reduction relation, and finally a lifted subset relation between states and their components, respectively.
-%
-Both $\alpha$ and $\sqsubseteq$ are parameterized on the choice of $\Addr$, for which the following soundness criterion should hold:
-\begin{mathpar}
-  \inferrule*[right={[Sound allocation]}]{\alpha(\mstate) \sqsubseteq \mastate}{\alpha(\alloc(\mstate)) \sqsubseteq \widehat{\alloc}(\mastate)}
-\end{mathpar}
-}
+More live addresses means a less effective GC, but the introspective pushdown analysis chose this tradeoff for a possibly smaller state space.
+%
+
 \section{Deriving CFA2}
@@
 There are two orthogonal features of the semantics, embodied by our semantics' two metafunctions ($\bind$ and $\lookup$, respectively in \autoref{fig:frame-semantics}).
-%
-The first of these is an addition to the stack-less context.
 
@@
 \end{theorem}
-This proof has the same structure as that of PDCFA.
+%This proof has the same structure as that of PDCFA.
 
@@
     &
-    [push prompt]
+    \textsc{[push prompt]}
     \\
@@
     &
-    [pop prompt]
+    \textsc{[pop prompt]}
     \\
@@
     &
-    [capture continuation]
+    \textsc{[capture continuation]}
     \\ & where $\mstore' = \joinone{\mstore}{\maddr}{\vcomp{\mkont}}$
@@
     &
-    [compose continuation]
+    \textsc{[compose continuation]}
   \end{tabular}  
@@
 %
-Unfortunately, since $\rt$ continuations contain a store, and continuations can now appear in the store, this introduces a circularity that could cause the analysis to never terminate.
+Unfortunately, since $\rt$ continuations contain a store, and continuations can now appear in the store, this introduces an unbounded data structure that could cause the analysis to never terminate.
 %
@@
 
-\newcommand{\replacectx}{\mathit{replace}\mstore}
-\newcommand{\addstore}{\mathit{add}\mstore}
-\begin{align*}
-  \approximate(\mkont, \mmktab, \maddr) &= (\replacectx(\mkont), \addstore(\mkont)) \\
-  \text{where }
+\newcommand{\replacectx}{\hat\mkont} %\mathit{replace}\mstore
+\newcommand{\addstore}{\hat\mmktab} % \mathit{add}\mstore
+\begin{center}
+  $\approximate(\mkont, \mmktab, \maddr) = (\replacectx(\mkont), \addstore(\mkont)) \text{ where}$\\
+  \begin{minipage}{0.45\linewidth}
+    \begin{align*}
       \replacectx(\kmt) &= \kmt \\
       \replacectx(\krt{((\mexpr,\menv),\ldots)}) &= \krt{((\mexpr,\menv),\maddr)} \\
-   \replacectx(\kcons{\mkframe}{\mkont}) &= \kcons{\mkframe}{\replacectx(\mkont)}
-  \\[2pt]
+      \replacectx(\kcons{\mkframe}{\mkont}) &=
+      \kcons{\mkframe}{\replacectx(\mkont)}
+    \end{align*}
+  \end{minipage}
+  \begin{minipage}{0.50\linewidth}
+    \begin{align*}
       \addstore(\kmt) &= \mmktab \\
@@
     \end{align*}
+  \end{minipage}
+\end{center}
 %
@@
 
-\iftr{%
-\begin{theorem}\label{thm:concrete-sr}
-  Tabular semantics simulates the standard semantics given fresh allocation.
-\end{theorem}
-\begin{proof}[Sketch]
-  We have the invariant that the continuation and meta-continuation have unique unrollings in addition to the table invariants from \autoref{sec:pdcfa},
-  since the distinct $\mstore$ at push and capture time makes entries in $\mktab$ unique, regardless of $\mmktab$.
-\end{proof}
-}
 \begin{theorem}\label{thm:sound-sr}
@@
 \end{theorem}
-\iftr{%
-\begin{proof}[Sketch]
-  This follows the same line of reasoning as the correctness of PDCFA.
-\end{proof}
-}
+
 \paragraph{Comparison to PDCFA CPS'd to remove {\tt shift} and {\tt reset}:}{
@@
 %
-In pushdown analysis, we track function return points with entire calling contexts rather than the $k$-CFA way of allocating the continuation in the heap at some lower-precision address, like a variable.
-%
-In our summarizing analysis of delimited composable control, we get the same added precision for {\tt reset}, {\tt shift}, and continuation invocation points.
+%In pushdown analysis, we track function return points with entire calling contexts rather than the $k$-CFA way of allocating the continuation in the heap at some lower-precision address, like a variable.
+%
+%In our summarizing analysis of delimited composable control, we get added precision for {\tt reset}, {\tt shift}, and continuation invocation points.
 %
@@
 %
-Our specialized analysis for delimited control, also with a monovariant allocation strategy, will predict the non-CPS'd version returns true.}
+Our specialized analysis for delimited control, will predict the non-CPS'd version returns true.}
 
@@
 %
-HORSs are a powerful abstraction target for which monadic second order logic (MSO) is decidable, but systematic constructions of models in HORS from arbitrary programming languages are still out of reach.
+%HORSs are a powerful abstraction target for which monadic second order logic (MSO) is decidable, but systematic constructions of models in HORS from arbitrary programming languages are still out of reach.
 
@@
 %
-We conjecture that the same recipe will apply to the most intricate control operators in production languages such as in \citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}.
-%
-We have an unproven prototype of the stack inspection capabilities provided by the {\tt with-continuation-mark}, {\tt current-continuation-marks} and {\tt continuation-mark-set->list} primitives, where a ``context'' then also includes what is the result of {\tt current-continuation-marks}, since functions can inspect only this amount of the stack.
+We conjecture that the same framework will apply to the most intricate control operators in production languages such as in \citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}.
+%
+\iftr{
+We have an unproven prototype of the stack inspection capabilities provided by the {\tt with-continuation-mark}, {\tt current-continuation-marks} and {\tt continuation-mark-set->list} primitives, where a ``context'' then also includes what is the result of {\tt current-continuation-marks}, since functions can inspect only this amount of the stack.}
 %%

### 2013-12-11 metatheory.tex
Trimmed more verbage. Adding supplemental document, though still incomplete.

\subsection{Metatheory}

Correctness follows directly from the invariants we prove of $\mktab$ and $\mmemo$ in the tabular semantics.
%
Allocation strategies are comparable if they produce equal addresses regardless of the differences in the state representations.
%
More formally, $\alloc$ and $\alloc^*$ are comparable if this implication holds:
\begin{mathpar}
  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \emptyset}}
            {\alloc(\tpl{\mpoint,\mstore,\mkont}) = \alloc^*(\tpl{\mpoint,\mstore,\mkont',\mktab,\mmemo})}
\end{mathpar}

The $\mathit{unroll}$ function interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, defined as the greatest fixed point of the following rules:
\begin{mathpar}
  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt, G}} \quad
  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \emptyset}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont', G}} \\
  \inferrule{\mctx \notin G  \\
             \mkont'' \in \mktab(\mctx) \\
             \mkont \in \unroll{\mktab}{\mkont'', \set{\mctx}\cup G}}
            {\mkont \in \unroll{\mktab}{\krt{\mctx}, G}}  
\end{mathpar}

We add $G$ to protect against unguarded corecursion in order for $\mathit{unroll}$ to be well-defined.
%
Interpreting a function with ill-founded recursion can lead to a table such that $\krt{\mctx} \in \mktab(\mctx)$.
%
The abstractions that we impose on our data might also cause this situation, even though the concrete execution might always terminate.

%%
The different tables we have encode information about execution history that we prove is invariant (\autoref{fig:inv}).
%
The first is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
%
The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
%
The final invariant, $\phi_{\memo}$, states that all paths starting at a function call either reach the memoized result, or if the path does not include a return, there is an extension that will.
%
The portion of these paths between call and return hal the calling context's continuation in the tail to justify using stack-irrelevance.

\begin{figure}
  \centering
  \begin{align*}
    \inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo}) &=
    \dom(\mmemo) \subseteq \dom(\mktab) \\
    &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab), \mkont''
    \in \mktab(\mpoint',\mstore'),
    \mkont' \in \unroll{\mktab}{\mkont'', \emptyset}. \\
    &\qquad \phi_{\reachable} \wedge \phi_{\memo} \\
    \text{where } \phi_{\reachable} &= \exists \mkont'' \in
    \unroll{\mktab}{\mkont,\emptyset}.
    \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont''} \\
    \phi_{\memo} &=
    \forall (\mval, \mstore'') \in \mmemo(\mpoint',\mstore'). \\
    &\qquad\exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore'',\mkont'}. \hastail(\mtrace,\mkont')
  \end{align*}
  \caption{Table invariants}
\label{fig:inv}
\end{figure}
\begin{lemma}[Table invariants in CESK$\mathit{\Xi{}M}$]\label{lem:tab-inv}
  If $\inv(\mstate)$ and $\mstate \stepto \mstate'$ then $\inv(\mstate')$.
\end{lemma}
\begin{proof}
  By cases on $\stepto$.
  \begin{byCases}
% variable lookup
    \case{\tpl{(\svar\mvar, \menv), \mstore, \mkont, \mktab, \mmemo}
          \stepto
          \tpl{\mval,\mstore,\mkont, \mktab, \mmemo} \text{ if }\mval \in \mstore(\menv(\mvar))}{
     Same continuation and tables, so holds by $\inv(\mstate)$.}
% application
    \case{\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont, \mktab, \mmemo}
          \stepto
          \tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}}{
    Let $\mctx, \mkont^\circ, \mkont'$ be arbitrary.
    Let $\mkont^*$ be the witness for $\mstate$.
    The next witness is $\kcons{\kar{\mexpri1,\menv}}{\mkont^*}$, by definition of $\mathit{unroll}$.}
% argument evaluation
    \case{\tpl{\mval, \mstore, \kcons{\kar{\mexpr,\menv}}{\mkont}, \mktab, \mmemo}
          \stepto
          \tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}}{
    Let $\mctx, \mkont^\circ, \mkont'$ be arbitrary.
    Let $\kcons{\kar{\mexpr,\menv}}{\mkont^*}$ be the witness for $\mstate$.
    The next witness is $\kcons{\kfn{\mval}}{\mkont^*}$, by definition of $\mathit{unroll}$.}
% function call
    \case{\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo}
          \stepto
          \tpl{\mpoint, \mstore', \krt{\mpoint, \mstore'}, \mktab', \mmemo}}{
     where
     $\begin{array}{l}
       \mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr}) \\
       \mstore' = \joinone{\mstore}{\maddr}{\mval} \\
       \mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{\mkont}
     \end{array}$ \\

    Let $\mctx, \mkont^\circ, \mkont'$ be arbitrary.
    Let $\kcons{\kfn{\mval}}{\mkont^*}$ be the witness for $\mstate$.
    The next witness is $\mkont^*$, by definition of $\mathit{unroll}$.}
% memo-lookup
    \case{\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo}
          \stepto
          \tpl{\mval_\mathit{result}, \mstore'', \mkont, \mktab', \mmemo}
          \text{ if } (\mval_\mathit{result},\mstore'') \in \mmemo(\mpoint,\mstore')}{
        same where clause as previous case.

    Let $\mctx, \mkont^\circ, \mkont'$ be arbitrary.
    By definition of $\inv$, and since $\tpl{\mpoint',\mstore',\mkont'} \stepto^* \tpl{\mval_{\mathit{result}}, \mstore'', \mkont'}$.
    For the memo entry to exist, there must have been at least one previous continuation in $\mktab(\mpoint',\mstore')$.
    We use $\inv$ with this continuation, combined with stack irrelevance to produce the trace involving $\mkont'$.
    }
% return
    \case{\tpl{\mval, \mstore, \krt{\mpoint,\mstore'}, \mktab, \mmemo}
          \stepto
          \tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}
          \text{ if } \mkont \in \mktab(\mpoint, \mstore')}{
    Let $\mctx, \mkont^\circ, \mkont'$ be arbitrary.
    $\mkont'$ is an acceptable witness for the first property.
    The path we construct to $\mstate'$ is sufficient for the second property.}
  \end{byCases}
\end{proof}

In order to prove this, we need a lemma that allows us to plug in any continuation for memoized results.
%
When traces (sequences of states such that adjacent states are related by $\stepto$) have a common tail, we can replace it with anything.

\begin{lemma}[Stack irrelevance in CESK]\label{lem:irrelevance}
  $\hastail(\mtrace,\mkont)$ implies $\replacetail(\mtrace,\mkont,\mkont')$ is a valid trace.
\end{lemma}
\begin{proof}
  By induction on $\mtrace$. Base cases trivially hold, and induction step follows from cases on $\mstate \stepto \mstate'$ and definition of $\replacetail$.
\end{proof}

\begin{mathpar}
  \inferrule{ }{\hastail(\epsilon,\mkont)} \quad
  \inferrule{ }{\hastail(\tpl{\mpoint,\mstore,\append{\mkont'}{\mkont}},\mkont)} \quad
  \inferrule{\hastail(\mtrace\mstate,\mkont) \quad
             \mstate \stepto \mstate' \quad
             \hastail(\mstate',\mkont)}
            {\hastail(\mtrace\mstate\mstate',\mkont)}
\end{mathpar}

\begin{align*}
  \replacetail(\tpl{\mpoint,\mstore,\append{\mkont'}{\mkont}},\mkont,\mkont'') &= \tpl{\mpoint,\mstore,\append{\mkont'}{\mkont''}} \\
  \replacetail(\epsilon,\mkont,\mkont'') &= \epsilon \\
  \replacetail(\mtrace\mstate,\mkont,\mkont') &= \replacetail(\mtrace,\mkont,\mkont')\replacetail(\mstate,\mkont,\mkont') \\
\end{align*}

\begin{theorem}[Correctness of Concrete Summarization]\label{thm:concrete-tabular}
  The standard semantics has a skipping bisimulation with the tabular semantics given comparable fresh allocation strategies.
\end{theorem}
\begin{proof}[Sketch]
  The tabular semantics skipping simulates that standard semantics by appealing to the table invariants to show the long ``recomputation'' path.
  Crucially, since the memo table is only set once before being used to bypass to the result, there must be only one possible result in the standard semantics. Otherwise, the tabular semantics would miss out on the other results for subsequent calls with the same context.
  The fresh allocation strategy ensures that relation is deterministic and thus function calls only reduce to one result.

  The standard semantics skipping simulates the tabular semantics because ``recomputing'' paths always have the calling stack as its tail, so we can step forward on the invariant's inductively built path until the result.
\end{proof}

Finally, we can show that the reductions we find between states paired with the shared tables define a sound and complete abstraction of the CESK machine with a comparable allocation strategy.

### 2013-12-11 supplemental.tex
Trimmed more verbage. Adding supplemental document, though still incomplete.

\documentclass[article]{llncs}

% preprint only
\pagestyle{plain}
\usepackage{amsmath,stmaryrd,listings,mathpartir,amssymb,graphicx}

\usepackage[usenames,dvipsnames]{color}

\usepackage{pfsteps} % for appendix

\newcommand{\Scribtexttt}[1]{{\texttt{#1}}}
\newcommand{\SColorize}[2]{\color{#1}{#2}}
\newcommand{\inColor}[2]{{\Scribtexttt{\SColorize{#1}{#2}}}}
\definecolor{PaleBlue}{rgb}{0.90,0.90,1.0}
\newcommand{\rackett}[1]{\inColor{black}{#1}}
\input{local-macros}
\usepackage[numbers,sectionbib]{natbib}
\usepackage[english]{babel}
% pdfpagescrop for preprint only
%\usepackage[hidelinks,pdfpagescrop={92 62 523 728}]{hyperref}
\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}
\usepackage{placeins,xr}

\externaldocument{paper-jfp}

\input{preamble}
\begin{document}
\title{The Essence of Summarization\\ in Pushdown Flow Analyses\\
       \textbf{Supplemental Material}}
\author{J. Ian Johnson and David Van Horn}

\section{PDCFA}
\input{metatheory}

\begin{theorem}\label{thm:concrete-sr}
  Tabular semantics simulates the standard semantics given fresh allocation.
\end{theorem}
\begin{proof}[Sketch]
  We have the invariant that the continuation and meta-continuation have unique unrollings in addition to the table invariants from paper's section 3,
  since the distinct $\mstore$ at push and capture time makes entries in $\mktab$ unique, regardless of $\mmktab$.
\end{proof}

\noindent{}Proof of paper's theorem ``Correctness of PDCFA''
\begin{proof}[Sketch]
  The invariants on the tables still hold. 
  The widening to share $\mmemo$ and $\mktab$ amongst all explored states reifies the search state into a $\sa{System}$. We can add a set of edges simply.
  If we add $\mmemo$ and $\mktab$ to all states in these edges, we get a reduction relation comparable to the unwidened tabular semantics.
  We prove that the infinite unrolling of the standard reduction relation stuttering bisimulates the relation reified from an infinite stepping of the widened tabular semantics.
  This only holds if the state selection from the frontier is fair, since unfair choices can leave some function calls unresolved.
  Formally,
  \begin{equation*}
    \setbuild{\mtrace}{\mtrace : i \le \omega, \mstate_0 \longmapsto^i \mstate} \simeq {\mathcal F}(\mstate_0)^\omega
  \end{equation*}
  where $\simeq$ is TODO
by coinduction and appeals to the table invariants.
  If a new function result is found after a call already used the memo table, the context of the memo user is used after memoizing, so it looks like that result was part of the original memo table lookup.
  The fair selection guarantees that all function results will be eventually discovered, so the memo table appeals eventually catch up to the found results.
\end{proof}

\begin{theorem}[No dangling pointers]\label{thm:gc-pointers}
  All addresses in $\Gamma(\mstate)$ are in the domain of $\Gamma(\mstate).\mstore$.
\end{theorem}

\begin{theorem}[Garbage is irrelevant]\label{thm:gc-concrete}
  If $\mstate \stepto^n \mstate'$ and $\mstate \stepto^n \mstate''$ then $\Gamma(\mstate') = \Gamma(\mstate'')$.
\end{theorem}

\begin{theorem}[Abstract GC is sound for garbage-free states]\label{thm:gc-sound}
  If $\mstate \stepto \mstate'$ and $\alpha(\Gamma(\mstate)) \sqsubseteq \mastate$ then there is a $\mastate'$ such that $\mastate \astepto \mastate'$ and $\alpha(\Gamma(\mstate')) \sqsubseteq \mastate'$.
\end{theorem}

This final proposition is not in the original work on abstract GC; they prove soundness with respect to a concrete semantics that performs GC after every step.
%
The moral correctness of the abstraction is that it never under-approximates \emph{relevant} portions of the state space --- in this case, the reachable subset of the store.
%
The definitions of $\alpha$, $\astepto$ and $\sqsubseteq$ are the simple pointwise abstractions of addresses through states and the reduction relation, and finally a lifted subset relation between states and their components, respectively.
%
Both $\alpha$ and $\sqsubseteq$ are parameterized on the choice of $\Addr$, for which the following soundness criterion should hold:
\begin{mathpar}
  \inferrule*[right={[Sound allocation]}]{\alpha(\mstate) \sqsubseteq \mastate}{\alpha(\alloc(\mstate)) \sqsubseteq \widehat{\alloc}(\mastate)}
\end{mathpar}

\section{Shift/reset}
\noindent{}Proof of paper's theorem 4 follows the same line of reasoning as the correctness of PDCFA.

% \noindent{}Proof of \autoref{thm:pdcfa-tabular}, which is more formally stated as (tabular $\pdstepto$, CESK $\stepto$): There is a well-founded equivalence bisimulation between $\pdstepto$ (quotiented to states satisfying $\inv$) and $\tred{\stepto}$.
% We define equivalence ($\simeq$) as
% \begin{mathpar}
%   \inferrule{\mkont \in \unroll{\mktab}{\mkont'} \\ \mkont \nequiv \kcons{\kfn{\mval}}{\mkont''}}
%             {\mtrace\tpl{\mpoint,\mstore,\mkont} \simeq \tpl{\mpoint,\mstore,\mkont,\mktab,\mmemo}} \\
% %
%   \inferrule{\mkont \in \unroll{\mktab}{\mkont'} \\
%              ((\mexpr,\menv[\mvar \mapsto \maddr]), \mstore\sqcup[\maddr \mapsto \set{\mval}]) \notin \dom(\mmemo)}
%             {\mtrace\tpl{\mval,\mstore,\kcons{\kfn{\slam{\mvar}{\mexpr}, \menv}{\mkont}}} \simeq \tpl{\mval,\mstore,\kcons{\kfn{\slam{\mvar}{\mexpr},\menv}}{\mkont'},\mktab,\mmemo}} \\
% %
%   \inferrule{\tpl{\mval', \mstore', \mkont} \notin \mtrace'' \\
%              \mkont \in \unroll{\mktab}{\mkont''} \\
%              ((\mexpr,\menv[\mvar \mapsto \maddr]), \mstore\sqcup[\maddr \mapsto \set{\mval}]) \in \dom(\mmemo)}
%             {\mtrace'\tpl{\mval, \mstore, \kcons{\kfn{\slam{\mvar}{\mexpr}, \menv}}{\mkont}}\mtrace''
%               \simeq \tpl{\mval, \mstore, \kcons{\kfn{\slam{\mvar}{\mexpr}, \menv}}{\mkont''}, \mktab, \mmemo}}
% \end{mathpar}
% This means that states which ``line up'' regardless of history are related, and that states that follow a function call that has not yet completed are related to the function call state where we already have the memoized results.
% %
% Each step in this case will get ``closer to'' the result to memoize, as given by the length of the path we get from the invariant on the memo table.
% \begin{proof}
%   Let $\mtrace,\mtrace',\mstate_\Xi$ be arbitrary such that $\mtrace \simeq \mstate_\Xi$ and $\mtrace \tred{\stepto} \mtrace'$.
%   By cases on $\mtrace \simeq \mstate_\Xi$:
%   \begin{byCases}
%     \case{\text{First rule}}{One step related by simple case analysis.}
%     \case{\text{Second rule}}{One step related by definitions of $\mathit{unroll}$, $\stepto$ and $\pdstepto$.}
%     \case{\text{Third rule}}{
%     By cases on $\mtrace\mstate \tred{\stepto} \mtrace\mstate\mstate'$:
%     \begin{byCases}
%       \case{\mtrace\tpl{(\svar\mvar, \menv), \mstore, \mkont} \stepto
%         \mtrace\mstate\tpl{\mval,\mstore,\mkont}}{...}
%       % application
%       \case{\mtrace\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore,
%           \mkont} \stepto \mtrace\mstate\tpl{(\mexpri0, \menv),
%           \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}}}{...}
%       % argument evaluation
%       \case{\mtrace\tpl{\mval, \mstore,
%           \kcons{\kar{\mexpr,\menv}}{\mkont}} \stepto
%         \mtrace\mstate\tpl{(\mexpr, \menv), \mstore,
%           \kcons{\kfn{\mval}}{\mkont}}}{...}
%       % function call
%       \case{\mtrace\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont}}
%         \stepto \mtrace\mstate\tpl{(\mexpr,
%           \extm{\menv}{\mvar}{\maddr}),
%           \joinone{\mstore}{\maddr}{\mval}, \mkont}}{...}
%     \end{byCases}}
%   \end{byCases}
%   By \autoref{lem:tab-inv}

%   Soundness is a corollary of \autoref{lem:tab-inv}.
% \end{proof}

% \noindent{}Proof of \autoref{thm:global}.
% \begin{proof}
% %% TODO
%   The invariants of the memo and k-tables are independent of the states that carry them, so lifting them to a global position does not affect their information.
% \end{proof}

% \noindent{}Proof of \autoref{thm:refinement}.
% \begin{proof}
%   todo
% \end{proof}

% \noindent{}Proof of \autoref{thm:cfa2}.
% \begin{proof}
%   todo
% \end{proof}

% \noindent{}Proof of \autoref{thm:concrete-sr}.
% \begin{proof}
%   todo
% \end{proof}

% \noindent{}Proof of \autoref{thm:sound-sr}
% \begin{proof}
%   todo
% \end{proof}

\end{document}

### 2013-12-13 content.tex
Shrinking.

@@
 %
-They are hence called ``pushdown analyses.''
-%
-\footnote{We will refer to the classic finite model analyses as ``regular analyses'' after regular languages.}
-%
-CFA2 and PDCFA have difficult details to easily apply to an off-the-shelf semantics --- especially if they feature non-local control transfer that breaks the pushdown model.
+They are hence called ``pushdown analyses.''%
+%
+\footnote{We refer to finite model analyses as ``regular analyses'' after regular languages.}
+%
+CFA2 and PDCFA have difficult details to easily apply to an off-the-shelf semantics---especially if they feature non-local control transfer that breaks the pushdown model.
 %%
@@
 For a concrete semantics, we require the allocation meta-function $\alloc :
-\State \to \Addr$ to return fresh addresses ($\alloc(\tpl{(\mexpr,\menv),\mstore,\mkont}) \notin \dom(\mstore)$), but any choice of address is sound~\citep{dvanhorn:Might2009Posteriori}.
+\State \to \Addr$ to return fresh addresses ($\alloc(\tpl{(\mexpr,\menv),\mstore,\mkont}) \notin \dom(\mstore)$), but any strategy is sound~\citep{dvanhorn:Might2009Posteriori}.
 %
@@
 %
-A finite allocation strategy makes the continuation the only unbounded data structure in the machine, which follows a stack discipline with a finite ``stack alphabet'' --- thus the pushdown system interpretation is apt.
+A finite allocation strategy makes the continuation the only unbounded data structure in the machine, which follows a stack discipline with a finite ``stack alphabet''---thus the pushdown system interpretation is apt.
 %
@@
 %
-In this case, the context includes the function, the environment, and the store.
+In this case, the context includes the function, environment, and store.
 %
@@
 %
-We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores --- this leads to a recursive data-structure that we are trying to avoid.
+We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores---this leads to a recursive data-structure that we are trying to avoid.
 %
@@
 %
+Here $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$.
+%
 We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
@@
 \begin{figure}
-      \begin{tabular}{rcl}
+      \begin{tabular}{rcccl}
         $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$
         & $\Rightarrow$ &
-        $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$ \\
-        & & \multicolumn{1}{c}{$\Downarrow$} \\
-        \begin{minipage}{0.5\linewidth}
+        $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$
+        & $\Rightarrow$ &
+        $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
+      \end{tabular}
       \begin{center}
-            {\small
       \begin{tabular}{|l|c|c|c|}
@@
         \hline
-              \end{tabular}} \\[8pt]
-            {\small
+        \end{tabular}
       \begin{tabular}{|l|c|}
@@
         \hline
-              \end{tabular}}
-          \end{center}
-
-      \end{minipage} & &
-        $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
       \end{tabular}
-    \begin{center} {\small
       \begin{tabular}{|l|c|c|}
@@
         \hline
-        \end{tabular}} \\[10pt]
- Where $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$
+      \end{tabular}
       \end{center}
@@
 %
-Address reachability is in terms of a base ``touches'' function, $\touches$, that is different for our continuations.
+Address reachability is in terms of a base ``touches'' function, $\touches$.
 %
@@
 
-A tabular semantics for the CES$\xi$K machine follows the previous section almost entirely, except that $\mktab$ now stores $(\mkont, \mframe)$ pairs, and returns reinstate the saved $\mframe$:
+A tabular semantics for the CES$\xi$K machine follows the previous section, except that $\mktab$ now stores $(\mkont, \mframe)$ pairs, and returns reinstate the saved $\mframe$:
 \begin{align*}
@@
 %%
-CFA2 also includes what they called ``transitive summaries'' to deal with tail calls.
-%
-We create an $\rt$ continuation at every function call.
-%
-This view contends with tail calls, but we can identify tail calls easily --- any call with an $\rt$ as its continuation is a tail call.
-%
-Tail calls are important in language implementations for space complexity reasons \citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these concerns are less important since there is no unbounded recursion.
-%
-The repeated popping of $\rt$ inserted by what otherwise were tail-calls is synonymous with CFA2's transitive summaries.
+%% CFA2 also includes what they called ``transitive summaries'' to deal with tail calls.
+%% %
+%% We create an $\rt$ continuation at every function call.
+%% %
+%% This view contends with tail calls, but we can identify tail calls easily---any call with an $\rt$ as its continuation is a tail call.
+%% %
+%% Tail calls are important in language implementations for space complexity reasons \citep{ianjohnson:clinger:tail-calls:1998}, but in an analysis, these concerns are less important since there is no unbounded recursion.
+%% %
+%% The repeated popping of $\rt$ inserted by what otherwise were tail-calls is synonymous with CFA2's transitive summaries.
 
@@
 %
-For this, we extend $\alloc$'s specification to allow more arguments --- the private store, and a set of keys that need cons cells allocated for them --- and produce a map from keys to pairs of addresses.
+For this, we extend $\alloc$'s specification to allow more arguments---the private store, and a set of keys that need cons cells allocated for them---and produce a map from keys to pairs of addresses.
 %
@@
 
-At return boundaries, the memo table must add the result to all the represented contexts:
+At returns, the memo table must add the result to all represented contexts:
 \begin{align*}
@@
 % \end{lstlisting}
+\begin{small}
 \begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{let*}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{x}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{x}\RktPn{)}\RktPn{]}
@@
 \mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{0}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g0v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktVal{1}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{g1v}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\le$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g0v}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{g1v}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\end{SingleColumn}\end{RktBlk}\end{SCodeFlow}
-
+\end{small}
 PDCFA with a monovariant allocation strategy will predict the CPS'd version returns true or false.
@@
 %
-The important difference when we move higher-order is that the model construction to feed these methods is an additional problem --- the one we solve here.
+The important difference when we move higher-order is that the model construction to feed these methods is an additional problem---the one we solve here.
 %%
@@
 %%
-Our goal in the future is to show that this technique is even more widely applicable than shift and reset.
-%
-We conjecture that the same framework will apply to the most intricate control operators in production languages such as in \citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}.
-%
-\iftr{
-We have an unproven prototype of the stack inspection capabilities provided by the {\tt with-continuation-mark}, {\tt current-continuation-marks} and {\tt continuation-mark-set->list} primitives, where a ``context'' then also includes what is the result of {\tt current-continuation-marks}, since functions can inspect only this amount of the stack.}
-%%
-
-%%
-Memoization with respect to the entire store is safe, if not a little wasteful; there are portions of the store that can change and be irrelevant to the result of a function call.
-%
-There are techniques to ignore differences in the heap when they are irrelevant, dubbed ``sparseness analysis'' that have been successful for large-scale first-order analysis~\citep{ianjohnson:DBLP:conf/pldi/OhHLLY12}, but they rely on a pre-analysis that knows the control-flow graph.
-%
-Alternatively, there is an analogous concept in concrete semantics called ``self-adjusting computation''~\citep{ianjohnson:DBLP:conf/afp/AcarL08} that has been used for higher-order programs that looks like the right starting point for systematically constructing sparse analyses.
+%% Our goal in the future is to show that this technique is even more widely applicable than shift and reset.
+%% %
+%% We conjecture that the same framework will apply to the most intricate control operators in production languages such as in \citet{ianjohnson:Flatt:2007:ADC:1291151.1291178}.
+%% %
+%% \iftr{
+%% We have an unproven prototype of the stack inspection capabilities provided by the {\tt with-continuation-mark}, {\tt current-continuation-marks} and {\tt continuation-mark-set->list} primitives, where a ``context'' then also includes what is the result of {\tt current-continuation-marks}, since functions can inspect only this amount of the stack.}
+%% %%
+
+%% %%
+%% Memoization with respect to the entire store is safe, if not a little wasteful; there are portions of the store that can change and be irrelevant to the result of a function call.
+%% %
+%% There are techniques to ignore differences in the heap when they are irrelevant, dubbed ``sparseness analysis'' that have been successful for large-scale first-order analysis~\citep{ianjohnson:DBLP:conf/pldi/OhHLLY12}, but they rely on a pre-analysis that knows the control-flow graph.
+%% %
+%% Alternatively, there is an analogous concept in concrete semantics called ``self-adjusting computation''~\citep{ianjohnson:DBLP:conf/afp/AcarL08} that has been used for higher-order programs that looks like the right starting point for systematically constructing sparse analyses.
 %

### 2013-12-13 content.tex
Less relevant technical bits removed for space.

--- 2013-12-11 content.tex
@@
 %
-Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, say, 0-CFA would provide.
+Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model like 0-CFA would provide.
 %%
@@
 
-CFA2 is the first published analysis of an untyped higher-order programming language that could properly match calls and returns.
-%
-Vardoulakis and Shivers had a clear goal of harnessing the extra information a pushdown model provides to produce a high-precision analysis that works well in practice.
+The creators of CFA2 had a clear goal of harnessing the extra information a pushdown model provides to produce a high-precision analysis that works well in practice.
 %
@@
 %
-Their criteria for a binding never escaping is like the criteria of register-allocatable bindings in the Orbit Scheme compiler~\citep{ianjohnson:kranz:thesis:1988}.
+%Their criteria for a binding never escaping is like the criteria of register-allocatable bindings in the Orbit Scheme compiler~\citep{ianjohnson:kranz:thesis:1988}.
 %
@@
 %
-That is, since bindings in the abstract represent several values, we don't want to reference a variable \texttt{x} in two different places and have it resolve to two different values.
-%
-In AAM, a variable reference non-deterministically steps to all possible values associated with that variable, because there are several possible values in the store for allocated bindings.
-%
-Here we want to say that once \texttt{x} is considered to stand for value $v$, then all subsequent references of \texttt{x} should be $v$.
-%
-If they are not all considered $v$, it looks as if \texttt{x} was rebound; it hasn't, and thus it is a ``fake rebinding.''
-%
-CFA2 does not step to all values on variable reference, but instead carries all its values around in superposition until they need to be observed at, say, a function call.
-%
-We give a simplified semantics that is more along the AAM style, but CFA2's approach can easily be recovered from it.\footnote{Our Redex model implements the fake rebinding strategy CFA2 itself employs.}
+Bindings in the abstract represent several values, but we don't want to reference a variable \texttt{x} in different places in a function and have it resolve to different values
+\footnote{In AAM, a variable reference non-deterministically steps to all possible values associated with that variable.}.
+%
+If subsequent variable references are not the some, it looks as if the variable was rebound; it hasn't, and thus it is a ``fake rebinding.''
+%
+%CFA2 does not step to all values on variable reference, but instead carries all its values around in superposition until they need to be observed at, say, a function call.
+%
+%We give a simplified semantics that is more along the AAM style, but CFA2's approach can easily be recovered from it.\footnote{Our Redex model implements the fake rebinding strategy CFA2 itself employs.}
 %%
@@
 %
-We create an $\rt$ continuation at every function call.
-%
-This view contends with tail calls, but we can identify tail calls easily --- any call with an $\rt$ as its continuation is a tail call.
+Our memoization view contends with tail calls, but we can identify tail calls easily --- those with an $\rt$ continuation.
 %
@@
 %
-The additional escape information that allows us to elide allocation in the heap introduces a second level of reasoning, where the burden of proof falls on the escape analysis.
-%
-The escape analysis adds information to use in the proof, that if $\mvar$ never escapes, then every reference to allocations of $\mvar$'s binding will always be able to use $\mframe$ for the lookup.
-%
-With this proposition, we can prove that the CES$\xi$K machine that always allocates in the store has a bisimulation with the refined machine that exploits escape analysis to avoid unnecessary changes to the store.
+%The additional escape information that allows us to elide allocation in the heap introduces a second level of reasoning, where the burden of proof falls on the escape analysis.
+%
+%The escape analysis adds information to use in the proof, that if $\mvar$ never escapes, then every reference to allocations of $\mvar$'s binding will always be able to use $\mframe$ for the lookup.
+%
+%With this proposition, we can prove that the CES$\xi$K machine that always allocates in the store has a bisimulation with the refined machine that exploits escape analysis to avoid unnecessary changes to the store.
 %
@@
 
+%% TODO: Cut here next.
 There is contention among programming language researchers whether \rackett{call/cc} should be a language primitive, since it captures the entire stack, leading to space leaks \citep{ianjohnson:kiselyov:against-callcc}.

### 2013-12-13 abstract.tex
Started a section on with-continuation-mark, but had to punt due to extension vs copy adding more complexity than I could fit.

@@
 %
-The machinery has no concrete counterpart --- there is little to no operational intuition as to why it works.
+The machinery has no concrete counterpart---there is little to no operational intuition as to why it works.
 %
@@
 %
-We use this interpretation to derive a novel pushdown analysis for delimited, composable control.
+We use this interpretation to derive novel pushdown analyses for delimited, composable control.

### 2013-12-13 content.tex
Got down to page limit. Cut the proof of non-termination in problematic semantics. FIAT SEMANTICS

@@
 
-\paragraph{Example:} if we use this semantics to run through the example of the previous section, we find that with the addition of stack frames, there is no confusion about \texttt{z}'s value either, so if the operator weren't $\le$ but instead $<$ or $+$, we could still constant fold that away.
-%%
-
-%%
-The stack frame introduces a refinement over the original CESK machine, so we first prove it correct with respect to the concrete semantics.
-%
-We follow up with a similar proof to that of PDCFA, which is that the CES$\xi$K machine has a bisimulation with its tabular semantics, with comparable allocation strategies.
+\paragraph{Example:} revising the earlier example, we find that with the addition of stack frames, there is no confusion about \texttt{z}'s value either.
+%
+If the operator were instead $<$ or $+$, we could still constant fold that away.
+%%
+
+%%
+The stack frame introduces a refinement over the original CESK machine, so we first prove it correct with respect to the concrete semantics, in a similar fashion to PDCFA.
+%
 \begin{theorem}[Correctness of stack frame refinement]\label{thm:refinement}
@@
 %% TODO: Cut here next.
-There is contention among programming language researchers whether \rackett{call/cc} should be a language primitive, since it captures the entire stack, leading to space leaks \citep{ianjohnson:kiselyov:against-callcc}.
-%
-Alternative control operators have been proposed that delimit how much of the stack to capture, such as $\%$ (read ``prompt'') and capture operator ${\mathcal F}$ (read ``control'') \citep{ianjohnson:felleisen:control:1988}, or \texttt{reset} (equivalent to $\%$) and \texttt{shift} \citep{ianjohnson:danvy:filinski:delim:1990}.
+There is contention among programming language researchers whether \rackett{call/cc} should be a language primitive\citep{ianjohnson:kiselyov:against-callcc}.
+%
+Alternative control operators have been proposed that delimit how much of the stack to capture, such as $\%$ (read ``prompt'') and capture operator ${\mathcal F}$ (read ``control'') \citep{ianjohnson:felleisen:control:1988}, or \texttt{reset} and \texttt{shift} \citep{ianjohnson:danvy:filinski:delim:1990}.
 %
@@
 %
-Stack replacement is easily modeled in a regular analysis using the AAM approach, and Vardoulakis and Shivers showed it can be done in a pushdown approach.
-%
-Stack extension, however, poses a new challenge for pushdown analysis, since one application of a composable continuation means pushing an unbounded amount of frames onto the stack.
-%
-Vardoulakis' and Shivers' approach does not immediately apply in this situation, since their technique drops all knowledge of the stack at a continuation's invocation site; extension, however, must preserve it.
+Stack extension poses a new challenge for pushdown analysis, since one application of a composable continuation means pushing an unbounded amount of frames onto the stack.
+%
+Vardoulakis' and Shivers' technique drops all knowledge of the stack at a continuation's invocation site; extension must preserve it.
 %%
@@
 
-Turning this into a table-based semantics involves making prompts a point of indirection, just like function calls.
+The table-based semantics makes prompts a point of indirection, just like function calls.
 %
@@
 %
-Unfortunately, since $\rt$ continuations contain a store, and continuations can now appear in the store, this introduces an unbounded data structure that could cause the analysis to never terminate.
-%
-For example, a loop that continually captures continuations would introduce unboundedly many continuation values due to the ever-growing store.
-%
-This means the store is not finite height, and there may be no fixed point.
-
-\paragraph{A fix to non-termination:} we allow for a tunable approximation by using $\alloc$.
+\paragraph{Problem:} now $\rt$ continuations contain a store, and captured continuations can appear in the store.
+%
+This introduces an unbounded data structure that can cause the analysis to never terminate.
+
+\paragraph{Solution:} we allow for a tunable approximation by using $\alloc$.
 %
@@
 
-\section{Conclusion and future work}
+\section{Conclusion}
 

### 2014-02-19 content.tex
minor paper fixes. Switched to sigplan.

@@
 
-Programs in higher-order languages heavily use function calls and method dispatch as part of their control flow.
-%
-Until recently, flow analyses for higher-order languages could not handle return flow precisely \citep{ianjohnson:vardoulakis-lmcs11, dvanhorn:Earl2010Pushdown}, which leads to several spurious paths (and thus false positives) due to the pervasiveness of function or method calls and subsequent returns.
+Programs in higher-order languages heavily use function calls and method dispatch for control flow.
+%
+Standard flow analyses' imprecise handling of returns damages all specific analyses' precision.
+%
+Recent techniques match calls and returns precisely \citep{ianjohnson:vardoulakis-lmcs11, dvanhorn:Earl2010Pushdown} and build smaller models more quickly than a standard 0-CFA (evaluation predicted 2-5 times more constant bindings).
 %
@@
 %
-\footnote{We refer to finite model analyses as ``regular analyses'' after regular languages.}
+\footnote{We refer to finite model analyses as ``regular analyses'' after the regular languages of traces they realize.}
 %
@@
 %%
-There is a systematic process for transforming off-the-shelf programming language semantics into a form amenable to \emph{regular} analysis that has been widely applied with great success to production programming languages (a technique called abstracting abstract machines, or AAM~\citep{dvanhorn:VanHorn2010Abstracting}).
+There is a systematic process for transforming off-the-shelf programming language semantics into a form amenable to \emph{regular} analysis that has been widely applied to production programming languages (a technique called abstracting abstract machines, or AAM~\citep{dvanhorn:VanHorn2010Abstracting}).
 %
@@
 %
-We contend that the machinery that we employ in the abstract should have a concrete counterpart that maintains the meaning of the language so that we can see their effect without introducing approximations which might mask correctness issues.
-%
-Abstraction should be a simple process that is ``obviously correct.''
+Premature abstraction can mask correctness issues with complicated machinery we employ in our analyses\footnote{Oh so many personal experiences}.
+%
+We contend that analysis machinery should have an executable concrete counterpart that maintains the meaning of the language for effective testing.
+%
+Abstraction should then be a simple process that is ``obviously correct.''
 %
@@
 %
-The semantic spaces for the machine are modified thusly:
+The semantic spaces for the machine are modified as follows:
 
@@
 %
-Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store \textsc{[memo lookup]}.
+Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store (\todo{show in figure})
 %%
@@
   $\mstate \stepto \mstate' \text{ where } \maddr = \alloc(\mstate)$ \\
-  \begin{tabular}{r|ll}
+  \begin{tabular}{r|l}%{r|ll}
     \hline
@@
     $\tpl{\mval,\mstore,\mkont, \mktab, \mmemo}$ if $\mval \in \mstore(\menv(\mvar))$
-    &
-    \textsc{[variable lookup]}
+%    & \textsc{[variable lookup]}
     \\
@@
     $\tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}$
-    &
-    \textsc{[function eval]}
+%    & \textsc{[function eval]}
     \\
@@
     $\tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}$
-    &
-    \textsc{[argument eval]}
+%    & \textsc{[argument eval]}
     \\
@@
           \mmemo}$
-    &
-    \textsc{[un-memoized call]}
+%    & \textsc{[un-memoized call]}
      \\ & \quad if $(\mpoint, \mstore') \notin \dom(\mmemo)$
@@
             \mktab',
-            \mmemo}$
-    & \textsc{[memo lookup]}
+            \mmemo}$ % & \textsc{[memo lookup]}
     \\ & \quad if $(\mval_\mathit{result},\mstore'') \in \mmemo(\mpoint,\mstore')$
@@
     $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}$
-    &
-    \textsc{[memoize/return]}
+%    & \textsc{[memoize/return]}
     \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
@@
 %
-0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
+0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thus predicts this program evaluates to true or false.
 %
@@
   \,
-  \begin{minipage}{0.32\linewidth}
     $\touchesk{\mkont}(\mktab) = \setbuild{\maddr}{\mkont \rightsquigarrow_\mktab^*\maddr}$
+\\[4pt]
+  \begin{minipage}{0.5\linewidth}
+    $\touchesf(\kar{\mexpr,\menv}) = \mathit{range}(\menv|_{\mathit{fv}(\mexpr)})$
   \end{minipage}
-  \begin{minipage}{0.36\linewidth}
-    $\touchesf(\kar{\mexpr,\menv}) = \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)})$
-  \end{minipage}
-  \begin{minipage}{0.3\linewidth}
+  \begin{minipage}{0.4\linewidth}
     $\touchesf(\kfn{\mval}) = \touches(\mval)$
@@
       \text{where } \mathit{root} &= \touchesk{\mkont}(\mktab) \cup
-      \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)})
+      \mathit{range}(\menv|_{\mathit{fv}(\mexpr)})
     \end{align*}
@@
 %
-This resulted in more than just the call/return matching of the previous section, which is why we are showing the two separately.
+This resulted in the following features in addition to just the call/return matching of \autoref{sec:pdcfa}:
 %
@@
 %
-If subsequent variable references are not the some, it looks as if the variable was rebound; it hasn't, and thus it is a ``fake rebinding.''
+If subsequent variable references are not the same, it looks as if the variable was rebound; it hasn't, and thus it is a ``fake rebinding.''
 %
@@
 \begin{align*}
-  \tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} &\stepto
+  \tpl{\mval,\mstore,\mframe,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} &\stepto
   \tpl{\mpoint,
@@
     $\tpl{(\mexpr, \menv), \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
-    &
-    \textsc{[push prompt]}
+%    & \textsc{[push prompt]}
     \\
@@
     $\tpl{\mval, \mstore, {\mkont}, {\mmkont}}$
-    &
-    \textsc{[pop prompt]}
+%    & \textsc{[pop prompt]}
     \\
@@
     $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore',\kmt,\mmkont}$
-    &
-    \textsc{[capture continuation]}
+%    & \textsc{[capture continuation]}
     \\ & where $\mstore' = \joinone{\mstore}{\maddr}{\vcomp{\mkont}}$
@@
     $\tpl{\mval, \mstore, \mkont', \mkapp{\mkont}{\mmkont}}$
-    &
-    \textsc{[compose continuation]}
+%    & \textsc{[compose continuation]}
   \end{tabular}  
@@
     &
-    $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab,\mmemo}$ if $(\mval',\mstore') \in \mmemo(\mkont',\mval,\mstore)$
+    $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab,\mmemo}$ if $(\mval',\mstore') \in \mmemo(\vcomp{\mkont'},\mval,\mstore)$
   \end{tabular}  

### 2014-02-21 content.tex
Drafted a why-we-matter section.

@@
 To exercise the framework further, we give a novel analysis for delimited, composable control. \\
+
+\section{Models in higher-order analysis}
+
+Constructing models and analyzing models are different processes.
+%
+The goal of this work is to simplify the program $\to$ model process.
+%
+First-order languages have the luxury of a fully-precise control flow graph that can be extracted from a program AST in linear time.
+%
+In a sense, the first-order world gets their models ``for free.''
+%
+Higher-order recursion schemes (HORS) are a powerful model for simply typed functional programs with good model-checking properties~\citep{dvanhorn:Ong2006ModelChecking}, but the program $\to$ HORS process is narrowly applicable.
+%
+It is still a matter of mystery for HORS to support programming language features like mutable state and untyped computation.
+%%
+As a middle ground, we show a simple way to transform one's programming language semantics into a what can be viewed as a pushdown model constructor.
+%%
+
+%%
+When striving for precision, the interconnectedness of data-flow and control-flow analysis in higher-order languages extend upwards to the interconnectedness of model construction and model-checking.
+%
+When we tune the abstractions, the result of translating semantics can range from control-flow analysis~\citep{dvanhorn:VanHorn2010Abstracting} to a symbolic execution engine~\citep{dvanhorn:TobinHochstadt2012Higherorder}.
+%
+These are all forward-executing techniques, so the analogy does not fully extend to all analytical technologies.
+%
+We find the power and simplicity in the constructions enticing nonetheless.
+%
+The resulting artifacts can be used as models in the first-order tools~\citep{ianjohnson:bandera}
 
@@
 %
-A memoizing program will call a function in a non-tail context in order to obtain the result, store the result to an ``enough context to obtain the same result'' key in a table (in a pure setting, just the function and arguments), and continue with the result.
+A memoizing program will call a function in a non-tail context in order to obtain the result, store the result to an \emph{enough context to obtain the same result} key in a table (in a pure setting, just the function and arguments), and continue with the result.
 %

### 2014-02-24 content.tex
Small changes. About to remove gratuitous parentheses.

@@
 %%
-This paper gives a common, simple framework to derive CFA2, PDCFA and its garbage-collecting extension in a manner similar to AAM.
-%
-To exercise the framework further, we give a novel analysis for delimited, composable control. \\
+This paper gives a common, simple methodology to derive CFA2, PDCFA and its garbage-collecting extension in a manner similar to AAM.
+%
+Our presentation is meant to be a tutorial to this related literature, with the additional promise of a novel analysis.
+%
+We don't presume a language of formal semantics such that our technique applies to all formalisms in such a language.
+%
+Such an approach is unlikely to capture all ``real language's'' formalizations.
+%
+An additional abstraction also detracts from the operational intuitions and simplicity of the methodology.
 
@@
 %
-The goal of this work is to simplify the program $\to$ model process.
+The goal of this work is to simplify the \textit{program} $\to$ \textit{model} process.
 %
@@
 %
-Higher-order recursion schemes (HORS) are a powerful model for simply typed functional programs with good model-checking properties~\citep{dvanhorn:Ong2006ModelChecking}, but the program $\to$ HORS process is narrowly applicable.
+Higher-order recursion schemes (HORS) are a powerful model for simply typed functional programs with good model-checking properties~\citep{dvanhorn:Ong2006ModelChecking}, but the \textit{program} $\to$ \textit{HORS} process is narrowly applicable.
 %
@@
 %%
-When striving for precision, the interconnectedness of data-flow and control-flow analysis in higher-order languages extend upwards to the interconnectedness of model construction and model-checking.
+When striving for precision, the interconnectedness of data-flow and control-flow analysis in higher-order languages extends upwards to the interconnectedness of model construction and model-checking.
 %
@@
 %
-The resulting artifacts can be used as models in the first-order tools~\citep{ianjohnson:bandera}
+The resulting artifacts can be used as models in the first-order tools, as demonstrated by the Java model-checker Bandera~\citep{ianjohnson:bandera}.
 
@@
 %
-We can adopt the technique for higher precision in the common case but still handle difficult cases such as first-class control.
+The technique is adaptable for higher precision in the common case but to still handle difficult cases such as first-class control.
 %
@@
 \begin{align*}
-  \mexpr \in \Expr = \svar{\mvar} \alt \sapp{\mexpr}{\mexpr} \alt \slam{\mvar}{\mexpr}
-\end{align*}
-\begin{minipage}[b]{.55\linewidth}
-  \begin{align*}
+  \mexpr \in \Expr &= \svar{\mvar} \alt \sapp{\mexpr}{\mexpr} \alt \slam{\mvar}{\mexpr} \\
   \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \\
@@
   \mkont \in \Kont &::= \kmt \alt \kcons{\mkframe}{\mkont} \\
-  \mkframe \in \Frame &::= \kar{\mexpr,\menv} \alt \kfn{\mval}
-  \end{align*}\hrule height 0pt\end{minipage}
-\begin{minipage}[b]{.40\linewidth}
-  \begin{align*}
+  \mkframe \in \Frame &::= \kar{\mexpr,\menv} \alt \kfn{\mval} \\
   \menv \in \Env &= \Var \to \Addr \\
@@
   \maddr \in \Addr &\text{ an infinite set}
-  \end{align*}\hrule height 0pt\end{minipage}
+\end{align*}
 }
@@
 %%
+%
+The $\kmt$ continuation is the empty continuation.
+%
+The continuation frames represent an \textbf{ar}gument left to evaluate, and \textbf{f}u\textbf{n}ction left to call.
+%
+%
 The $\Addr$ space is what controls the precision of the model.
@@
 %
-A finite allocation strategy makes the continuation the only unbounded data structure in the machine, which follows a stack discipline with a finite ``stack alphabet''---thus the pushdown system interpretation is apt.
+A finite allocation strategy makes the continuation the only unbounded data structure in the machine.
+%
+The semantics' treatment of continuations follows a stack discipline, and since it has a finite ``stack alphabet'' (space of frames), a pushdown system interpretation is apt.
 %
@@
 % variable lookup
-    $\tpl{(\svar\mvar, \menv), \mstore, \mkont}$
+    $\tpl{\svar\mvar, \menv, \mstore, \mkont}$
     &
@@
 \end{figure}
+
+\paragraph{Notation:} we make judicious use of the isomorphism between associations of $n$-tuples to spare the paper from excessive parentheses, even though our implementations are in a Lisp-descendent language.
 

### 2014-02-25 content.tex
Fixed formatting. Added to metatheory section. Need to prove the correctness theorem for PDCFA and fix the example(s).

@@
   \mexpr \in \Expr &= \svar{\mvar} \alt \sapp{\mexpr}{\mexpr} \alt \slam{\mvar}{\mexpr} \\
-  \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \\
+  \mstate \in \State &= \Expr \times \Env \times \Store \times \Kont \\
   \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \\
@@
 For a concrete semantics, we require the allocation meta-function $\alloc :
-\State \to \Addr$ to return fresh addresses ($\alloc(\tpl{(\mexpr,\menv),\mstore,\mkont}) \notin \dom(\mstore)$), but any strategy is sound~\citep{dvanhorn:Might2009Posteriori}.
+\State \to \Addr$ to return fresh addresses ($\alloc(\tpl{\mexpr,\menv,\mstore,\mkont}) \notin \dom(\mstore)$), but any strategy is sound~\citep{dvanhorn:Might2009Posteriori}.
 %
@@
 % application
-    $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont}$
-    &
-    $\tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}}$
+    $\tpl{\sapp{\mexpri0}{\mexpri1}, \menv, \mstore, \mkont}$
+    &
+    $\tpl{\mexpri0, \menv, \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}}$
     \\
@@
     &
-    $\tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}}$
+    $\tpl{\mexpr, \menv, \mstore, \kcons{\kfn{\mval}}{\mkont}}$
     \\
@@
     &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\mval}, \mkont}$
+    $\tpl{\mexpr, \extm{\menv}{\mvar}{\maddr}, \joinone{\mstore}{\maddr}{\mval}, \mkont}$
   \end{tabular}
@@
 
-\paragraph{Notation:} we make judicious use of the isomorphism between associations of $n$-tuples to spare the paper from excessive parentheses, even though our implementations are in a Lisp-descendent language.
+\paragraph{Notation:} we make judicious use of the isomorphism between associations of tuples to spare the paper from excessive parentheses, even though our implementations are in a Lisp-descendent language.
 
@@
 \begin{align*}
-  \mstate \in \State &= (\Expr \times \Env) \times \Store \times \Kont \times \KTab \times \Memo \\
-  \mkont \in \Kont &::= \kmt \alt \krt{(\mexpr,\menv),\mstore} \alt \kcons{\mkframe}{\mkont} \\
-  \mktab \in \KTab &= (\Expr \times \Env)\times \Store \to \wp(\Kont) \\
-  \mmemo \in \Memo &= (\Expr \times \Env)\times \Store \to \wp(\Value \times \Store)
+  \mstate \in \State &= \Expr \times \Env \times \Store \times \Kont \times \KTab \times \Memo \\
+  \mkont \in \Kont &::= \kmt \alt \krt{\mexpr,\menv,\mstore} \alt \kcons{\mkframe}{\mkont} \\
+  \mktab \in \KTab &= Expr \times \Env\times \Store \to \wp(\Kont) \\
+  \mmemo \in \Memo &= \Expr \times \Env\times \Store \to \wp(\Value \times \Store)
 \end{align*}
@@
 % variable lookup
-    $\tpl{(\svar\mvar, \menv), \mstore, \mkont, \mktab, \mmemo}$
+    $\tpl{\svar\mvar, \menv, \mstore, \mkont, \mktab, \mmemo}$
     &
@@
 % application
-    $\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}$
+    $\tpl{\sapp{\mexpri0}{\mexpri1}, \menv, \mstore, \mkont, \mktab, \mmemo}$
+    &
+    $\tpl{\mexpri0, \menv, \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}$
 %    & \textsc{[function eval]}
@@
     &
-    $\tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}$
+    $\tpl{\mexpr, \menv, \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}$
 %    & \textsc{[argument eval]}
@@
 % function call
-    $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo}$
+    $\tpl{\mval,\mstore,\kcons{\kfn{\slam{\mvar}{\mexpr},\menv}}{\mkont},\mktab,\mmemo}$
     & % actually do call,
@@
     &
-    $\tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}$
+    $\tpl{\mval, \mstore, \mkont, \mktab, M'}$
 %    & \textsc{[memoize/return]}
     \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
+    \\ where & $\mmemo' = \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}$
   \end{tabular}
@@
 \begin{align*}
-  \mastate \in \sa{State} &= (\Expr \times \Env) \times \Store \times \Kont \\
-  \System &= \wp(\sa{State}) \times \wp(\sa{State}) \times \KTab \times \Memo \\
-  {\mathcal F} &: \sa{\State} \to \System \to \System \\
-  {\mathcal F}(\mastate_0)(S, \emptyset, \mktab, \mmemo) &= (S, \emptyset, \mktab, \mmemo) \\
-  {\mathcal F}(\mastate_0)(S, F\uplus\set{\mastate}, \mktab, \mmemo) &= (S \cup S', F\cup F', \mktab', \mmemo') \\
+  \mastate \in \sa{State} &= \Expr \times \Env \times \Store \times \Kont \\
+  \System &= \wp(\sa{State}) \times \wp(\sa{State} \times \sa{State}) \times \KTab \times \Memo \\
+  {\mathcal F} &: \sa{\State} \to \System \to \System
+\end{align*}
+\begin{align*}
+  {\mathcal F}(\mastate_0)(\emptyset, R, \mktab, \mmemo) &= (\emptyset, R, \mktab, \mmemo) \\
+  {\mathcal F}(\mastate_0)(F\uplus\set{\mastate}, R, \mktab, \mmemo) &= (F\cup F', R \cup R', \mktab', \mmemo') \\
   \text{where } I &= \setbuild{\mstate'}{\wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
-                \wn(\tpl{(\mexpr,\menv),\mstore,\mkont},\mktab,\mmemo)
-                   &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
-\end{align*}
-\begin{minipage}{.55\linewidth}
-  \begin{align*}
+                \wn(\tpl{\mexpr,\menv,\mstore,\mkont},\mktab,\mmemo)
+                   &= \tpl{\mexpr,\menv,\mstore,\mkont,\mktab,\mmemo}
+\end{align*}
+\begin{align*}
+    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge \mastate\nequiv\tpl{\mval,\mstore,\krt{\mctx}}} \cup \set{\mastate_0}\\
     \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \mktab', \_) \in I} \\
-    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \mmemo') \in I}
-  \end{align*}
-\end{minipage}
-\begin{minipage}{.40\linewidth}
-  \begin{align*}
-    S' &= \setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \cup \set{\mastate_0}\\
-    F' &= S' \setminus S
-  \end{align*}
-\end{minipage}
+    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \mmemo') \in I} \\
+    F' &= \setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \setminus (\pi_1 R \cup \pi_2 R)
+  \end{align*}
 }
@@
 %%
-A $\System$ embodies the states seen and at which store, $S$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
+A $\System$ embodies the reductions seen disregarding the administrative pop steps, $R$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
+%
+We define $\pi_1$ and $\pi_2$ as the pointwise lifting of the left and right projections out of a pair, respectively.
 % Note: Only if widened store.{
@@
 %
-The least fixed-point of ${\mathcal F}(\tpl{(\mexpr,\bot),\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
+The least fixed-point of ${\mathcal F}(\tpl{\mexpr,\bot,\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
 %
 Systematic techniques for a performant implementation can be found in \citet{ianjohnson:oaam:icfp2013}.
+%
 
@@
 \begin{center}
-  \includegraphics[scale=0.5]{example0}
+  \todo{fix example}
+  %\includegraphics[scale=0.5]{example0}
 \end{center}
@@
 \begin{figure}
-      \begin{tabular}{rcccl}
-        $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$
-        & $\Rightarrow$ &
-        $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$
-        & $\Rightarrow$ &
-        $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
-      \end{tabular}
-      \begin{center}
-      \begin{tabular}{|l|c|c|c|}
-        \hline
-        \# & $\mktab$ & $\mmemo$ \\
-        \hline
-        0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
-        1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
-        \hline
-        \end{tabular}
-      \begin{tabular}{|l|c|}
-        \hline
-        \# & $\mctx$ \\
-        \hline
-        0  & $(({\tt x}, \iota), \mstore_1)$ \\
-        1  & $(({\tt x}, \iota), \mstore_3)$ \\
-        \hline
-      \end{tabular}
-      \begin{tabular}{|l|c|c|}
-        \hline
-        \# & $\mstore$ & $\mkont$ \\
-        \hline
-        0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\iota)}]$ & $\mkont_0$ \\
-        1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
-        2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)}, \iota)$ \\
-        3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
-        4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
-        \hline
-      \end{tabular}
-      \end{center}
+      % \begin{tabular}{rcccl}
+      %   $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$
+      %   & $\Rightarrow$ &
+      %   $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$
+      %   & $\Rightarrow$ &
+      %   $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
+      % \end{tabular}
+      % \begin{center}
+      % \begin{tabular}{|l|c|c|c|}
+      %   \hline
+      %   \# & $\mktab$ & $\mmemo$ \\
+      %   \hline
+      %   0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
+      %   1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
+      %   \hline
+      %   \end{tabular}
+      % \begin{tabular}{|l|c|}
+      %   \hline
+      %   \# & $\mctx$ \\
+      %   \hline
+      %   0  & $(({\tt x}, \iota), \mstore_1)$ \\
+      %   1  & $(({\tt x}, \iota), \mstore_3)$ \\
+      %   \hline
+      % \end{tabular}
+      % \begin{tabular}{|l|c|c|}
+      %   \hline
+      %   \# & $\mstore$ & $\mkont$ \\
+      %   \hline
+      %   0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\iota)}]$ & $\mkont_0$ \\
+      %   1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
+      %   2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)}, \iota)$ \\
+      %   3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
+      %   4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
+      %   \hline
+      % \end{tabular}
+      % \end{center}
+\todo{fix example}
       \caption{Example analysis evaluation}
@@
 \iftr{\input{metatheory}}
-
-\begin{theorem}[Correctness of PDCFA]\label{thm:global}
-  With a fair state selection from the frontier, the widened tabular semantics reflects a relation that has a bisimulation with the standard semantics.
-\end{theorem}
 
@@
     \begin{align*}
-      \Gamma(\tpl{(\mexpr,\menv),\mstore,\mkont, \mktab, \mmemo}) &= \tpl{(\mexpr,\menv),\mstore|_{\reaches(\mathit{root}, \mstore)}, \mkont, \mktab, \mmemo} \\
+      \Gamma(\tpl{\mexpr,\menv,\mstore,\mkont, \mktab, \mmemo}) &= \tpl{\mexpr,\menv,\mstore|_{\reaches(\mathit{root}, \mstore)}, \mkont, \mktab, \mmemo} \\
       \text{where } \mathit{root} &= \touchesk{\mkont}(\mktab) \cup
@@
   \reaches(\mathit{root}, \mstore) = \setbuild{b}{\maddr \in \mathit{root}, \maddr \rightsquigarrow_\mstore^* b}
-   \qquad
+   \quad
                \inferrule{\mval \in \mstore(\maddr) \\
@@
 % variable lookup
-    $\tpl{(\svar[\mlab]{\mvar}, \menv), \mstore, \mframe, \mkont}$
+    $\tpl{\svar[\mlab]{\mvar}, \menv, \mstore, \mframe, \mkont}$
     &
@@
     &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore', \mframe', \kcons{\kpush{\mframe}}{\mkont}}$
+    $\tpl{\mexpr, \extm{\menv}{\mvar}{\maddr}, \mstore', \mframe', \kcons{\kpush{\mframe}}{\mkont}}$
     \\ & where $(\mstore',\mframe') = \bind(\mstore,\maddr,\mvar,\mval)$
@@
 %
+We will call the non-escaping conditions ``local.''
+%
 %Their criteria for a binding never escaping is like the criteria of register-allocatable bindings in the Orbit Scheme compiler~\citep{ianjohnson:kranz:thesis:1988}.
 %
-They use the information that a binding never escapes to not bother allocating it in the heap.
+CFA2 uses the information that a binding never escapes to not bother allocating it in the heap.
 %
@@
 %
+Model-checking literature has \emph{explicit} constructs to locally bind the result of this non-deterministic choice: \emph{choice variables}~\citep{ianjohnson:DBLP:conf/spin/CookKS05, ianjohnson:DBLP:journals/cacm/DilligDA10}.
+%
 %CFA2 does not step to all values on variable reference, but instead carries all its values around in superposition until they need to be observed at, say, a function call.
@@
 %%
-We show only the significantly modified rules of the semantics in figure \ref{fig:frame-semantics}.
+We show only the significantly modified rules of the semantics in \autoref{fig:frame-semantics}.
 %
@@
    \left\{\begin{array}{ll}
-            (\mstore, \snglm{\maddr}{\mval}) &\text{ if } \mvar \text{ allocations never escape} \\
+            (\mstore, \snglm{\maddr}{\mval}) &\text{ if } \mvar \text{ local} \\ % allocations never escape
             (\joinone{\mstore}{\maddr}{\mval}, \snglm{\maddr}{\mval}) &\text{ otherwise} \\
@@
           \setbuild{(\extm{\mframe}{\maddr}{\set{\mval}}, \mval)}
-                   {\mval \in \mframe(\maddr)} & \text{if } \mlab \text{ non-escaping} \\
+                   {\mval \in \mframe(\maddr)} & \text{if } \mlab \text{ local} \\ %non-escaping
           \setbuild{(\mframe,\mval)}{\mval \in \mstore(\maddr)} & \text{ otherwise}
@@
 %
-As soon as the non-determinism has been determined, we could extend the stack frame so any subsequent reference means what it meant previously in the function\footnote{This is reminiscent of how equality of unknowns is tracked in \citet{ianjohnson:DBLP:journals/cacm/DilligDA10}}.
+As soon as the non-determinism has been determined, we could extend the stack frame so any subsequent reference means what it meant previously in the function.
 %
@@
 \begin{align*}
-  \tpl{\mval,\mstore,\mframe,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} &\stepto
+  \tpl{\mval,\mstore,\mframe,\kcons{\kfn{\slam{\mvar}{\mexpr},\menv}}{\mkont},\mktab,\mmemo} &\stepto
   \tpl{\mpoint,
@@
  \tpl{\mval, \mstore, \mframe, \krt{\mpoint,\mstore'}, \mktab, \mmemo} &\stepto
- \tpl{\mval, \mstore, \mframe', \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}
+ \tpl{\mval, \mstore, \mframe', \mkont, \mktab, \mmemo'}
   \\ & \quad\text{if } (\mkont,\mframe') \in \mktab(\mpoint, \mstore')
+  \\ \text{where } & \mmemo' = \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}
 \end{align*}
@@
 \begin{theorem}[Correctness of stack frame refinement]\label{thm:refinement}
-  The CES$\xi$K machine has a bisimulation with the CESK machine with fresh allocation.
+  The CES$\xi$K machine has a stuttering bisimulation with the CESK machine with fresh allocation.
 \end{theorem}
@@
 %
+We stutter past the stack frame replacement steps.
+%
 %That is, if we always allocated bindings in the heap.
@@
 \begin{theorem}[Correctness of CFA2]\label{thm:cfa2}
-  The widened tabular semantics reifies a relation that has a bisimulation with standard semantics.
+  The widened tabular semantics reflects a relation that has a bisimulation with the CES$\xi$K machine.
 \end{theorem}
@@
   \mexpr \in \Expr &::= \swcm{\mexpr}{\mexpr}{\mexpr} \alt \sccms \alt \scmsl{\mexpr}{\mexpr} \alt \ldots \\
-  \mctx \in \Context &::= ((\mexpr, \menv), \mstore, \mmarkset) \\
+  \mctx \in \Context &::= (\mexpr, \menv, \mstore, \mmarkset) \\
   \mkframe \in \Frame &::= \kwcmk{\mexpr, \mexpr, \menv} \alt
@@
 % wcm eval key
-    $\tpl{(\swcm{\mexpr_k}{\mexpr_v}{\mexpr}, \menv), \mstore, \mkont^\mmarks, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr_k, \menv), \mstore, \kcons{\kwcmk{\mexpr_v,\mexpr,\menv}^\mmarks}{\mkont}, \mktab, \mmemo}$
+    $\tpl{\swcm{\mexpr_k}{\mexpr_v}{\mexpr}, \menv, \mstore, \mkont^\mmarks, \mktab, \mmemo}$
+    &
+    $\tpl{\mexpr_k, \menv, \mstore, \kcons{\kwcmk{\mexpr_v,\mexpr,\menv}^\mmarks}{\mkont}, \mktab, \mmemo}$
     \\
@@
     &
-    $\tpl{(\mexpr_v, \menv), \mstore, \kcons{\kwcmv{\mval_k,\mexpr,\menv}^\mmarks}{\mkont}, \mktab, \mmemo}$
+    $\tpl{\mexpr_v, \menv, \mstore, \kcons{\kwcmv{\mval_k,\mexpr,\menv}^\mmarks}{\mkont}, \mktab, \mmemo}$
     \\    
@@
     &
-    $\tpl{(\mexpr, \menv'), \mstore', \krt{\mctx}^\mmarks,\mktab',\mmemo}$
+    $\tpl{\mexpr, \menv', \mstore', \krt{\mctx}^\mmarks,\mktab',\mmemo}$
     \\ where & $\menv' = \extm{\menv}{\mvar}{\maddr}$ \\
     & $\mstore' = \joinone{\mstore}{\maddr}{\mval}$ \\
-    & $\mctx = ((\mexpr,\menv'),\mstore', \domarkset(\mstate))$ \\
+    & $\mctx = (\mexpr,\menv',\mstore', \domarkset(\mstate))$ \\
     & $\mktab' = \joinone{\mktab}{\mctx}{\mkont}$
@@
 % Current continuation marks
-    $\tpl{(\sccms,\menv),\mstore,\mkont,\mktab,\mmemo}$
+    $\tpl{\sccms,\menv,\mstore,\mkont,\mktab,\mmemo}$
     &
@@
 % cms->list mark set
-    $\tpl{(\scmsl{\mexpr_\mmarks}{\mexpr_k}, \menv), \mstore, \mkont^\mmarks, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr_\mmarks, \menv), \mstore, \kcons{\kcmsls{\mexpr_k,\menv}^\mmarks}{\mkont}, \mktab, \mmemo}$
+    $\tpl{\scmsl{\mexpr_\mmarks}{\mexpr_k}, \menv, \mstore, \mkont^\mmarks, \mktab, \mmemo}$
+    &
+    $\tpl{\mexpr_\mmarks, \menv, \mstore, \kcons{\kcmsls{\mexpr_k,\menv}^\mmarks}{\mkont}, \mktab, \mmemo}$
     \\
@@
     &
-    $\tpl{(\mexpr_k, \menv), \mstore, \kcons{\kcmslk{\mval}^\mmarks}{\mkont}, \mktab, \mmemo}$
+    $\tpl{\mexpr_k, \menv, \mstore, \kcons{\kcmslk{\mval}^\mmarks}{\mkont}, \mktab, \mmemo}$
     \\
@@
 \\[2pt]
-  \rewritectx((\mexpr,\menv),\mstore,(\mmarks, \mstore')) &= ((\mexpr,\menv),\mstore,(\mmarks\sqcup_{!{}}[\mval_k \mapsto \mval_v], \mstore'))
+  \rewritectx(\mexpr,\menv,\mstore,(\mmarks, \mstore')) &= (\mexpr,\menv,\mstore,(\mmarks\sqcup_{!{}}[\mval_k \mapsto \mval_v], \mstore'))
 \\[2pt]
@@
 % Reset
-    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont}$
-    &
-    $\tpl{(\mexpr, \menv), \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
+    $\tpl{\sreset{\mexpr}, \menv, \mstore, \mkont, \mmkont}$
+    &
+    $\tpl{\mexpr, \menv, \mstore, \kmt, \mkapp{\mkont}{\mmkont}}$
 %    & \textsc{[push prompt]}
@@
 % Shift
-    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont}$
-    &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore',\kmt,\mmkont}$
+    $\tpl{\sshift{\mvar}{\mexpr}, \menv, \mstore, \mkont, \mmkont}$
+    &
+    $\tpl{\mexpr, \extm{\menv}{\mvar}{\maddr}, \mstore',\kmt,\mmkont}$
 %    & \textsc{[capture continuation]}
@@
 \begin{align*}
-  \mctx \in \Context &::= ((\mexpr,\menv),\mstore) \alt (\vcomp{\mkont},\mval,\mstore) \\
+  \mctx \in \Context &::= (\mexpr,\menv,\mstore) \alt (\vcomp{\mkont},\mval,\mstore) \\
   \mkont \in \Kont &= \kmt \alt \krt{\mctx} \alt \kcons{\mkframe}{\mkont} \\
@@
 %
-\paragraph{Problem:} now $\rt$ continuations contain a store, and captured continuations can appear in the store.
+\paragraph{Problem:} now captured continuations can appear in the store, and $\rt$ continuations contain a store.
 %
@@
 % Reset
-    $\tpl{(\sreset{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
+    $\tpl{\sreset{\mexpr}, \menv, \mstore, \mkont, \mmkont, \mktab, \mmemo}$
     &
@@
 % Shift
-    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \joinone{\mstore}{\maddr}{\vcomp{\mkont}},\kmt,\mmkont,\mktab,\mmemo}$
+    $\tpl{\sshift{\mvar}{\mexpr}, \menv, \mstore, \mkont, \mmkont, \mktab, \mmemo}$
+    &
+    $\tpl{\mexpr, \extm{\menv}{\mvar}{\maddr}, \joinone{\mstore}{\maddr}{\vcomp{\mkont}},\kmt,\mmkont,\mktab,\mmemo}$
     \\
@@
     \begin{align*}
-      \msctx \in \SContext &::= ((\mexpr,\menv),\maddr) \\
+      \msctx \in \SContext &::= (\mexpr,\menv,\maddr) \\
       \mctx \in \Context &::= \msctx \alt
-      ((\mexpr,\menv),\mstore,\mmktab) \\ &\alt (\vcomp{\mskont},
+      (\mexpr,\menv,\mstore,\mmktab) \\ &\alt (\vcomp{\mskont},
       \mval, \mstore, \mmktab)
@@
     \hline
-    $\tpl{(\sshift{\mvar}{\mexpr}, \menv), \mstore, \mkont, \mmkont, \mmktab, \mktab, \mmemo}$
-    &
-    $\tpl{(\mexpr, \extm{\menv}{\mvar}{\maddr}), \mstore',\kmt,\mmkont, \mmktab', \mktab,\mmemo}$
+    $\tpl{\sshift{\mvar}{\mexpr}, \menv, \mstore, \mkont, \mmkont, \mmktab, \mktab, \mmemo}$
+    &
+    $\tpl{\mexpr, \extm{\menv}{\mvar}{\maddr}, \mstore',\kmt,\mmkont, \mmktab', \mktab,\mmemo}$
     \\
@@
       \replacectx(\kmt) &= \kmt \\
-      \replacectx(\krt{((\mexpr,\menv),\ldots)}) &= \krt{((\mexpr,\menv),\maddr)} \\
+      \replacectx(\krt{(\mexpr,\menv,\ldots)}) &= \krt{(\mexpr,\menv,\maddr)} \\
       \replacectx(\kcons{\mkframe}{\mkont}) &=
@@
       \addstore(\kmt) &= \mmktab \\
-      \addstore(\krt{((\mexpr,\menv),\mstore,\mmktab')}) &= \joinone{\mmktab}{\maddr}{\mstore}\sqcup \mmktab' \\
-      \addstore(\krt{((\mexpr,\menv),\maddr')}) &= \joinm{\mmktab}{\maddr}{\mmktab(\maddr')} \\
+      \addstore(\krt{(\mexpr,\menv,\mstore,\mmktab')}) &= \joinone{\mmktab}{\maddr}{\mstore}\sqcup \mmktab' \\
+      \addstore(\krt{(\mexpr,\menv,\maddr')}) &= \joinm{\mmktab}{\maddr}{\mmktab(\maddr')} \\
       \addstore(\kcons{\mkframe}{\mkont}) &= \addstore(\mkont)
@@
 \begin{align*}
-  \returns(\mmktab, \mktab, ((\mexpr,\menv), \mstore, \mmktab')) &= \mktab((\mexpr,\menv), \mstore, \mmktab') \\
-  \returns(\mmktab, \mktab, ((\mexpr,\menv), \maddr)) &=
-    \bigcup\setbuild{\mktab((\mexpr,\menv),\mstore,\mmktab')}{(\mstore \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
+  \returns(\mmktab, \mktab, (\mexpr,\menv, \mstore, \mmktab')) &= \mktab(\mexpr,\menv, \mstore, \mmktab') \\
+  \returns(\mmktab, \mktab, (\mexpr,\menv, \maddr)) &=
+    \bigcup\setbuild{\mktab(\mexpr,\menv,\mstore,\mmktab')}{(\mstore \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
 \end{align*}
@@
 \begin{align*}
-  \memoize(\mmktab, \mktab, ((\mexpr,\menv), \mstore', \mmktab'), \mathit{out}) &=
-  [((\mexpr,\menv), \mstore',\mmktab') \mapsto \set{{\mathit{out}}}] \\
-  \memoize(\mmktab, \mktab, ((\mexpr,\menv), \maddr), \mathit{out}) &= [\mctx \mapsto \set{\mathit{out}} : \mctx \in \dom(\mktab) \cap C] \\
-    \text{where } C &= \setbuild{((\mexpr,\menv), \mstore', \mmktab')}{\mstore' \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
+  \memoize(\mmktab, \mktab, (\mexpr,\menv, \mstore', \mmktab'), \mathit{out}) &=
+  [(\mexpr,\menv, \mstore',\mmktab') \mapsto \set{{\mathit{out}}}] \\
+  \memoize(\mmktab, \mktab, (\mexpr,\menv, \maddr), \mathit{out}) &= [\mctx \mapsto \set{\mathit{out}} : \mctx \in \dom(\mktab) \cap C] \\
+    \text{where } C &= \setbuild{(\mexpr,\menv, \mstore', \mmktab')}{\mstore' \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
 \end{align*}
@@
 \begin{theorem}\label{thm:sound-sr}
-  Widened tabular semantics reflects a relation that simulates standard semantics.
+  The widened tabular semantics reflects a relation that \emph{simulates} the shift/reset machine.
 \end{theorem}

### 2014-02-25 metatheory.tex
Fixed formatting. Added to metatheory section. Need to prove the correctness theorem for PDCFA and fix the example(s).

@@
 
-Correctness follows directly from the invariants we prove of $\mktab$ and $\mmemo$ in the tabular semantics.
+Our correctness claim is that a PDCFA system reflects the subset of the CESK reduction semantics that is reachable from the initial state.
 %
-Allocation strategies are comparable if they produce equal addresses regardless of the differences in the state representations.
+We reify the reduction relation from a system in the following way:
+
+\begin{align*}
+  \reify &: \sa{System} \to \wp(\sa{State}_\mathit{CESK} \times \sa{State}_\mathit{CESK}) \\
+  \reify(F, R, \mktab, \mmemo) &=
+  \setbuild{(\tpl{\mpoint,\mstore,\append{\mkont}{\mkont''}},
+             \tpl{\mpoint',\mstore',\append{\mkont'}{\mkont''}})}
+           {\\ &\phantom{= \{}(\tpl{\mpoint,\mstore,\mkont}, \tpl{\mpoint',\mstore',\mkont'}) \in R,
+            \\ &\phantom{= \{}\mkont'' \in \tails(\mkont, \mktab)} \\
+\text{where } \tails(\kmt, \mktab) &= \set{\kmt} \\
+              \tails(\krt{\mctx}, \mktab) &= \setbuild{\mkont' \in \unroll{\mktab}{\mkont}}{\mkont \in \mktab(\mctx)} \\
+              \tails(\kcons{\mkframe}{\mkont}, \mktab) &= \tails(\mkont,\mktab)
+\end{align*}
+
+The $\mathit{unroll}$ relation interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, defined inductively by the following rules:
+\begin{mathpar}
+  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt}} \quad
+  \inferrule{\mkont \in \unroll{\mktab}{\mkont'}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont'}} \\
+  \inferrule{\mkont' \in \mktab(\mctx) \\
+             \mkont \in \unroll{\mktab}{\mkont'}}
+            {\mkont \in \unroll{\mktab}{\krt{\mctx}}}  
+\end{mathpar}
+
+For a pushdown analysis, we can simply take the least fixed-point of ${\mathcal F}$, but in general we can only say a witnessed reduction after $n$ steps in one case will in the other case witness the same reduction in some $m$ steps.
 %
-More formally, $\alloc$ and $\alloc^*$ are comparable if this implication holds:
+In the limit, the CESK reduction relation unfolds to the reification of the PDCFA system, and vice versa.
+%
+One small caveat is that state selection from $F$ must be fair: for all states $\mastate \in F$ there is some $m$ such that $\mastate$ not in $F$ component of ${\mathcal F}(\mastate_0)^m(\tpl{F,R,\mktab,\mmemo})$.
+%
+A queue representation of $F$ suffices to guarantee fairness.
+
+\begin{theorem}[Correctness of Summarization]\label{thm:concrete-tabular}
+  $\forall \mexpr$ closed, let $\mstate_0 = \tpl{\mexpr,\bot\bot,\kmt}$ in
+  $\forall n \in \nat$:
+  \begin{itemize}
+  \item{if $(\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},n,\emptyset)$ then
+      $\exists m \in \nat. (\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^m(\emptyset,\emptyset,\bot,\bot))$}
+  \item{
+      if $(\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^n(\emptyset,\emptyset,\bot,\bot))$ then
+      $\exists m \in \nat. (\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},m,\emptyset)$}
+  \end{itemize}
+  given comparable allocation strategies and fair state selection.
+\end{theorem}
+
+We unfold a reduction relation from a starting state in the following way:
+\begin{align*}
+  \reachrestrict(\mstate_0, \stepto, 0, R) &= R \\
+  \reachrestrict(\mstate_0, \stepto, n+1, R) &= \reachrestrict(\mstate_0, \stepto, n, R') \\
+  \textit{where } R' &= \setbuild{(\mstate_0,\mstate')}{\mstate_0 \stepto \mstate'}
+  \\ &\cup \setbuild{(\mstate,\mstate')}{(\_,\mstate) \in R, \mstate \stepto \mstate'}
+\end{align*}
+
+Allocation strategies $\alloc$ and $\alloc^*$ are comparable if they produce equal addresses regardless of the differences in the state representations:
 \begin{mathpar}
-  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \emptyset}}
+  \inferrule{\mkont \in \unroll{\mktab}{\mkont'}}
             {\alloc(\tpl{\mpoint,\mstore,\mkont}) = \alloc^*(\tpl{\mpoint,\mstore,\mkont',\mktab,\mmemo})}
@@
 
-The $\mathit{unroll}$ function interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, defined as the greatest fixed point of the following rules:
-\begin{mathpar}
-  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt, G}} \quad
-  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \emptyset}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont', G}} \\
-  \inferrule{\mctx \notin G  \\
-             \mkont'' \in \mktab(\mctx) \\
-             \mkont \in \unroll{\mktab}{\mkont'', \set{\mctx}\cup G}}
-            {\mkont \in \unroll{\mktab}{\krt{\mctx}, G}}  
-\end{mathpar}
+We build toward a proof of correctness by giving invariant characterizations of the $\mktab$ and $\mmemo$ tables (\autoref{fig:inv}), which further depend on a ``context irrelevance'' lemma about the CESK machine.
+%
+Our development is partially verified in Coq, available with the Redex models in \texttt{model.v}; propositions marked with ${}^*$ are mechanically checked.
 
-We add $G$ to protect against unguarded corecursion in order for $\mathit{unroll}$ to be well-defined.
-%
-Interpreting a function with ill-founded recursion can lead to a table such that $\krt{\mctx} \in \mktab(\mctx)$.
-%
-The abstractions that we impose on our data might also cause this situation, even though the concrete execution might always terminate.
-
-%%
-The different tables we have encode information about execution history that we prove is invariant (\autoref{fig:inv}).
-%
-The first is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
-%
-The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
-%
-The final invariant, $\phi_{\memo}$, states that all paths starting at a function call either reach the memoized result, or if the path does not include a return, there is an extension that will.
-%
-The portion of these paths between call and return hal the calling context's continuation in the tail to justify using stack-irrelevance.
-
-\begin{figure}
-  \centering
-  \begin{align*}
-    \inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo}) &=
-    \dom(\mmemo) \subseteq \dom(\mktab) \\
-    &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab), \mkont''
-    \in \mktab(\mpoint',\mstore'),
-    \mkont' \in \unroll{\mktab}{\mkont'', \emptyset}. \\
-    &\qquad \phi_{\reachable} \wedge \phi_{\memo} \\
-    \text{where } \phi_{\reachable} &= \exists \mkont'' \in
-    \unroll{\mktab}{\mkont,\emptyset}.
-    \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont''} \\
-    \phi_{\memo} &=
-    \forall (\mval, \mstore'') \in \mmemo(\mpoint',\mstore'). \\
-    &\qquad\exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore'',\mkont'}. \hastail(\mtrace,\mkont')
-  \end{align*}
-  \caption{Table invariants}
-\label{fig:inv}
-\end{figure}
-\begin{lemma}[Table invariants in CESK$\mathit{\Xi{}M}$]\label{lem:tab-inv}
-  If $\inv(\mstate)$ and $\mstate \stepto \mstate'$ then $\inv(\mstate')$.
+\begin{lemma}[Context irrelevance${}^*$]\label{lem:stack-irrelevance}
+  For all CESK traces $\mtrace$ and continuations $\mkont$ such that $\hastail(\mtrace,\mkont)$,
+  for any $\mkont'$, $\replacetail(\mtrace,\mkont,\mkont')$ is also a valid trace.
 \end{lemma}
-\begin{proof}
-  By cases on $\stepto$.
-  \begin{byCases}
-% variable lookup
-    \case{\tpl{(\svar\mvar, \menv), \mstore, \mkont, \mktab, \mmemo}
-          \stepto
-          \tpl{\mval,\mstore,\mkont, \mktab, \mmemo} \text{ if }\mval \in \mstore(\menv(\mvar))}{
-     Same continuation and tables, so holds by $\inv(\mstate)$.}
-% application
-    \case{\tpl{(\sapp{\mexpri0}{\mexpri1}, \menv), \mstore, \mkont, \mktab, \mmemo}
-          \stepto
-          \tpl{(\mexpri0, \menv), \mstore, \kcons{\kar{\mexpri1,\menv}}{\mkont}, \mktab, \mmemo}}{
-    Let $\mctx, \mkont^\circ, \mkont'$ be arbitrary.
-    Let $\mkont^*$ be the witness for $\mstate$.
-    The next witness is $\kcons{\kar{\mexpri1,\menv}}{\mkont^*}$, by definition of $\mathit{unroll}$.}
-% argument evaluation
-    \case{\tpl{\mval, \mstore, \kcons{\kar{\mexpr,\menv}}{\mkont}, \mktab, \mmemo}
-          \stepto
-          \tpl{(\mexpr, \menv), \mstore, \kcons{\kfn{\mval}}{\mkont}, \mktab, \mmemo}}{
-    Let $\mctx, \mkont^\circ, \mkont'$ be arbitrary.
-    Let $\kcons{\kar{\mexpr,\menv}}{\mkont^*}$ be the witness for $\mstate$.
-    The next witness is $\kcons{\kfn{\mval}}{\mkont^*}$, by definition of $\mathit{unroll}$.}
-% function call
-    \case{\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo}
-          \stepto
-          \tpl{\mpoint, \mstore', \krt{\mpoint, \mstore'}, \mktab', \mmemo}}{
-     where
-     $\begin{array}{l}
-       \mpoint = (\mexpr, \extm{\menv}{\mvar}{\maddr}) \\
-       \mstore' = \joinone{\mstore}{\maddr}{\mval} \\
-       \mktab' = \joinone{\mktab}{(\mpoint, \mstore')}{\mkont}
-     \end{array}$ \\
-
-    Let $\mctx, \mkont^\circ, \mkont'$ be arbitrary.
-    Let $\kcons{\kfn{\mval}}{\mkont^*}$ be the witness for $\mstate$.
-    The next witness is $\mkont^*$, by definition of $\mathit{unroll}$.}
-% memo-lookup
-    \case{\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo}
-          \stepto
-          \tpl{\mval_\mathit{result}, \mstore'', \mkont, \mktab', \mmemo}
-          \text{ if } (\mval_\mathit{result},\mstore'') \in \mmemo(\mpoint,\mstore')}{
-        same where clause as previous case.
-
-    Let $\mctx, \mkont^\circ, \mkont'$ be arbitrary.
-    By definition of $\inv$, and since $\tpl{\mpoint',\mstore',\mkont'} \stepto^* \tpl{\mval_{\mathit{result}}, \mstore'', \mkont'}$.
-    For the memo entry to exist, there must have been at least one previous continuation in $\mktab(\mpoint',\mstore')$.
-    We use $\inv$ with this continuation, combined with stack irrelevance to produce the trace involving $\mkont'$.
-    }
-% return
-    \case{\tpl{\mval, \mstore, \krt{\mpoint,\mstore'}, \mktab, \mmemo}
-          \stepto
-          \tpl{\mval, \mstore, \mkont, \mktab, \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}}
-          \text{ if } \mkont \in \mktab(\mpoint, \mstore')}{
-    Let $\mctx, \mkont^\circ, \mkont'$ be arbitrary.
-    $\mkont'$ is an acceptable witness for the first property.
-    The path we construct to $\mstate'$ is sufficient for the second property.}
-  \end{byCases}
-\end{proof}
-
-In order to prove this, we need a lemma that allows us to plug in any continuation for memoized results.
-%
-When traces (sequences of states such that adjacent states are related by $\stepto$) have a common tail, we can replace it with anything.
-
-\begin{lemma}[Stack irrelevance in CESK]\label{lem:irrelevance}
-  $\hastail(\mtrace,\mkont)$ implies $\replacetail(\mtrace,\mkont,\mkont')$ is a valid trace.
-\end{lemma}
-\begin{proof}
-  By induction on $\mtrace$. Base cases trivially hold, and induction step follows from cases on $\mstate \stepto \mstate'$ and definition of $\replacetail$.
-\end{proof}
-
+Where $\hastail$ (``has tail'') is
 \begin{mathpar}
@@
 \end{mathpar}
-
+and $\replacetail$ (``replace tail'') is
 \begin{align*}
@@
 \end{align*}
+Proof by induction on $\mtrace$ and cases on $\stepto$.
 
-\begin{theorem}[Correctness of Concrete Summarization]\label{thm:concrete-tabular}
-  The standard semantics has a skipping bisimulation with the tabular semantics given comparable fresh allocation strategies.
-\end{theorem}
-\begin{proof}[Sketch]
-  The tabular semantics skipping simulates that standard semantics by appealing to the table invariants to show the long ``recomputation'' path.
-  Crucially, since the memo table is only set once before being used to bypass to the result, there must be only one possible result in the standard semantics.
-Otherwise, the tabular semantics would miss out on the other results for subsequent calls with the same context.
-  The fresh allocation strategy ensures that relation is deterministic and thus function calls only reduce to one result.
+The first invariant is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
+%
+Second, there is a path from the starting state of the current continuation to the current state.
+%
+Third, similar paths exist for all continuations stored in $\mktab$.
+%
+Finally, there is some continuation that stays in the tail of the trace of a memoization entry from starting context to final results (so we can interchange continuations with context irrelevance to justify memo table uses).
+%
+%The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
+%
+%The final invariant, $\phi_{\memo}$, states that all paths starting at a function call either reach the memoized result, or if the path does not include a return, there is an extension that will.
+%
+%The portion of these paths between call and return hal the calling context's continuation in the tail to justify using stack-irrelevance.
 
-  The standard semantics skipping simulates the tabular semantics because ``recomputing'' paths always have the calling stack as its tail, so we can step forward on the invariant's inductively built path until the result.
-\end{proof}
+\begin{figure}
+  \centering
+  \begin{align*}
+    \startstate(\kmt) &= \tpl{\mexpr_\mathit{initial}, \bot, \bot, \kmt} \\
+    \startstate(\krt{\mctx}) &= \tpl{\mctx, \kmt} \\    
+    \startstate(\kcons{\mkframe}{\mkont}) &= \startstate(\mkont)
+ \\[2pt]
+    \inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo}) &=
+    \dom(\mmemo) \subseteq \dom(\mktab) \\ % domain containment
+    % path to current kont
+    &\wedge \startstate(\mkont) \stepto^* \tpl{\mpoint, \mstore,\append{\mkont}{\kmt}} \\
+    % ktable meaning
+    &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab), \mkont''
+    \in \mktab(\mpoint',\mstore'). \\
+    &\quad\startstate(\mkont'') \stepto^* \tpl{\mpoint',\mstore',\append{\mkont''}{\kmt}} \\
+    % memo table meaning
+    &\forall(\mpoint',\mstore') \in \dom(\mmemo).
+      \forall (\mval,\mstore'') \in \mmemo(\mpoint',\mstore').
+      \\ & \quad
+       \exists \mkont, \mtrace\equiv\tpl{\mpoint',\mstore',\mkont}\stepto^* \tpl{\mval,\mstore'',\mkont}.
+         \hastail(\mtrace,\mkont)
+  \end{align*}
+  \caption{Table invariants}
+\label{fig:inv}
+\end{figure}
+\begin{lemma}[Table invariants in CESK$\mathit{\Xi{}M}^*$]\label{lem:tab-inv}
+  If $\inv(\mstate)$ and $\mstate \stepto \mstate'$ then $\inv(\mstate')$.
+\end{lemma}
+  By cases on $\stepto$.
 
-Finally, we can show that the reductions we find between states paired with the shared tables define a sound and complete abstraction of the CESK machine with a comparable allocation strategy.
+As the later developments in this paper are of a similar character, we will describe correctness claims as a system reflecting a relation that has a bisimulation with the original semantics.

### 2014-02-25 content.tex
Minor intro cleanup.

@@
 %
-Such an approach is unlikely to capture all ``real language's'' formalizations.
-%
-An additional abstraction also detracts from the operational intuitions and simplicity of the methodology.
+An overly formal approach is unlikely to capture all ``real language's'' formalizations.
+%
+The additional abstraction also detracts from the operational intuitions and simplicity of the methodology.
 

### 2014-02-27 content.tex
Changed to 10pt as per OOPSLA submission guidelines. Changed formatting to fit figures.

@@
 %
-The online quality of the algorithm is what makes it suitable for higher-order languages.
+The online quality of the algorithm is what makes it suitable for higher-order languages, since CFA essentially interprets the program for each discovered state to build the model on the fly.
 %
@@
   \mexpr \in \Expr &= \svar{\mvar} \alt \sapp{\mexpr}{\mexpr} \alt \slam{\mvar}{\mexpr} \\
-  \mstate \in \State &= \Expr \times \Env \times \Store \times \Kont \\
+  \mstate \in \State &::= \tpl{\mexpr,\menv,\mstore,\mkont} \\
   \mval \in \Value &::= \vclo{\slam{\mvar}{\mexpr}}{\menv} \\
@@
 \begin{align*}
-  \mstate \in \State &= \Expr \times \Env \times \Store \times \Kont \times \KTab \times \Memo \\
+  \mstate \in \State &::= \tpl{\mexpr,\menv,\mstore,\mkont,\mktab,\mmemo} \\
   \mkont \in \Kont &::= \kmt \alt \krt{\mexpr,\menv,\mstore} \alt \kcons{\mkframe}{\mkont} \\
@@
 \begin{align*}
-  \mastate \in \sa{State} &= \Expr \times \Env \times \Store \times \Kont \\
+  \mastate \in \sa{State} &::= \tpl{\mexpr,\menv,\mstore,\mkont} \\
   \System &= \wp(\sa{State}) \times \wp(\sa{State} \times \sa{State}) \times \KTab \times \Memo \\
@@
 The least fixed-point of ${\mathcal F}(\tpl{\mexpr,\bot,\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
+%
+Note the ordering on systems is pointwise, so the bottom element is $(\emptyset,\emptyset,\bot,\bot)$, which we'll call $\mathit{sys}_0$.
 %
@@
 \\[4pt]
-  \begin{minipage}{0.5\linewidth}
+  \begin{minipage}{0.55\linewidth}
     $\touchesf(\kar{\mexpr,\menv}) = \mathit{range}(\menv|_{\mathit{fv}(\mexpr)})$
@@
 %  \inferrule{\mkont \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}
-  \quad
+  \\
   \inferrule{\mkont \in \mktab(\mctx)}{\krt{\mctx} \rightsquigarrow_\mktab \mkont}
@@
 
-\begin{align*}
-  \reaches(\mathit{root}, \mstore) = \setbuild{b}{\maddr \in \mathit{root}, \maddr \rightsquigarrow_\mstore^* b}
-   \quad
+\begin{center}
+  $\reaches(\mathit{root}, \mstore) = \setbuild{b}{\maddr \in \mathit{root}, \maddr \rightsquigarrow_\mstore^* b}$
+\end{center}
+\begin{mathpar}
   \inferrule{\mval \in \mstore(\maddr) \\
@@
   {\maddr \rightsquigarrow_\mstore b}
-\end{align*}
+\end{mathpar}
 
@@
     &
-    $\tpl{\mval,\mstore,\mframe',\mkont}$ if $(\mframe', \mval) \in \lookup(\mstore,\mframe,\menv(\mvar),\mlab)$
+    $\tpl{\mval,\mstore,\mframe',\mkont}$
+    \\ & if $(\mframe', \mval) \in \lookup(\mstore,\mframe,\menv(\mvar),\mlab)$
     \\
@@
 %
-Look-up would then always try the stack frame before falling back on the heap (this must be limited to immutable addresses).
+Look-up would then always try the stack frame before falling back on the heap, though this must be limited to immutable addresses.
 %%
@@
 %    & \textsc{[capture continuation]}
-    \\ & where $\mstore' = \joinone{\mstore}{\maddr}{\vcomp{\mkont}}$
+    \\ where & $\mstore' = \joinone{\mstore}{\maddr}{\vcomp{\mkont}}$
     \\
@@
 \begin{align*}
-  \returns(\mmktab, \mktab, (\mexpr,\menv, \mstore, \mmktab')) &= \mktab(\mexpr,\menv, \mstore, \mmktab') \\
-  \returns(\mmktab, \mktab, (\mexpr,\menv, \maddr)) &=
-    \bigcup\setbuild{\mktab(\mexpr,\menv,\mstore,\mmktab')}{(\mstore \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
+  &\returns(\mmktab, \mktab, (\mexpr,\menv, \mstore, \mmktab')) = \mktab(\mexpr,\menv, \mstore, \mmktab') \\
+  &\returns(\mmktab, \mktab, (\mexpr,\menv, \maddr)) =
+    \bigcup\setbuild{\mktab(\mexpr,\menv,\mstore,\mmktab')}{\mstore \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
 \end{align*}
@@
 At returns, the memo table must add the result to all represented contexts:
-\begin{align*}
-  \memoize(\mmktab, \mktab, (\mexpr,\menv, \mstore', \mmktab'), \mathit{out}) &=
-  [(\mexpr,\menv, \mstore',\mmktab') \mapsto \set{{\mathit{out}}}] \\
-  \memoize(\mmktab, \mktab, (\mexpr,\menv, \maddr), \mathit{out}) &= [\mctx \mapsto \set{\mathit{out}} : \mctx \in \dom(\mktab) \cap C] \\
-    \text{where } C &= \setbuild{(\mexpr,\menv, \mstore', \mmktab')}{\mstore' \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
+\newcommand{\res}{m}
+\begin{align*}
+  &\memoize(\mmktab, \mktab, (\mexpr,\menv, \mstore', \mmktab'), \res) =
+  [(\mexpr,\menv, \mstore',\mmktab') \mapsto \set{{\res}}] \\
+  &\memoize(\mmktab, \mktab, (\mexpr,\menv, \maddr), \res) = [\mctx \mapsto \set{\res} : \mctx \in \dom(\mktab) \cap C] \\
+  &\text{where } C = \setbuild{(\mexpr,\menv, \mstore', \mmktab')}{\mstore' \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
 \end{align*}
@@
 
-\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{[}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{z}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{h}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{h}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{f}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{z}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{fv}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{fv}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{i}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{i}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{]}\RktPn{)}
+\mbox{\hphantom{\Scribtexttt{xxxxxxx}}}\RktPn{[}\RktSym{g}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{z}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{h}\RktPn{)}
+\\\mbox{\hphantom{\Scribtexttt{xxxxxxxxxxx}}}\RktPn{(}\RktSym{h}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{f}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{z}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{fv}\RktPn{)}
+\\\mbox{\hphantom{\Scribtexttt{xxxxxxxxxxxxxxxxxxxx}}}\RktPn{(}\RktSym{id}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{fv}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\lambda$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{i}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{i}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{)}\RktPn{]}\RktPn{)}
 

### 2014-02-27 metatheory.tex
Changed to 10pt as per OOPSLA submission guidelines. Changed formatting to fit figures.

@@
-\subsection{Metatheory}
+\subsection{Correctness}
 
@@
 \begin{align*}
-  \reify &: \sa{System} \to \wp(\sa{State}_\mathit{CESK} \times \sa{State}_\mathit{CESK}) \\
+  \reify &: \mathit{System} \to \wp(\mathit{State}_\mathit{CESK} \times \mathit{State}_\mathit{CESK}) \\
   \reify(F, R, \mktab, \mmemo) &=
@@
 
-For a pushdown analysis, we can simply take the least fixed-point of ${\mathcal F}$, but in general we can only say a witnessed reduction after $n$ steps in one case will in the other case witness the same reduction in some $m$ steps.
+For a pushdown analysis, we can simply take the least fixed-point of ${\mathcal F}(\mastate_0)$, but in general we can only say a witnessed reduction after $n$ steps, in one case, will in the other case witness the same reduction in some $m$ steps.
 %
@@
   \item{if $(\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},n,\emptyset)$ then
-      $\exists m \in \nat. (\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^m(\emptyset,\emptyset,\bot,\bot))$}
+      $\exists m \in \nat. (\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^m(\mathit{sys}_0))$}
   \item{
-      if $(\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^n(\emptyset,\emptyset,\bot,\bot))$ then
+      if $(\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^n(\mathit{sys}_0))$ then
       $\exists m \in \nat. (\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},m,\emptyset)$}
@@
   \inferrule{ }{\hastail(\epsilon,\mkont)} \quad
-  \inferrule{ }{\hastail(\tpl{\mpoint,\mstore,\append{\mkont'}{\mkont}},\mkont)} \quad
+  \inferrule{ }{\hastail(\tpl{\mpoint,\mstore,\append{\mkont'}{\mkont}},\mkont)} \\
   \inferrule{\hastail(\mtrace\mstate,\mkont) \quad
@@
     \startstate(\kcons{\mkframe}{\mkont}) &= \startstate(\mkont)
- \\[2pt]
-    \inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo}) &=
+  \end{align*}
+  \begin{mathpar}
+    \inferrule{
         \dom(\mmemo) \subseteq \dom(\mktab) \\ % domain containment
     % path to current kont
-    &\wedge \startstate(\mkont) \stepto^* \tpl{\mpoint, \mstore,\append{\mkont}{\kmt}} \\
+        \startstate(\mkont) \stepto^* \tpl{\mpoint, \mstore,\append{\mkont}{\kmt}} \\
     % ktable meaning
-    &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab), \mkont''
+        \forall (\mpoint',\mstore') \in \dom(\mktab), \mkont'
         \in \mktab(\mpoint',\mstore'). \\
-    &\quad\startstate(\mkont'') \stepto^* \tpl{\mpoint',\mstore',\append{\mkont''}{\kmt}} \\
+        \quad\startstate(\mkont') \stepto^* \tpl{\mpoint',\mstore',\append{\mkont'}{\kmt}} \\
     % memo table meaning
-    &\forall(\mpoint',\mstore') \in \dom(\mmemo).
+      \forall(\mpoint',\mstore') \in \dom(\mmemo).
       \forall (\mval,\mstore'') \in \mmemo(\mpoint',\mstore').
-      \\ & \quad
+      \\ \quad
        \exists \mkont, \mtrace\equiv\tpl{\mpoint',\mstore',\mkont}\stepto^* \tpl{\mval,\mstore'',\mkont}.
          \hastail(\mtrace,\mkont)
-  \end{align*}
+      }
+    {\inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo})}
+  \end{mathpar}
   \caption{Table invariants}
@@
 As the later developments in this paper are of a similar character, we will describe correctness claims as a system reflecting a relation that has a bisimulation with the original semantics.
+%
+The correctness theorem is a corollary of the following stronger statement:
+\begin{lemma}[Strengthened correctness \textbf{TODO}]
+  $\forall \mexpr$ closed, let $\mstate_0 = \tpl{\mexpr,\bot\bot,\kmt}$ in
+  $\forall n \in \nat$:
+  \begin{itemize}
+  \item{if $(\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},n,R)$ then
+      $\exists m \in \nat. (\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^m(\mathit{sys}_0))$}
+  \item{
+      if $(\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^n(\mathit{sys}_0))$ then
+      $\exists m \in \nat. (\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},m,\emptyset)$}
+  \end{itemize}
+  given comparable allocation strategies and fair state selection.
+\end{lemma}

### 2014-03-20 content.tex
Bugfix in system definition. Initial state was added to R' (type error) instead of F'.
Filling out correctness proof in metatheory. Still needs a bit more work (and discussion of order invariants)

@@
 %%
-Testing new ideas in analysis for improving precision or increasing performance can be a difficult venture.
+Testing new ideas in analysis for improving precision or performance can be a difficult venture.
 %
@@
 %
-Higher-order recursion schemes (HORS) are a powerful model for simply typed functional programs with good model-checking properties~\citep{dvanhorn:Ong2006ModelChecking}, but the \textit{program} $\to$ \textit{HORS} process is narrowly applicable.
+Higher-order recursion schemes (HORS) are a powerful model for simply typed functional programs with good model-checking properties~\citep{dvanhorn:Ong2006ModelChecking}, but the \textit{program} $\to$ \textit{HORS} process is still narrowly applicable.
 %
@@
 %%
-As a middle ground, we show a simple way to transform one's programming language semantics into a what can be viewed as a pushdown model constructor.
+As a middle ground, we show a simple way to transform one's programming language semantics into what can be viewed as a pushdown model constructor.
 %%
@@
 %
-The online quality of the algorithm is what makes it suitable for higher-order languages, since CFA essentially interprets the program for each discovered state to build the model on the fly.
+The online quality of the algorithm is what makes it suitable for higher-order languages, since CFA essentially interprets the program for each discovered state to build the model on-the-fly.
 %
@@
 %
-This was shown for the \rackett{call/cc} operator in~\citet{ianjohnson:Vardoulakis2011Pushdown}.
+This was shown for the \rackett{call/cc} operator by~\citet{ianjohnson:Vardoulakis2011Pushdown}.
 %
@@
 %
-Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model like 0-CFA would provide.
+Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, \eg, 0-CFA would provide.
 %%
@@
 %
+Once the machine semantics is in this form, simple point-wise abstraction leads to the summarization algorithms that we see in the literature.
+
+\paragraph{The problem with incomplete memo information}
 Consider a non-deterministic semantics where a function call reaches a result and memoizes, a second call uses the memo table (storing its stack), and then the original call reaches another result (since the worklist allows for such parallel reduction).
@@
 %%
-Once the machine semantics is in this form, simple point-wise abstraction leads to the summarization algorithms that we see in the literature.
-%
+\paragraph{Overview}
 In \autoref{sec:pdcfa} we derive a cousin of PDCFA.
@@
 % return
-    $\tpl{\mval, \mstore, \krt{\mpoint,\mstore'}, \mktab, \mmemo}$
+    $\tpl{\mval, \mstore, \krt{\mctx}, \mktab, \mmemo}$
     &
@@
 %    & \textsc{[memoize/return]}
-    \\ & \quad if $\mkont \in \mktab(\mpoint, \mstore')$
-    \\ where & $\mmemo' = \joinone{\mmemo}{(\mpoint, \mstore')}{(\mval,\mstore)}$
+    \\ & \quad if $\mkont \in \mktab(\mctx)$
+    \\ where & $\mmemo' = \joinone{\mmemo}{\mctx}{(\mval,\mstore)}$
   \end{tabular}
@@
 %
-As it is, the CESK machine simulates the tabular machine, but not the other way around;
+As it is, the CESK machine approximates the tabular machine, but not the other way around;
 %
@@
 \begin{align*}
-    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge \mastate\nequiv\tpl{\mval,\mstore,\krt{\mctx}}} \cup \set{\mastate_0}\\
+    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge \mastate\nequiv\tpl{\mval,\mstore,\krt{\mctx}}} \\
     \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \mktab', \_) \in I} \\
     \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \mmemo') \in I} \\
-    F' &= \setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \setminus (\pi_1 R \cup \pi_2 R)
+    F' &= (\setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \cup \set{\mastate_0})\setminus (\pi_1 R \cup \pi_2 R)
   \end{align*}
@@
     \begin{align*}
-      \frac{\mstate \stepto \mstate'}{\mstate \stepto \Gamma(\mstate')}
+      \mstate \stepto \Gamma(\mstate)
     \end{align*}
@@
 \item{stack allocation for some bindings in an additional $\mframe \in \Store$}
-\item{strong updates on stack frames for resolved addresses}
+\item{strong updates on stack frames for looked-up addresses}
 \end{enumerate}
@@
 \begin{theorem}[Correctness of stack frame refinement]\label{thm:refinement}
-  The CES$\xi$K machine has a stuttering bisimulation with the CESK machine with fresh allocation.
+  The CES$\xi$K machine has a stuttering bisimulation refinement with the CESK machine with fresh allocation.
 \end{theorem}

### 2014-03-20 metatheory.tex
Filling out correctness proof in metatheory. Still needs a bit more work (and discussion of order invariants)

@@
            {\\ &\phantom{= \{}(\tpl{\mpoint,\mstore,\mkont}, \tpl{\mpoint',\mstore',\mkont'}) \in R,
-            \\ &\phantom{= \{}\mkont'' \in \tails(\mkont, \mktab)} \\
-\text{where } \tails(\kmt, \mktab) &= \set{\kmt} \\
-              \tails(\krt{\mctx}, \mktab) &= \setbuild{\mkont' \in \unroll{\mktab}{\mkont}}{\mkont \in \mktab(\mctx)} \\
-              \tails(\kcons{\mkframe}{\mkont}, \mktab) &= \tails(\mkont,\mktab)
+            \\ &\phantom{= \{}\mkont'' \in \tails{\mktab}{\mkont}} \\
+\text{where } \tails{\mktab}{\kmt} &= \set{\kmt} \\
+              \tails{\mktab}{\krt{\mctx}} &= \setbuild{\mkont' \in \unroll{\mktab}{\mkont}}{\mkont \in \mktab(\mctx)} \\
+              \tails{\mktab}{\kcons{\mkframe}{\mkont}} &= \tails{\mktab}{\mkont}
+\\[2pt]
+              \unroll{\mktab}{\mkont} &= \setbuild{\append{\mkont}{\mkont'}}{\mkont' \in \tails{\mktab}{\mkont}}
+\\[2pt]
+              \append{\kmt}{\mkont} &= \append{\krt{\mctx}}{\mkont} = \mkont \\
+              \append{\kcons{\mkframe}{\mkont}}{\mkont'} &= \kcons{\mkframe}{(\append{\mkont}{\mkont'})}
 \end{align*}
 
-The $\mathit{unroll}$ relation interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, defined inductively by the following rules:
-\begin{mathpar}
-  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt}} \quad
-  \inferrule{\mkont \in \unroll{\mktab}{\mkont'}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont'}} \\
-  \inferrule{\mkont' \in \mktab(\mctx) \\
-             \mkont \in \unroll{\mktab}{\mkont'}}
-            {\mkont \in \unroll{\mktab}{\krt{\mctx}}}  
-\end{mathpar}
+The $\mathit{unroll}$ set interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, and $\mathit{tails}$ is defined mutually to produce all continuations that could be unrolled from the tail of a truncated continuation.
 
@@
 %
-One small caveat is that state selection from $F$ must be fair: for all states $\mastate \in F$ there is some $m$ such that $\mastate$ not in $F$ component of ${\mathcal F}(\mastate_0)^m(\tpl{F,R,\mktab,\mmemo})$.
+One small caveat is that state selection from $F$ must be fair: for all states $\mastate \in F$ there is some $m$ such that $\mastate$ not in $F$ component of ${\mathcal F}(\mastate_0)^m(\tpl{F,R,\mktab,\mmemo})$, which means $\mastate$ was selected and processed in one of those $m$ steps.
 %
@@
   $\forall \mexpr$ closed, let $\mstate_0 = \tpl{\mexpr,\bot\bot,\kmt}$ in
-  $\forall n \in \nat$:
+  $\forall n \in \nat,\mstate,\mstate' \in \State_\mathit{CESK}$:
   \begin{itemize}
-  \item{if $(\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},n,\emptyset)$ then
+  \item{if $(\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},n)$ then
       $\exists m \in \nat. (\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^m(\mathit{sys}_0))$}
@@
       if $(\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^n(\mathit{sys}_0))$ then
-      $\exists m \in \nat. (\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},m,\emptyset)$}
+      $\exists m \in \nat. (\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},m)$}
   \end{itemize}
@@
 \begin{align*}
-  \reachrestrict(\mstate_0, \stepto, 0, R) &= R \\
-  \reachrestrict(\mstate_0, \stepto, n+1, R) &= \reachrestrict(\mstate_0, \stepto, n, R') \\
-  \textit{where } R' &= \setbuild{(\mstate_0,\mstate')}{\mstate_0 \stepto \mstate'}
-  \\ &\cup \setbuild{(\mstate,\mstate')}{(\_,\mstate) \in R, \mstate \stepto \mstate'}
+  \reachrestrict(\mstate_0, \stepto, 0) &= \setbuild{(\mstate_0,\mstate)}{\mstate_0 \stepto \mstate} \\
+  \reachrestrict(\mstate_0, \stepto, n+1) &= \stepextend(\reachrestrict(\mstate_0,\stepto,n)) \\
+  \textit{where } \stepextend(R) &= R \cup \setbuild{(\mstate,\mstate')}{(\_,\mstate) \in R, \mstate \stepto \mstate'}
 \end{align*}
@@
 %
-Our development is partially verified in Coq, available with the Redex models in \texttt{model.v}; propositions marked with ${}^*$ are mechanically checked.
+Our development is partially verified in Coq in \texttt{model.v}, available with the Redex models; propositions marked with ${}^*$ are mechanically checked.
 
@@
 \end{align*}
-Proof by induction on $\mtrace$ and cases on $\stepto$.
-
+\begin{proof}
+  By induction on $\mtrace$ and cases on $\stepto$.
+\end{proof}
 The first invariant is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
@@
 \end{lemma}
+\begin{proof}
   By cases on $\stepto$.
-
+\end{proof}
 As the later developments in this paper are of a similar character, we will describe correctness claims as a system reflecting a relation that has a bisimulation with the original semantics.
 %
-The correctness theorem is a corollary of the following stronger statement:
-\begin{lemma}[Strengthened correctness \textbf{TODO}]
-  $\forall \mexpr$ closed, let $\mstate_0 = \tpl{\mexpr,\bot\bot,\kmt}$ in
-  $\forall n \in \nat$:
-  \begin{itemize}
-  \item{if $(\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},n,R)$ then
-      $\exists m \in \nat. (\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^m(\mathit{sys}_0))$}
-  \item{
-      if $(\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^n(\mathit{sys}_0))$ then
-      $\exists m \in \nat. (\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},m,\emptyset)$}
-  \end{itemize}
-  given comparable allocation strategies and fair state selection.
-\end{lemma}
+We argue the correctness theorem by induction on $n$:
+\begin{proof}
+  Base case, first bullet: $m = 2$, since $m = 1$ adds the initial state to the frontier, and in $m = 2$ it is the only choice to process.
+By cases on $\stepto$ the reductions are in the reification.
+\\
+  Base case, second bullet: vacuously true, since the relation is empty. \\
+  Induction step, first bullet: 
+IH: $\forall (\mstate,\mstate') \in \mathit{unfold}(\mstate_0,\stepto,n). \exists m. (\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^m(\mathit{sys}_0))$
+Given $(\mstate,\mstate') \in \mathit{unfold}(\mstate_0,\stepto,n+1) = \mathit{unfold1}(\mathit{unfold}(\mstate_0,\stepto,n))$
+WTS $\exists m. (\mstate,\mstate') \in \mathit{reify}({\mathcal F}(\mstate_0)^m(\mathit{sys}_0))$
+Split on $(\mstate,\mstate') \in \mathit{unfold}(\mstate_0,\stepto,n)$ or $(\mstate,\mstate') \in \setbuild{(\mstate,\mstate')}{(_,\mstate) \in \mathit{unfold}(\mstate_0,\stepto,n), \mstate \stepto \mstate'}$
+\paragraph{Case 1:} Holds by IH.
+\paragraph{Case 2:} there is some $\mstate''$ such that $(\mstate'',\mstate) \in \mathit{unfold}(\mstate_0,\stepto,n)$.
+        By IH, there is an m such that $(\mstate'',\mstate) \in \mathit{reify}({\mathcal F}(\mstate_0)^m(\mathit{sys}_0))$
+        Let $\mktab$ be the ktable component of ${\mathcal F}(\mstate_0)^m(\mathit{sys}_0)$.
+        So $(\mstate'',\mstate)$ is of the form $(\tpl{p, \mstore, \append{\mkont}{\mkont''}},\tpl{p',\mstore',\append{\mkont'}{\mkont''}})$ where $\mkont'' \in \tails{\Xi}{\mkont}$.
+   Let $\mathit{sys} = {\mathcal F}(\mstate_0)^m(\mathit{sys}_0)$
+   By cases on $\tpl{p',\mstore',\mkont'} \in \mathit{sys}.F$.
+\\
+    Case in: let $\ell$ be the fairness witness. We follow the same argument for $\mathit{sys}={\mathcal F}(\mstate_0)^{m+\ell}$ as
+              the following case.
+\\
+    Case not in: it must be that $\tpl{p',\mstore',\mkont'}$ was selected and ``processed.''
+                 By cases on $\mstate \stepto \mstate'$.
+\\
+        Case $p'=(\mvar,\menv)$:
+        $\mstate'$ of the form $\tpl{\mval,\mstore',\append{\mkont'}{\mkont''}}$ where $\mval \in \mstore(\menv(\mvar))$
+         Thus $(\mstate,\mstate') \in \mathit{reify}(\mathit{sys})$
+\\
+        Case $p'=((\mexpri0\ \mexpri1),\menv)$:
+        $\mstate'$ of the form $\tpl{(\mexpri0,\menv),\mstore',\append{\mkont'}{\mkont''}}$
+         Thus $(\mstate,\mstate') \in \mathit{reify}(\mathit{sys})$
+\\
+        Case $p'=v, \append{\mkont'}{\mkont''}=\kcons{\mkframe}{\mkont^*}$ (that is, $\mkont' = \kcons{\mkframe}{\mkont^{**}} and \append{\mkont^{**}}{\mkont''}=\mkont^*$ or $\mkont'=\krt{\mctx}$ and $\mkont''=\kcons{\mkframe}{\mkont^*}$)
+         For $\mkont'$ case 2: $\mkont''$ must be unrolled from $\krt{\mctx} in \mktab$.
+                         there is a $\mkont'{}^* \in \mktab(\mctx)$ such that $\mkont'' = \mkont'$
+\\
+         Case $\mkframe=\kar{\mexpr,\menv}$:
+          $\mstate'$ of the form  $\tpl{(\mexpr,\menv),\mstore',\mkont^*}$
+          For $\mkont'$ case 1: $(\tpl{p',\mstore',\mkont'}, \tpl{(\mexpr,\menv),\mstore',\kcons{\kfn{\mval}}{\mkont^{**}}}) \in \mathit{sys}.R$, so $(\mstate,\mstate') \in \mathit{reify}(\mathit{sys})$
+          Case 2: $\mkont'{}^* = \kcons{\kar{\mexpr,\menv}}{\mkont'{}^{**}}$ and
+          $(\tpl{p',\mstore',\mkont'{}^*},\tpl{(\mexpr,\menv),\mstore',\kcons{\kfn{\mval}}{\mkont'{}^{**}}}) \in sys'.R$
+\\
+         Case $\mkframe=\kfn{\vclo{\slam{x}{\mexpr}}{\menv}}$
+          Similar.
+\\
+%%
+
+%%
+  Induction step, second bullet:
+Let $\mathit{sys} = {\mathcal F}(\mstate0)^n(\mathit{sys}_0)$.
+Let $\mathit{sys}' = {\mathcal F}(\mathit{sys})$
+IH: $\forall (\mstate,\mstate') \in \mathit{reify}(\mathit{sys}). \exists m. (\mstate,\mstate') \in \mathit{unfold}(\mstate_0,\stepto,m)$
+Given $(\mstate,\mstate') \in \mathit{reify}(\mathit{sys}')$
+WTS $\exists m. (\mstate,\mstate') \in \mathit{unfold}(\mstate_0,\stepto,m)$.
+
+We know $(\mstate,\mstate')$ of form $(\tpl{p,\mstore,\append{\mkont}{\mkont''}},\tpl{p',\mstore',\append{\mkont'}{\mkont''}})$ where
+$(\tpl{p,\mstore,\mkont},\tpl{p',\mstore',\mkont'}) \in \mathit{sys}'.R$ and $\mkont'' \in \tails{\mathit{sys}'.\mktab}{\mkont}$.
+Split on edge in $\mathit{sys}.R$ or in $\mathit{sys}'.R$ (and not in $\mathit{sys}.R$).
+\\
+Case in: $\mkont''$ is irrelevant because it's in the tail, and there must be some continuation in $\tails{\mathit{sys}.\mktab}{\mkont}$ by $\inv$.
+         Thus holds by IH.
+\\
+Case not in:
+  Must be that $\tpl{p,\mstore,\mkont}$ was the selected state, and that
+  $\tpl{p,\mstore,\mkont,\mathit{sys}.\mktab,\mathit{sys}.\mmemo} \stepto \tpl{p',\mstore',\mkont',\mktab',\mmemo'}$ where $\mktab' \sqsubseteq \mathit{sys}'.\mktab$ and $\mmemo' \sqsubseteq \mathit{sys}'.\mmemo$.
+  Split on $\tpl{p,\mstore,\mkont}$ the initial state or
+           there is a $\tpl{p^*,\mstore^*,\mkont^*}$ such that $(\tpl{p^*,\mstore^*,\mkont^*}, \tpl{p,\mstore,\mkont}) \in \mathit{sys}.R$.
+\\
+  Case initial: $m$ is 0.
+\\
+  Case else:
+   Need to show there is a path using $\mkont''$ whereever $\tpl{p,\mstore,\mkont}$ is, and the length should be an upper bound for $m$.
+   This brings in the table invariants and the order invariant which allows us to get to the starting point for 
+   $\tpl{p^*,\mstore^*,\mkont^*}$ and replace with the latest frame to get to where we are.
+\end{proof}

### 2014-03-24 content.tex
Added more to the argument of correctness for delimited control, plus a discussion of similarity with CFA2's approach for call/cc.

@@
 The additional abstraction also detracts from the operational intuitions and simplicity of the methodology.
+%
+We still present the definitions and proofs in a manner that should be well-positioned for mechanization.
+%
+This paper has been partially mechanized in the Coq proof assistant, with proofs in the supplemental materials\footnote{\url{http://github.com/ianj/concrete-summaries} in \texttt{model.v}}.
 
@@
 %
-All of the semantics modeled in this paper are implemented in full detail in PLT Redex~\citep{ianjohnson:Felleisen:2009:SEP:1795772} and available online.\footnote{\url{http://github.com/ianj/concrete-summaries}}
+All of the semantics modeled in this paper are implemented in full detail in PLT Redex~\citep{ianjohnson:Felleisen:2009:SEP:1795772} and available in the supplemental materials.
 
@@
 %
+We separate contexts stored in meta-continuations from those in continuations to avoid losing precision from abstracting continuations' contexts.
+%
+The table for continuations is treated the same way as in the previous semantics for standard function call and return.
+%
 The new semantic spaces are then
@@
   \mathit{KRange} &= \wp(\Kont \times \MKont) \cup \wp(\Kont) \\
-  \mktab \in \KTab &= \Context \to \mathit{KRange} \\
+  (\mktab_\mkont,\mktab_\mmkont) \in \KTab &= (\Context \to \wp(\Kont)) \\
+                                        &\times (\Context \to \wp(\Kont \times \MKont)) \\
   \mmemo \in \Memo &= \Context \to \wp(\Value \times \Store)
@@
 % Reset
-    $\tpl{\sreset{\mexpr}, \menv, \mstore, \mkont, \mmkont, \mktab, \mmemo}$
-    &
-    $\tpl{\mpoint, \mstore, \kmt, \kprompt{\mctx}, \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}, \mmemo}$
+    $\tpl{\sreset{\mexpr}, \menv, \mstore, \mkont, \mmkont, (\mktab_\mkont,\mktab_\mmkont), \mmemo}$
+    &
+    $\tpl{\mpoint, \mstore, \kmt, \kprompt{\mctx}, (\mktab_\mkont,\joinone{\mktab_\mmkont}{\mctx}{(\mkont,\mmkont)}), \mmemo}$
     \\
@@
 % Pop prompt
-    $\tpl{\mval, \mstore, \kmt, \kprompt{\mctx}, \mktab, \mmemo}$
+    $\tpl{\mval, \mstore, \kmt, \kprompt{\mctx}, \mktab\equiv(\mktab_\mkont,\mktab_\mmkont), \mmemo}$
     &
     $\tpl{\mval, \mstore, {\mkont}, {\mmkont}, \mktab, \joinone{\mmemo}{\mctx}{(\mval,\mstore)}}$
-    if $(\mkont,\mmkont) \in \mktab(\mctx)$
+    if $(\mkont,\mmkont) \in \mktab_\mkont(\mctx)$
     \\
@@
 % continuation call
-    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont, \mktab, \mmemo}$
-    &
-    $\tpl{\mval, \mstore, \mkont', \kprompt{\mctx}, \mktab',\mmemo}$
+    $\tpl{\mval,\mstore,\kcons{\kfn{\vcomp{\mkont'}}}{\mkont}, \mmkont, (\mktab_\mkont,\mktab_\mmkont), \mmemo}$
+    &
+    $\tpl{\mval, \mstore, \mkont', \kprompt{\mctx}, (\mktab_\mkont,\mktab_\mmkont'),\mmemo}$
     \\ & \quad if $\mctx \notin \dom(\mmemo)$
@@
     & or \\
-    & $\tpl{\mval', \mstore', \mkont, \mmkont, \mktab',\mmemo}$
+    & $\tpl{\mval', \mstore', \mkont, \mmkont, (\mktab_\mkont,\mktab_\mmkont'),\mmemo}$
     \\ & \quad if $(\mval',\mstore') \in \mmemo(\mctx)$
@@
     where & $\mctx = (\vcomp{\mkont'}, \mval, \mstore)$ \\
-          & $\mktab' = \joinone{\mktab}{\mctx}{(\mkont,\mmkont)}$
+          & $\mktab_\mmkont' = \joinone{\mktab_\mmkont}{\mctx}{(\mkont,\mmkont)}$
     \\
@@
 % Return
-   $\tpl{\mval, \mstore, \krt{\mctx}, \mmkont, \mmktab, \mktab, \mmemo}$
+   $\tpl{\mval, \mstore, \krt{\mctx}, \mmkont, \mmktab, \mktab\equiv(\mktab_\mkont,\mktab_\mmkont), \mmemo}$
    &
    $\tpl{\mval, \mstore, \mkont, \mmkont, \mmktab, \mktab, \mmemo'}$
-   if $\mkont \in \returns(\mmktab, \mktab, \mctx)$
-   \\
-   where & $\mmemo' = \mmemo \sqcup \memoize(\mmktab, \mktab, \mctx, (\mval, \mstore, \mmktab))$
+   if $\mkont \in \returns(\mmktab, \mktab_\mkont, \mctx)$
+   \\
+   where & $\mmemo' = \mmemo \sqcup \memoize(\mmktab, \mktab_\mkont, \mctx, (\mval, \mstore, \mmktab))$
   \end{tabular}
@@
 \begin{align*}
-  &\returns(\mmktab, \mktab, (\mexpr,\menv, \mstore, \mmktab')) = \mktab(\mexpr,\menv, \mstore, \mmktab') \\
-  &\returns(\mmktab, \mktab, (\mexpr,\menv, \maddr)) =
-    \bigcup\setbuild{\mktab(\mexpr,\menv,\mstore,\mmktab')}{\mstore \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
+  &\returns(\mmktab, \mktab_\mkont, (\mexpr,\menv, \mstore, \mmktab')) = \mktab_\mkont(\mexpr,\menv, \mstore, \mmktab') \\
+  &\returns(\mmktab, \mktab_\mkont, (\mexpr,\menv, \maddr)) =
+    \bigcup\setbuild{\mktab_\mkont(\mexpr,\menv,\mstore,\mmktab')}{\mstore \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
 \end{align*}
@@
 \begin{align*}
-  &\memoize(\mmktab, \mktab, (\mexpr,\menv, \mstore', \mmktab'), \res) =
+  &\memoize(\mmktab, \mktab_\mkont, (\mexpr,\menv, \mstore', \mmktab'), \res) =
   [(\mexpr,\menv, \mstore',\mmktab') \mapsto \set{{\res}}] \\
-  &\memoize(\mmktab, \mktab, (\mexpr,\menv, \maddr), \res) = [\mctx \mapsto \set{\res} : \mctx \in \dom(\mktab) \cap C] \\
+  &\memoize(\mmktab, \mktab_\mkont, (\mexpr,\menv, \maddr), \res) = [\mctx \mapsto \set{\res} : \mctx \in \dom(\mktab_\mkont) \cap C] \\
   &\text{where } C = \setbuild{(\mexpr,\menv, \mstore', \mmktab')}{\mstore' \in \mmktab(\maddr), \mmktab' \sqsubseteq \mmktab}
@@
 \end{theorem}
+
+This theorem amounts to just the first bullet of \autoref{thm:concrete-tabular}.
+%
+The definition of ${\mathcal F}$ is similar in that only $\mktab$ and $\mmemo$ are widened, but states in $R$ now also contain $\mmkont$s and $\mmktab$s.
+%
+The definition of $\reify$ must change to account for the different target state space, the new context form and the continuation closure:
+\\
+
+  $\reify(F, R, \mktab, \mmemo) =$
+\begin{align*}
+  &\quad \{ (\tpl{\mpoint,\mstore,\mkont_0,\mmkont_0},
+              \tpl{\mpoint',\mstore',\mkont_1,\mmkont_1})\ :\ 
+\\ &\qquad E\equiv(\tpl{\mpoint,\mstore,\mkont,\mmkont,\mmktab}, \tpl{\mpoint',\mstore',\mkont',\mmkont',\mmktab'}) \in R\text,
+\\ &\qquad((\mkont_0,\mmkont_0),(\mkont_1,\mmkont_1)) \in \mathit{conts}_\mktab^\mmktab(E) \}
+\end{align*}
+
+The $\mathit{conts}$ function needs to inspect the reduction actually used to know if the meta-continuation is pushed, popped, or stays the same.
+%
+We'll classify the different reductions as such in the in the definition of $\mathit{conts}$ to not obfuscate the intended meaning.
+%
+We'll use $(\mkont,\kprompt{\mctx}) \overset{+}{\to} (\mkont',\kprompt{\mctx'})$ for a pushed continuation from a meta-continuation $\kprompt{\mctx}$ to $\kprompt{\mctx'}$, which happens in the \texttt{reset} and continuation invocation rules.
+%
+Similarly, the prompt-popping rule is marked as $(\kmt,\kprompt{\mctx}) \overset{-}{\to} (\mkont,\kprompt{\mctx'})$, since it replaces the $\kmt$ with some $\mkont$ and changes the prompt's context.
+%
+We don't need to inspect the second context since its unrolling will include the pushed/popped continuations.
+
+\begin{center}
+  \begin{tabular}{l}
+    $\mathit{conts}_\mktab^\mmktab((\mkont,\mmkont) \overset{+}{\to} (\mkont',\mmkont')) =$ \\
+    \quad$\setbuildb{((\mkont_0,\mmkont^*),(\mkont_1,\mkapp{\mkont_0}{\mmkont^*}))}
+    {
+      \begin{array}{l}
+        \mkont_0 \in \unrollK{\mmktab}{\mktab_\mkont}{\mkont}, \\
+        \mkont_1 \in \unrollK{\mmktab}{\mktab_\mkont}{\mkont'}, \\
+        \mmkont^* \in \unrollC{\mktab_\mmkont}{\mmkont}
+      \end{array}}$
+    \\
+%
+    $\mathit{conts}_\mktab^\mmktab((\mkont,\mmkont) \to (\mkont',\mmkont)) =$ \\
+    \quad$\setbuildb{((\append{\mkont}{\mkont''},\mmkont), (\append{\mkont'}{\mkont''}, \mmkont))}
+    {\begin{array}{l}
+        \mkont'' \in \tails{\mktab_\mkont}{\mkont}, \\
+        \mmkont \in \unrollC{\mktab}{\mmkont}
+      \end{array}}$
+    \\
+%
+    $\mathit{conts}_\mktab^\mmktab((\kmt,\kprompt{\mctx}) \overset{-}{\to} (\mkont,\kprompt{\mctx'})) =$ \\
+    \quad$\setbuildb{((\kmt,\mkapp{\mkont'}{\mmkont'}),(\mkont',\mmkont'))}{
+    \begin{array}{l}
+      (\mkont,\mmkont) \in \mktab_\mmkont(\mctx), \\
+      \mkont' \in \unrollK{\mmktab}{\mktab_\mkont}{\mkont}, \\
+      \mmkont' \in \unrollC{\mktab_\mmkont}{\mmkont}
+    \end{array}}$
+  \end{tabular}
+\end{center}
+
+\begin{align*}
+  \unrollK{\mmktab}{\mktab_\mkont}{\mkont} &= \setbuild{\append{\mkont}{\mkont'}}{\mkont' \in \tailscont{\mktab_\mkont}{\mmktab}{\mkont}}
+\\[2pt]
+  \tailscont{\mktab_\mkont}{\mmktab}{\mkont}(\kmt) &= \set{\kmt} \\
+  \tailscont{\mktab_\mkont}{\mmktab}{\mkont}(\kcons{\mkframe}{\mkont}) &= \tailscont{\mktab_\mkont}{\mmktab}{\mkont} \\
+  \tailscont{\mktab_\mkont}{\mmktab}{\mkont}(\krt{\mctx}) &= \bigcup\limits_{\mkont \in \returns(\mmktab,\mktab_\mkont,\mctx)}{\unrollK{\mmktab}{\mktab_\mkont}{\mkont}}
+\\[2pt]
+  \unrollC{\mktab}{\kmt} &= \set{\kmt} \\
+  \unrollC{(\mktab_\mkont,\mktab_\mmkont)}{\kprompt{\mctx}} &=
+    \setbuild{\mkapp{\mkont'}{\mmkont'}}{(\mkont,\mmkont) \in \mktab_\mmkont(\mctx),
+      \\&\phantom{= \{}\quad \mkont' \in \unrollK{\mmktab}{\mktab_\mkont}{\mkont},
+      \\&\phantom{= \{}\quad \mmkont' \in \unrollC{\mktab}{\mmkont}}
+\end{align*}
 
@@
 
+\paragraph{Comparison to CFA2 and \texttt{call/cc}}
+Although we make no formal comparison of the constructions in this paper that those of CFA2, we can point out similarities in the approaches.
+%
+We claim that what we showed in this section marks a strict generalization of the methods in \citet{dvanhorn:Vardoulakis2011Pushdown}.
+%
+CFA2 makes a specific choice of parameters for context-sensitivity and heap abstraction.
+%
+If we specialize this work to a monovariant allocation strategy and a global store widening, the utility of a continuation closure $\mmktab$ becomes moot, since the store is monotonically increasing.
+%
+In other words, any distinction between contexts that we might hope to achieve given a $\mmktab$ will fail since the $\mmktab' \sqsubseteq \mmktab$ proposition in the $\returns$ meta-function will always be true.
+%
+With this realization, we can look closer at what happens when invoking a continuation.
+%
+Once execution reaches the $\krt{\mctx}$ tail of an installed captured continuation, $\returns$ will ensure flow continues to absolutely every context that captured that continuation.
+%
+For the readers making the comparison themselves, lines 28 and 29 of CFA2's workset algorithm embody this flow explosion rule.
+%
+To paraphrase, whenever execution reaches the invocation of a captured continuation, every call of the function that captures the continuation is marked as returning with the value passed to the continuation (remember, non-compositional continuations act like returns and not calls).
+%
+Since we semantically distinguish captured continuations from non-captured, we don't need any judgement about the heap-allocation of certain bindings that CFA2 uses.
+
 \section{Related Work}

### 2014-03-24 metatheory.tex
Added more to the argument of correctness for delimited control, plus a discussion of similarity with CFA2's approach for call/cc.

--- 2014-02-27 metatheory.tex
@@
 %
-One small caveat is that state selection from $F$ must be fair: for all states $\mastate \in F$ there is some $m$ such that $\mastate$ not in $F$ component of ${\mathcal F}(\mastate_0)^m(\tpl{F,R,\mktab,\mmemo})$.
+One small caveat is that state selection from $F$ must be fair: for all states $\mastate \in F$ there is some $m \in \nat$ such that $\mastate$ not in $F$ component of ${\mathcal F}(\mastate_0)^m(F,R,\mktab,\mmemo)$, which means $\mastate$ was selected and processed in one of those $m$ steps.
 %
@@
 \begin{align*}
-  \reachrestrict(\mstate_0, \stepto, 0, R) &= R \\
-  \reachrestrict(\mstate_0, \stepto, n+1, R) &= \reachrestrict(\mstate_0, \stepto, n, R') \\
-  \textit{where } R' &= \setbuild{(\mstate_0,\mstate')}{\mstate_0 \stepto \mstate'}
-  \\ &\cup \setbuild{(\mstate,\mstate')}{(\_,\mstate) \in R, \mstate \stepto \mstate'}
+  \reachrestrict(\mstate_0, \stepto, 0) &= \setbuild{(\mstate_0,\mstate)}{\mstate_0 \stepto \mstate} \\
+  \reachrestrict(\mstate_0, \stepto, n+1) &= \stepextend(\reachrestrict(\mstate_0,\stepto,n)) \\
+  \textit{where } \stepextend(R) &= R \cup \setbuild{(\mstate,\mstate')}{(\_,\mstate) \in R, \mstate \stepto \mstate'}
 \end{align*}
+For the rest of this section, we will leave the $\mstate_0$ and $\stepto_\mathit{CESK}$ arguments to $\mathit{unfold}$ implicit.
 
@@
 \end{align*}
-Proof by induction on $\mtrace$ and cases on $\stepto$.
-
+\begin{proof}
+  By induction on $\mtrace$ and cases on $\stepto$.
+\end{proof}
 The first invariant is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
 %
-Second, there is a path from the starting state of the current continuation to the current state.
+Second, the continuations that weave through $\mktab$ have successively larger contexts, and of course the contexts in continuations are also mapped in $\mktab$.
 %
-Third, similar paths exist for all continuations stored in $\mktab$.
+Additionally, the current continuation will be able to satisfy the order property if it is stored in $\mktab$.
+%
+Third, there is a path from the starting state of the current continuation to the current state.
+%
+Fourth, similar paths exist for all continuations stored in $\mktab$.
 %
 Finally, there is some continuation that stays in the tail of the trace of a memoization entry from starting context to final results (so we can interchange continuations with context irrelevance to justify memo table uses).
+%
+The properties that append $\kmt$ ensure that the continuations are proper CESK continuations.
+We could quantify over all continuations instead of using $\kmt$, but the context irrelevance lemma makes that unnecessary.
+The third and fourth properties allow us to build up paths piecemeal as we need them.
 %
@@
   \begin{mathpar}
+    \inferrule{ }{\klectx{\mktab}{\kmt}{\mctx}} \quad
+    \inferrule{\mstore \sqsubseteq \mstore' \quad (\mpoint,\mstore) \in \dom(\mktab)}{\klectx{\mktab}{\krt{\mpoint,\mstore}}{(\mpoint,\mstore')}} \quad
+    \inferrule{\klectx{\mktab}{\mkont}{\mctx}}{\klectx{\mktab}{\kcons{\mkframe}{\mkont}}{\mctx}}
+\\
     \inferrule{
         \dom(\mmemo) \subseteq \dom(\mktab) \\ % domain containment
+    % order in \Xi.
+        \forall \mctx \in \dom(\mktab), \mkont' \in \mktab(\mctx). \klectx{\mktab}{\mkont'}{\mctx} \quad
+    % order on current cont
+        \klectx{\mktab}{\mkont}{(\mpoint,\mstore)} \\
     % path to current kont
@@
     % memo table meaning
-      \forall(\mpoint',\mstore') \in \dom(\mmemo).
-      \forall (\mval,\mstore'') \in \mmemo(\mpoint',\mstore').
+      \forall(\mpoint',\mstore') \in \dom(\mmemo), (\mval,\mstore'') \in \mmemo(\mpoint',\mstore').
       \\ \quad
@@
 \end{lemma}
+\begin{proof}
   By cases on $\stepto$.
-
+\end{proof}
 As the later developments in this paper are of a similar character, we will describe correctness claims as a system reflecting a relation that has a bisimulation with the original semantics.
 %
-The correctness theorem is a corollary of the following stronger statement:
-\begin{lemma}[Strengthened correctness \textbf{TODO}]
-  $\forall \mexpr$ closed, let $\mstate_0 = \tpl{\mexpr,\bot\bot,\kmt}$ in
-  $\forall n \in \nat$:
-  \begin{itemize}
-  \item{if $(\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},n,R)$ then
-      $\exists m \in \nat. (\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^m(\mathit{sys}_0))$}
-  \item{
-      if $(\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^n(\mathit{sys}_0))$ then
-      $\exists m \in \nat. (\mstate,\mstate') \in \reachrestrict(\mstate_0,\stepto_\mathit{CESK},m,\emptyset)$}
-  \end{itemize}
-  given comparable allocation strategies and fair state selection.
-\end{lemma}
+We argue the correctness theorem by induction on $n$:
+\begin{proof}
+  Base case, first bullet: $m = 2$, since $m = 1$ adds the initial state to the frontier, and in $m = 2$ it is the only choice to process.
+By cases on $\stepto$ the reductions are in the reification.
+\\
+  Base case, second bullet: vacuously true, since the relation is empty. \\
+  Induction step, first bullet: 
+%IH: $\forall (\mstate,\mstate') \in \mathit{unfold}(\mstate_0,\stepto,n). \exists m. (\mstate,\mstate') \in \reify({\mathcal F}(\mstate_0)^m(\mathit{sys}_0))$
+given $(\mstate,\mstate') \in \mathit{unfold}(n+1) = \mathit{unfold}_1(\mathit{unfold}(n))$,
+by cases on $(\mstate,\mstate') \in \mathit{unfold}(n)$ or $(\mstate,\mstate') \in \setbuild{(\mstate,\mstate')}{(\_,\mstate) \in \mathit{unfold}(n), \mstate \stepto \mstate'}$:
+\\
+Case 1: Holds by IH.
+\\
+Case 2: there is some $\mstate''$ such that $(\mstate'',\mstate) \in \mathit{unfold}(n)$.
+        By IH, there is an $m$ such that (let $\mathit{sys} = {\mathcal F}(\mstate_0)^m(\mathit{sys}_0)$) $(\mstate'',\mstate) \in \mathit{reify}(\mathit{sys})$.
+        So $(\mstate'',\mstate) \equiv (\tpl{\mpoint, \mstore, \append{\mkont}{\mkont''}},\tpl{\mpoint',\mstore',\append{\mkont'}{\mkont''}})$ where $\mkont'' \in \tails{\mktab}{\mkont}$.
+
+   \noindent By cases on $\tpl{\mpoint',\mstore',\mkont'} \in \mathit{sys}.F$:
+\\
+    Case in: let $\ell$ be the fairness witness. We follow the same argument for $\mathit{sys}={\mathcal F}(\mstate_0)^{m+\ell}(\mathit{sys}_0)$ as
+              the following case.
+\\
+    Case not in: it must be that $\tpl{\mpoint',\mstore',\mkont'}$ was selected and ``processed.''
+                 By cases on $\mstate \stepto \mstate'$:
+\\
+        Case $\tpl{(\mvar,\menv),\mstore',\append{\mkont'}{\mkont''}} \stepto \tpl{\mval,\mstore',\append{\mkont'}{\mkont''}}$ where $\mval \in \mstore(\menv(\mvar))$:
+        $\mkont''$ is irrelevant, thus $(\mstate,\mstate') \in \mathit{reify}(\mathit{sys})$
+\\
+        Case $\tpl{((\mexpri0\ \mexpri1),\menv),\mstore',\append{\mkont'}{\mkont''}} \stepto \tpl{(\mexpri0, \menv),\mstore',\append{\mkont'}{\mkont''}}$,
+        thus $(\mstate,\mstate') \in \mathit{reify}(\mathit{sys})$
+\\
+        Case $p'\equiv\mval, \append{\mkont'}{\mkont''}\equiv\kcons{\mkframe}{\mkont^*}$ (that is, $\mkont' \equiv \kcons{\mkframe}{\mkont^{**}}$ and $\append{\mkont^{**}}{\mkont''}=\mkont^*$ or $\mkont'\equiv\krt{\mctx}$ and $\mkont''=\kcons{\mkframe}{\mkont^*}$)
+         For $\mkont'$ case 2: $\mkont''$ must be unrolled from $\krt{\mctx} \in \mathit{sys}.\mktab$.
+                         Thus there is a $\mkont'{}^* \in \mathit{sys}.\mktab(\mctx)$ such that $\mkont'' = \mkont'$
+\\
+         Case $\mkframe\equiv\kar{\mexpr,\menv}$, $\mstate' \equiv \tpl{(\mexpr,\menv),\mstore',\mkont^*}$:
+          For $\mkont'$ case 1: $(\tpl{\mpoint',\mstore',\mkont'}, \tpl{(\mexpr,\menv),\mstore',\kcons{\kfn{\mval}}{\mkont^{**}}}) \in \mathit{sys}.R$, so $(\mstate,\mstate') \in \mathit{reify}(\mathit{sys})$.
+\\
+          Case 2: $\mkont'{}^* \equiv \kcons{\kar{\mexpr,\menv}}{\mkont'{}^{**}}$ and
+          \\ $(\tpl{\mpoint',\mstore',\mkont'{}^*},\tpl{(\mexpr,\menv),\mstore',\kcons{\kfn{\mval}}{\mkont'{}^{**}}}) \in \mathit{sys'}.R$
+\\
+         Case $\mkframe\equiv\kfn{\vclo{\slam{x}{\mexpr}}{\menv}}$
+          Similar.
+\\
+%%
+
+%%
+\noindent  Induction step, second bullet:
+Let $\mathit{sys} = {\mathcal F}(\mstate_0)^n(\mathit{sys}_0)$.
+Let $\mathit{sys}' = {\mathcal F}(\mathit{sys})$
+%IH: $\forall (\mstate,\mstate') \in \mathit{reify}(\mathit{sys}). \exists m. (\mstate,\mstate') \in \mathit{unfold}(\mstate_0,\stepto,m)$
+Given $(\mstate,\mstate') \in \mathit{reify}(\mathit{sys}')$,
+%WTS $\exists m. (\mstate,\mstate') \in \mathit{unfold}(\mstate_0,\stepto,m)$.
+%
+we know $(\mstate,\mstate') \equiv (\tpl{\mpoint,\mstore,\append{\mkont}{\mkont''}},\tpl{\mpoint',\mstore',\append{\mkont'}{\mkont''}})$ where
+\\$(\tpl{\mpoint,\mstore,\mkont},\tpl{\mpoint',\mstore',\mkont'}) \in \mathit{sys}'.R$, if $p \equiv \mval$ then $\mkont \nequiv \krt{\mctx}$ and $\mkont'' \in \tails{\mathit{sys}'.\mktab}{\mkont}$.
+
+\noindent By cases on $\mstate$ initial:
+Case $\mstate = \mstate_0$: a sufficient $m$ is 0.
+\\
+Case else: it must be that there is a $(\tpl{\mpoint^*,\mstore^*,\mkont^*},\tpl{\mpoint,\mstore,\mkont}) \in \mathit{sys}.R$.
+ By IH, there is an $m$ such that for some 
+ $\mkont_o \in \tails{\mathit{sys}.\mktab}{\mkont^*}$,
+ $(\tpl{\mpoint^*,\mstore^*,\append{\mkont^*}{\mkont_o}},\tpl{\mpoint,\mstore,\append{\mkont}{\mkont_o}}) \in \mathit{unfold}(m)$.
+
+By cases on edge in $\mathit{sys}.R$ or in $\mathit{sys}'.R$ (and not in $\mathit{sys}.R$).
+\\
+Case in: $\mkont''$ is irrelevant because it's in the tail, and there must be some continuation in $\tails{\mathit{sys}.\mktab}{\mkont}$ by $\inv$.
+         Thus holds by IH.
+\\
+Case not in:
+  It must be that $\tpl{\mpoint,\mstore,\mkont}$ was the selected state, and that
+  $\tpl{\mpoint,\mstore,\mkont,\mathit{sys}.\mktab,\mathit{sys}.\mmemo} \stepto \tpl{\mpoint',\mstore',\mkont',\mktab',\mmemo'}$ where $\mktab' \sqsubseteq \mathit{sys}'.\mktab$ and $\mmemo' \sqsubseteq \mathit{sys}'.\mmemo$.
+
+  By cases on this reduction:
+\\
+  Case $\tpl{(\mvar,\menv),\mstore,\mkont,\_,\_} \stepto \tpl{\mval,\mstore,\mkont,\_,\_}$ where $\mval \in \mstore(\menv(\mvar))$:
+\\
+  Case $\tpl{((\mexpri0\ \mexpri1),\menv),\mstore,\mkont,\_,\_} \stepto \tpl{(\mexpri0,\menv),\mstore,\mkont,\_,\_}$:
+\\
+  Case $\tpl{\mval,\mstore,\kcons{\kar{\mexpr,\menv}}{\mkont},\_,\_} \stepto \tpl{(\mexpr,\menv),\mstore,\kcons{\kfn{\mval}}{\mkont},\_,\_}$:
+\\
+  Case $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} \stepto \tpl{\mctx,\krt{\mctx},\mktab',\mmemo}$
+  where $\maddr = \alloc^*(\mstate)$, $\menv' = \extm{\menv}{\mvar}{\maddr}$, $\mstore' = \joinone{\mstore}{\maddr}{\set{\mval}}$, $\mctx = ((\mexpr,\menv'),\mstore')$, and $\mctx \notin \dom(\mmemo)$:
+\\
+  Split on $\tpl{\mpoint,\mstore,\mkont}$ the initial state or
+           there is a $\tpl{\mpoint^*,\mstore^*,\mkont^*}$ such that $(\tpl{\mpoint^*,\mstore^*,\mkont^*}, \tpl{\mpoint,\mstore,\mkont}) \in \mathit{sys}.R$.
+\\
+  Case initial: $m$ is 0.
+\\
+  Case else:
+   Need to show there is a path using $\mkont''$ whereever $\tpl{\mpoint,\mstore,\mkont}$ is, and the length should be an upper bound for $m$.
+   This brings in the table invariants and the order invariant which allows us to get to the starting point for 
+   $\tpl{\mpoint^*,\mstore^*,\mkont^*}$ and replace with the latest frame to get to where we are.
+\end{proof}

### 2014-03-24 metatheory.tex
Proof for PDCFA finished?

@@
         By IH, there is an $m$ such that (let $\mathit{sys} = {\mathcal F}(\mstate_0)^m(\mathit{sys}_0)$) $(\mstate'',\mstate) \in \mathit{reify}(\mathit{sys})$.
+\\
         So $(\mstate'',\mstate) \equiv (\tpl{\mpoint, \mstore, \append{\mkont}{\mkont''}},\tpl{\mpoint',\mstore',\append{\mkont'}{\mkont''}})$ where $\mkont'' \in \tails{\mktab}{\mkont}$.
@@
          For $\mkont'$ case 2: $\mkont''$ must be unrolled from $\krt{\mctx} \in \mathit{sys}.\mktab$.
-                         Thus there is a $\mkont'{}^* \in \mathit{sys}.\mktab(\mctx)$ such that $\mkont'' = \mkont'$
+                         Thus there is a $\kcons{\mkframe}{\mkont^{**}} \in \mathit{sys}.\mktab(\mctx)$ such that $\mkont^* \in \unroll{\mathit{sys}.\mktab}{\mkont^{**}}$
                          \\
          Case $\mkframe\equiv\kar{\mexpr,\menv}$, $\mstate' \equiv \tpl{(\mexpr,\menv),\mstore',\mkont^*}$:
-          For $\mkont'$ case 1: $(\tpl{\mpoint',\mstore',\mkont'}, \tpl{(\mexpr,\menv),\mstore',\kcons{\kfn{\mval}}{\mkont^{**}}}) \in \mathit{sys}.R$, so $(\mstate,\mstate') \in \mathit{reify}(\mathit{sys})$.
+          By definition of ${\mathcal F}$, $(\tpl{\mpoint',\mstore',\mkont'}, \tpl{(\mexpr,\menv),\mstore',\kcons{\kfn{\mval}}{\mkont^{**}}}) \in \mathit{sys}.R$, so $(\mstate,\mstate') \in \mathit{reify}(\mathit{sys})$.
 \\
-          Case 2: $\mkont'{}^* \equiv \kcons{\kar{\mexpr,\menv}}{\mkont'{}^{**}}$ and
-          \\ $(\tpl{\mpoint',\mstore',\mkont'{}^*},\tpl{(\mexpr,\menv),\mstore',\kcons{\kfn{\mval}}{\mkont'{}^{**}}}) \in \mathit{sys'}.R$
-\\
-         Case $\mkframe\equiv\kfn{\vclo{\slam{x}{\mexpr}}{\menv}}$
-          Similar.
+         Case $\mkframe\equiv\kfn{\vclo{\slam{x}{\mexpr}}{\menv}}$: Similar.
 \\
@@
 \\
-Case else: it must be that there is a $(\tpl{\mpoint^*,\mstore^*,\mkont^*},\tpl{\mpoint,\mstore,\mkont}) \in \mathit{sys}.R$.
- By IH, there is an $m$ such that for some 
- $\mkont_o \in \tails{\mathit{sys}.\mktab}{\mkont^*}$,
+Case else: it must be that there is a state that steps to $\tpl{\mpoint,\mstore,\mkont}$, with the edge represented in $\mathit{sys}.R$.
+%
+ $(\tpl{\mpoint^*,\mstore^*,\mkont^*},\tpl{\mpoint,\mstore,\mkont}) \in \mathit{sys}.R$.
+%
+ By IH, for fixed arbitrary $\mkont_o \in \tails{\mathit{sys}.\mktab}{\mkont^*}$, there is an $m$ such that
  $(\tpl{\mpoint^*,\mstore^*,\append{\mkont^*}{\mkont_o}},\tpl{\mpoint,\mstore,\append{\mkont}{\mkont_o}}) \in \mathit{unfold}(m)$.
+ We know that the context of $\mkont$ must be in the domain of $\mathit{sys}.\mktab$ due to the invariant.
+\\
+\newcommand{\ctxof}{\mathit{ctx\text{-}of}}
+ \textbf{The intricate part of the proof:}
+ If $\mkont''$ uses any truncated continuations from $\mathit{sys}'.\mktab$ that are not in $\mathit{sys}.\mktab$, they must all be in the entry for the context in $\mkont$, and they must all be at the top of $\mkont''$ with no other truncated continuations from other contexts' entries (due to the order invariant).
+ In other words, $\mkont''= \append{\append{\mkont_0}{\append{\ldots}{\mkont_n}}}{\mkont_o}$ for some $\vec{\mkont} \in \mathit{sys}'.\mktab(\ctxof(\mkont))^*$ and $\mkont_o \in \tails{\mathit{sys}.\mktab}{\mkont}$.
+We mean $\ctxof$ to extract the context stored in a continuation's $\mathbf{rt}$.
+%
+If the continuation does not have an $\mathbf{rt}$, then $\mkont''$ must be $\kmt$, and the argument becomes trivial.
+ By induction on this list of $\vec{\mkont}$, we can use the invariants to build a path piecemeal.
+\\
+ Specifically we argue \emph{a fortiori} the following:
+\\
+ For all $\vec{\mkont} \in \mathit{sys}'.\mktab(\ctxof(\mkont))^*$,
+ all $(\mstate,\mstate') \in \reify(\mathit{sys}')$,
+ all $\mkont_o \in \tails{\mathit{sys}.\mktab}{\mkont}$,
+ and all $\mpoint^*,\mstore^*,\mkont^*,\mpoint',\mstore',\mkont',$ if $(\mstate,\mstate')$ equals
+\\
+$(\tpl{\mpoint^*,\mstore^*,\append{\mkont^*}{\append{\vec{\mkont}}{\mkont_o}}},\tpl{\mpoint',\mstore',\append{\mkont'}{\append{\vec{\mkont}}{\mkont_o}}})$ and
+\\
+ $\tails{\mathit{sys}'.\mktab}{\mkont^*} = \tails{\mathit{sys}'.\mktab}{\mkont}$, then there is some natural $m$, such that $(\mstate,\mstate') \in \mathit{unfold}(m)$.
+\\
+\textit{Proof.} The argument follows by induction on $\vec{\mkont}$:
+\\
+Base case ($\vec{\mkont}$ empty): We use the (outer) induction hypothesis with $\mkont_o$ to get an $m$.
+Then since $\mstate \stepto \mstate'$, we conclude $(\mstate,\mstate') \in \mathit{unfold}(m+1)$.
+\\
+Induction step ($\vec{\mkont} \equiv \mkont_0 : \vec{\mkont}'$):
+By the invariant and context irrelevance, there is a path $\mtrace\equiv\tpl{\ctxof(\mkont_0), \append{\vec{\mkont}}{\mkont_o}} \stepto^* \mstate_\mathit{base}$ where $\mstate_\mathit{base} = \tpl{\ctxof(\mkont),\append{\mkont_0}{\append{\vec{\mkont}}{\mkont_o}}}$.
+Also by the invariant, there is a path $\mtrace' \equiv \mstate_\mathit{base} \stepto^* \mstate$.
+We use the last edge of $\mtrace$ and the induction hypothesis to get an $m$ to unfold through the second path to show $(\mstate,\mstate') \in \mathit{unfold}(m + |\mtrace'| + 1)$.
+%We case split on whether $\mkont_0$ has a context equal to the context of $\mkont$.
 
-By cases on edge in $\mathit{sys}.R$ or in $\mathit{sys}'.R$ (and not in $\mathit{sys}.R$).
-\\
-Case in: $\mkont''$ is irrelevant because it's in the tail, and there must be some continuation in $\tails{\mathit{sys}.\mktab}{\mkont}$ by $\inv$.
-         Thus holds by IH.
-\\
-Case not in:
-  It must be that $\tpl{\mpoint,\mstore,\mkont}$ was the selected state, and that
-  $\tpl{\mpoint,\mstore,\mkont,\mathit{sys}.\mktab,\mathit{sys}.\mmemo} \stepto \tpl{\mpoint',\mstore',\mkont',\mktab',\mmemo'}$ where $\mktab' \sqsubseteq \mathit{sys}'.\mktab$ and $\mmemo' \sqsubseteq \mathit{sys}'.\mmemo$.
+% By cases on edge in $\mathit{sys}.R$ or in $\mathit{sys}'.R$ (and not in $\mathit{sys}.R$).
+% \\
+% Case in: $\mkont''$ is irrelevant because it's in the tail, and there must be some continuation in $\tails{\mathit{sys}.\mktab}{\mkont}$ by $\inv$.
+%          Thus holds by IH.
+% \\
+% Case not in:
+%   It must be that $\tpl{\mpoint,\mstore,\mkont}$ was the selected state, and that
+%   $\tpl{\mpoint,\mstore,\mkont,\mathit{sys}.\mktab,\mathit{sys}.\mmemo} \stepto \tpl{\mpoint',\mstore',\mkont',\mktab',\mmemo'}$ where $\mktab' \sqsubseteq \mathit{sys}'.\mktab$ and $\mmemo' \sqsubseteq \mathit{sys}'.\mmemo$.
 
-  By cases on this reduction:
-\\
-  Case $\tpl{(\mvar,\menv),\mstore,\mkont,\_,\_} \stepto \tpl{\mval,\mstore,\mkont,\_,\_}$ where $\mval \in \mstore(\menv(\mvar))$:
-\\
-  Case $\tpl{((\mexpri0\ \mexpri1),\menv),\mstore,\mkont,\_,\_} \stepto \tpl{(\mexpri0,\menv),\mstore,\mkont,\_,\_}$:
-\\
-  Case $\tpl{\mval,\mstore,\kcons{\kar{\mexpr,\menv}}{\mkont},\_,\_} \stepto \tpl{(\mexpr,\menv),\mstore,\kcons{\kfn{\mval}}{\mkont},\_,\_}$:
-\\
-  Case $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} \stepto \tpl{\mctx,\krt{\mctx},\mktab',\mmemo}$
-  where $\maddr = \alloc^*(\mstate)$, $\menv' = \extm{\menv}{\mvar}{\maddr}$, $\mstore' = \joinone{\mstore}{\maddr}{\set{\mval}}$, $\mctx = ((\mexpr,\menv'),\mstore')$, and $\mctx \notin \dom(\mmemo)$:
-\\
-  Split on $\tpl{\mpoint,\mstore,\mkont}$ the initial state or
-           there is a $\tpl{\mpoint^*,\mstore^*,\mkont^*}$ such that $(\tpl{\mpoint^*,\mstore^*,\mkont^*}, \tpl{\mpoint,\mstore,\mkont}) \in \mathit{sys}.R$.
-\\
-  Case initial: $m$ is 0.
-\\
-  Case else:
-   Need to show there is a path using $\mkont''$ whereever $\tpl{\mpoint,\mstore,\mkont}$ is, and the length should be an upper bound for $m$.
-   This brings in the table invariants and the order invariant which allows us to get to the starting point for 
-   $\tpl{\mpoint^*,\mstore^*,\mkont^*}$ and replace with the latest frame to get to where we are.
+%   By cases on this reduction:
+% \\
+%   Case $\tpl{(\mvar,\menv),\mstore,\mkont,\_,\_} \stepto \tpl{\mval,\mstore,\mkont,\_,\_}$ where $\mval \in \mstore(\menv(\mvar))$:
+% \\
+%   Case $\tpl{((\mexpri0\ \mexpri1),\menv),\mstore,\mkont,\_,\_} \stepto \tpl{(\mexpri0,\menv),\mstore,\mkont,\_,\_}$:
+% \\
+%   Case $\tpl{\mval,\mstore,\kcons{\kar{\mexpr,\menv}}{\mkont},\_,\_} \stepto \tpl{(\mexpr,\menv),\mstore,\kcons{\kfn{\mval}}{\mkont},\_,\_}$:
+% \\
+%   Case $\tpl{\mval,\mstore,\kcons{\kfn{\vclo{\slam{\mvar}{\mexpr}}{\menv}}}{\mkont},\mktab,\mmemo} \stepto \tpl{\mctx,\krt{\mctx},\mktab',\mmemo}$
+%   where $\maddr = \alloc^*(\mstate)$, $\menv' = \extm{\menv}{\mvar}{\maddr}$, $\mstore' = \joinone{\mstore}{\maddr}{\set{\mval}}$, $\mctx = ((\mexpr,\menv'),\mstore')$, and $\mctx \notin \dom(\mmemo)$:
+% \\
+%   Split on $\tpl{\mpoint,\mstore,\mkont}$ the initial state or
+%            there is a $\tpl{\mpoint^*,\mstore^*,\mkont^*}$ such that $(\tpl{\mpoint^*,\mstore^*,\mkont^*}, \tpl{\mpoint,\mstore,\mkont}) \in \mathit{sys}.R$.
+% \\
+%   Case initial: $m$ is 0.
+% \\
+%   Case else:
+%   \textbf{TODO:}
+%    Need to show there is a path using $\mkont''$ whereever $\tpl{\mpoint,\mstore,\mkont}$ is, and the length should be an upper bound for $m$.
+%    This brings in the table invariants and the order invariant which allows us to get to the starting point for 
+%    $\tpl{\mpoint^*,\mstore^*,\mkont^*}$ and replace with the latest frame to get to where we are.
 \end{proof}

### 2014-03-25 content.tex
Proof for PDCFA finished?
Fixed up (still contrived) example. Added Racket implementation. Small fix to reify and F.

@@
 We still present the definitions and proofs in a manner that should be well-positioned for mechanization.
+%
+The reader unfamiliar with or uninterested in correctness proofs should still be able to grok the technique and apply it with little difficulty.
 %
@@
 %
-if we use a finite allocation strategy, then we don't explore some non-deterministic paths that lead to different results for a function call.
+if we use a finite allocation strategy for example, then we don't explore some non-deterministic paths that lead to different results for a function call.
 %
@@
 We use a meta-function $\wn$, ``wide to narrow,'' to shift between the two representations of states.
+
+\paragraph{Technical aside} There are minor differences with PDCFA's algorithm due to our choice to process one state at a time.
+If we process the entire frontier in one step, then we need to have additional tables of ``changes to $\mktab$ and $\mmemo$'' that need to be matched up in the following step.
+Here ``matched up'' means if both tables have the same context in their domain, we must add the state that has the memoized result(s) with the saved continuation(s).
+If we don't, then some memoized results may not flow to newly reached calling contexts.
+This corresponds with PDCFA's worksets of edges.
 %
@@
 \begin{align*}
-    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge \mastate\nequiv\tpl{\mval,\mstore,\krt{\mctx}}} \\
+    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge (\mastate,\mastate') \text{ not a call using } \mmemo} \\
     \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \mktab', \_) \in I} \\
@@
 %
+The set of reductions does not include uses of the memo table since those do not correspond to true steps that the original machine takes.
+%
 We define $\pi_1$ and $\pi_2$ as the pointwise lifting of the left and right projections out of a pair, respectively.
@@
 
-\paragraph{Example:} Let's consider a monovariant allocation strategy (the address for each binding is the variable name itself, so $\menv$ is always the identity function $\iota$) to run the following example:
+\paragraph{Translation from Greek}
+ ${\mathcal F}$'s definition may be implemented with destructive updates to the components of $\mathit{sys}$ instead of the big joins and unions.
+%
+ The definition of $\stepto$ is useful for defining the semantics, but for an implementation we would use a
+ function \texttt{step} that takes a state and produces all next states.
+%
+ We give a completely self-contained implementation in Racket of ${\mathcal F}$ along with \texttt{step} in \autoref{fig:real-impl}.
+%
+ The only free variable is \texttt{alloc}, which can be instantiated with (for example)
+\\[2pt]
+ \includegraphics{calloc}
+\\
+ for concrete allocation, and
+\\[2pt]
+ \includegraphics{malloc}
+\\
+ for monovariant allocation.
+%
+\\[2pt]
+\begin{figure}
+  \includegraphics[scale=0.8]{impl}
+  \caption{Complete Racket implementation}
+\label{fig:real-impl}
+\end{figure}
+
+\paragraph{Example:} Let's consider a monovariant allocation strategy to run the following example.
+Note that monovariance means the address for each binding is the variable name itself, so $\menv$ is always the identity function. We therefore omit it.
 
 \begin{center}
-  \todo{fix example}
-  %\includegraphics[scale=0.5]{example0}
+  \includegraphics[scale=0.5]{example0}
 \end{center}
@@
 %
-Here $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$.
+Here $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))})$.
 %
@@
 \begin{figure}
-      % \begin{tabular}{rcccl}
-      %   $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$
-      %   & $\Rightarrow$ &
-      %   $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$
-      %   & $\Rightarrow$ &
-      %   $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
-      % \end{tabular}
-      % \begin{center}
-      % \begin{tabular}{|l|c|c|c|}
-      %   \hline
-      %   \# & $\mktab$ & $\mmemo$ \\
-      %   \hline
-      %   0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
-      %   1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
-      %   \hline
-      %   \end{tabular}
-      % \begin{tabular}{|l|c|}
-      %   \hline
-      %   \# & $\mctx$ \\
-      %   \hline
-      %   0  & $(({\tt x}, \iota), \mstore_1)$ \\
-      %   1  & $(({\tt x}, \iota), \mstore_3)$ \\
-      %   \hline
-      % \end{tabular}
-      % \begin{tabular}{|l|c|c|}
-      %   \hline
-      %   \# & $\mstore$ & $\mkont$ \\
-      %   \hline
-      %   0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\iota)}]$ & $\mkont_0$ \\
-      %   1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
-      %   2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)}, \iota)$ \\
-      %   3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
-      %   4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
-      %   \hline
-      % \end{tabular}
-      % \end{center}
-\todo{fix example}
-      \caption{Example analysis evaluation}
+  \centering
+      \begin{tabular}{c}
+        $\begin{array}{l}\includegraphics[scale=0.5]{example1}\end{array}$
+        \\ $\Downarrow$ \\
+        $\begin{array}{l}\includegraphics[scale=0.5]{example2}\end{array}$
+        \\ $\Downarrow$ \\
+        $\begin{array}{l}\includegraphics[scale=0.5]{example3}\end{array}$
+      \end{tabular}
+      \begin{center}
+% KTables and Memo tables
+      \begin{tabular}{|l|c|c|c|}
+        \hline
+        \# & $\mktab$ & $\mmemo$ \\
+        \hline
+        0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{(0,\mstore_1)}]$ \\
+        1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{(0,\mstore_3),(1,\mstore_3)}]$ \\
+        \hline
+        \end{tabular}
+% contexts
+      \begin{tabular}{|l|c|}
+        \hline
+        \# & $\mctx$ \\
+        \hline
+        0  & $({\tt x}, \mstore_1)$ \\
+        1  & $({\tt x}, \mstore_3)$ \\
+        \hline
+      \end{tabular}
+% stores and continuations
+      \begin{tabular}{|l|c|c|}
+        \hline
+        \# & $\mstore$ & $\mkont$ \\
+        \hline
+        0 & $[{\tt id} \mapsto \set{\slam{\mvar}{\mvar}}]$ & $\mkont_0$ \\
+        1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
+        2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)})$ \\
+        3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
+        4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
+        \hline
+      \end{tabular}
+      \end{center}
+      Where $\mkont_0 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}}$
+      \caption{Example evaluation}
       \label{fig:ex-eval}
 \end{figure}
-% %% FIXME stores in memo-table
-% \begin{enumerate}
-% \item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
-%     \mstore_0[\texttt{x} \mapsto \set{0}]$ and
-%     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
-%          $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
-%     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
-% \item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
-% \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
-% \item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
-%       $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
-%       (call this $\mkont_2$).}
-% \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
-%       (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
-% \item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
-% \item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
-% \end{enumerate}
+%% FIXME stores in memo-table
+\begin{enumerate}
+\item{\texttt{(id 0)} steps to \texttt{x} at continuation $\krt{\mctx_0}$ with store $\mstore_1$ and continuation table $\mktab_0$, still with empty memo table.}
+\item{\texttt{0} at $\krt{\mctx_0}$ steps to $0$ at $\mkont_0$ and new memo table $\mmemo_0$.}
+\item{\texttt{0} at $\mkont_0$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} with store now $\mstore_2$}
+\item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} with continuation $\mkont_2$.}
+\item{\texttt{(id 1)} steps to \texttt{x} with store $\mstore_3$ and and continuation table $\mktab_1$.}
+\item{\texttt{0} or \texttt{1} at $\krt{\mctx_1}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ with memo table $\mmemo_1$.}
+\item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
+\end{enumerate}
 
@@
 %
-Returns are un-stuck by an imperative summarization algorithm driving the state exploration.
+Returns are un-stuck by an external summarization algorithm driving the state exploration.
 %
@@
 
-\paragraph{Example:} revising the earlier example, we find that with the addition of stack frames, there is no confusion about \texttt{z}'s value either.
+\paragraph{Example:} revisiting the earlier example, we find that with the addition of stack frames, there is no confusion about \texttt{z}'s value either.
 %
@@
 
+\textbf{TODO:} address FLOPS reviewer's concerns regarding Cook's transformation and Jones' use for analysis.
+
 \section{Conclusion}

### 2014-03-25 metatheory.tex
Fixed up (still contrived) example. Added Racket implementation. Small fix to reify and F.

@@
              \tpl{\mpoint',\mstore',\append{\mkont'}{\mkont''}})}
-           {\\ &\phantom{= \{}(\tpl{\mpoint,\mstore,\mkont}, \tpl{\mpoint',\mstore',\mkont'}) \in R,
+           {\\ &\phantom{= \{}(\tpl{\mpoint^*,\mstore^*,\mkont^*}, \tpl{\mpoint',\mstore',\mkont'}) \in R,
+            \\ &\phantom{= \{}\text{if } \mpoint^*\equiv\mval \text{ and } \mkont^* \equiv \krt{\mctx} \text{ then }
+            \\ &\phantom{= \{else }\text{let } (\tpl{\mpoint,\mstore,\mkont},\tpl{\mpoint^*,\mstore^*,\mkont^*}) \in R,
+            \\ &\phantom{= \{}\text{else let } \tpl{\mpoint,\mstore,\mkont} = \tpl{\mpoint^*,\mstore^*,\mkont^*}
             \\ &\phantom{= \{}\mkont'' \in \tails{\mktab}{\mkont}} \\
@@
 
+Any return edge (\ie, the left state has a value and a $\krt{\mctx}$ continuation) is administrative, so the reification stutters those out by linking the predecessors of return states with their successors.
+%
 The $\mathit{unroll}$ set interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, and $\mathit{tails}$ is defined mutually to produce all continuations that could be unrolled from the tail of a truncated continuation.
@@
 \end{proof}
+
+To reiterate, this correctness proof is strong enough for any allocation strategy, including fresh (for concrete execution) and finite (for ``pushdown analysis'').
+%
+The finite case is special since the $\System$ space becomes finite, which means there is some $m$ such that ${\mathcal F}^m(\mastate_0)$ is a fixed point, by Kleene's fixed point theorem.
+%
+In other words, the fixed point soundly and completely represents all paths reachable from the initial state in the CESK machine with an unbounded stack but finite address space.
+%
+This form of the CESK machine has a shallow translation into a pushdown system, but we needn't go there.

### 2014-03-26 content.tex
More related work, spell checking.

@@
 The additional abstraction also detracts from the operational intuitions and simplicity of the methodology.
-%
-We still present the definitions and proofs in a manner that should be well-positioned for mechanization.
-%
-The reader unfamiliar with or uninterested in correctness proofs should still be able to grok the technique and apply it with little difficulty.
-%
-This paper has been partially mechanized in the Coq proof assistant, with proofs in the supplemental materials\footnote{\url{http://github.com/ianj/concrete-summaries} in \texttt{model.v}}.
+%%
+
+%%
+We present the definitions and proofs in a manner that should be well-positioned for mechanization.
+%
+The reader unfamiliar with or uninterested in correctness proofs should still be able to understand the technique and apply it with little difficulty.
+%
+This paper has been partially mechanized in the Coq proof assistant~\citep{Filliâtre97thecoq}, with proofs in the supplemental materials\footnote{\url{http://github.com/ianj/concrete-summaries} in \texttt{model.v}}.
 
@@
 
-CFA2 is an online version of a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with PDCFA's crucial ``$\epsilon$-closure graph''.
+CFA2 is an online version of a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with PDCFA's crucial ``$\epsilon$-closure graph.''
 %
@@
 %
-(1) the state space contains all function callsites to project out a control-flow graph, and %
+(1) the state space contains all function call-sites to project out a control-flow graph, and %
 (2) the worklist semantics of execution makes it possible to appeal to an incomplete memo table, so the saved stacks allow new results to flow to calls that used incomplete memo information.
@@
 
-\paragraph{Notation:} we make judicious use of the isomorphism between associations of tuples to spare the paper from excessive parentheses, even though our implementations are in a Lisp-descendent language.
+\paragraph{Notation:} we make judicious use of the isomorphism between associations of tuples to spare the paper from excessive parentheses, even though our implementations are in a Lisp-descendant language.
 
@@
 %
-All the extra components at a context $\mctx$ get stored in a separate table, $\mktab$, so that updates to the memo table will propagate to all callers that depend on the result of $\mctx$.
+All the extra components at a context $\mctx$ get stored in a separate table, $\mktab$.
+%
+Updates to the memo table will propagate to all callers that depend on the result of $\mctx$.
 %%
@@
 %
-We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores---this leads to a recursive data-structure that we are trying to avoid.
+We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores---this leads to a recursive data-structure that can cause the analysis to diverge.
 %
@@
   \mkont \in \Kont &::= \kmt \alt \krt{\mexpr,\menv,\mstore} \alt \kcons{\mkframe}{\mkont} \\
-  \mktab \in \KTab &= Expr \times \Env\times \Store \to \wp(\Kont) \\
+  \mktab \in \KTab &= \Expr \times \Env\times \Store \to \wp(\Kont) \\
   \mmemo \in \Memo &= \Expr \times \Env\times \Store \to \wp(\Value \times \Store)
 \end{align*}
-
-\paragraph{The role of summaries:} notice the additional $\Memo$ component.
-%
-A ``summary edge'' in CFA2 or equivalently an $\epsilon$-edge in PDCFA is an edge from the source of a push edge to the target of the matching pop edge.
-%
-``Matching'' here means there is a path through machine reductions that don't change the stack, or through summary edges.
-%
-In our model, we only ``push'' when we call a function, and we only ``pop'' when we return with a value.
-%
-There is an analogy to something more operational: summary edges embody memoization.
-%
-Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store (\todo{show in figure})
-%%
 
@@
 
-This semantics is only correct for fresh allocation strategies, which in turn make the reduction relation deterministic.
+\paragraph{The role of summaries:} notice the additional $\Memo$ component.
+%
+A ``summary edge'' in CFA2 or equivalently an $\epsilon$-edge in PDCFA is an edge from the source of a push edge to the target of the matching pop edge.
+%
+``Matching'' here means there is a path through machine reductions that don't change the stack, or through summary edges.
+%
+In our model, we only ``push'' when we call a function, and we only ``pop'' when we return with a value.
+%
+There is an analogy to something more operational: summary edges embody memoization.
+%
+Outside the pure world, functions don't simply return values --- they can also impact the outer environment.
+%
+In our setting of the CESK machine, we also memoize the state of the store.
+%
+Once we reach a function call for which we have witnessed a result, we can simply jump to that result (with the current continuation).
+%
+This is the second case of the function call rule in \autoref{fig:summary-semantics}
+%%
+
+\paragraph{A fly in the ointment} This semantics is only correct for fresh allocation strategies, which in turn make the reduction relation deterministic.
 %
@@
 We use a meta-function $\wn$, ``wide to narrow,'' to shift between the two representations of states.
+
+{
+\setlength{\abovedisplayskip}{0pt}
+\setlength{\belowdisplayskip}{4pt}
+\setlength{\abovedisplayshortskip}{0pt}
+\setlength{\belowdisplayshortskip}{8pt}
+\begin{align*}
+  \mastate \in \sa{State} &::= \tpl{\mexpr,\menv,\mstore,\mkont} \\
+  \System &= \wp(\sa{State}) \times \wp(\sa{State} \times \sa{State}) \times \KTab \times \Memo \\
+  {\mathcal F} &: \sa{\State} \to \System \to \System
+\end{align*}
+\begin{align*}
+  {\mathcal F}(\mastate_0)(\emptyset, R, \mktab, \mmemo) &= (\emptyset, R, \mktab, \mmemo) \\
+  {\mathcal F}(\mastate_0)(F\uplus\set{\mastate}, R, \mktab, \mmemo) &= (F\cup F', R \cup R', \mktab', \mmemo') \\
+  \text{where } I &= \setbuild{\mstate'}{\wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
+                \wn(\tpl{\mexpr,\menv,\mstore,\mkont},\mktab,\mmemo)
+                   &= \tpl{\mexpr,\menv,\mstore,\mkont,\mktab,\mmemo}
+\end{align*}
+\begin{align*}
+    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge (\mastate,\mastate') \text{ not a call using } \mmemo} \\
+    \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \mktab', \_) \in I} \\
+    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \mmemo') \in I} \\
+    F' &= (\setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \cup \set{\mastate_0})\setminus (\pi_1 R \cup \pi_2 R)
+  \end{align*}
+}
+
+A $\System$ embodies the reductions seen disregarding the administrative pop steps, $R$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
+%
+The set of reductions does not include uses of the memo table since those do not correspond to true steps that the original machine takes.
+%
+We define $\pi_1$ and $\pi_2$ as the point-wise lifting of the left and right projections out of a pair, respectively.
+%
+The least fixed-point of ${\mathcal F}(\tpl{\mexpr,\bot,\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
+%
+Note the ordering on systems is point-wise, so the bottom element is $(\emptyset,\emptyset,\bot,\bot)$, which we'll call $\mathit{sys}_0$.
+%
+Systematic techniques for a performant implementation can be found in \citet{ianjohnson:oaam:icfp2013}.
+%
 
@@
 This corresponds with PDCFA's worksets of edges.
-%
-
-{
-\setlength{\abovedisplayskip}{0pt}
-\setlength{\belowdisplayskip}{4pt}
-\setlength{\abovedisplayshortskip}{0pt}
-\setlength{\belowdisplayshortskip}{8pt}
-\begin{align*}
-  \mastate \in \sa{State} &::= \tpl{\mexpr,\menv,\mstore,\mkont} \\
-  \System &= \wp(\sa{State}) \times \wp(\sa{State} \times \sa{State}) \times \KTab \times \Memo \\
-  {\mathcal F} &: \sa{\State} \to \System \to \System
-\end{align*}
-\begin{align*}
-  {\mathcal F}(\mastate_0)(\emptyset, R, \mktab, \mmemo) &= (\emptyset, R, \mktab, \mmemo) \\
-  {\mathcal F}(\mastate_0)(F\uplus\set{\mastate}, R, \mktab, \mmemo) &= (F\cup F', R \cup R', \mktab', \mmemo') \\
-  \text{where } I &= \setbuild{\mstate'}{\wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
-                \wn(\tpl{\mexpr,\menv,\mstore,\mkont},\mktab,\mmemo)
-                   &= \tpl{\mexpr,\menv,\mstore,\mkont,\mktab,\mmemo}
-\end{align*}
-\begin{align*}
-    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge (\mastate,\mastate') \text{ not a call using } \mmemo} \\
-    \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \mktab', \_) \in I} \\
-    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \mmemo') \in I} \\
-    F' &= (\setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \cup \set{\mastate_0})\setminus (\pi_1 R \cup \pi_2 R)
-  \end{align*}
-}
-%%
-
-%%
-A $\System$ embodies the reductions seen disregarding the administrative pop steps, $R$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
-%
-The set of reductions does not include uses of the memo table since those do not correspond to true steps that the original machine takes.
-%
-We define $\pi_1$ and $\pi_2$ as the pointwise lifting of the left and right projections out of a pair, respectively.
+%%
+
 % Note: Only if widened store.{
@@
 %The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.}
-%
-The least fixed-point of ${\mathcal F}(\tpl{\mexpr,\bot,\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
-%
-Note the ordering on systems is pointwise, so the bottom element is $(\emptyset,\emptyset,\bot,\bot)$, which we'll call $\mathit{sys}_0$.
-%
-Systematic techniques for a performant implementation can be found in \citet{ianjohnson:oaam:icfp2013}.
 %
@@
 
-\paragraph{Example:} Let's consider a monovariant allocation strategy to run the following example.
+\paragraph{Example:} Let's consider a monovariant allocation strategy to run the following example\footnote{The example is first-order for brevity, not technical difficulty.}.
 Note that monovariance means the address for each binding is the variable name itself, so $\menv$ is always the identity function. We therefore omit it.
@@
 The example evaluation is in \autoref{fig:ex-eval}.
-%
-Here $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))})$.
 %
@@
 %
-The ``reaches'' metafunction ($\reaches$) is a transitive closure of ``touches'' as it weaves through the store, starting from some root set of addresses:
+The ``reaches'' meta-function ($\reaches$) is a transitive closure of ``touches'' as it weaves through the store, starting from some root set of addresses:
 
@@
 %
-More live addresses means a less effective GC, but the introspective pushdown analysis chose this tradeoff for a possibly smaller state space.
+More live addresses means a less effective GC, but the introspective pushdown analysis chose this trade-off for a possibly smaller state space.
 %
@@
 \end{enumerate}
-There are two orthogonal features of the semantics, embodied by our semantics' two metafunctions ($\bind$ and $\lookup$, respectively in \autoref{fig:frame-semantics}).
+There are two orthogonal features of the semantics, embodied by our semantics' two meta-functions ($\bind$ and $\lookup$, respectively in \autoref{fig:frame-semantics}).
 
@@
 %
+The refinement simply removes the local stack frame from the state and $\mathbf{push}$ frames from the stack.
+%
 %That is, if we always allocated bindings in the heap.
@@
 
-The two metafunctions, $\domark$ and $\domarkset$ are what do the heavy-lifting of adding marks to a continuation and reifying the mark set into a map of keys to lists.
+The two meta-functions, $\domark$ and $\domarkset$ are what do the heavy-lifting of adding marks to a continuation and reifying the mark set into a map of keys to lists.
 %
@@
 
-Second, the $\domarkset$ metafunction builds a map of keys to lists, and builds a private store for the list constructions.
+Second, the $\domarkset$ meta-function builds a map of keys to lists, and builds a private store for the list constructions.
 %
@@
 
-\autoref{fig:shift-reset-table0} has what one might naturally write using just this information.
-%
-\paragraph{Problem:} now captured continuations can appear in the store, and $\rt$ continuations contain a store.
-%
-This introduces an unbounded data structure that can cause the analysis to never terminate.
-
-\paragraph{Solution:} we allow for a tunable approximation by using $\alloc$.
-%
-Captured continuations' store component in the $\rt$ context gets replaced with an address of possible stores that it would approximate.
-
 \begin{figure*}
@@
 
+\autoref{fig:shift-reset-table0} has what one might naturally write using just this information.
+%
+\paragraph{Problem:} now captured continuations can appear in the store, and $\rt$ continuations contain a store.
+%
+This introduces an unbounded data structure that can cause the analysis to never terminate.
+
+\paragraph{Solution:} we allow for a tunable approximation by using $\alloc$.
+%
+Captured continuations' store component in the $\rt$ context gets replaced with an address of possible stores that it would approximate.
+
 The new indirection possibility changes $\Context$ to also include $\maddr$ where there was previously a $\mstore$, and the approximations of continuations live in the store is relevant to contexts now.
@@
 %
-This operation makes lookups more costly since we have to find closures bounded above by the current closure, rather than one exact closure.
+This operation makes look-ups more costly since we have to find closures bounded above by the current closure, rather than one exact closure.
 
@@
 
+\paragraph{Notice} implementers may skip to the next bolded subsection.
+%%
+
+%%
 This theorem amounts to just the first bullet of \autoref{thm:concrete-tabular}.
@@
 
+\paragraph{Pushdown models and memoization}
+The idea of relating pushdown automata with memoization is not new.
+%
+In 1971, Stephen Cook~\citep{DBLP:conf/ifip/Cook71} devised a transformation to simulate 2-way (on a fixed input) \emph{deterministic} pushdown automata in time linear in the size of the input, that uses the same ``context irrelevance'' idea to skip from state $q$ seen before to a corresponding first state that produced a smaller stack than $q$ was seen with.
+%
+Such a state is an instance of what are called \emph{terminator} states.
+%
+A \emph{terminator} state is simply a state that performs a pop operation.
+%
+Six years later, Neil D.
+Jones\citep{Jones:1977:NLT} simplified the transformation instead to \emph{interpret} a stack machine program to work \emph{on-the-fly} still on a deterministic machine, but with the same idea of using memo tables to remember corresponding terminator states.
+%
+Thirty-six years after that, at David Schmidt's Festschrift, Robert Gl\"uck extended the technique to two-way \emph{non-deterministic} pushdown automata, and showed that the technique can be used to recognize context-free languages in the standard ${\mathcal O}(n^3)$ time~\citep{DBLP:journals/corr/Gluck13}.
+%
+Gl\"uck's technique (as yet, correctness unproven) uses the meta-language's stack with a deeply recursive interpretation function to preclude the use of a frontier and something akin to $\mktab$\footnote{See \texttt{gluck.rkt} in supplementary materials for \autoref{fig:real-impl} in Gl\"uck's style}.
+%
+By exploring the state space \emph{depth-first}, the interpreter can find all the different terminators a state can reach one-by-one by destructively updating the memo table with the ``latest'' terminator found.
+%
+The trade-offs with this technique are that it does not obviously scale to first-class control, and the search can overflow the stack when interpreting moderate-sized programs.
+%
+We have not performed an extensive evaluation to test the latter point, however.
+%
+A minor disadvantage is that it is also not a fair evaluation strategy when allocation is unbounded.
+%
+The technique can nevertheless be a viable alternative for languages with simple control-flow mechanisms.
+%
+It has close similarities to ``Big-CFA2'' in Vardoulakis' dissertation~\citep{vardoulakis-diss12} (there is no detection of infinite loops, which may simply be an oversight in the dissertation).
+%%
+
+%%
+\paragraph{CFA2 and PDCFA}
 The immediately related work is that of PDCFA \citep{dvanhorn:Earl2010Pushdown, dvanhorn:Earl2012Introspective}, CFA2~\citep{ianjohnson:vardoulakis-lmcs11, ianjohnson:Vardoulakis2011Pushdown}, and AAM~\citep{dvanhorn:VanHorn2010Abstracting}, the first two of which we recreated in full detail.
@@
 %
-Our semantics adapts to \rackett{call/cc} by removing the meta-continuation operations, and thus this work supercedes theirs.
+Our semantics adapts to \rackett{call/cc} by removing the meta-continuation operations, and thus this work supersedes theirs.
 %
@@
 %%
+\paragraph{Stack inspection}
 Stack inspecting flow analyses also exist, but operate on pre-constructed regular control-flow graphs~\citep{ianjohnson:bartoletti2004stack}, so the CFGs cannot be trimmed due to the extra information at construction time, leading to less precision.
@@
 %%
+\paragraph{Analysis of pushdown automata}
 Pushdown models have existed in the first-order static analysis literature~\citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}\citep{ianjohnson:reps:pushdown:1995}, and the first-order model checking literature \citep{ianjohnson:bouajiani:esparza:pushdown:1997}, for some time.
@@
 The important difference when we move higher-order is that the model construction to feed these methods is an additional problem---the one we solve here.
-%%
-
-%%
+%
+Additionally, the algorithms employed in these works expect a complete description of the model up front, rather than work with a modified \texttt{step} function (also called \texttt{post}), such as in ``on-the-fly'' model-checking algorithms for finite state systems~\citep{DBLP:conf/tacas/SchwoonE05}.
+%%
+
+%%
+\paragraph{Derivation from abstract machines}
 The trend of deriving static analyses from abstract machines does not stop at flow analyses.
@@
 
-\textbf{TODO:} address FLOPS reviewer's concerns regarding Cook's transformation and Jones' use for analysis.
-
 \section{Conclusion}
@@
 %
+We make a modest attempt to re-explain the techniques in a common framework and extend it in a novel and instructive way.
+%
 It is our view that systematic approaches to applying the techniques are pivotal to scaling them to production programming languages.
 %
-We showed how to systematically construct executable concrete semantics that pointwise abstract to known algorithms for pushdown analyses of higher-order languages.
+We showed how to systematically construct executable concrete semantics that point-wise abstract to known algorithms for pushdown analyses of higher-order languages.
 %
@@
 % Greater control of the stack gives us the ability to drill deep into context-sensitive security properties and make precise predictions.
+
+% LocalWords:  pushdown automata CFA PDCFA AAM dvanhorn VanHorn Coq
+% LocalWords:  formalisms formalizations AST HORS interconnectedness
+% LocalWords:  analytical Bandera ianjohnson summarization Sharir PLT
+% LocalWords:  PDCFA's composable Redex memoize memoizing CESK tuples
+% LocalWords:  memoizes worklist isomorphism memoization indirecting
+% LocalWords:  amongst performant memoized monovariant monovariance
+% LocalWords:  GC reachability tunable storeable computable
